### 2018-12-03 range-shift-lit.Rmd
modify csl for "inline" annotations and make range-shift-lit into .Rmd

---
title: "Range shift literature review"
author: Philina English
date: "December 3, 2018"
bibliography: pbs-rangeshifts.bib
csl: apa-annotated-bibliograph-annotation-inline.csl 
output: pdf_document
---

```{r}

```

[@parmesan_globally_2003]

[@thorson_model-based_2016]

[@pinsky2013]

# References