### 2018-12-03 code_notes.Rmd
update code_notes

---
title: "Code Notes"
author: "Philina English"
date: "October 24, 2018"
output: html_document
---

# installing key packages

```{r}
# install.packages("tidyverse")
# install.packages(c("devtools", "roxygen2", "testthat", "knitr"))
# install.packages("devtools")
# library(devtools)
# devtools::install_github("nwfsc-assess/geostatistical_delta-GLMM")
# install.packages("INLA", repos=c(getOption("repos"), INLA="https://inla.r-inla-download.org/R/stable"), dep=TRUE)
```

# ggplot2

```{r}
library(ggplot2)
 + coord_flip() # switches the x and y axes eg. if you want horizontal boxplots, long labels that are hard to get them to fit without overlapping on x-axis
 + coord_quickmap()

```

# Bibliography

```{r}
library(rbbt)
bbt_cite()
bbt_bib_zotero()
```

```{r}
library(RefManageR)
biblio <- ReadZotero(user = 863479, .params = list(collection = 'items'))
```

```{r eval=False}
library(tidyverse)
tidyverse_update()
```