### 2018-12-04 range-shift-lit.Rmd
add a couple ref from Sunday's paper

---
title: "Range shift literature review"
author: Philina English
date: "December 3, 2018"
bibliography: pbs-rangeshifts.bib
csl: apa-annotated-bibliograph-annotation-inline.csl 
output: html_document
---

[@parry_climate_2007]

[@burrows2011]

[@parmesan_globally_2003]

[@poloczanska2013]

[@thorson_model-based_2016]

[@pinsky2013]

# References