### 2018-12-05 range-shift-lit.Rmd
interim commit on lit review notes

---
title: "Range shift literature review"
author: Philina English
date: "December 3, 2018"
bibliography: pbs-rangeshifts.bib
csl: apa-annotated-bibliograph-annotation-inline.csl 
output: html_document
---

##Changing climate
[@parry_climate_2007]

### Changes in the marine environment
[@burrows2011]

##--> Many organisms must "adapt, move, or die"
--when the environment changes, if they are to persist, organisms must be able to cope with change, change themselves, or move ([@maggini_are_2011])

[@root_fingerprints_2003]

###When species move: shifting distributions
####General patterns 
--many terrestrial species from a range of taxanomic groups have moved to higher latitudes and elevations ([@hickling_distributions_2006])

--despite similar climate velocities, within regions, some species have rapidly extended their geographical distributions, while others have moved less, or have even moved in the opposite direction

[@parmesan_globally_2003]

[@poloczanska2013]

[@sunday_species_2015]

- but climate velocity was found to explain shifts better than species traits for a range of marine taxa ([@pinsky2013]); however methods used in this study may not have accounted fully for changes in the spatial distribution of survey effort in all cases ([@thorson_model-based_2016])

###Stages of distribution change and the importance of species traits
[@bates_defining_2014]

[@poyry_species_2009]
###Modeling responses to environmental gradients
[@maggini_are_2011]

[@pinsky2013]

# References