### 2018-12-05 range-shift-lit.Rmd
more citations added

---
title: "Range shift literature review"
author: Philina English
date: "December 3, 2018"
bibliography: pbs-rangeshifts.bib
csl: apa-annotated-bibliograph-annotation-inline.csl 
output: html_document
---

##Changing climate
[@parry_climate_2007]

### Changes in the marine environment
[@burrows2011]

##In response organisms "adapt, move, or die"
--when the environment changes, if they are to persist, organisms must be able to cope with change, change themselves, or move ([@maggini_are_2011])

[@root_fingerprints_2003]

###When species move: shifting distributions
####General patterns 
--many terrestrial species from a range of taxanomic groups have moved to higher latitudes and elevations ([@hickling_distributions_2006,@chen_rapid_2011])

--despite similar climate velocities, within regions, some species have rapidly extended their geographical distributions, while others have moved less, or have even moved in the opposite direction

[@parmesan_globally_2003]

[@poloczanska2013]

[@sunday_species_2015]

--even statistically and biologically significant models based on species traits can have limited exlanatory power and therefore conservation value ([@angert_species_2011])

###Stages of distribution change and the importance of species traits
[@bates_defining_2014]

[@poyry_species_2009]

###Modeling responses to environmental change

--distances moved were correlated with warming and generally sufficient to track temperatures latitudinally, but not elevationally [@chen_rapid_2011]

--climate velocity was found to explain shifts better than species traits for a range of marine taxa ([@pinsky2013]); however methods used in this study may not have accounted fully for changes in the spatial distribution of survey effort in all cases ([@thorson_model-based_2016])

--sea surface temperature was not correlated with range shifts [@przeslawski_using_2012]

[@maggini_are_2011]

# References