### 2019-01-22 manual-geostat-TMB.Rmd
annotate manual test

---
title: "manual-geostat-TMB"
author: "Philina English"
date: "1/21/2019"
output: html_document
---

#### Load packages

```{r}
library(dplyr)
library(gfplot)
library(ggplot2)
```

#### Retrieve data
```{r}
survey <- "SYN QCS"

d <- gfplot::get_survey_sets("pacific cod", ssid = 1)
 # d <- dplyr::filter(d, !(year == 2014 & survey_abbrev == "SYN WCHG")) # something went wrong in 2014

```

#### Tidy data and fill in missing depth values
```{r}

col <- if (grepl("SYN", survey)) "density_kgpm2" else "density_ppkm2" # grep says if SYN is in survey returns logical (l), use kgpm2

dat <- gfplot:::tidy_survey_sets(d, survey, years = seq(1, 1e6), # say which surveys and years to include
    density_column = col) # make utm = x,y rename density, remove

#glimpse(dat)
  if (mean(dat$present) < 0.05) stop("Not enough data.") # stops if too little data

#  .scale <- if (grepl("SYN", survey)) 1000 else 1 # for computational stability # may be needed for INLA
#  dat <- dplyr::mutate(dat, density = density * .scale) # may be needed for INLA

  if (any(is.na(dat$depth)))
    dat <- gfplot:::interp_survey_bathymetry(dat)$data
    dat <- gfplot:::scale_survey_predictors(dat) # choose interp data if needed, and scale
```

#### Make prediction grid for plotting
```{r}
#  if (grepl("SYN", survey)) {
    grid_locs <- gfplot:::make_prediction_grid(
      dplyr::filter(dat, year == max(dat$year)), survey = survey,
      cell_width = 2)$grid # cell width of 2 km
    grid_locs <- dplyr::rename(grid_locs, depth = akima_depth)
#  } else {
#    grid_locs <- if (surv == "HBLL OUT N") gfplot::hbll_n_grid$grid else gfplot::hbll_s_grid$grid
 # }
  grid_locs$year <- NULL

#  formula <- if (include_depth) {
#    stats::as.formula(density ~ 0 + as.factor(year))
#  } else {
#    stats::as.formula(density ~ 0 + as.factor(year) + depth_scaled + depth_scaled2)
# }
```

#### Spatial model
```{r}
  spde <- sdmTMB::make_spde(dat$X, dat$Y, n_knots = 100)

  m <- sdmTMB::sdmTMB(
            formula = density ~ 0 + as.factor(year) + depth_scaled + depth_scaled2,
            data = dat, time = "year", 
            spde = spde, 
            family = sdmTMB::tweedie(link = "log"),
            anisotropy = FALSE, 
            silent = FALSE)

  names(m) 
```

#### Save preditions from model
```{r}
  predictions <- stats::predict(m, newdata = grid_locs)$data
# predictions$resids <- residuals(m) # randomized quantile residuals but not working
  glimpse(predictions) # est = predicted values

# A short function for plotting our predictions:
plot_map <- function(dat, column = "est") {
  ggplot(dat, aes_string("X", "Y", fill = column)) +
    geom_raster() +
    facet_wrap(~year) +
    coord_fixed()
}
```

#### Check residuals
```{r}
#ggplot(predictions, aes(X, Y, col = resids)) + scale_colour_gradient2() +
#  geom_point() + facet_wrap(~year)

#hist(predictions$resids)
#qqnorm(predictions$resids);abline(a = 0, b = 1)
```

#### Plot overall model
```{r}
plot_map(predictions, "exp(est)") +
  scale_fill_viridis_c(trans = "sqrt") +
  ggtitle("Prediction (fixed effects + all random effects)")
```

#### Plot fixed effects
```{r}

plot_map(predictions, "exp(est_fe)") +
  ggtitle("Prediction (fixed effects only)") +
  scale_fill_viridis_c(trans = "sqrt")
```

#### Plot spatial random effects
```{r}
plot_map(predictions, "est_re_s") +
  ggtitle("Spatial random effects only") +
  scale_fill_gradient2()
```

#### Plot spatiotemporal random effects
```{r}
plot_map(predictions, "est_re_st") +
  ggtitle("Spatiotemporal random effects only") +
  scale_fill_gradient2()

```