### 2019-02-20 analysis/m_tweedie_TMB.Rmd
mixed effect tweedie in TMB

---
title: "Mixed effect tweedie model in TMB"
output: html_notebook
---

```{r}
# use tmb.template snippet
```

```{r set-up}
## Load TMB 
library(TMB)
library(tweedie)

#library(statmod)

# Make C++ file
# TMB::template("m_tweedie_tmb.cpp")

## Compile and load the model
compile("src/m_tweedie_tmb.cpp")
dyn.load(dynlib("src/m_tweedie_tmb"))
```

# Simulate data with a tweedie distribution and mixed-effects...

```{r}
# Simulate predictors

n_groups = 10
n_within = 10
n_total = n_groups*n_within

group_sd = 1

group <- rep( 1:n_groups, each = n_within)
group_beta <- rnorm(length(unique(group)), mean = 0, sd = group_sd)

x <- runif(n_total, -2, 2)
a <- 3
b <- 1

power <- 1.5 # must be between 1 and 2 for compound poisson-gamma
phi <- 1 # dispersion must be positive
eta <- a + group_beta[group] + b * x 
mu <- exp(eta)

# Simulate response
y = rtweedie(n_total, power = power, mu = mu, phi = phi)

d = data.frame(y, x, group)

# plot simulated data 
plot(x, y)
hist(y, breaks=30)
```

# plot data by group to see if it looks right
```{r}
library(ggplot2)
library(viridis)

ggplot(d, aes(x, y, group = as.factor(group), color = as.factor(group))) + geom_point() + geom_smooth(method = 'loess', formula = 'y ~ x', span = 3, se = FALSE) + scale_color_viridis_d()
```

```{r}
## Data and parameters
data <- list("g_i" = group-1, "y_i" = y, "x" = x)
parameters <- list("a" = 0, "b" = 0, "log_phi" = log(1), "p" = 0, "log_SDZ"= 2, "z_g" = rep(0, length(unique(group))))
random = c("z_g")
#if(REML==TRUE ) Random = union( Random, "a")

## Make a function object
obj <- MakeADFun(data, parameters, random = random, DLL="m_tweedie_tmb")

## Call function minimizer
opt <- nlminb(obj$par, obj$fn, obj$gr)

## Get parameter uncertainties and convergence diagnostics
sdr <- sdreport(obj)
sdr
```

```{r}
## Get likelihood profiles for individual parameters:
a_profile <- tmbprofile(obj, "a")
b_profile <- tmbprofile(obj, "b")
confint(a_profile)
confint(b_profile)

plot(a_profile)
plot(b_profile)
```