### 2019-03-13 analysis/tmb-sensor-explore/3-fit-do-salinity.Rmd
Switch R files to Rmd to save the plots for a presentation

# Fit spatiotemporal models to dissolved oxygen and salinity

```{r}
library(sdmTMB)
library(tidyverse)
d_fit <- readRDS(here::here("analysis/tmb-sensor-explore/sensor-data-processed.rds"))
d_fit$log_do_mlpl <- log(d_fit$do_mlpl)

d <- filter(d_fit, !is.na(do_mlpl), !is.na(salinity_psu))
spde <- make_spde(d$X, d$Y, n_knots = 150)
unique(d$year)
plot_spde(spde)

hist(d$do_mlpl)
hist(log(d$do_mlpl))
min(d$do_mlpl)

m_do1 <- sdmTMB(d, log_do_mlpl ~ 1 + depth_scaled,
  time = "year", spde = spde, family = gaussian(link = "identity"), silent = TRUE)

m_do1$model$par
exp(m_do1$model$par[["ln_phi"]])

get_diag <- function(m, response = "log_do_mlpl") {
  predictions1 <- predict(m)
  predictions1$data$residuals <- residuals(m)

  plot_map <- function(dat, column = "est") {
    ggplot(dat, aes_string("X", "Y", colour = column)) +
      geom_point() +
      coord_fixed()
  }

  g <- plot_map(predictions1$data, "est") +
    scale_colour_viridis_c() +
    ggtitle("Prediction (fixed effects + all random effects)")
  print(g)

  g <- plot_map(predictions1$data, "est_fe") +
    ggtitle("Prediction (fixed effects only)") +
    scale_colour_viridis_c()
  print(g)

  g <- plot_map(predictions1$data, "est_re_s") +
    ggtitle("Spatial random effects only") +
    scale_colour_gradient2()
  print(g)

  g <- ggplot(predictions1$data, aes_string("est", response)) +
    geom_point() +
    facet_wrap(~year) +
    coord_fixed() +
    geom_abline()
  print(g)

  g <- ggplot(predictions1$data, aes(est, residuals)) +
    geom_point() +
    geom_smooth()
  print(g)

  g <- ggplot(predictions1$data, aes(depth_scaled, residuals)) +
    geom_point() +
    geom_smooth() +
    facet_wrap(~year)
  print(g)

  g <- ggplot(predictions1$data, aes(X, Y, colour = residuals)) +
    geom_point() +
    coord_fixed() +
    scale_color_gradient2()
  print(g)

  aic <- function(m) {
    k <- length(m$model$par)
    nll <- m$model$objective
    2 * k - 2 * (-nll)
  }

  r2 <- cor(predictions1$data$est, predictions1$data$log_do_mlpl)^2
  print(r2)

  print(aic(m))
}

get_diag(m_do1)

m_do2 <- sdmTMB(d, log_do_mlpl ~ 1 + poly(depth_scaled, 2),
  time = "year", spde = spde, family = gaussian(link = "identity"), silent = TRUE)
get_diag(m_do2)

m_do3 <- sdmTMB(d, log_do_mlpl ~ 1 + poly(depth_scaled, 3),
  time = "year", spde = spde, family = gaussian(link = "identity"), silent = TRUE)
get_diag(m_do3)

.grid <- sdmTMB::qcs_grid %>% filter(year == 2017)
predictions3_nd <- predict(m_do3, newdata = .grid)

plot_map <- function(dat, column = "est") {
  ggplot(dat, aes_string("X", "Y", fill = column)) +
    geom_raster() +
    facet_wrap(~year) +
    coord_fixed()
}

plot_map(predictions3_nd$data, "exp(est)") +
  scale_fill_viridis_c(option = "C") +
  ggtitle("Prediction (fixed effects + all random effects)") +
  geom_point(data = d, aes(X, Y, size = do_mlpl), inherit.aes = FALSE, pch = 21) +
  scale_size_area(max_size = 10)

plot_map(predictions3_nd$data, "est_fe") +
  ggtitle("Prediction (fixed effects only)") +
  scale_fill_viridis_c(trans = "sqrt")

plot_map(predictions3_nd$data, "est_re_s") +
  ggtitle("Spatial random effects only") +
  scale_fill_gradient2()

# Salinity:

hist(d$salinity_psu, breaks = 40)
hist(log(d$salinity_psu))
hist(exp(d$salinity_psu), breaks = 40)

m_salinity3 <- sdmTMB(d, salinity_psu ~ 1 + poly(depth_scaled, 3),
  time = "year", spde = spde, family = gaussian(link = "identity"), silent = TRUE)
get_diag(m_salinity3, response = "salinity_psu")

m_salinity4 <- sdmTMB(d,
  salinity_psu ~ 1 + depth_scaled * log_do_mlpl,
  time = "year", spde = spde, family = gaussian(link = "identity"), silent = TRUE)
get_diag(m_salinity4, response = "salinity_psu")

m_salinity5 <- sdmTMB(d,
  salinity_psu ~ 1 + poly(depth_scaled, 3) * log_do_mlpl,
  time = "year", spde = spde, family = gaussian(link = "identity"), silent = TRUE)
get_diag(m_salinity5, response = "salinity_psu")

m_salinity5 <- sdmTMB(d,
  salinity_psu ~ 1 + poly(depth_scaled, 3) * temperature_c,
  time = "year", spde = spde, family = gaussian(link = "identity"), silent = TRUE)
get_diag(m_salinity5, response = "salinity_psu")

m_salinity5 <- sdmTMB(d,
  salinity_psu ~ 1 + poly(depth_scaled, 3) * log_do_mlpl + temperature_c,
  time = "year", spde = spde, family = gaussian(link = "identity"), silent = TRUE)
get_diag(m_salinity5, response = "salinity_psu")

# much harder to fit
```