### 2019-03-13 analysis/tmb-sensor-explore/4-mountain-rw-examples.Rmd

# Fit time-varying quadratic models with depth and temperature

```{r}
library(dplyr)
library(ggplot2)
survey <- "SYN QCS"

fit_mountains <- function(survey_set_data, type = c("temperature", "depth"), n_knots = 100) {

  type <- match.arg(type)

  d <- survey_set_data
  # sd <- gfplot::get_sensor_data_trawl(ssid = 1, spread_attributes = FALSE)
  sd <- readRDS(here::here("analysis/tmb-sensor-explore/dat-sensor-trawl.rds"))
  sd <- sd %>%
    select(-count, -start_time, -end_time, -min, -max) %>%
    mutate(attribute = tolower(attribute)) %>%
    distinct() %>%
    reshape2::dcast(fishing_event_id + year + ssid + survey_desc ~ attribute, value.var = "avg")

  col <- if (grepl("SYN", survey)) "density_kgpm2" else "density_ppkm2"
  dat <- gfplot:::tidy_survey_sets(d, survey, years = seq(1, 1e6),
    density_column = col)

  dat <- left_join(dat, select(sd, temperature_c, year, fishing_event_id), 
    by = c("year", "fishing_event_id")) %>%
    rename(temperature = temperature_c)

  nrow(dat)
  dat <- filter(dat, !is.na(temperature), !is.na(depth))
  nrow(dat)

  spde <- sdmTMB::make_spde(dat$X, dat$Y, n_knots = n_knots)

  dat <- mutate(dat,
    temperature_mean = mean(temperature),
    temperature_sd = sd(temperature),
    depth_mean = mean(depth),
    depth_sd = sd(depth),
    depth_scaled = (depth - mean(depth)) / sd(depth),
    temperature_scaled = (temperature - mean(temperature)) / sd(temperature),
    temperature_scaled2 = temperature_scaled^2,
    depth_scaled2 = depth_scaled^2,
  )

  if (type == "depth") {
    formula <-  ~ 0 + depth_scaled + depth_scaled2
  }
  if (type == "temperature"){
    formula <-  ~ 0 + temperature_scaled + temperature_scaled2
  }
  m <- sdmTMB::sdmTMB(
    formula = density ~ 0 + as.factor(year),
    time_varying = formula,
    data = dat, time = "year",
    ar1_fields = FALSE,
    spde = spde, include_spatial = TRUE,
    family = sdmTMB::tweedie(link = "log"),
    anisotropy = FALSE,
    silent = TRUE)

  get_y_hat <- function(b0, b1, b2, year, predictor, mean_column, sd_column) {
    x_pred <- seq(min(dat[[predictor]]), max(dat[[predictor]]), length.out = 300)
    data.frame(
      x = x_pred * dat[[sd_column]][[1]] + dat[[mean_column]][[1]],
      y_hat = exp(b0 + b1 * x_pred + b2 * x_pred^2),
      year = year)
  }

  r <- m$tmb_obj$report()
  r$b_rw_t
  n_t <- nrow(r$b_rw_t)
  yrs <- sort(unique(dat$year))

  pred_temperature <- purrr::map_df(seq_len(n_t), function(.t) {
    get_y_hat(r$b_j[.t], r$b_rw_t[.t,1], r$b_rw_t[.t,2], yrs[.t], 
      sd_column = if (type == "depth") "depth_sd" else "temperature_sd" ,
      mean_column = if (type == "depth") "depth_mean" else "temperature_mean", 
      predictor = if (type == "depth") "depth_scaled" else "temperature_scaled")
  })

  pred_temperature
}

plot_mountains <- function(dat, type = c("depth", "temperature")) {
  type <- match.arg(type)
  mountain_scaler <- max(dat$y_hat)
  mountain_scaler <- mountain_scaler * 10000
  ggplot(dat, aes_string(x = "x", 
    ymax = "100000*y_hat + year*mountain_scaler", ymin = "0 + year*mountain_scaler",
    group = "year", colour = "year", fill = "year")) +
    geom_ribbon(lwd = 0.5, alpha = 0.2) +
    xlab(if (type == "depth") "Depth" else "Temperature") +
    scale_color_viridis_c(option = "C") +
    scale_fill_viridis_c(option = "C") +
    ylab("Predicted density") +
    gfplot::theme_pbs() +
    coord_cartesian(expand = FALSE) + 
    theme(axis.text.y = element_blank())
}
```

```{r, fig.asp=1.3, fig.width=5}
# d <- gfplot::get_survey_sets("pacific cod", ssid = 1)
d <- readRDS(here::here("../gfsynopsis/report/data-cache/petrale-sole.rds"))
f <- fit_mountains(d$survey_sets, type = "temperature")
plot_mountains(f, type = "temperature")
f <- fit_mountains(d$survey_sets, type = "depth")
plot_mountains(f, type = "depth")

d <- readRDS(here::here("../gfsynopsis/report/data-cache/pacific-cod.rds"))
f <- fit_mountains(d$survey_sets, type = "temperature")
plot_mountains(f, type = "temperature")
f <- fit_mountains(d$survey_sets, type = "depth")
plot_mountains(f, type = "depth")

d <- readRDS(here::here("../gfsynopsis/report/data-cache/pacific-ocean-perch.rds"))
f <- fit_mountains(d$survey_sets, type = "temperature")
plot_mountains(f, type = "temperature")
f <- fit_mountains(d$survey_sets, type = "depth")
plot_mountains(f, type = "depth")

d <- readRDS(here::here("../gfsynopsis/report/data-cache/rougheye-blackspotted-rockfish-complex.rds"))
f <- fit_mountains(d$survey_sets, type = "depth")
plot_mountains(f, type = "depth")
```