### 2019-06-08 analysis/VOCC/dist-vocc.Rmd
calculate age at maturity with random effect of year

---
title: "Distanced-based VOCC analysis"
author: "Philina English"
date: '2019-05-14'
output: html_document
---

```{r setup, include=FALSE}
# install.packages("yaImpute") # install package for k-nearest neighbour (kNN) search
# install.packages("raster") # required for calcslope + all raster functions
# install.package(gfplot) # required for some plotting and data functions

library(dplyr)
library(ggplot2)
library(sdmTMB)
library(gfranges)

# Basic map function
plot_map <- function(dat, column = "est") {
  ggplot(dat, aes_string("X", "Y", fill = column)) +
    geom_raster() +
    facet_wrap(~year) +
    coord_fixed() + theme(axis.title=element_blank())
}
```

Retrieve raw data
```{r}
all_depth <- readRDS("../tmb-sensor-explore/data/dat-sensor-trawl-all-depth.rds")

# add and scale predictors after filtering for survey(s) of interest
data <- all_depth$data %>%
  filter(depth_max > 10) %>% # removes data if sensor stayed too near surface
  # filter(ssid != 1) %>%
  # filter(ssid != 3) %>%
  # filter(ssid != 4) %>%
  # filter(ssid != 16) %>%
  gfplot:::scale_survey_predictors()

data <- data[data$fishing_event_id != 481861, ]
# add any new predictors
# data <- data %>%
#   dplyr::mutate(
#   DOY = lubridate::yday(date),
#   shallow = ifelse(depth > 35, 0, 1)
# )
```

Or use old data 
```{r eval=FALSE}
data <- readRDS("../tmb-sensor-explore/data/sensor-data-processed.rds")
data <- data %>%
  mutate(depth = depth_m) %>%
  filter(depth_m > 10) %>% # removes data if sensor stayed too near surface
  gfplot:::scale_survey_predictors()

data <- data[data$fishing_event_id != 481861, ]
```

Create prediction grids for each survey area so that only years with data are included
```{r}
# choose base year(s) to create grid from
dummy_year <- c(2004, 2005)

# create grids for each ssid separately
ssid <- 1
survey_abbrev <- "SYN QCS"
nd_1 <- spatiotemporal_grid(data, ssid, survey_abbrev, dummy_year)
unique(nd_1$year)

ssid <- 3
survey_abbrev <- "SYN HS"
nd_3 <- spatiotemporal_grid(data, ssid, survey_abbrev, dummy_year)
unique(nd_3$year)

ssid <- 4
survey_abbrev <- "SYN WCVI"
nd_4 <- spatiotemporal_grid(data, ssid, survey_abbrev, dummy_year)
unique(nd_4$year)

ssid <- 16
survey_abbrev <- "SYN WCHG"
nd_16 <- spatiotemporal_grid(data, ssid, survey_abbrev, dummy_year = 2006)
unique(nd_16$year)

nd <- rbind(nd_1, nd_3, nd_4, nd_16)
nd <- nd %>% dplyr::mutate(shallow = ifelse(depth > 35, 0, 1))
```

Run sdmTMB model (if not done previously).
```{r eval=FALSE}
# # choose the spatial range to model
dat <- data %>% dplyr::filter(year > 2003) # %>% select(year, ssid, X, Y, temperature_c, depth_scaled, depth_scaled2)
#nd <- nd_1 # %>% filter(ssid == 4)

spde <- sdmTMB::make_spde(dat$X, dat$Y, n_knots = 500)
sdmTMB::plot_spde(spde)

m_temp <- sdmTMB::sdmTMB(dat,
  temperature_c ~ 0 + as.factor(year),
  time_varying = ~ 0 + depth_scaled + depth_scaled2,
  time = "year", spde = spde,
  family = gaussian(link = "identity"),
  ar1_fields = TRUE, # maybe TRUE is better for all areas combined?
  include_spatial = TRUE,
  silent = FALSE
)

# stopifnot(m_temp$model$convergence == 0L)
# m_temp
# # Warning messages:
# #   1: In doTryCatch(return(expr), name, parentenv, handler) :
# #   restarting interrupted promise evaluation
#
# saveRDS(m_temp, file = "data/m_temp_post2003.rds")
saveRDS(m_temp, file = "data/m_temp_allyears.rds")
```

Calculate predicted values
```{r}
m_temp <- readRDS("../VOCC/data/m_temp_allyears.rds")

# can also be filtered to exclude certain years
# m_temp <- readRDS("../VOCC/data/m_temp_allpost2003.rds")
# nd <- nd %>% dplyr::filter(year > 2003)

temp <- predict(m_temp, newdata = nd)
temperature <- temp %>% 
  # filter(ssid!=3) %>% 
  filter(ssid!=4) %>% 
  filter(ssid!=16) %>% 
  filter(year>2004)

glimpse(temperature)

temp <- plot_map(temperature, "est") +
  scale_fill_viridis_c(trans = "sqrt", option = "D") +
  ggtitle("Bottom temperature") + theme(legend.position = "none") #c(0.9, 0), legend.justification = c(0.9, 0), legend.title = element_blank() ) 
print(temp)
```

Explore subsets ot the spatial range 
```{r eval=FALSE}
predicted1 <- predictions %>% filter(ssid == 1)
predicted3 <- predictions %>% filter(ssid == 3)
predicted4 <- predictions %>% filter(ssid == 4)
predicted16 <- predictions %>% filter(ssid == 16)
predicted1n3 <- predictions %>% filter(ssid != 16) %>% filter(ssid != 4) %>% filter(year > 2004)

p <- plot_map(predicted1n3, "est") +
  scale_fill_viridis_c(trans = "sqrt", option = "C") +
  ggtitle("Prediction (fixed effects + all random effects)")
print(p)
```

```{r}
out <- make_vector_data(temperature,
  ssid = c(1,3),
  start_time = 2005,
  #skip_time = 2011,
  input_cell_size = 2,
  scale_fac = 1,
  delta_t_total = 10,
  delta_t_step = 2,
  indices = c(1, 1, 1, 2, 2, 2, 2),
  thresholds = c(0.5) # vector of plus/minus threshold(s) to define climate match.
)
# glimpse(out)
```

```{r}
gvocc <- plot_vocc(out,
  low_col = "white",
  mid_col = "white",
  high_col = "grey87",
  vec_aes = "distance",
  vec_lwd_range = c(0.2, 0.5),
  max_vec_plotted = 100,
  fill_col = "var_1_e",
  fill_label = "Mean\ntemperature\n2013-2017",
  raster_alpha = 1,
  vec_alpha = 0.25,
  axis_lables = FALSE,
  viridis_option = "D",
  transform_col = no_trans
)
gvocc <- gvocc + ggtitle("VOCC vectors for <0.5°C change")
gvocc
```

```{r}
gtrend <- plot_vocc(out,
  fill_col = "units_per_decade",
  fill_label = "°C per decade",
  raster_alpha = 1,
  vec_aes = NULL,
  transform_col = no_trans
)
gtrend <- gtrend + ggtitle("Bottom temperature trend 2005-2017")
gtrend
```

```{r}
# png(file = "accasp-wg-fig.png",   # The directory you want to save the file in
# res = 600,
# units = 'in',
# width = 10, # The width of the plot in inches
# height = 7) # The height of the plot in inches

gridExtra::grid.arrange(gtrend, gvocc, nrow = 1)
# dev.off()
```

Step through pairs of years?

```{r}

step1 <- make_vector_data(temperature,
  ssid = c(1, 3),
  start_time = 2007,
  end_time = 2017,
  skip_time = c(2011,2013),
  input_cell_size = 2,
  scale_fac = 2,
  delta_t_total = 10,
  delta_t_step = 2,
  indices = c(1,1,2,2), 
  thresholds = c(.75) # vector of plus/minus threshold(s) to define climate match.
)

gvocc <- plot_vocc(step1,
  low_col = "white",
  mid_col = "white",
  high_col = "grey97",
  vec_aes = "distance",
  vec_lwd_range = c(0.5, 0.6),
  max_vec_plotted = 100,
  fill_col = "var_1_e",
  fill_label = "Mean temperature\n2015 & 2017",
  raster_alpha = 1,
  vec_alpha = 0.8
)
gvocc
```

# Biomass data

```{r}
library(dplyr)
library(ggplot2)
library(gfplot)
library(gfdata)

events <- readRDS("analysis/VOCC/data/event-data-pop-1n3.rds")
fish <- readRDS("analysis/VOCC/data/bio-data-pop-1n3.rds")

biomass <- split_catch_maturity (events, fish, 
  survey = c("SYN HS","SYN QCS"), 
  years = NULL,
  cutoff_quantile = c(.9995),
  plot = TRUE
)

saveRDS(biomass, file = "analysis/VOCC/data/maturity-biomass-data-pop-1n3-re.rds")

# disagreement between catch_weight and bio sample for single fishing event with 17 large POP. 
# events[events$fishing_event_id==1506954,]$catch_weight
# fish[fish$fishing_event_id==1506954,]$length
# tidy_sets[tidy_sets$fishing_event_id==1506954,]$catch_weight

```

```{r}
biomass <- readRDS("data/maturity-biomass-data-pop-1n3-re.rds")
#doy <- data %>% select(fishing_event_id, DOY)
#biomass <- left_join(biomass, doy)
```

```{r}
p <- ggplot(biomass, aes(X, Y, size = density, colour = density)) + 
  geom_point(alpha=0.5) + scale_colour_viridis_c(trans = "sqrt", option = "C", guide = "legend") + facet_wrap(~year)
print(p)

p <- ggplot(biomass, aes(X, Y, size = adult_density, colour = adult_density)) + 
  geom_point(alpha=0.5) + scale_colour_viridis_c(trans = "sqrt", option = "C", guide = "legend") + facet_wrap(~year)
print(p)

p <- ggplot(biomass, aes(X, Y, size = imm_density, colour = imm_density))  +
  geom_point(alpha=0.5) + scale_colour_viridis_c(trans = "sqrt", option = "C", guide = "legend") + facet_wrap(~year)
print(p)
```

Run sdmTMB model (if not done previously).
```{r eval=FALSE}
# # choose the spatial range to model
dat <- biomass %>% gfplot:::interp_survey_bathymetry() 
dat <- dat$data %>% gfplot:::scale_survey_predictors()
#nd <- nd %>% filter(ssid != 16) %>% filter(ssid != 4)

spde <- sdmTMB::make_spde(dat$X, dat$Y, n_knots = 500)
sdmTMB::plot_spde(spde)

adult_biomass <- sdmTMB::sdmTMB(dat,
  adult_density ~ 0 + as.factor(year),
  time_varying = ~ 0 + depth_scaled + depth_scaled2,
  time = "year", spde = spde,
  family = tweedie(link = "log"),
  ar1_fields = TRUE, 
  include_spatial = TRUE,
  silent = FALSE
)

imm_biomass <- sdmTMB::sdmTMB(dat,
  imm_density ~ 0 + as.factor(year),
  time_varying = ~ 0 + depth_scaled + depth_scaled2,
  time = "year", spde = spde,
  family = tweedie(link = "log"),
  ar1_fields = FALSE, 
  include_spatial = TRUE,
  silent = FALSE
)

saveRDS(adult_biomass, file = "data/m_adult_biomass.rds")
saveRDS(imm_biomass, file = "data/m_imm_biomass.rds")
```

Calculate predicted values
```{r}
adult_biomass <- readRDS("data/m_adult_biomass.rds")
imm_biomass <- readRDS("data/m_imm_biomass.rds")

nd <- nd %>% mutate(year = as.integer(year)) %>%
  # dplyr::filter(year > 2003) %>% 
  filter(ssid!=3) %>% 
  filter(ssid!=4) %>% 
  filter(ssid!=16) 
#unique(biomass$year)

adult_predictions <- predict(adult_biomass, newdata = nd)
imm_predictions <- predict(imm_biomass, newdata = nd)
```

```{r}
predictions <- adult_predictions
predictions$est_exp <- exp(predictions$est)*10000

breaks <- scales::trans_breaks(fourth_root_power[["transform"]], fourth_root_power[["inverse"]], n=5) 
labels <- function(x) { format(x, digits=1) }

p_adult <- plot_map(predictions, "est_exp") + 
  scale_fill_viridis_c(trans = fourth_root_power, breaks=breaks, labels = labels, option = "C") +
  ggtitle("Pacific Ocean Perch mature biomass") + theme(legend.position = c(0.9, 0), legend.justification = c(0.9, 0), legend.title = element_blank() ) 
print(p_adult)
```

```{r}
adult_predictions$est_exp <- exp(adult_predictions$est)*10000
adult <- make_vector_data(adult_predictions,
  ssid = c(1,3),
  start_time = 2005,
  #skip_time = 2011,
  input_cell_size = 2,
  scale_fac = 1,
  delta_t_total = 10,
  delta_t_step = 2,
  indices = c(1, 1, 1, 2, 2, 2, 2),
  variable_names = "est_exp",
  thresholds = c(0.001) # vector of plus/minus threshold(s) to define match.
)
#glimpse(adult )
```

```{r eval=FALSE}
gvobiomass <- plot_vocc(adult,
  # low_col = "yellow",
  # mid_col = "grey",
  # high_col = "black",
  # vec_aes = "units_per_decade",
  # vec_lwd_range = c(0.7,0.005 ),
  # vec_alpha = 0.7,
  # max_vec_plotted = 30,
  fill_col = "var_1_e",
  fill_label = "Mean mature \nbiomass (kg/ha)\n in 2011-2017",
  raster_alpha = 1,
  viridis_option = "C",
  vec_aes = NULL,
  transform_col = fourth_root_power
)
gvobiomass <- gvobiomass +
  ggtitle("Pacific Ocean Perch")

gvobiomass 
```

```{r}
biotrend_ad <- plot_vocc(adult,
  fill_col = "units_per_decade",
  fill_label = "Mature \nbiomass trend \n(kg/ha/decade)\nfor 2005-2017",
  raster_alpha = 1,
  vec_aes = NULL,
  transform_col = fourth_root_power
)
biotrend_ad
```

```{r}
predictions <- imm_predictions
predictions$est_exp <- exp(predictions$est)*10000

p <- plot_map(predictions, "est_exp") + 
  scale_fill_viridis_c(trans = fourth_root_power, breaks=breaks, labels = labels, option = "C") +
  ggtitle("Prediction (fixed effects + all random effects)")
print(p)
```

```{r}
imm_predictions$est_exp <- exp(imm_predictions$est)*10000

out_im <- make_vector_data(imm_predictions,
  ssid = c(1,3),
  start_time = 2005,
  skip_time = 2011,
  input_cell_size = 2,
  scale_fac = 1,
  delta_t_total = 10,
  delta_t_step = 2,
  indices = c(1, 1, 1, 2, 2, 2),
  variable_names = "est_exp",
  thresholds = c(0.0001) # vector of plus/minus threshold(s) to define match.
)
# glimpse(out_im)
```

```{r}
biotrend_im <- plot_vocc(out_im,
  fill_col = "units_per_decade",
  fill_label = "Immature \nbiomass trend \n(kg/ha/decade)\nfor 2005-2017",
  raster_alpha = 1,
  vec_aes = NULL,
  transform_col = fourth_root_power
)
biotrend_im
```
```{r}
png(file = "POP-maturity-year-re.png",   # The directory you want to save the file in
res = 600,
units = 'in',
width = 8, # The width of the plot in inches
height = 10)

plot_mat_ogive(m)
dev.off()
```

```{r}
png(file = "exploratory-pop-figs-trans.png",   # The directory you want to save the file in
res = 600,
units = 'in',
width = 14, # The width of the plot in inches
height = 12) # The height of the plot in inches

gridExtra::grid.arrange(temp, gvocc, gtrend, 
  # p_adult,
    gvobiomass, 
  biotrend_ad, biotrend_im, nrow = 2)
dev.off()
```