### 2019-06-12 analysis/VOCC/biomass-trends.Rmd
biomass plots rmd

---
title: "Biomass trend maps"
author: "Philina English"
date: '2019-06-07'
output: html_document
---

```{r}
species <- "Quillback Rockfish"
species <- "Yelloweye Rockfish"
species <- "North Pacific Spiny Dogfish" 
species <- "Silvergray Rockfish"
species <- "Arrowtooth Flounder"
```

```{r}
spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))
model_ssid <- c(1,3)
```

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(gfplot)
library(gfdata)
library(sdmTMB)
library(gfranges)
```

```{r, message=FALSE, warning=FALSE}
survey_sets <- readRDS(paste0("raw/event-data-", spp, ""))
fish <- readRDS(paste0("raw/bio-data-", spp, ""))
bath <- readRDS("data/bathymetry-data")
# for POP, exclude sample with unusually large fish that are impossible given recorded catch weight
# fish <- fish[fish$fishing_event_id!=1506954,] 
```

Create prediction grids for each survey area so that only years with data are included
```{r}
# choose base year(s) to create grid from
dummy_year <- c(2005,2006)
dat <- bath$data %>% scale_survey_predictors()
# create grids for each ssid separately
ssid <- 1
survey_abbrev <- "SYN QCS"
nd_1 <- spatiotemporal_grid(dat, ssid, survey_abbrev, dummy_year)
unique(nd_1$year)

ssid <- 3
survey_abbrev <- "SYN HS"
nd_3 <- spatiotemporal_grid(dat, ssid, survey_abbrev, dummy_year)
unique(nd_3$year)

ssid <- 4
survey_abbrev <- "SYN WCVI"
nd_4 <- spatiotemporal_grid(dat, ssid, survey_abbrev, dummy_year)
unique(nd_4$year)

ssid <- 16
survey_abbrev <- "SYN WCHG"
nd_16 <- spatiotemporal_grid(dat, ssid, survey_abbrev, dummy_year)
unique(nd_16$year)

nd_all <- rbind(nd_1, nd_3, nd_4, nd_16) %>%
  mutate(year = as.numeric(year)) 
#nd <- nd %>% dplyr::mutate(shallow = ifelse(depth > 35, 0, 1))
```

Calculate maturity weighted biomass
```{r, eval=FALSE}
maturity <- split_catch_maturity (survey_sets, fish, bath,
  survey = c("SYN HS","SYN QCS"), 
  years = NULL,
  cutoff_quantile = c(.9995),
  plot = TRUE
)

saveRDS(maturity$data, file = paste0("data/data-by-maturity-", spp, ".rds"))
```

```{r, eval=FALSE}
png(file = paste0("figs/", spp, "-mass-model.png"),   # The directory you want to save the file in
res = 600,
units = 'in',
width = 8, # The width of the plot in inches
height = 10)

maturity$mass_model

dev.off()
print(maturity$mass_model)
```

```{r, eval=FALSE}
png(file = paste0("figs/", spp, "-maturity-year-re.png"),   # The directory you want to save the file in
res = 600,
units = 'in',
width = 8, # The width of the plot in inches
height = 10)

maturity$maturity

dev.off()
print(maturity$maturity)
```

Choose spatial zone
```{r}
biomass <- readRDS(paste0("data/data-by-maturity-", spp, ".rds"))
data <- biomass %>% gfranges:::scale_survey_predictors()

nd <- nd_all %>% filter(ssid %in% model_ssid) 
data <- data %>% filter(ssid %in% model_ssid)
```

Run sdmTMB model (if not done previously).
```{r eval=FALSE}
spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 500)
sdmTMB::plot_spde(spde)

if (any(names(data)=="adult_density")) {
adult_biomass <- sdmTMB::sdmTMB(data,
  adult_density ~ 0 + as.factor(year),
  time_varying = ~ 0 + depth_scaled + depth_scaled2,
  time = "year", spde = spde,
  family = tweedie(link = "log"),
  ar1_fields = TRUE, 
  include_spatial = TRUE,
  enable_priors = TRUE,
  # control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
  silent = FALSE
)

imm_biomass <- sdmTMB::sdmTMB(data,
  imm_density ~ 0 + as.factor(year),
  time_varying = ~ 0 + depth_scaled + depth_scaled2,
  time = "year", spde = spde,
  family = tweedie(link = "log"),
  ar1_fields = FALSE, 
  include_spatial = TRUE,
  enable_priors = TRUE,
  silent = FALSE
)

saveRDS(adult_biomass, file = paste0("data/model-mature-biomass-", spp, "-1n3.rds"))

saveRDS(imm_biomass, file = paste0("data/model-imm-biomass-", spp, "-1n3.rds"))

} else {

total_biomass <- sdmTMB::sdmTMB(dat,
  density ~ 0 + as.factor(year),
  time_varying = ~ 0 + depth_scaled + depth_scaled2,
  time = "year", spde = spde,
  family = tweedie(link = "log"),
  ar1_fields = TRUE, 
  include_spatial = TRUE,
  silent = FALSE
)

saveRDS(total_biomass, file = paste0("data/model-total-biomass-", spp, "-1n3.rds"))

}
```

Calculate predicted values for available models
#FIXME: tryCatch not working here, try this?
```{r, error=TRUE}
if (length(unique(fish$maturity_code))>2) {
    adult_biomass <- readRDS(paste0("data/model-mature-biomass-", spp, "-1n3.rds"))
    nd$year<- as.integer(nd$year)
    adult_predictions <- predict(adult_biomass, newdata = nd)
    adult_predictions$est_exp <- exp(adult_predictions$est)*10000
    max_raster <- quantile(adult_predictions$est_exp, .9999)
    max_adult <-  round(max(adult_predictions$est_exp))
  } else {
    rm(adult_predictions)
    total_biomass <- readRDS(paste0("data/model-total-biomass-", spp, "-1n3.rds"))
    nd$year<- as.integer(nd$year)
    predictions <- predict(total_biomass, newdata = nd)
    predictions$total_bio <- exp(predictions$est)*10000
    max_raster <- quantile(predictions$total_bio, .9999)
    max_bio <-  round(max(predictions$total_bio))
  }
```

```{r, error=TRUE}
if (exists("adult_predictions")) {
p_adult <- plot_facet_map(adult_predictions, "est_exp", 
  transform_col = fourth_root_power, 
  raster_limits = c(0, max_raster)) + 
  labs(fill="kg/ha") +
  ggtitle(paste0("", species, " mature biomass (max =",max_adult,"kg/ha)"))
print(p_adult) 
} else { 
  p_adult <- plot_facet_map(predictions, "total_bio", 
    #raster_limits = c(0, max_raster),
    transform_col = fourth_root_power) + 
  labs(fill="kg/ha") +
  ggtitle(paste0("", species, " total biomass (max = ",max_bio,"kg/ha)"))
print(p_adult) 
  }
```

```{r, message=FALSE, warning=FALSE}
if (exists("adult_predictions")) {

out_mature1 <- make_vector_data(adult_predictions,
  ssid = c(1),
  start_time = 2003,
  #start_time = 2005,
  #skip_time = 2011,
  input_cell_size = 2,
  scale_fac = 1,
  delta_t_total = 10,
  delta_t_step = 2,
  indices = c(1, 1, 1, 1, 1, 2, 2, 2, 2),
  #indices = c(1, 1, 1, 2, 2, 2, 2),
  variable_names = "est_exp",
  thresholds = c(0.001) # vector of plus/minus threshold(s) to define match.
)

out_mature3 <- make_vector_data(adult_predictions,
  ssid = c(3),
  start_time = 2005,
  #skip_time = 2011,
  input_cell_size = 2,
  scale_fac = 1,
  delta_t_total = 10,
  delta_t_step = 2,
  indices = c(1, 1, 1, 2, 2, 2, 2),
  variable_names = "est_exp",
  thresholds = c(0.001) # vector of plus/minus threshold(s) to define match.
)

out_mature <- rbind(out_mature1, out_mature3)
} else { rm(out_mature)}
```

```{r, message=FALSE, warning=FALSE}
out_total <- make_vector_data(predictions,
  ssid = c(1,3),
  start_time = 2005,
  #skip_time = 2011,
  input_cell_size = 2,
  scale_fac = 1,
  delta_t_total = 10,
  delta_t_step = 2,
  indices = c(1, 1, 1, 2, 2, 2, 2),
  variable_names = "total_bio",
  thresholds = c(0.001) # vector of plus/minus threshold(s) to define match.
)
```

```{r eval=FALSE}
current_biomass <- plot_vocc(out_total,
  fill_col = "var_1_e",
  fill_label = "Mean mature \nbiomass (kg/ha)\n in 2011-2017",
  raster_alpha = 1,
  viridis_option = "C",
  vec_aes = NULL,
  transform_col = fourth_root_power
)
current_biomass <- current_biomass +
  ggtitle(paste0("", species, " total biomass"))

current_biomass
```

```{r, error=TRUE}
if (exists("out_mature")) {
data <- out_mature1
data <- out_mature
biotrend_ad <- plot_vocc(data,
  fill_col = "units_per_decade",
  fill_label = "Mature \nbiomass trend \n(kg/ha/decade)\nfor 2003-2017",
  raster_alpha = 1,
  vec_aes = NULL,
  transform_col = fourth_root_power
)
biotrend_ad
} else {
  biotrend_ad <- plot_vocc(out_total,
  fill_col = "units_per_decade",
  fill_label = "Biomass trend \n(kg/ha/decade)\nfor 2003-2017",
  raster_alpha = 1,
  vec_aes = NULL,
  transform_col = fourth_root_power
)
biotrend_ad
}
```

```{r, error=TRUE}
if (length(unique(fish$maturity_code))>2) {
    imm_biomass <- readRDS(paste0("data/model-imm-biomass-", spp, "-1n3.rds"))

    nd$year<- as.integer(nd$year)
    imm_predictions <- predict(imm_biomass, newdata = nd)
    imm_predictions$est_exp <- exp(imm_predictions$est)*10000

    predictions <- adult_predictions
    predictions$prop_imm <- imm_predictions$est_exp/(adult_predictions$est_exp+imm_predictions$est_exp)
    predictions$total_bio <- imm_predictions$est_exp + adult_predictions$est_exp

  }else{ 
    rm(imm_predictions) 
}
```

```{r}
if (exists("imm_predictions")) {
p_imm <- plot_facet_map(imm_predictions, "est_exp", transform_col = fourth_root_power, 
  raster_limits = c(0, max_raster)) + 
  labs(fill="kg/ha") +
  ggtitle(paste0("", species, " immature biomass"))
p_imm
} else {
  p_imm <- grid::grid.rect(gp=grid::gpar(col="white"))
}
```

```{r, message=FALSE, warning=FALSE}
if (exists("imm_predictions")) {

out_im1 <- make_vector_data(imm_predictions,
  ssid = c(1),
  start_time = 2003,
  #start_time = 2005,
  #skip_time = 2011,
  input_cell_size = 2,
  scale_fac = 1,
  delta_t_total = 10,
  delta_t_step = 2,
  indices = c(1, 1, 1, 1, 1, 2, 2, 2, 2),
  #indices = c(1, 1, 1, 2, 2, 2, 2),
  variable_names = "est_exp",
  thresholds = c(0.0001) # vector of plus/minus threshold(s) to define match.
)

out_im3 <- make_vector_data(imm_predictions,
  ssid = c(3),
  start_time = 2005,
  #skip_time = 2011,
  input_cell_size = 2,
  scale_fac = 1,
  delta_t_total = 10,
  delta_t_step = 2,
  indices = c(1, 1, 1, 2, 2, 2, 2),
  variable_names = "est_exp",
  thresholds = c(0.0001) # vector of plus/minus threshold(s) to define match.
)
# glimpse(out_im)

out_imm <- rbind(out_im1, out_im3)
} else { rm(out_imm)}
```

```{r}
if (exists("out_imm")) {
data <- out_im1
data <- out_imm
biotrend_im <- plot_vocc(data,
  fill_col = "units_per_decade",
  fill_label = "Immature \nbiomass trend \n(kg/ha/decade)\nfor 2003-2017",
  raster_alpha = 1,
  vec_aes = NULL,
  transform_col = fourth_root_power
)
biotrend_im
} else { biotrend_im <- grid::grid.rect(gp=grid::gpar(col="white"))}
```

#FIXME: failing
```{r}
if (exists("imm_predictions")) {

p_prop_imm <- plot_facet_map(predictions, "prop_imm") + 
  labs(fill="% by mass") +
  ggtitle(paste0("", species, " proportion immature"))
print(p_prop_imm)

} else { p_prop_imm <- grid::grid.rect(gp=grid::gpar(col="white"))}
```

Calculate predicted values for temperature
```{r, eval=FALSE}
m_temp <- readRDS("../VOCC/data/m_temp_allyears.rds")
temperature <- predict(m_temp, newdata = nd_all)
temperature <- filter(temperature, ssid %in% model_ssid)
```

```{r, eval=FALSE}
temp <-plot_facet_map(temperature, "est", transform_col = no_trans, viridis_option = "D") + 
  labs(fill="°C") +
  ggtitle("Bottom temperature")
print(temp)
```

```{r, eval=FALSE}
png(file = paste0("figs/exploratory-", spp, "-figs.png"),   # The directory you want to save the file in
res = 600,
units = 'in',
width = 14, # The width of the plot in inches
height = 12) # The height of the plot in inches

gridExtra::grid.arrange(
    p_adult, p_imm, 
    p_prop_imm, #current_biomass, 
    biotrend_ad, biotrend_im, 
    temp,
    nrow = 2)
dev.off()
```