### 2019-06-26 analysis/VOCC/biomass-trends.Rmd
make "trawled" time-varying

---
title: "Biomass trend maps"
author: "Philina English"
date: '2019-06-07'
output: html_document
---

Choose one value for each parameter:

Species
```{r}
species <- "Quillback Rockfish"
# species <- "Yelloweye Rockfish"
# priors <- TRUE

species <- "North Pacific Spiny Dogfish" # note: using all data for maturity thresholds
species <- "Silvergray Rockfish"
species <- "Arrowtooth Flounder"
species <- "Lingcod"
```

# Choose covariates  
```{r}
covariates <- "+muddy+mixed+rocky"
covariates <- "+mixed+rocky"

covariates <- "+trawled+muddy+rocky+mixed"
covariates <- "+trawled+mixed+rocky"
covariates <- "+trawled+mixed"
```

Is a new spatial needed?
```{r}
update_spatial_model <- TRUE
priors <- FALSE
#priors <- TRUE

# Choose region to update model for:
region <- "Both odd year surveys"
region <- "West Coast Vancouver Island"
region <- "West Coast Haida Gwaii"

```

If not...
```{r}
update_spatial_model <- FALSE
priors <- FALSE
# priors <- TRUE
region <- "All synoptic surveys"

if (region == "All synoptic surveys") {
  survey <- c("SYN QCS", "SYN HS", "SYN WCVI", "SYN WCHG")
  model_ssid <- c(1, 3, 4, 16)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}
```

Run all subsequent code...

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(gfplot)
library(gfdata)
library(sdmTMB)
library(gfranges)

covs <- gsub("\\+", "-", covariates)
spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))

# folder to hold figs for this species
dir.create(file.path("figs", spp))
dir.create(file.path("data", spp))
fish <- readRDS(paste0("raw/bio-data-", spp, ""))

if (region == "Both odd year surveys") {
  survey <- c("SYN QCS", "SYN HS")
  model_ssid <- c(1, 3)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "West Coast Vancouver Island") {
  survey <- c("SYN WCVI")
  model_ssid <- c(4)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "West Coast Haida Gwaii") {
  survey <- c("SYN WCHG")
  model_ssid <- c(16)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

```

```{r}

# if(region == "Queen Charlotte Sound") {
# survey <- c("SYN QCS")
# model_ssid <- c(1)
# ssid_string <- paste0(model_ssid, collapse = "n")
# years <- NULL
# }
#
# if(region == "Hecate Strait") {
# survey <- c("SYN HS")
# model_ssid <- c(3)
# ssid_string <- paste0(model_ssid, collapse = "n")
# years <- NULL
# }
```

Make prediction grids for all surveys and years, if not done before.
```{r, message=FALSE, warning=FALSE}
try(nd_all <- readRDS("data/nd_all_synoptic.rds"))

if (!exists("nd_all")) {

  # choose base year(s) to create grid from
  dummy_year <- c(2005, 2006)

  bath <- readRDS("data/bathymetry-data")
  dat <- bath$data %>% scale_survey_predictors()

  # create grids for each ssid separately
  ssid <- 1
  survey_abbrev <- "SYN QCS"
  nd_1 <- spatiotemporal_grid(dat, ssid, survey_abbrev, dummy_year)
  unique(nd_1$year)

  ssid <- 3
  survey_abbrev <- "SYN HS"
  nd_3 <- spatiotemporal_grid(dat, ssid, survey_abbrev, dummy_year)
  unique(nd_3$year)

  ssid <- 4
  survey_abbrev <- "SYN WCVI"
  nd_4 <- spatiotemporal_grid(dat, ssid, survey_abbrev, dummy_year)
  unique(nd_4$year)

  ssid <- 16
  survey_abbrev <- "SYN WCHG"
  nd_16 <- spatiotemporal_grid(dat, ssid, survey_abbrev, dummy_year)
  unique(nd_16$year)

  nd_all <- rbind(nd_1, nd_3, nd_4, nd_16) %>% 
    mutate(year = as.numeric(year))

  newcov <- readRDS("data/new_covariates.rds") %>% dplyr::select(X, Y, trawled, muddy, rocky, mixed)
  nd_all <- left_join(nd_all, newcov, by=c("X", "Y"))

  # nd <- nd %>% dplyr::mutate(shallow = ifelse(depth > 35, 0, 1))

  saveRDS(nd_all, file = paste0("data/nd_all_synoptic.rds"))
}
```

Calculate maturity weighted biomass if not done before
```{r}
rm(biomass)
try(biomass <- readRDS(paste0("data/", spp, "/data-by-maturity-", spp, "-", ssid_string, ".rds")))

if (!exists("biomass")) {
  survey_sets <- readRDS(paste0("raw/event-data-", spp, ""))
  fish <- readRDS(paste0("raw/bio-data-", spp, ""))
  bath <- readRDS("data/bathymetry-data")
  # for POP, exclude sample with unusually large fish that are impossible given recorded catch weight
  # fish <- fish[fish$fishing_event_id!=1506954,]

  maturity <- split_catch_maturity(survey_sets, fish, bath,
    # survey = c("SYN QCS","SYN HS","SYN WCVI","SYN WCHG"),
    # survey = c("SYN HS", "SYN WCHG"),
    survey = survey,
    years = years,
    cutoff_quantile = c(.9995),
    plot = TRUE
  )

  saveRDS(maturity$data, file = paste0("data/", spp, "/data-by-maturity-", spp, "-", ssid_string, ".rds"))
}
```

```{r}
if (exists("maturity")) {
  print(maturity$mass_model)
  ggsave(
    file = paste0("figs/", spp, "/", spp, "-mass-model-", ssid_string, ".png"),
    dpi = 600,
    width = 8, # The width of the plot in inches
    height = 10
  )
}

# try(print(maturity$mass_model))
```

```{r}
if (exists("maturity")) {
  png(
    file = paste0("figs/", spp, "/", spp, "-maturity-", ssid_string, ".png"), # The directory you want to save the file in
    res = 600,
    units = "in",
    width = 8, # The width of the plot in inches
    height = 10
  )
  print(maturity$maturity)
  dev.off()
}

try(print(maturity$maturity))
```

Run sdmTMB model (if not done previously).
```{r}
if (update_spatial_model) {
  biomass <- readRDS(paste0("data/", spp, "/data-by-maturity-", spp, "-", ssid_string, ".rds"))
  covars <- readRDS("data/event-covariates.rds")
  data <- dplyr::left_join(biomass,covars)

  # scale predictors before filtering to ensure mean and SD are global
  data <- data %>% gfranges:::scale_survey_predictors()

  data <- data %>% filter(ssid %in% model_ssid)

  # if (region == "All synoptic surveys") {
  #   spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 1200)
  # }

  if (region == "Both odd year surveys") {
    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 400)
  }

  if (region == "West Coast Vancouver Island") {
    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 200)
  }

  if (region == "West Coast Haida Gwaii") {
    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 200)
  }

  sdmTMB::plot_spde(spde)

  if (any(names(data) == "adult_density")) {

    adult_formula <- as.formula(paste("adult_density ~ 0 + as.factor(year)", covariates, ""))

    starttime1 <- Sys.time()
    adult_biomass <- sdmTMB::sdmTMB(data,
      adult_formula,
      time_varying = ~ 0 + depth_scaled + depth_scaled2 + trawled,
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = TRUE,
      include_spatial = TRUE,
      enable_priors = priors,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )
    endtime1 <- Sys.time()
    time1 <- round(starttime1 - endtime1)

    #saveRDS(adult_biomass, 
    # file = paste0("data/", spp, "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, time1, ".rds"))
    saveRDS(adult_biomass, 
      file = paste0("data/", spp, "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))

    imm_formula <- as.formula(paste("imm_density ~ 0 + as.factor(year)", covariates, ""))

    starttime2 <- Sys.time()
    imm_biomass <- sdmTMB::sdmTMB(data, 
      imm_formula,
      time_varying = ~ 0 + depth_scaled + depth_scaled2 + trawled,
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = FALSE,
      include_spatial = TRUE,
      enable_priors = priors,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )
    endtime2 <- Sys.time()
    time2 <- round(starttime2 - endtime2)

    #saveRDS(imm_biomass, 
    # file = paste0("data/", spp, "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, time2, ".rds"))
    saveRDS(imm_biomass, 
     file = paste0("data/", spp, "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))

  } else {
    starttime3 <- Sys.time()
    dens_formula <- as.formula(paste("density ~ 0 + as.factor(year)", covariates, ""))

    total_biomass <- sdmTMB::sdmTMB(dat,
      dens_formula,
      time_varying = ~ 0 + depth_scaled + depth_scaled2,
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = TRUE,
      include_spatial = TRUE,
      enable_priors = priors,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )
    endtime3 <- Sys.time()
    time3 <- round(starttime3 - endtime3)

    saveRDS(total_biomass, file = paste0("data/", spp, "/model-total-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))
  }
}
```

```{r eval=FALSE}
k <- length(adult_biomass$model$par)
nll <- adult_biomass$model$objective
AIC <- 2*k+2*nll
AIC #(-5500.438)
```

Calculate predicted values for available models, if needed
```{r, error=TRUE}
rm(adult_predictions)
rm(imm_predictions)
rm(predictions)

if (isFALSE(update_spatial_model)) {
  if (region != "All synoptic surveys") {
    try(adult_predictions <- readRDS(paste0("data/", spp, "/predictions-", spp, covs, "-", ssid_string, "-mature-biomass-prior-", priors, ".rds")))
    try(imm_predictions <- readRDS(paste0("data/", spp, "/predictions-", spp, covs, "-", ssid_string, "-imm-biomass-prior-", priors, ".rds")))
    try(predictions <- readRDS(paste0("data/", spp, "/predictions-", spp, covs, "-", ssid_string, "-total-biomass-prior-", priors, ".rds")))
  } else {
    # Bind together predictions from different models
    try ({
      pred1n3 <- readRDS(paste0("data/", spp, "/predictions-", spp, "-1n3-mature-biomass-prior-", priors, ".rds"))
      pred4 <- readRDS(paste0("data/", spp, "/predictions-", spp, "-4-mature-biomass-prior-", priors, ".rds"))
      pred16 <- readRDS(paste0("data/", spp, "/predictions-", spp, "-16-mature-biomass-prior-", priors, ".rds"))
      adult_predictions <- rbind(pred1n3, pred4, pred16) 
      # model_ssid <- unique(adult_predictions$ssid)
      # ssid_string <- paste0(model_ssid, collapse = "n")
    })

    try ({
      pred1n3_imm <- readRDS(paste0("data/", spp, "/predictions-", spp, "-1n3-imm-biomass-prior-", priors, ".rds"))
      pred4_imm <- readRDS(paste0("data/", spp, "/predictions-", spp, "-4-imm-biomass-prior-", priors, ".rds"))
      pred16_imm <- readRDS(paste0("data/", spp, "/predictions-", spp, "-16-imm-biomass-prior-", priors, ".rds"))
      imm_predictions <- rbind(pred1n3_imm, pred4_imm, pred16_imm) 
      # model_ssid <- unique(imm_predictions$ssid)
      # ssid_string <- paste0(model_ssid, collapse = "n")
    })

    try ({
      pred1n3_all <- readRDS(paste0("data/", spp, "/predictions-", spp, "-1n3-total-biomass-prior-", priors, ".rds"))
      pred4_all <- readRDS(paste0("data/", spp, "/predictions-", spp, "-4-total-biomass-prior-", priors, ".rds"))
      pred16_all <- readRDS(paste0("data/", spp, "/predictions-", spp, "-16-total-biomass-prior-", priors, ".rds"))
      predictions <- rbind(pred1n3_all, pred4_all, pred16_all) # ,
      #  model_ssid <- unique(predictions$ssid),
      # ssid_string <- paste0(model_ssid, collapse = "n")
    })

    if (!exists("adult_predictions")) paste("Maturity split biomass predictions are not available.")
    if (!exists("predictions")) paste("Total biomass predictions are not available.")
  }
} else {
  if (region == "All synoptic surveys") {
    paste("To update spatial models, you need to build and save each region's model separately.")
  }
}

if (!exists("adult_predictions")) {
    fish <- readRDS(paste0("raw/bio-data-", spp, ""))

  if (length(unique(fish$maturity_code)) > 2) {

    # adult_biomass <- readRDS(list.files(
    #   path = paste0("data/", spp, "/"),
    #   #pattern = paste0("mod-mat-biomass-", spp, "-", ssid_string, "-prior-", priors, "-*"),
    #   pattern = paste0("mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, "-*"),
    #   all.files = F, full.names = T
    # ))

     adult_biomass <- readRDS(paste0("data/", spp, "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))

    nd_all <- readRDS(paste0("data/nd_all_synoptic.rds"))
    nd <- nd_all %>% filter(ssid %in% model_ssid) %>% filter(year %in% unique(adult_biomass$data$year))
    nd <- na.omit(nd)
    nd$year <- as.integer(nd$year)

    adult_predictions <- predict(adult_biomass, newdata = nd)
    saveRDS(adult_predictions,
      file = paste0("data/", spp, "/predictions-", spp, covs,"-", ssid_string, "-mature-biomass-prior-", priors, ".rds")
    )

    # imm_biomass <- readRDS(list.files(
    #   path = paste0("data/", spp, "/"),
    #   #pattern = paste0("model-imm-biomass-", spp, "-", ssid_string, "-prior-", priors, "-*"),
    #   pattern = paste0("mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, "-*"),
    #   all.files = F, full.names = T
    # ))

    imm_biomass <- readRDS(paste0("data/", spp, "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))

    imm_predictions <- predict(imm_biomass, newdata = nd)
    saveRDS(imm_predictions,
      file = paste0("data/", spp, "/predictions-", spp, covs,"-", ssid_string, "-imm-biomass-prior-", priors, ".rds")
    )
  } else {
    rm(adult_predictions)
    rm(imm_predictions)

    if (!exists(predictions)) {
      nd_all <- readRDS(paste0("data/nd_all_synoptic.rds"))
      nd <- nd_all %>% filter(ssid %in% model_ssid)
      nd <- na.omit(nd)
      nd$year <- as.integer(nd$year)

      total_biomass <- readRDS(paste0("data/", spp, "/model-total-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"))
      nd$year <- as.integer(nd$year)
      predictions <- predict(total_biomass, newdata = nd)
      saveRDS(predictions, file = paste0("data/", spp, "/predictions-", spp, covs,"-", ssid_string, "-total-biomass-prior-", priors, ".rds"))
    }
  }
}
```

Bind together predictions from different models
```{r, eval=FALSE}
if (region == "All synoptic surveys") {
  if (isFALSE(update_spatial_model)) {
    try(
      pred1n3 <- readRDS(paste0("data/", spp, "/predictions-", spp, covs, "-1n3-mature-biomass-prior-", priors, ".rds")),
      pred4 <- readRDS(paste0("data/", spp, "/predictions-", spp, covs, "-4-mature-biomass-prior-", priors, ".rds")),
      pred16 <- readRDS(paste0("data/", spp, "/predictions-", spp, covs, "-16-mature-biomass-prior-", priors, ".rds")),

      adult_predictions <- rbind(pred1n3, pred4, pred16),
      model_ssid <- unique(adult_predictions$ssid),
      ssid_string <- paste0(model_ssid, collapse = "n")
    )

    try(
      pred1n3_imm <- readRDS(paste0("data/", spp, "/predictions-", spp, covs, "-1n3-imm-biomass-prior-", priors, ".rds")),
      pred4_imm <- readRDS(paste0("data/", spp, "/predictions-", spp, covs, "-4-imm-biomass-prior-", priors, ".rds")),
      pred16_imm <- readRDS(paste0("data/", spp, "/predictions-", spp, covs, "-16-imm-biomass-prior-", priors, ".rds")),

      imm_predictions <- rbind(pred1n3_imm, pred4_imm, pred16_imm),
      model_ssid <- unique(imm_predictions$ssid),
      ssid_string <- paste0(model_ssid, collapse = "n")
    )

    try(
      pred1n3_all <- readRDS(paste0("data/", spp, "/predictions-", spp, "-1n3-total-biomass-prior-", priors, ".rds")),
      pred4_all <- readRDS(paste0("data/", spp, "/predictions-", spp, "-4-total-biomass-prior-", priors, ".rds")),
      pred16_all <- readRDS(paste0("data/", spp, "/predictions-", spp, "-16-total-biomass-prior-", priors, ".rds")),

      predictions <- rbind(pred1n3_all, pred4_all, pred16_all),
      model_ssid <- unique(predictions$ssid),
      ssid_string <- paste0(model_ssid, collapse = "n")
    )
  } else {
    paste("To update spatial models, you need to build and save each region's model separately.")
  }
}

if (!exists(adult_predictions)) paste("Maturity split biomass predictions were not available.")
if (!exists(predictions)) paste("Total biomass predictions were not available.")
```

Caluclate variables of interest from predictions
```{r}
if (exists("adult_predictions")) {
  adult_predictions$est_exp <- exp(adult_predictions$est) * 10000
  imm_predictions$est_exp <- exp(imm_predictions$est) * 10000

  adult_predictions$total_bio <- imm_predictions$est_exp + adult_predictions$est_exp
  imm_predictions$prop_imm <- imm_predictions$est_exp / (adult_predictions$est_exp + imm_predictions$est_exp)

  max_raster <- quantile(adult_predictions$est_exp, .9999)
  max_adult <- signif(max(adult_predictions$est_exp), digits = 2)
} else {
  predictions$total_bio <- exp(predictions$est) * 10000
  max_raster <- quantile(predictions$total_bio, .9999)
  max_bio <- round(max(predictions$total_bio))
}
```

```{r, error=TRUE}
if (exists("adult_predictions")) {
  p_adult_all <- plot_facet_map(adult_predictions, "est_exp",
    raster_limits = c(0, max_raster),
    legend_position = "none",
    transform_col = fourth_root_power
  ) +
    labs(fill = "kg/ha") +
    ggtitle(paste0("", species, " mature biomass \n(max = ", max_adult, " kg/ha)"))
  print(p_adult_all)
} else {
  p_adult_all <- plot_facet_map(predictions, "total_bio",
    # raster_limits = c(0, max_raster),
    transform_col = fourth_root_power
  ) +
    labs(fill = "kg/ha") +
    ggtitle(paste0("", species, " total biomass \n(max = ", max_bio, " kg/ha)"))
  print(p_adult_all)
}
```

```{r}
if (exists("adult_predictions")) {
  # if (length(model_ssid)>1){
  if (1 %in% model_ssid) {
    adult_predictions1 <- adult_predictions %>% filter(ssid == 1)
  } else {
    adult_predictions1 <- NULL
  }

  if (3 %in% model_ssid) {
    adult_predictions3 <- adult_predictions %>% filter(ssid == 3)
  } else {
    adult_predictions3 <- NULL
  }

  if (4 %in% model_ssid) {
    adult_predictions4 <- adult_predictions %>% filter(ssid == 4)
  } else {
    adult_predictions4 <- NULL
  }

  if (16 %in% model_ssid) {
    adult_predictions16 <- adult_predictions %>% filter(ssid == 16)
  } else {
    adult_predictions16 <- NULL
  }

  ad_prediction_list <- list(adult_predictions1, adult_predictions3, adult_predictions4, adult_predictions16)

  ad_prediction_list <- ad_prediction_list[!sapply(ad_prediction_list, is.null)]

  p_adult <- list()
  for (i in seq_len(length(model_ssid))) {
    p_adult[[i]] <- plot_facet_map(ad_prediction_list[[i]], "est_exp",
      raster_limits = c(0, max_raster),
      # legend_position = "none"
      transform_col = fourth_root_power
    ) +
      labs(fill = "kg/ha") +
      ggtitle(paste0("", species, " mature biomass \n(max = ", max_adult, " kg/ha)"))
    p_adult[[i]]
  }
  print(p_adult)
}
# }
```

```{r, message=FALSE, warning=FALSE}
if (exists("adult_predictions")) {
  predictions <- adult_predictions
  # model_ssid <- unique(predictions$ssid)
  # ssid_string <- paste0(model_ssid, collapse = "n")
}

if (1 %in% model_ssid) {
  out_mature1 <- make_vector_data(predictions,
    ssid = c(1),
    start_time = 2003,
    # start_time = 2005,
    skip_time = 2004,
    input_cell_size = 2,
    scale_fac = 1,
    delta_t_total = 10,
    delta_t_step = 2,
    indices = c(1, 1, 1, 1, 2, 2, 2, 2),
    # indices = c(1, 1, 1, 2, 2, 2, 2),
    variable_names = "est_exp",
    thresholds = c(0.001)
  )
} else {
  out_mature1 <- NULL
}

if (3 %in% model_ssid) {
  out_mature3 <- make_vector_data(predictions,
    ssid = c(3),
    start_time = 2005,
    input_cell_size = 2,
    scale_fac = 1,
    delta_t_total = 10,
    delta_t_step = 2,
    indices = c(1, 1, 1, 2, 2, 2, 2),
    variable_names = "est_exp",
    thresholds = c(0.001)
  )
} else {
  out_mature3 <- NULL
}

if (4 %in% model_ssid) {
  out_mature4 <- make_vector_data(predictions,
    ssid = c(4),
    start_time = 2004,
    input_cell_size = 2,
    scale_fac = 1,
    delta_t_total = 10,
    delta_t_step = 2,
    indices = c(1, 1, 1, 1, 2, 2, 2, 2),
    variable_names = "est_exp",
    thresholds = c(0.001)
  )
} else {
  out_mature4 <- NULL
}

if (16 %in% model_ssid) {
  out_mature16 <- make_vector_data(predictions,
    ssid = c(16),
    start_time = 2004,
    skip_time = 2007,
    input_cell_size = 2,
    scale_fac = 1,
    delta_t_total = 10,
    delta_t_step = 2,
    indices = c(1, 1, 1, 2, 2, 2, 2),
    variable_names = "est_exp",
    thresholds = c(0.001)
  )
} else {
  out_mature16 <- NULL
}
out_mature <- rbind(out_mature1, out_mature3, out_mature4, out_mature16)

saveRDS(out_mature,
  file = paste0("data/", spp, "/trends-", spp, covs, "-", ssid_string, "-mature-biomass-priors-", priors, ".rds")
)

if (!exists("adult_predictions")) {
  out_total <- out_mature
  saveRDS(out_total,
    file = paste0("data/", spp, "/trends-", spp, covs, "-", ssid_string, "-total-biomass-priors-", priors, ".rds")
  )
}
```

Attempting to plot of current mature biomass with trawl footprint...
```{r }
# if (exists("out_mature")) {
#   
#   trawl_footprint <- sf::st_read(dsn="data/trawl-footprint",layer="Trawl_footprint")
#   proj.to <- "+proj=utm +zone=9 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs"
# 
#   trawl_footprint <- sf::st_transform(trawl_footprint, crs = proj.to)
# 
#   
#   current_biomass <- plot_vocc(out_mature,
#     fill_col = "var_1_e",
#     fill_label = "Mean mature \nbiomass (kg/ha)\n in 2011-2017",
#     raster_alpha = 1,
#     viridis_option = "C",
#     vec_aes = NULL,
#     transform_col = fourth_root_power
#   )
#   current_biomass <- current_biomass +
#   #ggplot(trawl_footprint) + #, aes(x=x/1000, y=y/1000)
#     geom_sf(data= trawl_footprint, aes(geometry=geometry/1000), inherit.aes = FALSE) #+
#     #coord_sf(datum = NULL) 
#   
#   +
#     ggtitle(paste0("", species, " total biomass"))  
# 
#   current_biomass
# }
```

```{r, error=TRUE}
try(out_mature <- readRDS(paste0("data/", spp, "/trends-", spp, covs, "-", ssid_string, "-mature-biomass-priors-", priors, ".rds")))

if (exists("out_mature")) {
  data <- out_mature
  biotrend_ad <- plot_vocc(data,
    fill_col = "units_per_decade",
    fill_label = "Mature \nbiomass \ntrend \n(kg/ha/decade)",
    raster_alpha = 1,
    vec_aes = NULL,
    # axis_lables = TRUE,
    white_zero = TRUE,
    transform_col = fourth_root_power,
    legend_position = c(0.2, 0.25)
  ) + ggtitle(paste0("", species, " ", min(unique(adult_predictions[["year"]])), "-", max(unique(adult_predictions[["year"]]))))
  biotrend_ad
} else {
  try(out_total <- readRDS(paste0("data/", spp, "/trends-", spp, covs, "-", ssid_string, "-total-biomass-priors-", priors, ".rds")))

  biotrend_ad <- plot_vocc(out_total,
    fill_col = "units_per_decade",
    fill_label = "Biomass \ntrend \n(kg/ha/decade)",
    raster_alpha = 1,
    vec_aes = NULL,
    # axis_lables = TRUE,
    white_zero = TRUE,
    transform_col = fourth_root_power,
    legend_position = c(0.2, 0.25)
  ) + ggtitle(paste0("", species, " ", min(unique(predictions[["year"]])), "-", max(unique(predictions[["year"]]))))
  biotrend_ad
}
```

```{r, error=TRUE}
if (exists("imm_predictions")) {
  p_imm_all <- plot_facet_map(imm_predictions, "est_exp",
    transform_col = fourth_root_power,
    raster_limits = c(0, max_raster)
  ) +
    labs(fill = "kg/ha") +
    ggtitle(paste0("", species, " immature biomass \n "))
  p_imm_all
} else {
  p_imm_all <- grid::grid.rect(gp = grid::gpar(col = "white"))
}
```

```{r}
if (exists("adult_predictions")) {
  # if (length(model_ssid)>1){
  if (1 %in% model_ssid) {
    imm_predictions1 <- imm_predictions %>% filter(ssid == 1)
  } else {
    imm_predictions1 <- NULL
  }

  if (3 %in% model_ssid) {
    imm_predictions3 <- imm_predictions %>% filter(ssid == 3)
  } else {
    imm_predictions3 <- NULL
  }

  if (4 %in% model_ssid) {
    imm_predictions4 <- imm_predictions %>% filter(ssid == 4)
  } else {
    imm_predictions4 <- NULL
  }

  if (16 %in% model_ssid) {
    imm_predictions16 <- imm_predictions %>% filter(ssid == 16)
  } else {
    imm_predictions16 <- NULL
  }

  im_prediction_list <- list(imm_predictions1, imm_predictions3, imm_predictions4, imm_predictions16)

  im_prediction_list <- im_prediction_list[!sapply(im_prediction_list, is.null)]

  p_imm <- list()
  for (i in seq_len(length(model_ssid))) {
    p_imm[[i]] <- plot_facet_map(im_prediction_list[[i]], "est_exp",
      raster_limits = c(0, max_raster),
      # legend_position = "none",
      transform_col = fourth_root_power
    ) +
      labs(fill = "kg/ha") +
      ggtitle(paste0("", species, " immature biomass \n(max = ", max_adult, " kg/ha)"))
    p_imm[[i]]
  }
  print(p_imm)
}
# }
```

```{r, message=FALSE, warning=FALSE}
if (exists("imm_predictions")) {
  predictions <- imm_predictions

  if (1 %in% model_ssid) {
    out_im1 <- make_vector_data(predictions,
      ssid = c(1),
      start_time = 2003,
      # start_time = 2005,
      skip_time = 2004,
      input_cell_size = 2,
      scale_fac = 1,
      delta_t_total = 10,
      delta_t_step = 2,
      indices = c(1, 1, 1, 1, 2, 2, 2, 2),
      # indices = c(1, 1, 1, 2, 2, 2, 2),
      variable_names = "est_exp",
      thresholds = c(0.001)
    )
  } else {
    out_im1 <- NULL
  }

  if (3 %in% model_ssid) {
    out_im3 <- make_vector_data(predictions,
      ssid = c(3),
      start_time = 2005,
      input_cell_size = 2,
      scale_fac = 1,
      delta_t_total = 10,
      delta_t_step = 2,
      indices = c(1, 1, 1, 2, 2, 2, 2),
      variable_names = "est_exp",
      thresholds = c(0.001)
    )
  } else {
    out_im3 <- NULL
  }

  if (4 %in% model_ssid) {
    out_im4 <- make_vector_data(predictions,
      ssid = c(4),
      start_time = 2004,
      input_cell_size = 2,
      scale_fac = 1,
      delta_t_total = 10,
      delta_t_step = 2,
      indices = c(1, 1, 1, 1, 2, 2, 2, 2),
      variable_names = "est_exp",
      thresholds = c(0.001)
    )
  } else {
    out_im4 <- NULL
  }

  if (16 %in% model_ssid) {
    out_im16 <- make_vector_data(predictions,
      ssid = c(16),
      start_time = 2004,
      input_cell_size = 2,
      scale_fac = 1,
      delta_t_total = 10,
      delta_t_step = 2,
      indices = c(1, 1, 1, 1, 2, 2, 2, 2),
      variable_names = "est_exp",
      thresholds = c(0.001)
    )
  } else {
    out_im16 <- NULL
  }

  out_imm <- rbind(out_im1, out_im3, out_im4, out_im16)
  saveRDS(out_imm,
    file = paste0("data/", spp, "/trends-", spp, covs, "-", ssid_string, "-immature-biomass-priors-", priors, ".rds")
  )
} else {
  rm(out_imm)
}
```

```{r}
try(out_imm <- readRDS(paste0("data/", spp, "/trends-", spp, covs, "-", ssid_string, "-immature-biomass-priors-", priors, ".rds")))

if (exists("out_imm")) {
  data <- out_imm
  biotrend_im <- plot_vocc(data,
    fill_col = "units_per_decade",
    fill_label = "Immature \nbiomass \ntrend \n(kg/ha/decade)",
    raster_alpha = 1,
    vec_aes = NULL,
    white_zero = TRUE,
    transform_col = fourth_root_power,
    # axis_lables = TRUE,
    legend_position = c(0.2, 0.25)
  ) + ggtitle(paste0("", species, " ", min(unique(imm_predictions[["year"]])), "-", max(unique(imm_predictions[["year"]]))))
  biotrend_im
} else {
  biotrend_im <- grid::grid.rect(gp = grid::gpar(col = "white"))
}
```

```{r, error=TRUE}
if (exists("imm_predictions")) {
  p_prop_imm <- plot_facet_map(imm_predictions, "prop_imm") +
    labs(fill = "% \nby mass") +
    ggtitle(paste0("", species, " proportion immature"))
  print(p_prop_imm)
} else {
  p_prop_imm <- grid::grid.rect(gp = grid::gpar(col = "white"))
}
```

Retrieve (or calculate) predicted values for temperature
```{r, eval=FALSE}
# m_temp <- readRDS("../VOCC/data/m_temp_allyears.rds")
# temperature <- predict(m_temp, newdata = nd_all)
# saveRDS(temperature, file = "data/predicted_temp_allyears.rds")

temperature <- readRDS("data/predicted_temp_allyears.rds")
temperature <- filter(temperature, ssid %in% model_ssid)
```

```{r, eval=FALSE}
temp <- plot_facet_map(temperature, "est", transform_col = no_trans, viridis_option = "D") +
  labs(fill = "°C") +
  ggtitle("Bottom temperature")
print(temp)
```

```{r}
png(
  file = paste0("figs/", spp, "/trends-", spp, covs, "-", ssid_string, "-priors-", priors, ".png"), # The directory you want to save the file in
  res = 600,
  units = "in",
  width = 11, # The width of the plot in inches
  height = 8
)
gridExtra::grid.arrange(
  # p_adult, p_imm,
  biotrend_ad, biotrend_im,
  nrow = 1
)
dev.off()
```

```{r}
biomass_plots <- c(p_adult, p_imm)

png(
  file = paste0("figs/", spp, "/biomass-", spp, covs, "-", ssid_string, "-priors-", priors, ".png"), # The directory you want to save the file in
  res = 600,
  units = "in",
  width = 16, # The width of the plot in inches
  height = 8.5
)
gridExtra::grid.arrange(
  grobs = biomass_plots,
  nrow = 2
)
dev.off()
```

```{r}
png(
  file = paste0("figs/", spp, "/exploratory-", spp, covs, "-", ssid_string, "-priors-", priors, ".png"), # The directory you want to save the file in
  res = 600,
  units = "in",
  width = 9.5, # The width of the plot in inches
  height = 12.5
) # The height of the plot in inches

gridExtra::grid.arrange(
  p_adult_all, p_imm_all,
  biotrend_ad, biotrend_im,
  # current_biomass,
  p_prop_imm, temp,
  nrow = 3
)
dev.off()
```