### 2019-07-24 analysis/VOCC/3-variable-response-plots.Rmd
.Rmd to make all mountain plots for a species

---
title: "Plot responses to model covariates to inform threshold choices for VOCC"
author: "Philina English"
date: '2019-07-18'
output: html_document
---

Species
```{r}

# # Species run so far...
# species <- "Pacific Cod"
species <- "Sablefish"
# species <- "Silvergray Rockfish"
# species <- "Lingcod"
# species <- "North Pacific Spiny Dogfish" # note: using all data for maturity thresholds
# species <- "Quillback Rockfish"
# species <- "Pacific Ocean Perch"

# species <- "Yelloweye Rockfish"
# species <- "Arrowtooth Flounder"

```

Choose model details
```{r}
# covariates <- "+muddy+mixed+rocky"
# covariates <- "+mixed+rocky"
covariates <- "+muddy+any_rock"
# covariates <- "+trawled+muddy+rocky+mixed"
# covariates <- "+trawled+mixed+rocky"
# covariates <- "+trawled+mixed"
```

Run all subsequent code...
```{r global_options, include=FALSE}
covs <- gsub("\\+", "-", covariates)
spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))

# folder to hold figs for this species
dir.create(file.path("figs", spp))
dir.create(file.path("data", spp))

knitr::opts_chunk$set(
  fig.width = 11, fig.height = 8.5,
  fig.path = paste0("figs/", spp, "/"),
  echo = FALSE, warning = FALSE, message = FALSE
)
```

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(gfplot)
library(gfdata)
library(sdmTMB)
library(gfranges)
```

Dissolved O2 plots
```{r}
rm(
  adult_survey1, imm_survey1,
  adult_survey4, imm_survey4,
  adult_survey16, imm_survey16
)
priors <- TRUE
# covariates <- ""
try({
  survey <- c("SYN QCS", "SYN HS")
  model_ssid <- c(1, 3)
  ssid_string <- paste0(model_ssid, collapse = "n")

  adult_survey1 <- readRDS(paste0(
    "data/", spp,
    "/mod-mat-biomass-", spp, "-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))

  imm_survey1 <- readRDS(paste0(
    "data/", spp,
    "/mod-imm-biomass-", spp, "-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))
})
if (!exists("adult_survey1")) adult_survey1 <- NULL
if (!exists("imm_survey1")) imm_survey1 <- NULL

try({
  survey <- c("SYN WCVI")
  model_ssid <- c(4)
  ssid_string <- paste0(model_ssid, collapse = "n")

  adult_survey4 <- readRDS(paste0(
    "data/", spp,
    "/mod-mat-biomass-", spp, "-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))

  imm_survey4 <- readRDS(paste0(
    "data/", spp,
    "/mod-imm-biomass-", spp, "-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))
})
if (!exists("adult_survey4")) adult_survey4 <- NULL
if (!exists("imm_survey4")) imm_survey4 <- NULL

try({
  survey <- c("SYN WCHG")
  model_ssid <- c(16)
  ssid_string <- paste0(model_ssid, collapse = "n")

  adult_survey16 <- readRDS(paste0(
    "data/", spp,
    "/mod-mat-biomass-", spp, "-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))

  imm_survey16 <- readRDS(paste0(
    "data/", spp,
    "/mod-imm-biomass-", spp, "-do", covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))
})

if (!exists("adult_survey16")) adult_survey16 <- NULL
if (!exists("imm_survey16")) imm_survey16 <- NULL

model_list <- list(
  adult_survey1 = adult_survey1, imm_survey1 = imm_survey1,
  adult_survey4 = adult_survey4, imm_survey4 = imm_survey4,
  adult_survey16 = adult_survey16, imm_survey16 = imm_survey16
)

model_list <- model_list[!sapply(model_list, is.null)]
```

```{r}
d <- list()
do_plots <- list()
for (i in seq_len(length(model_list))) {
  d[[i]] <- time_varying_density(model_list[[i]], predictor = "do_mlpl")
  do_plots[[i]] <- plot_mountains(d[[i]], variable_label = "Dissolved O2", xlimits = c(0,8)) + ggtitle(names(model_list[i]))
}
print(do_plots)
```

```{r}
png(
  file = paste0("figs/", spp, "/do-", spp, covs, "-priors-TRUE.png"), 
  res = 600,
  units = "in",
  width = 8.5, 
  height = 11
)
gridExtra::grid.arrange(
  grobs = do_plots,
  ncol = 2,
  top = grid::textGrob(paste(species))
)
dev.off()
```

Temperature plots
```{r}
rm(
  adult_survey1, imm_survey1,
  adult_survey4, imm_survey4,
  adult_survey16, imm_survey16
)
priors <- TRUE
try({
  survey <- c("SYN QCS", "SYN HS")
  model_ssid <- c(1, 3)
  ssid_string <- paste0(model_ssid, collapse = "n")

  adult_survey1 <- readRDS(paste0(
    "data/", spp,
    "/mod-mat-biomass-", spp, "-temp", covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))

  imm_survey1 <- readRDS(paste0(
    "data/", spp,
    "/mod-imm-biomass-", spp, "-temp", covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))
})
if (!exists("adult_survey1")) adult_survey1 <- NULL
if (!exists("imm_survey1")) imm_survey1 <- NULL

try({
  survey <- c("SYN WCVI")
  model_ssid <- c(4)
  ssid_string <- paste0(model_ssid, collapse = "n")

  adult_survey4 <- readRDS(paste0(
    "data/", spp,
    "/mod-mat-biomass-", spp, "-temp", covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))

  imm_survey4 <- readRDS(paste0(
    "data/", spp,
    "/mod-imm-biomass-", spp, "-temp", covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))
})
if (!exists("adult_survey4")) adult_survey4 <- NULL
if (!exists("imm_survey4")) imm_survey4 <- NULL

try({
  survey <- c("SYN WCHG")
  model_ssid <- c(16)
  ssid_string <- paste0(model_ssid, collapse = "n")

  adult_survey16 <- readRDS(paste0(
    "data/", spp,
    "/mod-mat-biomass-", spp, "-temp", covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))

  imm_survey16 <- readRDS(paste0(
    "data/", spp,
    "/mod-imm-biomass-", spp, "-temp", covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))
})

if (!exists("adult_survey16")) adult_survey16 <- NULL
if (!exists("imm_survey16")) imm_survey16 <- NULL

temp_model_list <- list(
  adult_survey1 = adult_survey1, imm_survey1 = imm_survey1,
  adult_survey4 = adult_survey4, imm_survey4 = imm_survey4,
  adult_survey16 = adult_survey16, imm_survey16 = imm_survey16
)

temp_model_list <- temp_model_list[!sapply(temp_model_list, is.null)]
```

```{r}
d <- list()
temp_plots <- list()
for (i in seq_len(length(temp_model_list))) {
  d[[i]] <- time_varying_density(temp_model_list[[i]], predictor = "temp")
  #d[[i]] <- fixed_density(temp_model_list[[i]], predictor = "temp")
  temp_plots[[i]] <- plot_mountains(d[[i]], variable_label = "Temperature", xlimits = c(2,10)) + ggtitle(names(temp_model_list[i]))
}
print(temp_plots)
```

```{r}
png(
  file = paste0("figs/", spp, "/temp-", spp, covs, "-priors-TRUE.png"), 
  res = 600,
  units = "in",
  width = 8.5, 
  height = 11
)
gridExtra::grid.arrange(
  grobs = temp_plots,
  ncol = 2,
  top = grid::textGrob(paste(species))
)
dev.off()
```

Depth plots
```{r}
rm(
  adult_survey1, imm_survey1,
  adult_survey4, imm_survey4,
  adult_survey16, imm_survey16
)
priors <- FALSE
try({
  survey <- c("SYN QCS", "SYN HS")
  model_ssid <- c(1, 3)
  ssid_string <- paste0(model_ssid, collapse = "n")

  adult_survey1 <- readRDS(paste0(
    "data/", spp,
    "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))

  imm_survey1 <- readRDS(paste0(
    "data/", spp,
    "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))
})
if (!exists("adult_survey1")) adult_survey1 <- NULL
if (!exists("imm_survey1")) imm_survey1 <- NULL

try({
  survey <- c("SYN WCVI")
  model_ssid <- c(4)
  ssid_string <- paste0(model_ssid, collapse = "n")

  adult_survey4 <- readRDS(paste0(
    "data/", spp,
    "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))

  imm_survey4 <- readRDS(paste0(
    "data/", spp,
    "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))
})
if (!exists("adult_survey4")) adult_survey4 <- NULL
if (!exists("imm_survey4")) imm_survey4 <- NULL

try({
  survey <- c("SYN WCHG")
  model_ssid <- c(16)
  ssid_string <- paste0(model_ssid, collapse = "n")

  adult_survey16 <- readRDS(paste0(
    "data/", spp,
    "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))

  imm_survey16 <- readRDS(paste0(
    "data/", spp,
    "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))
})

if (!exists("adult_survey16")) adult_survey16 <- NULL
if (!exists("imm_survey16")) imm_survey16 <- NULL

depth_model_list <- list(
  adult_survey1 = adult_survey1, imm_survey1 = imm_survey1,
  adult_survey4 = adult_survey4, imm_survey4 = imm_survey4,
  adult_survey16 = adult_survey16, imm_survey16 = imm_survey16
)

depth_model_list <- depth_model_list[!sapply(depth_model_list, is.null)]
```

```{r}
d <- list()
depth_plots <- list()
for (i in seq_len(length(depth_model_list))) {
  d[[i]] <- time_varying_density(depth_model_list[[i]], predictor = "depth")
  d[[i]]$x <- exp(d[[i]]$x)
  depth_plots[[i]] <- plot_mountains(d[[i]], variable_label = "Depth", xlimits = c(0,1000)) + ggtitle(names(depth_model_list[i]))
}
print(depth_plots)
```

```{r}
png(
  file = paste0("figs/", spp, 
    "/depth-", spp, covs, "-priors-", priors, ".png"), 
  res = 600,
  units = "in",
  width = 8.5, 
  height = 11
)
gridExtra::grid.arrange(
  grobs = depth_plots,
  ncol = 2,
  top = grid::textGrob(paste(species))
)
dev.off()
```

Plot time-varying effect of trawling at mean depth
```{r}
trawl_plots <- list()
for (i in seq_len(length(depth_model_list))) {
  r <- depth_model_list[[i]]$tmb_obj$report()
  b3 <- r$b_rw_t[, 3]
  yrs <- sort(unique(depth_model_list[[i]]$data$year))
  b_j <- depth_model_list[[i]]$model$par[1:length(yrs)]

  SE <- 0.35 #Place-holder

  trawled <- data.frame(year = yrs, intercept = b_j, trawling = b3)

  # trawl_plots[[i]] <- ggplot(trawled, aes(year, y=exp(trawling) ) ) + 
  #   geom_point(colour ="red") + 
  #   #geom_pointrange(aes(ymin=exp(intercept-0.35), ymax=exp(intercept+0.35))) +
  #   ylab("Difference in kg/ha within trawled zone") + ggtitle(names(depth_model_list[i]))

  trawl_plots[[i]] <- ggplot(trawled, aes(year, exp(intercept))) + geom_point() +
    geom_point(aes(year, y = exp(intercept + trawling)), colour = "red") +
    geom_pointrange(aes(ymin = exp(intercept - (1.96 * SE)), ymax = exp(intercept + (1.96 * SE)))) + # using mean se on year effects
    ylim(min(exp(trawled$intercept - trawled$trawling - (1.96 * SE))), max(exp(trawled$intercept + trawled$trawling + (1.96 * SE)))) +
    ylab("Predicted kg/ha at mean depth") + ggtitle(names(depth_model_list[i]))
}
print(trawl_plots)
```

### Code for calculation SE on time-varying effects...

```{r eval=FALSE}
# FIXME: need to add to sdmTMB!
r <- m$tmb_obj$report()
summary(m)
yrs <- sort(unique(m$data$year))
.sd <- summary(m$sd_report)
.sd <- as.data.frame(.sd)
.sd$par <- row.names(.sd)
.sd <- .sd[grepl("^b_rw_t", .sd$par), ]
varnames <- colnames(model.matrix(m$time_varying, m$data))

rep(varnames, each = length(yrs))
.sd$parameter <- rep(varnames, each = length(yrs))
.sd$year <- rep(seq_along(yrs), each = length(varnames))
row.names(.sd) <- NULL
.sd$par <- NULL

names(.sd$Estimate) <- "estimate"
names(.sd[2]) <- "se"
.sd$ci <- .sd$se * 1.96
m$time_varying

nrow(.sd)

b3 <- r$b_rw_t[, 3]
yrs <- sort(unique(m$data$year))
b_j <- m$tmb_params$b_j[1:8]

trawled <- data.frame(year = yrs, intercept = b_j, trawling = b3)

```