### 2019-08-06 analysis/VOCC/4-vocc-by-spp.Rmd
linking vocc and species density code, still in development

---
title: "Biomass predictions and VOCC"
author: "Philina English"
date: '2019-07-18'
output: html_document
---

This file ...

Species

```{r}
# # Species run so far...
 species <- "Arrowtooth Flounder"
# species <- "Pacific Cod"
# species <- "Sablefish"
# species <- "Silvergray Rockfish"
#  species <- "Lingcod"
#  species <- "North Pacific Spiny Dogfish" # note: using all data for maturity thresholds
#  species <- "Quillback Rockfish"
#  species <- "Pacific Ocean Perch"
# species <- "Yelloweye Rockfish"
spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))
```

Region 
```{r}
# region <- "All synoptic surveys"
 region <- "West Coast Vancouver Island"
# region <- "Both odd year surveys"
# region <- "West Coast Haida Gwaii"
```

Make climate vectors for biannual changes between surveys
```{r}
trend <- FALSE
start_time <- 2008
end_time <- 2011

#climate <- "temperature"
temp_threshold <- c(0.75)
optimal_temp <- 6.5 # yellow-eye 4

#climate <- "do"
do_threshold <- c(0.25)
optimal_do <- 2 # yellow-eye 4

climate <- "do and temperature"
max_thresholds <- c(Inf, 0.75)
min_thresholds <- c(0.25, Inf)
optimal_value <- c(6.5, 2)
```

Make climate vectors for average changes and lower threshold values
```{r}
# trend <- TRUE
# 
# climate <- "temperature"
# temp_trend_threshold <- c(0.25)

# climate <- "DO"
# do_trend_threshold <- c(0.2)

# climate <- "both DO and temperature"
# min_trend_thresholds <- c(0.2, Inf)
# max_trend_thresholds <- c(Inf, 0.5)
```

Run all subsequent code...
```{r global_options, include=FALSE}
library(dplyr)
library(ggplot2)
library(gfplot)
library(gfdata)
library(sdmTMB)
library(gfranges)

knitr::opts_chunk$set(
  fig.width = 11, fig.height = 8.5,
  echo = FALSE, warning = FALSE, message = FALSE
)
```

```{r, message=FALSE, warning=FALSE}
if (region == "Both odd year surveys") {
  survey <- c("SYN QCS", "SYN HS")
  model_ssid <- c(1, 3)
  ssid_string <- paste0(model_ssid, collapse = "n")
  trend_indices_do <- c(1, 1, 1, 2, 2)
  trend_indices_temp <- c(1, 1, 1, 2, 2)
  years <- NULL
}

if (region == "West Coast Vancouver Island") {
  survey <- c("SYN WCVI")
  model_ssid <- c(4)
  ssid_string <- paste0(model_ssid, collapse = "n")
  trend_indices_do <- c(1, 1, 1, 2, 2)
  trend_indices_temp <- c(1, 1, 1, 2, 2, 2)
  years <- NULL
}

# if (region == "West Coast Haida Gwaii") {
#   survey <- c("SYN WCHG")
#   model_ssid <- c(16)
#   ssid_string <- paste0(model_ssid, collapse = "n")
#   trend_indices_do <- c(1, 1, 1, 2, 2)
#   trend_indices_temp <- c(1, 1, 1, 2, 2)
#   years <- NULL
# }
```

```{r}
# Set default cell sizes and scale
input_cell_size <- 2
scale_fac <- 1
delta_t_step <- 2
skip_time <- NULL

if (climate == "temperature") {
  variable_names <- c("temp")
  min_thresholds <- c(Inf)
  delta_threshold <- temp_threshold
  optimal_value <- optimal_temp

  if (trend) {
    max_thresholds <- temp_trend_threshold
    start_time <- 2008 # change if can add back in prior 2008 data
    end_time <- NULL
    delta_t_total <- 6 # change if can add back in prior 2008 data
    indices <- trend_indices_temp
  } else {
    max_thresholds <- temp_threshold
    indices <- c(1, 2)
    delta_t_total <- 2
  }
}

if (climate == "do") {
  variable_names <- c("do_est")
  max_thresholds <- c(Inf)
  delta_threshold <- -1*(do_threshold)
  optimal_value <- optimal_do

  if (trend) {
    min_thresholds <- do_trend_threshold
    start_time <- 2008
    skip_time <- 2016
    end_time <- NULL
    delta_t_total <- 6
    indices <- trend_indices_do
  } else {
    min_thresholds <- do_threshold
    indices <- c(1, 2)
    delta_t_total <- 2
  }
}

if (climate == "do and temperature") {
  variable_names <- c("do_est", "temp")

  if (trend) {
    min_thresholds <- min_trend_thresholds
    max_thresholds <- max_trend_thresholds
    start_time <- 2008
    skip_time <- 2016
    end_time <- NULL
    delta_t_total <- 6
    indices <- trend_indices_do
  } else {
    min_thresholds <- min_thresholds
    max_thresholds <- max_thresholds
    indices <- c(1, 2)
    delta_t_total <- 2
  }
}

min_string <- paste0(min_thresholds, collapse = "-")
max_string <- paste0(max_thresholds, collapse = "-")

optimal_string <- paste0(optimal_value, collapse = "-")
climate_string <- gsub(" ", "-", gsub("\\/", "-", tolower(climate_string)))
```

# Calculate climate vectors

```{r}
rm(vocc)
try({
  vocc <- readRDS(paste0("data/vocc-", ssid_string,"-", climate_string, "-", optimal_string, "-",
    min_string,"to", max_string, "-", start_time, "-", delta_t_total, ".rds"))
})

if (!exists("vocc")) {
pred_temp_do <- readRDS(here::here("analysis/tmb-sensor-explore/models/predicted-DO.rds"))

if (length(variable_names) > 1) {
  climate_predictions <- replicate(length(variable_names), pred_temp_do, simplify = FALSE)
} else {
  climate_predictions <- pred_temp_do
}

starttime <- Sys.time()
vocc <- make_vector_data(climate_predictions,
  variable_names = variable_names,
  ssid = model_ssid,
  start_time = start_time,
  end_time = end_time,
  skip_time = skip_time,
  input_cell_size = input_cell_size,
  scale_fac = scale_fac,
  delta_t_total = delta_t_total,
  delta_t_step = delta_t_step,
  indices = indices,
  min_thresholds = min_thresholds,
  max_thresholds = max_thresholds,
  round_fact = 10
)
endtime <- Sys.time()
time_vocc <- round(starttime - endtime)

vocc$var_1_min <- min_thresholds[1]
vocc$var_2_min <- min_thresholds[2]
vocc$var_1_max<- max_thresholds[1]
vocc$var_2_max<- max_thresholds[2]

saveRDS(vocc, file = paste0("data/vocc-", ssid_string,"-", climate_string, "-", optimal_string, "-",
    min_string,"to", max_string, "-", start_time, "-", delta_t_total, ".rds"))
}
# glimpse(vocc)
```

### Plot climate vectors
```{r}

if(climate== "temperature") {
vocc$diff <- vocc$units_per_decade/10*delta_t_total

vocc_plot <- plot_vocc(vocc,
  vec_aes = "distance",
  #max_vec_plotted = 100,
  min_vec_plotted = input_cell_size,
  NA_label = ".",
  fill_col = "diff",
  fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
  raster_alpha = 1,
  vec_alpha = 0.25,
  axis_lables = FALSE,
  transform_col = no_trans#,
  #raster_limits = c(5, 13)
)
vocc_plot + ggtitle(climate)
}

if (climate == "do") {
vocc$diff <- vocc$units_per_decade/10*delta_t_total

vocc_plot <- plot_vocc(vocc,
  vec_aes = "distance",
  #max_vec_plotted = 100,
  min_vec_plotted = input_cell_size,
  NA_label = ".",
  fill_col = "diff",
  fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
  raster_alpha = 1,
  high_fill = "Steel Blue 4",
  low_fill = "Red 3",
  vec_alpha = 0.25,
  axis_lables = FALSE,
  transform_col = no_trans#,
  #raster_limits = c(5, 13)
)
vocc_plot + ggtitle(climate)
}

if (climate == "do and temperature") {

vocc$temp_diff <- vocc$temp.units_per_decade/10*delta_t_total

vocc_plot_temp <- plot_vocc(vocc,
  vec_aes = "distance",
  #max_vec_plotted = 100,
  min_vec_plotted = input_cell_size,
  NA_label = ".",
  fill_col = "temp_diff",
  fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
  raster_alpha = 1,
  vec_alpha = 0.25,
  axis_lables = FALSE,
  transform_col = no_trans#,
  #raster_limits = c(5, 13)
)+ ggtitle("temperature")

vocc$do_diff <- vocc$do_est.units_per_decade/10*delta_t_total

vocc_plot_do <- plot_vocc(vocc,
  vec_aes = "distance",
  #max_vec_plotted = 100,
  min_vec_plotted = input_cell_size,
  NA_label = ".",
  fill_col = "do_diff",
  fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
  raster_alpha = 1,
  high_fill = "Steel Blue 4",
  low_fill = "Red 3",
  vec_alpha = 0.25,
  axis_lables = FALSE,
  transform_col = no_trans#,
  #raster_limits = c(5, 13)
) + ggtitle("do")

plot <- gridExtra::grid.arrange(
  vocc_plot_temp, vocc_plot_do,
  ncol = 2,
  top = grid::textGrob(paste(climate, "-", optimal_value, "-",
    min_string,"to", max_string, "-", start_time, "-", delta_t_total, ".rds"))
)

}

```

Get matching biomass data for the above time periods

Get all biomass predictions 
```{r}
priors <- FALSE
# priors <- TRUE

covariates <- "+muddy+any_rock"

covs <- gsub("\\+", "-", covariates)

rm(adult_predictions)
rm(imm_predictions)
rm(predictions)

try({
  adult_predictions <- readRDS(paste0("data/", spp,
  "/all-predictions-", spp, covs, "-mat-biomass-prior-", priors, ".rds"))
})

try({
  imm_predictions <- readRDS(paste0("data/", spp,
  "/all-predictions-", spp, covs, "-imm-biomass-prior-", priors, ".rds"))
})

try({
  predictions <- readRDS(paste0("data/", spp,
  "/all-predictions-", spp, covs, "-total-biomass-prior-", priors, ".rds"))
})

if (!exists("adult_predictions")) adult_predictions <- NULL
if (!exists("imm_predictions")) imm_predictions <- NULL
if (!exists("predictions")) predictions <- NULL

biomass_list <- list(
  adult = adult_predictions, 
  imm = imm_predictions, 
  total = predictions
)

biomass_list <- biomass_list[!sapply(biomass_list, is.null)]

# glimpse(biomass_list)
```

# Extract biomass values to match spatial and temporal range of climate vectors
```{r}
vocc_by_sp <- list()
for (i in seq_along(biomass_list)){

biomass_change <- make_trend_data(biomass_list[[i]],
  ssid = model_ssid,
  start_time = start_time,
  skip_time = skip_time,
  end_time = end_time,
  input_cell_size = input_cell_size,
  scale_fac = scale_fac,
  delta_t_total = delta_t_total,
  delta_t_step = delta_t_step,
  indices = indices,
  variable_names = "est_exp"
)

start <- biomass_change %>% 
  rename(start_density = var_1_s, end_density = var_1_e) %>% 
  select(icell, x, y, start_density, end_density)

end <- biomass_change %>% 
  rename(start_density = var_1_s, 
         end_density = var_1_e, 
         target_X = x, target_Y = y) %>% 
  select(icell, start_density, end_density, target_X, target_Y)

source_density <- biomass_change %>% rename(source_density = var_1_s) %>% select(source_density, x, y)

if (length(variable_names) == 1) {
  vectors <- vocc %>%
    mutate(
      start_cell = icell, 
      vector_id = paste0(id, "_to_", tid),
      initial_value = var_1_s, 
      source_end_value = var_1_e, 
      target_value = as.numeric(target_values)
    ) %>%
    filter(distance > 1)  %>%
    select(vector_id, start_cell, x, y, target_X, target_Y, distance, 
      initial_value, source_end_value, target_value, 
      km_per_decade, timespan)

  if (climate == "do") vectors <- vectors %>% filter(source_end_value < (optimal_do - do_threshold))
  if (climate == "temperature") vectors <- vectors %>% filter(source_end_value > (optimal_temp + temp_threshold))

} else {

  vectors <- vocc %>%
    mutate(
      start_cell = icell, vector_id = paste0(id, "_to_", tid),
      initial_value = s, source_end_value = e, target_value = target_values
    ) %>%
    filter(distance > 1) %>%
    select(vector_id, start_cell, x, y, target_X, target_Y, distance, 
      initial_value, source_end_value, target_value, 
      km_per_decade, timespan)

  source_end <- purrr::map_df(strsplit(vectors$source_end_value, " "), function(x) data.frame(t(as.numeric(x))))
  vectors <- cbind(vectors, source_end)

  for (j in seq_along(variable_names)){
  vectors <- vectors %>% filter( paste0("X",j) < (optimal_value[j] - min_thresholds[j]) )
  }

}

if (length(vectors[ ,2]) == 0) {
  try({other_coefs <- readRDS("data/vocc-by-spp-sdmTMB-coefs.RDS")})

  coefs$var <- "NA"
  coefs$beta <- "NA"
  coefs$se <- "NA"
  coefs$species <- spp 
  coefs$lifestage <- "NA"
  coefs$climate <- climate 
  coefs$min_thresholds <- min_thresholds
  coefs$max_thresholds <- max_thresholds
  coefs$optimum <- optimal_value
  coefs$start_time <- start_time
  coefs$timespan <- delta_t_total
  coefs$region <- region
  coefs$ssid_string <- ssid_string

  rbind(other_coefs, coefs)
  saveRDS(coefs, file = "data/vocc-by-spp-sdmTMB-coefs.RDS")

  stop("No vectors.")
}

vector_source <- left_join(vectors, start, by = c("x", "y")) %>% 
  mutate(cell_type = "source", end_climate = source_end_value, source_density = start_density)

vector_target <- left_join(vectors, end, by = c("target_X", "target_Y")) %>% 
  mutate(cell_type = "target", end_climate = target_value)
vector_target <- left_join(vector_target, source_density, by = c("x", "y")) #%>%  
  # filter(end_climate > 2)

# combine source and target cell data
vocc_by_sp[[i]] <- rbind(vector_source, vector_target) %>% 
  filter(!is.na(icell)) %>% 
# remove potential source cells with no fish  
  #FIXME: should their be a flexible min density of fish for a source cell?
  filter(source_density > 0.000001)

# get mean and sd for densities without cell duplication
icells <- vocc_by_sp[[i]] %>% distinct(icell, .keep_all = TRUE)
mean.change <- mean(icells$perc_change, na.rm = TRUE)
sd.change <- sd(icells$perc_change, na.rm = TRUE)

mean.density <- mean(icells$source_density, na.rm = TRUE)
sd.density <- sd(icells$source_density, na.rm = TRUE)

# calculate standardized density varibles
vocc_by_sp[[i]] <- vocc_by_sp[[i]] %>% 
  mutate(
    perc_change = (end_density - start_density)/start_density, 
    cell_type = as.factor(cell_type),
    std_change = (perc_change)/sd.change, 
    std_source_density = (source_density - mean.density)/(sd.density*2), 
    mean_change = mean.change, sd_change = sd.change,
    mean_source_density = mean.density, sd_density = sd.density,
# set distance intercept at cell size
    distance = (ifelse(distance > 100, 100, distance)) - input_cell_size, 
    distance2 = distance*distance) %>% 
  rename(X = x, Y = y)

if (length(variable_names) == 1) {
mean.climate <- mean(icells$end_climate, na.rm = TRUE)
sd.climate <- sd(icells$end_climate, na.rm = TRUE)

vocc_by_sp[[i]] <- vocc_by_sp[[i]] %>% 
  mutate(
    mean_climate = mean.climate, sd_climate = sd.climate,
    std_end_climate = (end_climate - mean.climate)/(sd.climate*2),
    std_end_climate2 = std_end_climate^2
    )

}else{

if (length(variable_names) == 2) {

vectors <- vectors %>% rename(
  source_end_climateA = X1, source_end_climateB = X2
  )

mean.climateA <- mean(icells$source_end_climateA, na.rm = TRUE)
sd.climateA <- sd(icells$source_end_climateA, na.rm = TRUE)
mean.climateB <- mean(icells$source_end_climateB, na.rm = TRUE)
sd.climateB <- sd(icells$source_end_climateB, na.rm = TRUE)

vocc_by_sp[[i]] <- vocc_by_sp[[i]] %>% 
  mutate(
    mean_climateA = mean.climateA, sd_climate = sd.climateA,
    std_end_climateA = (source_end_climateA - mean.climateA)/(sd.climateA*2),
    std_end_climateA2 = std_end_climateA^2,
    mean_climateB = mean.climateB, sd_climate = sd.climateB,
    std_end_climateB = (source_end_climateB - mean.climateB)/(sd.climateB*2),
    std_end_climateB2 = std_end_climateB^2
    )
} else {
 stop("Doesn't currently support testing vectors bases on more than two climate variables at once.")
}
}

}
names(vocc_by_sp) <- names(biomass_list)

# glimpse(icells)
# range(icells$perc_change)
# range(vocc_by_sp[[i]]$std_change)
# range(vocc_by_sp[[i]]$X)

# glimpse(vocc_by_sp[[i]])
# trim spatial outliers 
# Xmax <- quantile(vocc_by_sp[[i]]$X, 0.9995)
# Xmin <- quantile(vocc_by_sp[[i]]$X, 0.005)
# std_max <- quantile(vocc_by_sp[[i]]$std_change, 0.9995)
# 
# vocc_by_sp[[i]] <- vocc_by_sp[[i]] %>% filter ( X >= Xmin & X <= Xmax )

```

# Adult

## Adult plots

Check final climate-biomass relationship
```{r}
data <- vocc_by_sp[["adult"]]
#glimpse(data)

mature_plot <- ggplot(data, aes(x = end_climate, y = (end_density # - start_density
  ), colour = cell_type, #size = start_density, 
  group = cell_type)) +
  geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
  geom_smooth(method="lm", formula = y ~ x + I(x^2) + I(x^3)) +
  #   , aes(x = end_climate, y = (end_density #- start_density
  #  )), inherit.aes = FALSE) +
  scale_size_continuous(guide = "none") +
  scale_y_continuous(trans = fourth_root_power) +
  ylab(paste("Final biomass")) +
  xlab(paste("Final value of", climate,"for all cells")) +
  scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
  gfplot::theme_pbs() + theme(legend.position = c(0.8, 0.8)) + 
  ggtitle(paste(survey, "mature", species))

mature_plot
```

Check distance threshold of 50 looks reasonable
```{r}
data <- vocc_by_sp[["adult"]]
#data <- filter(vocc_by_sp[["adult"]], distance > 1 & distance < 100)

mature_plot2 <- ggplot(data, aes(x = distance, y = std_change, 
  group = cell_type, size = start_density, colour = end_climate)) +
  geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
  geom_smooth(method="lm", formula = y~x+I(x^2) )+
  facet_wrap(~cell_type) +#, scales = "free") +
  scale_size_continuous(guide = "none") +
  # scale_y_continuous(trans = fourth_root_power) +
  ylab(paste("Standardized change in biomass over", vectors$timespan[1], "years")) +
  xlab(paste("Distance traveled to maintain <", min_thresholds, " decline or", max_thresholds, "increase in", climate,"")) +
  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", climate)) + 
  # scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
  gfplot::theme_pbs() + theme(legend.position = c(0.8, 0.8)) + 
  ggtitle(paste(survey, "mature", species))

mature_plot2
```

```{r warnings=FALSE}
data <- vocc_by_sp[["adult"]]
#data <- filter(vocc_by_sp[["adult"]], distance < 50)
#data <- filter(vocc_by_sp[["adult"]], distance > 4 & distance < 50)

mature_plot3 <- ggplot(data, aes(x = source_density, y = std_change, 
    colour = end_climate, 
    alpha = 1/(distance+ input_cell_size), #size = 1/distance,
    group = cell_type)) +
  geom_jitter(width = 0.01) +
  geom_smooth(method="lm") +
  facet_wrap(~cell_type) + #, scales = "free") +
  scale_alpha_continuous(guide = "none") +
  scale_size_continuous(guide = "none") +
  scale_x_continuous(trans = sqrt) +
  scale_y_continuous(trans = sqrt) +
  ylab(paste("Standardized change in biomass over", vectors$timespan[1], "years")) +
  xlab(paste("Source density (dark points are closer to source)")) +
  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", climate)) + 
  gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
  ggtitle(paste(survey, "mature", species, "(mean biomass change =", round(data$mean_change[1], digits = 4), "kg/ha)"))

mature_plot3
```

```{r}
data <- vocc_by_sp[["adult"]]
data <- filter(vocc_by_sp[["adult"]], distance < 50)

mature_plot4 <- ggplot(filter(data, distance < 10), aes(source_density, y = std_change , 
    colour = end_climate, 
    alpha = 1/(distance+ input_cell_size), #size = 1/distance,
    group = cell_type)) +
  geom_jitter(width = 0.01) +
  geom_smooth(method="lm", aes(group = cell_type)) +
  facet_wrap(~cell_type) +#, scales = "free") +
  scale_alpha_continuous(guide = "none") +
  scale_size_continuous(guide = "none") +
  scale_x_continuous(trans = sqrt) +
  #scale_y_continuous(trans = fourth_root_power) +
  ylab(paste("Standardized change in biomass over", vectors$timespan[1], "years")) +
  xlab(paste("Source density (dark points are closer to source)")) +
  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", climate)) + 
  gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
  ggtitle(paste(survey, "mature", species, "(mean biomass change =", round(data$mean_change[1], digits = 4), "kg/ha)"))

mature_plot4
```

```{r}
data <- vocc_by_sp[["adult"]]
data <- filter(vocc_by_sp[["adult"]], distance < 50)

mature_plot5 <- ggplot(data, aes(x = end_climate - initial_value , y = #std_change
    (end_density - start_density)
  , colour = cell_type, size = start_density)) +
  geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
  scale_size_continuous(guide = "none") +
  #scale_y_continuous(trans = fourth_root_power) +
  ylab(paste("kg/ha change in biomass over", vectors$timespan[1], "years")) +
  xlab(paste("Change in", climate,"")) +
  scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
  gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
  ggtitle(paste(survey, "mature", species))
mature_plot5
```

## Adult models

### Regular mixed-model

```{r}
# test <- lmerTest::lmer(log(end_density) ~ cell_type * std_source_density + cell_type * distance + 
#             std_end_climate + std_end_climate2 + 
#             (1 | vector_id), 
#  #   offset = log(start_density), 
#  #   family = Gamma(link = "inverse")
#      data = data)
# hist(data$end_density)
# hist(data$std_change)
data <- vocc_by_sp[["adult"]]

test2 <- lmerTest::lmer(std_change ~ cell_type * std_source_density + 
    cell_type * distance + cell_type * distance2 +
    std_end_climate + std_end_climate2 + (1 | vector_id), REML = F, data = data)

summary(test2)
plot(test2)
```

Robust lmm
```{r eval=FALSE}
library(lme4)
data <- vocc_by_sp[["adult"]]

test3 <- robustlmm::rlmer(std_change ~ cell_type * std_source_density + 
    cell_type * distance + cell_type * distance2 + 
    std_end_climate + std_end_climate2 + (1 | vector_id), data = data)

summary(test3)
plot(test3, ask = FALSE)
```

### Boosted regression
```{r}
data <- vocc_by_sp[["adult"]]

m <- gbm::gbm(
  std_change ~ source_density + end_climate + distance + cell_type + X + Y , 
  data = data, 
  distribution = "gaussian",
  n.trees = 2000, 
  interaction.depth = 3, 
  shrinkage = 0.02
  )

# m <- gbm3::gbmt(end_density ~ start_density + source_density + end_climate + 
#   distance + cell_type + X + Y, data = data, distribution = gbm_dist("Gamma"))
```

```{r}
summary(m)
```

```{r}
plot(m,1)
plot(m,2)
plot(m,3)
plot(m,4)
plot(m,c(1,2))
plot(m,c(1,3))
plot(m,c(1,4))
plot(m,c(2,3))
plot(m,c(2,4))
plot(m,c(1,2,4))
plot(m,c(2,3,4))
plot(m,c(5,6))
```

### sdmTMB

```{r}
data <- vocc_by_sp[["adult"]]
# Xmax <- quantile(data$X, 0.9995)
# Xmin <- quantile(data$X, 0.005)
# std_max <- quantile(data$std_change, 0.9995)
# 
# data <- data %>% filter ( X >= Xmin & X <= Xmax )

xy <- as.data.frame(cbind(data$X, data$Y))

  if (region == "Both odd year surveys") {
    n_knots <- 200
  }

  if (region == "West Coast Vancouver Island") {
    n_knots <- 100
  }

  # if (region == "West Coast Haida Gwaii") {
  #   spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 60)
  # }

if (n_knots> nrow(unique(xy))) { n_knots <- nrow(unique(xy))}

  spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = n_knots)
  sdmTMB::plot_spde(spde)

```

```{r}
#data$end_climate2 <- data$end_climate^2
#data <- vocc_by_sp[["adult"]]
data$std_source_density2 <- data$std_source_density^2
test <- sdmTMB(data = data,
      std_change ~ cell_type * std_source_density + cell_type * std_source_density2 +
    cell_type * distance + cell_type * distance2 + 
      std_end_climate + std_end_climate2, 
      # std_change ~ cell_type * source_density + cell_type * distance  + end_climate + end_climate2,
      spde = spde, # control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = TRUE)
test
```

```{r}

p <- predict(test)
p$residuals <- residuals(test)
#idata$resids <- p$residuals
#glimpse(p)
ggplot(p, aes(est, residuals)) +
  geom_point(alpha=0.4) +
  ylim(-8,8) +
  geom_smooth()

plot_map_raster <- function(dat, column = "est") {
    ggplot(dat, aes_string("X", "Y", fill = column)) +
    geom_raster() +
    coord_fixed()+
    gfplot::theme_pbs() + 
    theme(axis.title.x = element_blank(), axis.title.y = element_blank())
}

plot_map_raster(p, "est")
plot_map_raster(p, "est_rf")
plot_map_raster(p, "omega_s")
plot_map_raster(p, "residuals")

```

# Immature

### Immature plots

Check final climate-biomass relationship
```{r}
idata <- vocc_by_sp[["imm"]]
#glimpse(data)

immature_plot1 <- ggplot(idata, aes(x = end_climate, y = (end_density # - start_density
  ), colour = cell_type, #size = start_density, 
  group = cell_type)) +
  geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
  geom_smooth(method="lm", formula = y ~ x + I(x^2) + I(x^3)) +
  #geom_smooth() +
  scale_size_continuous(guide = "none") +
  scale_y_continuous(trans = fourth_root_power) +
  ylab(paste("Final biomass")) +
  xlab(paste("Final value of", climate,"for all cells")) +
  scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
  gfplot::theme_pbs() + theme(legend.position = c(0.8, 0.8)) + 
  ggtitle(paste(survey, "immature", species))

immature_plot1
```

Check distance threshold of 50 looks reasonable
```{r warnings=FALSE}
idata <- vocc_by_sp[["imm"]]
#data <- filter(vocc_by_sp[["adult"]], distance > 1 & distance < 100)

immature_plot2 <- ggplot(idata, aes(distance, y = std_change, 
  group = cell_type, size = start_density, colour = end_climate)) +
  geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
  geom_smooth(method="lm", formula = y~x+I(x^2) )+
  facet_wrap(~cell_type) + #, scales = "free") +
  scale_size_continuous(guide = "none") +
  scale_y_continuous(trans = sqrt) +
  ylab(paste("Standardized change in biomass over", vectors$timespan[1], "years")) +
  xlab(paste("Distance traveled to maintain <", min_thresholds, " decline or", max_thresholds, "increase in", climate,"")) +
  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", climate)) + 
  # scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
  gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
  ggtitle(paste(survey, "immature", species))

immature_plot2
```

```{r warnings=FALSE}
idata <- vocc_by_sp[["imm"]]
#idata <- filter(vocc_by_sp[["imm"]], distance < 50)
#idata <- filter(vocc_by_sp[[""imm"]], distance > 4 & distance < 50)

immature_plot3 <- ggplot(idata, aes(x = source_density, y = std_change, 
    colour = end_climate, 
    alpha = 1/(distance+ input_cell_size), #size = 1/distance,
    group = cell_type)) +
  geom_jitter(width = 0.01) +
  geom_smooth(method="lm") +
  facet_wrap(~cell_type) + #, scales = "free") +
  scale_alpha_continuous(guide = "none") +
  scale_size_continuous(guide = "none") +
  scale_x_continuous(trans = sqrt) +
  scale_y_continuous(trans = sqrt) +
  ylab(paste("Standardized change in biomass over", vectors$timespan[1], "years")) +
  xlab(paste("Source density (dark points are closer to source)")) +
  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", climate)) + 
  gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
  ggtitle(paste(survey, "immature", species, "(mean biomass change =", round(data$mean_change[1], digits = 4), "kg/ha)"))

immature_plot3
```

```{r warnings=FALSE}
idata <- vocc_by_sp[["imm"]]
idata <- filter(vocc_by_sp[["imm"]], distance < 50)

immature_plot4 <- ggplot(filter(idata, distance < 10), aes(source_density, y = std_change , 
    colour = end_climate, 
    alpha = 1/(distance+ input_cell_size), #size = 1/distance,
    group = cell_type)) +
  geom_jitter(width = 0.01) +
  geom_smooth(method="lm") +
  facet_wrap(~cell_type) +#, scales = "free") +
  scale_alpha_continuous(guide = "none") +
  scale_size_continuous(guide = "none") +
  scale_x_continuous(trans = sqrt) +
  scale_y_continuous(trans = sqrt) +
  ylab(paste("Standardized change in biomass over", vectors$timespan[1], "years")) +
  xlab(paste("Source density (dark points are closer to source)")) +
  scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", climate)) + 
  gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
  ggtitle(paste(survey, "immature", species, "(mean biomass change =", round(data$mean_change[1], digits = 4), "kg/ha)"))

immature_plot4
```

```{r}
idata <- vocc_by_sp[["imm"]]
idata <- filter(vocc_by_sp[["imm"]], distance < 50)

immature_plot5 <- ggplot(idata, aes(x = end_climate - initial_value , y = (end_density - start_density
  ), colour = cell_type, size = start_density)) +
  geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
  scale_size_continuous(guide = "none") +
  #scale_y_continuous(trans = fourth_root_power) +
  ylab(paste("kg/ha change in biomass over", vectors$timespan[1], "years")) +
  xlab(paste("Change in", climate,"")) +
  scale_colour_viridis_d(begin = 0.1, end = 0.5) + 
  gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) + 
  ggtitle(paste(survey, "immature", species))
immature_plot5
```

## Immature models

### Regular mixed-model
```{r}
idata <- vocc_by_sp[["imm"]]

# Xmax <- quantile(idata$X, 0.9995)
# Xmin <- quantile(idata$X, 0.005)
# std_max <- quantile(idata$std_change, 0.9995)
# 
# idata <- idata %>% filter ( X >= Xmin & X <= Xmax )
#  
itest <- lmerTest::lmer(std_change ~ 
      cell_type * std_source_density + 
      cell_type * distance + cell_type * distance2 +
      std_end_climate + std_end_climate2 + (1 | vector_id), 
      REML = F, data = idata)
summary(itest)
plot(itest)
```
```{r eval=FALSE}
library(lme4)
itest3 <- robustlmm::rlmer(std_change ~ cell_type * std_source_density + 
    cell_type * distance + cell_type * distance2 + 
    std_end_climate + std_end_climate2 + (1 | vector_id), data = idata)
summary(itest3)
plot(itest3, ask = FALSE)
```

### Boosted regression

```{r}
idata <- vocc_by_sp[["imm"]]

# Xmax <- quantile(idata$X, 0.9995)
# Xmin <- quantile(idata$X, 0.005)
# std_max <- quantile(idata$std_change, 0.9995)
# 
# idata <- idata %>% filter ( X >= Xmin & X <= Xmax )

mi <- gbm::gbm(
  std_change ~ source_density + end_climate + distance + cell_type + X + Y , 
  data = idata, 
  distribution = "gaussian",
  n.trees = 2000, 
  interaction.depth = 3, 
  shrinkage = 0.02
  )

# m <- gbm3::gbmt(end_density ~ start_density + source_density + end_climate + 
#   distance + cell_type + X + Y, data = data, distribution = gbm_dist("Gamma"))
```

```{r}
summary(mi)
```

```{r}
plot(mi,1)
plot(mi,2)
plot(mi,3)
plot(mi,4)
plot(mi,c(1,2))
plot(mi,c(1,3))
plot(mi,c(1,4))
plot(mi,c(2,3))
plot(mi,c(2,4))
plot(mi,c(1,2,4))
plot(mi,c(2,3,4))
plot(mi,c(5,6))
```

### sdmTMB

```{r}
idata <- vocc_by_sp[["imm"]]

# Xmax <- quantile(data$X, 0.9995)
# Xmin <- quantile(data$X, 0.005)
# std_max <- quantile(data$std_change, 0.9995)
# 
# data <- data %>% filter ( X >= Xmin & X <= Xmax )

ixy <- as.data.frame(cbind(idata$X, idata$Y))

  if (region == "Both odd year surveys") {
    n_knots <- 200
  }

  if (region == "West Coast Vancouver Island") {
    n_knots <- 100
  }

  # if (region == "West Coast Haida Gwaii") {
  #    n_knots <- 60
  # }

if (n_knots> nrow(unique(xy))) { n_knots <- nrow(unique(ixy))}

  ispde <- sdmTMB::make_spde(idata$X, idata$Y, n_knots = n_knots)
  sdmTMB::plot_spde(ispde)

```

```{r }
#glimpse(idata)

idata$std_source_density2 <- idata$std_source_density^2
itest <- sdmTMB(data = idata, 
      std_change ~ cell_type * std_source_density + cell_type * std_source_density2 + 
      cell_type * distance + cell_type * distance2 + 
      std_end_climate + std_end_climate2, 
      spde = ispde, silent = TRUE)
itest

```

```{r}
ip <- predict(itest)
ip$residuals <- residuals(itest)
#idata$resids <- p$residuals
#glimpse(p)
ggplot(ip, aes(est, residuals)) +
  geom_point(alpha=0.4) +
  ylim(-8,8) +
  geom_smooth()

plot_map_raster <- function(dat, column = "est") {
    ggplot(dat, aes_string("X", "Y", fill = column)) +
    geom_raster() +
    coord_fixed()+
    gfplot::theme_pbs() + 
    theme(axis.title.x = element_blank(), axis.title.y = element_blank())
}

plot_map_raster(ip, "est")
plot_map_raster(ip, "est_rf")
plot_map_raster(ip, "omega_s")
plot_map_raster(ip, "residuals")
```

```{r}
m <- test

varnames <- colnames(m$tmb_data$X_ij)

coefs <- summary(m$sd_report)
coefs <- as.data.frame(coefs)
names(coefs) <- c("beta","se")
coefs$ci <- coefs$se*1.96
var <- varnames
var[1] <- "intercept"
coefs <- coefs[1:length(var), ]
row.names(coefs) <- NULL

coefs <-cbind(var, coefs)
#coefs$var
#coefs$beta
#coefs$se
coefs$species <- spp 
coefs$lifestage <- "adult"
coefs$climate <- climate 
coefs$min_thresholds <- min_thresholds
coefs$max_thresholds <- max_thresholds
coefs$optimum <- optimal_value
coefs$start_time <- start_time
coefs$timespan <- delta_t_total
coefs$region <- region
coefs$ssid_string <- ssid_string

glimpse(coefs)
rm(other_coefs)
try({other_coefs <- readRDS("data/vocc-by-spp-sdmTMB-coefs.RDS")})

if (exists("other_coefs")) {
coefs <- rbind(other_coefs, coefs)
}
saveRDS(coefs, file = "data/vocc-by-spp-sdmTMB-coefs.RDS")

data$mean_source_density[1]
range(data$source_density)
```

```{r}
m <- itest

varnames <- colnames(m$tmb_data$X_ij)

coefs <- summary(m$sd_report)
coefs <- as.data.frame(coefs)
names(coefs) <- c("beta","se")
coefs$ci <- coefs$se*1.96
var <- varnames
var[1] <- "intercept"
coefs <- coefs[1:length(var), ]
row.names(coefs) <- NULL
coefs <-cbind(var, coefs)

coefs$species <- spp 
coefs$lifestage <- "immature"
coefs$climate <- climate 
coefs$min_thresholds <- min_thresholds
coefs$max_thresholds <- max_thresholds
coefs$optimum <- optimal_value
coefs$start_time <- start_time
coefs$timespan <- delta_t_total
coefs$region <- region
coefs$ssid_string <- ssid_string

glimpse(coefs)
try({other_coefs <- readRDS("data/vocc-by-spp-sdmTMB-coefs.RDS")})
if (exists("other_coefs")) {
coefs <- rbind(other_coefs, coefs)
}
saveRDS(coefs, file = "data/vocc-by-spp-sdmTMB-coefs.RDS")

# intercept = mean std percent change in biomass for cell with avarage source biomass
idata$mean_source_density[1]
range(idata$source_density)
```

```{r}
# png(
#   file = paste0("figs/", spp, "/vocc-biomass-change", spp, ssid_string, ".png"),
#   res = 600,
#   units = "in",
#   width = 5,
#   height = 4
# )
# 
# do_vect_plot
# dev.off()
```