### 2019-08-22 analysis/VOCC/4-vocc-BACI.Rmd
exact matching option

---
title: "BACI of biomass change for VOCC source cells"
author: "Philina English"
date: '2019-08-15'
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 11, fig.height = 8.5,
  echo = FALSE, warning = FALSE, message = FALSE
)
options(scipen = 999)
options(digits = 4)
```

#### Set species
```{r echo=TRUE}
# # Species run so far...
# species <- "Arrowtooth Flounder"
# species <- "Pacific Cod"
# species <- "Sablefish"
species <- "Silvergray Rockfish"
# species <- "Lingcod"
# species <- "North Pacific Spiny Dogfish" # note: using all data for maturity thresholds
# species <- "Quillback Rockfish"
# species <- "Pacific Ocean Perch"
# species <- "Yelloweye Rockfish"
```

#### Get species-specific optimal values from mountain plots for most abundant year
```{r, echo=FALSE}
spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))

params <- readRDS(paste0("data/optimal-values-by-species.rds"))
params <- params[rev(order(as.Date(params$date))), ]
params <- params[!duplicated(params$species), ]

optimal_do <- round(params[params$species == species, ]$optimal_do, digits = 2)
optimal_temp <- round(params[params$species == species, ]$optimal_temp, digits = 2)

optimal_values <- c(optimal_do, optimal_temp)
```

#### Set vector type
```{r echo=TRUE}
 climate <- "temperature"
# climate <- "do"
# climate <- "do and temperature"
```

#### Make climate vectors for biannual changes between surveys
```{r echo=TRUE}
trend <- FALSE
start_time <- 2012
end_time <- 2015
temp_threshold <- c(0.5)
do_threshold <- c(0.1)
```

# OR

#### Make climate vectors for average changes and lower threshold values
```{r echo=TRUE}
# trend <- TRUE
# temp_trend_threshold <- c(0.25)
# do_trend_threshold <- c(0.2)
```

#### Set Region 
```{r echo=TRUE}
# region <- "West Coast Vancouver Island"
region <- "both odd year surveys"
# region <- "West Coast Haida Gwaii"
```

Run all subsequent code...
```{r, include=FALSE}
library(dplyr)
library(ggplot2)
library(gfplot)
# library(gfdata)
library(sdmTMB)
#library(itsadug)
library(gfranges)
```

```{r, message=FALSE, warning=FALSE, include=FALSE}
if (region == "both odd year surveys") {
  survey <- c("SYN QCS", "SYN HS")
  model_ssid <- c(1, 3)
  ssid_string <- paste0(model_ssid, collapse = "n")
  trend_indices_do <- c(1, 1, 1, 2, 2)
  trend_indices_temp <- c(1, 1, 1, 2, 2)
  years <- NULL
}

if (region == "West Coast Vancouver Island") {
  survey <- c("SYN WCVI")
  model_ssid <- c(4)
  ssid_string <- paste0(model_ssid, collapse = "n")
  trend_indices_do <- c(1, 1, 1, 2, 2)
  trend_indices_temp <- c(1, 1, 1, 2, 2, 2)
  years <- NULL
}

if (region == "West Coast Haida Gwaii") {
  survey <- c("SYN WCHG")
  model_ssid <- c(16)
  ssid_string <- paste0(model_ssid, collapse = "n")
  trend_indices_do <- c(1, 1, 1, 2, 2)
  trend_indices_temp <- c(1, 1, 1, 2, 2)
  years <- NULL
}
```

```{r, include=FALSE}
# Set default cell sizes and scale
input_cell_size <- 2
scale_fac <- 1
delta_t_step <- 2
skip_time <- NULL

if (climate == "temperature") {
  variable_names <- c("temp")
  min_thresholds <- c(Inf)
  #delta_threshold <- temp_threshold
  optimal_values <- optimal_temp

  if (trend) {
    max_thresholds <- temp_trend_threshold
    start_time <- 2008 # change if can add back in prior 2008 data
    end_time <- NULL
    delta_t_total <- 6 # change if can add back in prior 2008 data
    indices <- trend_indices_temp
  } else {
    max_thresholds <- temp_threshold
    indices <- c(1, 2)
    delta_t_total <- 2
  }
}

if (climate == "do") {
  variable_names <- c("do_est")
  max_thresholds <- c(Inf)
  #delta_threshold <- -1 * (do_threshold)
  optimal_values <- optimal_do

  if (trend) {
    min_thresholds <- do_trend_threshold
    start_time <- 2008
    skip_time <- 2016
    end_time <- NULL
    delta_t_total <- 6
    indices <- trend_indices_do
  } else {
    min_thresholds <- do_threshold
    indices <- c(1, 2)
    delta_t_total <- 2
  }
}

if (climate == "do and temperature") {
  variable_names <- c("do_est", "temp")

  if (trend) {
    min_thresholds <- c(do_trend_threshold, Inf)
    max_thresholds <- c(Inf, temp_trend_threshold)
    start_time <- 2008
    skip_time <- 2016
    end_time <- NULL
    delta_t_total <- 6
    indices <- trend_indices_do
  } else {
    min_thresholds <- c(do_threshold, Inf)
    max_thresholds <- c(Inf, temp_threshold)
    indices <- c(1, 2)
    delta_t_total <- 2
  }
}

min_string <- paste0(min_thresholds, collapse = "-")
max_string <- paste0(max_thresholds, collapse = "-")

optimal_string <- paste0(optimal_values, collapse = "-")
climate_string <- gsub(" ", "-", gsub("\\/", "-", tolower(climate)))
```

# Calculate climate vectors

```{r}
rm(vocc)
try({
  vocc <- readRDS(paste0(
    "data/vocc-", ssid_string, "-", climate_string, "-",
    min_string, "to", max_string, "-", start_time, "-", delta_t_total, ".rds"
  ))
})

if (!exists("vocc")) {
  pred_temp_do <- readRDS(here::here("analysis/tmb-sensor-explore/models/predicted-DO.rds"))

  if (length(variable_names) > 1) {
    climate_predictions <- replicate(length(variable_names), pred_temp_do, simplify = FALSE)
  } else {
    climate_predictions <- pred_temp_do
  }

  starttime <- Sys.time()
  vocc <- make_vector_data(climate_predictions,
    variable_names = variable_names,
    ssid = model_ssid,
    start_time = start_time,
    end_time = end_time,
    skip_time = skip_time,
    input_cell_size = input_cell_size,
    scale_fac = scale_fac,
    delta_t_total = delta_t_total,
    delta_t_step = delta_t_step,
    indices = indices,
    min_thresholds = min_thresholds,
    max_thresholds = max_thresholds,
    round_fact = 10
  )
  endtime <- Sys.time()
  time_vocc <- round(starttime - endtime)

  vocc$var_1_min <- min_thresholds[1]
  vocc$var_2_min <- min_thresholds[2]
  vocc$var_1_max <- max_thresholds[1]
  vocc$var_2_max <- max_thresholds[2]

  saveRDS(vocc, file = paste0(
    "data/vocc-", ssid_string, "-", climate_string, "-",
    min_string, "to", max_string, "-", start_time, "-", delta_t_total, ".rds"
  ))
}
# glimpse(vocc)
```

Plot all possible vectors
```{r echo=FALSE}
if (climate == "temperature") {
  vocc$diff <- vocc$units_per_decade / 10 * delta_t_total

  vocc_plot <- plot_vocc(vocc,
    vec_aes = "distance",
    min_vec_plotted = input_cell_size,
    NA_label = ".",
    fill_col = "diff",
    fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
    raster_alpha = 1,
    white_zero = TRUE,
    vec_alpha = 0.25,
    axis_lables = FALSE,
    transform_col = no_trans 
  ) + ggtitle(climate)
}

if (climate == "do") {
  vocc$diff <- vocc$units_per_decade / 10 * delta_t_total

  vocc_plot <- plot_vocc(vocc,
    vec_aes = "distance",
    min_vec_plotted = input_cell_size,
    NA_label = ".",
    fill_col = "diff",
    fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
    raster_alpha = 1,
    white_zero = TRUE,
    high_fill = "Steel Blue 4",
    low_fill = "Red 3",
    vec_alpha = 0.25,
    axis_lables = FALSE,
    transform_col = no_trans 
  ) + ggtitle(climate)
}

if (climate == "do and temperature") {
  vocc$temp_diff <- vocc$temp.units_per_decade / 10 * delta_t_total

  vocc_plot_temp <- plot_vocc(vocc,
    vec_aes = "distance",
    min_vec_plotted = input_cell_size,
    NA_label = ".",
    fill_col = "temp_diff",
    fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
    raster_alpha = 1,
    white_zero = TRUE,
    vec_alpha = 0.25,
    axis_lables = FALSE,
    transform_col = no_trans 
  ) + ggtitle("temperature")

  vocc$do_diff <- vocc$do_est.units_per_decade / 10 * delta_t_total

  vocc_plot_do <- plot_vocc(vocc,
    vec_aes = "distance",
    min_vec_plotted = input_cell_size,
    NA_label = ".",
    fill_col = "do_diff",
    fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
    raster_alpha = 1,
    white_zero = TRUE,
    high_fill = "Steel Blue 4",
    low_fill = "Red 3",
    vec_alpha = 0.25,
    axis_lables = FALSE,
    transform_col = no_trans 
  ) + ggtitle("do")

  vocc_plot <- gridExtra::grid.arrange(
    vocc_plot_temp, vocc_plot_do,
    ncol = 2,
    top = grid::textGrob(paste("do of ", optimal_values[1], "less", min_thresholds[1], "or temperature of ", optimal_values[2], "plus", max_thresholds[2], " over ", start_time, "-", start_time + delta_t_total, ""))
  )
}
vocc_plot
```

Filter vectors to match optimal range for chosen species
```{r}
vectors <- trim_vector_data(vocc, variable_names, optimal_values, min_thresholds, max_thresholds,
  cell_size = input_cell_size, dist_intercept = input_cell_size/2,
  max_dist = 120
)

# if (length(vectors[, 2]) == 0) {
#   other_coefs <- readRDS(paste0("data/", climate_string, "-vocc-by-spp-sdmTMB-coefs.RDS"))
#   coefs <- other_coefs[1, ]
#   coefs$var <- "NA"
#   coefs$beta <- "NA"
#   coefs$se <- "NA"
#   coefs$model <- "NA"
#   coefs$species <- spp
#   coefs$lifestage <- "NA"
#   coefs$climate <- climate
#   coefs$min_thresholds <- min_string
#   coefs$max_thresholds <- max_string
#   coefs$optimum <- optimal_string
#   coefs$start_time <- start_time
#   coefs$timespan <- delta_t_total
#   coefs$region <- region
#   coefs$ssid_string <- ssid_string
# 
#   coefs <- rbind(other_coefs, coefs) %>% distinct()
#   saveRDS(coefs, file = paste0("data/", climate_string, "-vocc-by-spp-sdmTMB-coefs.RDS"))
# 
#   stop("No vectors. Conditions saved to 'data/vocc-byspp-sdmTMB-coefs'.")
# }
# glimpse(vocc)
# glimpse(vectors)
```

### Plot remaining climate vectors for regions where change exceeds thresholds
```{r echo=FALSE, warning=FALSE}

if (climate == "temperature") {
  vectors$diff <- vectors$units_per_decade / 10 * delta_t_total

  vocc_plot <- plot_vocc(vectors,
    vec_aes = "distance",
    min_vec_plotted = input_cell_size,
    NA_label = ".",
    fill_col = "diff",
    fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
    raster_alpha = 1,
    white_zero = TRUE,
    vec_alpha = 0.25,
    axis_lables = FALSE,
    transform_col = no_trans 
  ) + ggtitle(climate)
}

if (climate == "do") {
  vectors$diff <- vectors$units_per_decade / 10 * delta_t_total

  vocc_plot <- plot_vocc(vectors,
    vec_aes = "distance",
    min_vec_plotted = input_cell_size,
    NA_label = ".",
    fill_col = "diff",
    fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
    raster_alpha = 1,
    white_zero = TRUE,
    high_fill = "Steel Blue 4",
    low_fill = "Red 3",
    vec_alpha = 0.25,
    axis_lables = FALSE,
    transform_col = no_trans 
  ) + ggtitle(climate)
}

if (climate == "do and temperature") {
  vectors$temp_diff <- vectors$temp.units_per_decade / 10 * delta_t_total

  vocc_plot_temp <- plot_vocc(vectors,
    vec_aes = "distance",
    min_vec_plotted = input_cell_size,
    NA_label = ".",
    fill_col = "temp_diff",
    fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
    raster_alpha = 1,
    white_zero = TRUE,
    vec_alpha = 0.25,
    axis_lables = FALSE,
    transform_col = no_trans 
  ) + ggtitle("temperature")

  vectors$do_diff <- vectors$do_est.units_per_decade / 10 * delta_t_total

  vocc_plot_do <- plot_vocc(vectors,
    vec_aes = "distance",
    min_vec_plotted = input_cell_size,
    NA_label = ".",
    fill_col = "do_diff",
    fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
    raster_alpha = 1,
    white_zero = TRUE,
    high_fill = "Steel Blue 4",
    low_fill = "Red 3",
    vec_alpha = 0.25,
    axis_lables = FALSE,
    transform_col = no_trans 
  ) + ggtitle("do")

  vocc_plot <- gridExtra::grid.arrange(
    vocc_plot_temp, vocc_plot_do,
    ncol = 2,
    top = grid::textGrob(paste("DO of ", optimal_values[1], "less", min_thresholds[1], "or temperature of ", optimal_values[2], "plus", max_thresholds[2], " over ", start_time, "-", start_time + delta_t_total, ""))
  )
}
vocc_plot
```
# Add species biomass data

#### Get all biomass predictions 
```{r, echo=FALSE, warning=FALSE}
priors <- FALSE
# priors <- TRUE

covariates <- "+muddy+any_rock"

covs <- gsub("\\+", "-", covariates)

rm(adult_predictions)
rm(imm_predictions)
rm(predictions)

try({
  adult_predictions <- readRDS(paste0(
    "data/", spp,
    "/all-predictions-", spp, covs, "-mat-biomass-prior-", priors, ".rds"
  ))
})

try({
  imm_predictions <- readRDS(paste0(
    "data/", spp,
    "/all-predictions-", spp, covs, "-imm-biomass-prior-", priors, ".rds"
  ))
})

try({
  predictions <- readRDS(paste0(
    "data/", spp,
    "/all-predictions-", spp, covs, "-total-biomass-prior-", priors, ".rds"
  ))
})

if (!exists("adult_predictions")) adult_predictions <- NULL
if (!exists("imm_predictions")) imm_predictions <- NULL
if (!exists("predictions")) predictions <- NULL

biomass_list <- list(
  adult = adult_predictions,
  imm = imm_predictions,
  total = predictions
)

biomass_list <- biomass_list[!sapply(biomass_list, is.null)]

# glimpse(biomass_list)
```

#### Extract biomass values to match spatial and temporal range of climate vectors

```{r}
#i <- biomass_list[[1]]

vocc_by_sp <- lapply(biomass_list, function(i) {
  biomass_change <- make_trend_data(i,
    ssid = model_ssid,
    start_time = start_time,
    skip_time = skip_time,
    end_time = end_time,
    input_cell_size = input_cell_size,
    scale_fac = scale_fac,
    delta_t_total = delta_t_total,
    delta_t_step = delta_t_step,
    indices = indices,
    variable_names = "est_exp"
  )

  biomass_covs <- i %>% select(X, Y, depth, trawled, muddy, any_rock, omega_s) %>% distinct()

  biomass_change <- biomass_change %>% 
    rename(start_density = est_exp_s, end_density = est_exp_e) %>%
    dplyr::select(icell, x, y, start_density, end_density)
    # dplyr::select(-N, -time_step, - units_per_decade, -slope)

  biomass  <- left_join(biomass_change,  biomass_covs, by = c("x" = "X", "y"="Y"))

  # keep just one row for each cell that exceeded the end climate threshold 
  source_cells <- vectors %>% distinct(icell, .keep_all = TRUE) %>% 
    select(-target_X, -target_Y, -target_values, -tid)

  source_biomass <- left_join(source_cells, biomass, by = c("icell", "x", "y")) %>%
    mutate(cell_type = "source") %>% filter(!is.na(start_density))

  # identify all possible control cells, assume distance = 0 given 
  control_cells <- anti_join(vocc, vectors, by = c("icell", "x", "y")) %>% distinct(icell, .keep_all = TRUE) %>% 
    mutate(distance = 0) %>% 
    select(-target_X, -target_Y, -target_values, -tid)

  control_biomass <- left_join(control_cells, biomass, by = c("icell", "x", "y")) %>%
    mutate(cell_type = "control") %>% filter(!is.na(start_density))

#browser()
  alldata <- rbind(source_biomass, control_biomass) 
  alldata <- mutate(alldata, treat = ifelse(cell_type=="source", 1, 0)) %>% 
    select(-var_2_min, -var_2_max) %>% 
    rename(vect_dist = distance)
  alldata$treat <- as.factor(alldata$treat)
  ggplot(alldata, aes(x=cell_type, y=log(start_density))) + geom_boxplot()

  ## still including this trimming to reduce size of dataset 
  control_min <- min(alldata$start_density[alldata$treat=="0"])
  source_min <- min(alldata$start_density[alldata$treat=="1"])
  # control_min <- quantile(alldata$start_density[alldata$treat=="0"], 0.001)
  # source_min <- quantile(alldata$start_density[alldata$treat=="1"], 0.001)
  group_min <- max(control_min, source_min)

  control_max <- max(alldata$start_density[alldata$treat=="0"])
  source_max <- max(alldata$start_density[alldata$treat=="1"])
  # control_max <- quantile(alldata$start_density[alldata$treat=="0"], 0.95)
  # source_max <- quantile(alldata$start_density[alldata$treat=="1"], 0.95)
  group_max <- min(control_max, source_max)
  # 
   alldata <- filter(alldata, start_density >= group_min) #%>% filter(start_density <= group_max)
  # 
  # num_source <- nrow(alldata[alldata$treat=="1", ])
  #

  # alldata <- filter(alldata, start_density >= control_min)

  # low_biomass <- quantile(alldata$start_density, 0.1)
  # alldata <- filter(alldata, start_density >= low_biomass )

  # if (order == "decreasing") 
  #alldata <-   alldata[rev(order(alldata$start_density)), ]
  # 
  # if (order == "increasing") 
 # alldata <- alldata[rev(order(alldata$start_density)), ]

  # need to be standardized? seems not
  alldata$depth2 <- alldata$depth^2
  alldata$start_density.st <- round(arm::rescale(log(alldata$start_density))*5)/5
  alldata$start_density_st <- round(arm::rescale(log(alldata$start_density)), 1)
  alldata$depth.st <- arm::rescale(log(alldata$depth)) 
  alldata$muddy.st <- arm::rescale(alldata$muddy) 
  alldata$any_rock.st <- arm::rescale(alldata$any_rock) 
  #alldata$omega_s.st <- arm::rescale(alldata$omega_s)

 ## Using mysterious cem optimization ... doesn't work   
  # fit_formula <- paste0("treat ~ start_density_st") 
  # fit_formula <- paste0("treat ~ start_density_st + depth.st") 
  # fit_formula <- paste0("treat ~ start_density_st + depth + trawled + muddy + any_rock + omega_s + x + y") 
  # fit_formula <- paste0("treat ~ start_density_st + depth.st + muddy.st + any_rock.st") #+ + omega_s
  # 
  # # for (j in (variable_names)) {
  # #   fit_formula <- paste0("", fit_formula, " + ", j, "_s")
  # #   }
  # 
  # # # options("optmatch_max_problem_size" = Inf)
  # # matched <- MatchIt::matchit(as.formula(fit_formula), data = alldata, method = "optimal", ratio = 1)
  # #fit_formula <- paste0("treat ~ depth.st")
  # matched <- MatchIt::matchit(as.formula(fit_formula),
  #   method = "cem",
  #   distance = "probit",
  #   discard="control",
  #   data = alldata)

  # Using exact match and rounding
  fit_formula <- paste0("treat ~ depth.st +  muddy.st + any_rock.st") #+ trawled + omega_s
  for (j in (variable_names)) {
    fit_formula <- paste0("", fit_formula, " + ", j, "_s")
    }

  matched <- MatchIt::matchit(as.formula(fit_formula), exact = c("start_density.st"),
    method = "nearest", m.order = "random",
    #method = "full",
    distance = "probit",
    discard="control",
    data = alldata)

  mdata <- MatchIt::match.data(matched, distance = "pscore")

  summary(matched)
  plot(matched)

  sources <- mdata[row.names(matched$match.matrix), ]
  sources$origobs <- row.names(sources)   
  sources$matchobs <- matched$match.matrix[,1]

  controls <- mdata[matched$match.matrix, ]
  controls$origobs <- row.names(controls)   
  controls$matchobs <- matched$match.matrix[,1]

  # controls2 <- mdata[matched$match.matrix, ]
  # controls2$origobs <- row.names(controls)   
  # controls2$matchobs <- matched$match.matrix[,2]

  mdata <- rbind(sources, controls) %>% tidyr::drop_na() %>% filter(matchobs!="-1")

  #mdata <- mdata %>% rename(distance = vect_dist) #%>% select(-start_density.st, -depth.st, -muddy.st, -any_rock.st, -omega_s.st)

  matched_time0 <- mdata %>% mutate(after = 0, density = start_density) #%>% dplyr::select(-start_density, -end_density)
  matched_time1 <- mdata %>% mutate(after = 1, density = end_density) #%>% dplyr::select(-start_density, -end_density)

  for (j in seq_along(variable_names)) {
    climate <- paste0(variable_names[j])
    matched_time0 <- matched_time0 %>% 
      mutate((!!climate) := (UQ(rlang::sym(paste0(variable_names[j], "_s")))))
    matched_time1 <- matched_time1 %>% 
      mutate((!!climate) := (UQ(rlang::sym(paste0(variable_names[j], "_e")))))
   }

  vocc_by_sp <- rbind(matched_time0, matched_time1)

  # calculate standardized climate variables
  for (j in seq_along(variable_names)) {
    mean.var <- paste0("mean.", variable_names[j])
    sd.var <- paste0("sd.", variable_names[j])
    std.var <- paste0("std_", variable_names[j])
    std.var2 <- paste0("std_", variable_names[j], "2")

    vocc_by_sp <- mutate(
      vocc_by_sp,
      (!!mean.var) := mean(UQ(rlang::sym(paste0(variable_names[j]))), na.rm = TRUE),
      (!!sd.var) := sd(UQ(rlang::sym(paste0(variable_names[j]))), na.rm = TRUE),
      (!!std.var) := ((UQ(rlang::sym(paste0(variable_names[j]))) - 
          mean(UQ(rlang::sym(paste0(variable_names[j]))), na.rm = TRUE)) / 
          (sd(UQ(rlang::sym(paste0(variable_names[j]))), na.rm = TRUE))*2),
      (!!std.var2) := (( UQ(rlang::sym(paste0(variable_names[j]))) - mean(UQ(rlang::sym(paste0(variable_names[j]))), na.rm = TRUE) ) / (sd(UQ(rlang::sym(paste0(variable_names[j]))), na.rm = TRUE)*2))^2
    )
  }

  # rename spatial coordinates for inclusion in models and plots without naming conflicts
  vocc_by_sp <- vocc_by_sp %>%
    rename(X = x, Y = y)

  vocc_by_sp
})
```

## Adults

Prep data
```{r}
data <- vocc_by_sp [["adult"]]
# View(data)

data$log_density <-  log(data$density)
data$vect_dist2 <- data$vect_dist^2

data <- mutate(data, time = ifelse(after==1, "end", "begin"))
data$matchobs <- as.factor(data$matchobs)
data$origobs  <- as.factor(data$origobs)   
```

Plot data
```{r}
ggplot(data, 
#ggplot(filter(data, as.numeric(matchobs) <= 550), 
  aes(y=Y, x=X, shape=as.factor(treat), colour=matchobs)) +
  geom_point(alpha=0.4, size =4) +
  scale_shape_manual(values= c(4, 1)) +
  scale_colour_viridis_d(option="C", guide = FALSE) +
  theme_bw()

lapply(variable_names, function(j) {
ggplot(data, aes(y=log_density, x=UQ(rlang::sym(paste(j))), colour=time)) +
  geom_point(alpha=0.35) +
  scale_colour_viridis_d(option="C") +
  theme_bw()
})

ggplot(data, aes(y=log_density, x=depth, colour=cell_type)) +
  geom_point(alpha=0.35) +
  scale_colour_viridis_d(option="C") +
  facet_wrap(~time) +
  theme_bw()

ggplot(data, aes(log_density, fill = cell_type)) +
  geom_density(alpha=0.5) +
  scale_fill_viridis_d(option="C") +
  facet_wrap(~time)+
  theme_bw()

ggplot(data, aes(log_density, fill = cell_type)) +
  geom_histogram(alpha=0.45) +
  scale_fill_viridis_d(option="C") +
  facet_wrap(~time)+
  theme_bw()

ggplot(data, aes(y=log_density, x=cell_type, fill=cell_type)) +
  geom_boxplot(alpha=0.5) +
  scale_fill_viridis_d(option="C") +
  facet_wrap(~time)+
  theme_bw()
```

Basic BACI
```{r}
for (j in (variable_names)) {
  baci_formula <- paste0("log_density~after + cell_type + after:cell_type + n_targets + vect_dist + vect_dist2 + std_", j, "+ std_", j, "2 + (1|icell) + (1|matchobs)")
}

m <- lmerTest::lmer(as.formula(baci_formula), data=data)
summary(m)
```

```{r}
sjPlot::plot_model(m, type = "std2", order.terms=c(1,2,8,3,4,5,6,7), title = paste(species, "adult"))
```

GAMM
```{r}

for (j in (variable_names)) {
  gamm_formula <- paste0("log_density ~ time * cell_type + s(vect_dist) + s(std_", j, ") + s(X,Y)")
}

gam4 <- gamm4::gamm4(as.formula(gamm_formula), random = ~(1|matchobs) + (1|origobs), data=data) # + s(X,Y, bs="re")

# for (j in (variable_names)) {
#   gamm_formula <- paste0("log_density ~ time * cell_type + s(vect_dist) + s(std_", j, ") + te(X,Y)")
# }
# gam <- mgcv::gamm(as.formula(gamm_formula), random = ~(1|matchobs) + (1|origobs), data=data) # + s(X,Y, bs="re")

plot(gam4$gam,pages=1)

summary(gam4$mer) ## underlying mixed model
anova(gam4$gam)

# itsadug::plot_parametric(gam4$gam, pred=list(cell_type=c("control", "source"), after=c("0","1")), 
#   parametricOnly = TRUE, main="") + scale_y_reverse()

plot_parametric(gam4$gam, pred=list(time=c("begin","end"), cell_type=c("control", "source")), 
   parametricOnly = TRUE, main="", xlab ="log_density")

summary(gam4$gam) ## summary of gam
```

## Immatures

Prep data
```{r}
idata <- vocc_by_sp[["imm"]]
#View(data)
idata$log_density <-  log(idata$density)
idata$vect_dist2 <- idata$vect_dist^2

idata <- mutate(idata, time = ifelse(after==1, "end", "begin"))
idata$matchobs <- as.factor(idata$matchobs)
idata$origobs  <- as.factor(idata$origobs)   
```

Plot data
```{r}
ggplot(filter(data, as.numeric(matchobs) <= 150), aes(y=Y, x=X, shape=as.factor(treat), colour=matchobs)) +
  geom_point(alpha=0.4, size =4) +
  scale_shape_manual(values= c(4, 1)) +
  scale_colour_viridis_d(option="C", guide = FALSE) +
  theme_bw()

lapply(variable_names, function(j) {
ggplot(idata, aes(y=log_density, x=UQ(rlang::sym(paste(j))), colour=time)) +
  geom_point(alpha=0.35) +
  scale_colour_viridis_d(option="C") 
})
ggplot(idata, aes(log_density, fill = cell_type)) +
  geom_density(alpha=0.5) +
  scale_fill_viridis_d(option="C") +
  facet_wrap(~time)

ggplot(idata, aes(log_density, fill = cell_type)) +
  geom_histogram(alpha=0.5) +
  scale_fill_viridis_d(option="C") +
  facet_wrap(~time)

ggplot(idata, aes(y=log_density, x=cell_type, fill=cell_type)) +
  geom_boxplot(alpha=0.5) +
  scale_fill_viridis_d(option="C") +
  facet_wrap(~time)
```

Basic BACI
```{r}
mi <- lmerTest::lmer(as.formula(baci_formula), data=idata)
summary(mi)
```

```{r}
sjPlot::plot_model(mi, type = "std2", title = paste(species, "immature"))
```

GAMM
```{r}
# for (j in (variable_names)) {
#   gamm_formula <- paste0("log_density ~ time * cell_type + s(vect_dist) + s(std_", j, ") + s(X,Y)")
# }

igam4 <- gamm4::gamm4(as.formula(gamm_formula), random = ~(1|matchobs) + (1|origobs), data=idata) # + s(X,Y, bs="re")

plot(igam4$gam,pages=1)

summary(igam4$gam) ## summary of gam
summary(igam4$mer) ## underlying mixed model
anova(igam4$gam)

# itsadug::plot_parametric(gam4$gam, pred=list(cell_type=c("control", "source"), after=c("0","1")), 
#   parametricOnly = TRUE, main="") + scale_y_reverse()

#my_cols <- c("#E69F00", "#56B4E9")
plot_parametric(igam4$gam, pred=list(time=c("begin","end"), cell_type=c("control", "source")), 
   parametricOnly = TRUE, main="", xlab ="log_density"
  #, groups = cell_type, gcolor = my_cols, color = my_cols["time"]
  ) 
```

Misc exploring gamm 
```{r eval=FALSE}
# mi <- mgcv::gamm(log_density~ as.factor(after) + cell_type + after:cell_type + std_temp + std_temp2 + vect_dist + vect_dist2 + s(as.factor(idata$icell), bs="re") + s(as.factor(idata$matchobs), bs="re"), data=idata)
# summary(mi)
# 
# mi <- mgcv::gamm(log_density~ as.factor(after) * cell_type + s(std_temp) + s(vect_dist) + s(as.factor(idata$matchobs), bs="re") , data=idata) # + s(X,Y, bs="re")
# summary(mi)
# mi
# 
# 
# idata$after <- as.factor(idata$after)
# idata$matchobs <- as.factor(idata$matchobs)
# 
# m4b <- gamm4::gamm4(log_density~ after * cell_type + s(std_temp) + s(vect_dist) + #s(X) + s(Y) + 
#     s(X,Y), random = ~(1|matchobs), data=idata) # + s(X,Y, bs="re")
# 
# 
# plot(m4b$gam,pages=1)
# 
# summary(m4b$gam) ## summary of gam
# summary(m4b$mer) ## underlying mixed model
# anova(m4b$gam) 
# 
# gamtabs(m1)
# 
# #library(itsadug)
# plot_parametric(m4b$gam, pred=list(cell_type=c("control", "source"), after=c("0","1")))
# 
# library(mgcv)
```