### 2019-08-28 analysis/VOCC/1-biomass-models.Rmd
fix switch for amalgamating catch data across surveys when calculating length at maturity

---
title: "Spatiotemporal models of species biomass split by maturity class"
author: "Philina English"
date: '2019-07-18'
output: html_document
params: 
    species: "Arrowtooth Flounder"

---

Choose one value for each parameter:

Species
```{r}
species <- params$species

# # Species run so far...
# species <- "Arrowtooth Flounder"
#species <- "Pacific Cod"
# species <- "Sablefish"
# species <- "Silvergray Rockfish"
# species <- "Lingcod"
# species <- "North Pacific Spiny Dogfish" # using all data for maturity thresholds
# species <- "Quillback Rockfish" # no maturity classes and WCHG not work
# species <- "Pacific Ocean Perch" 
 species <- "Yelloweye Rockfish"
```

Region 
```{r}
 region <- "All synoptic surveys"
# region <- "Both odd year surveys"
# region <- "West Coast Vancouver Island"
# region <- "West Coast Haida Gwaii"
```

Choose model details
```{r}
priors <- FALSE
# priors <- TRUE
# covariates <- "+muddy+any_rock"
covariates <- ""
covs <- gsub("\\+", "-", "+muddy+any_rock")
#covs <- gsub("\\+", "-", covariates)
# covariates <- "+muddy+mixed+rocky"
# covariates <- "+mixed+rocky"
# covariates <- "+trawled+muddy+rocky+mixed"
# covariates <- "+trawled+mixed+rocky"
# covariates <- "+trawled+mixed"
```

Run all subsequent code...
```{r global_options, include=FALSE}
# covs <- gsub("\\+", "-", covariates)
spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))

# folder to hold figs for this species
dir.create(file.path("figs", spp))
dir.create(file.path("data", spp))

knitr::opts_chunk$set(fig.width=11, fig.height=8.5, 
                      fig.path=paste0("figs/", spp, "/"),
                      echo=FALSE, warning=FALSE, message=FALSE)
```

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(gfplot)
library(gfdata)
library(sdmTMB)
library(gfranges)

fish <- readRDS(paste0("raw/bio-data-", spp, ""))

if (region == "Both odd year surveys") {
  survey <- c("SYN QCS", "SYN HS")
  model_ssid <- c(1, 3)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "West Coast Vancouver Island") {
  survey <- c("SYN WCVI")
  model_ssid <- c(4)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "West Coast Haida Gwaii") {
  survey <- c("SYN WCHG")
  model_ssid <- c(16)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "All synoptic surveys") {
  survey <- c("SYN QCS", "SYN HS", "SYN WCVI", "SYN WCHG")
  model_ssid <- c(1, 3, 4, 16)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}
```

Make prediction grids for all surveys and years, if not done before.
```{r, message=FALSE, warning=FALSE}
try(nd_all <- readRDS("data/nd_all_synoptic.rds"))

if (!exists("nd_all")) {

  # choose base year(s) to create grid from
  dummy_year <- c(2005, 2006)

  bath <- readRDS("data/bathymetry-data")
  #dat <- bath$data %>% scale_survey_predictors()
  dat <- bath$data %>% mutate(raw_depth = depth, depth = log(raw_depth)) 
  dat <- gfranges::scale_predictors(dat, predictors = c(quo(depth)))

  # create grids for each ssid separately
  ssid <- 1
  survey_abbrev <- "SYN QCS"
  nd_1 <- spatiotemporal_grid(dat, ssid, survey_abbrev, dummy_year)
  unique(nd_1$year)

  ssid <- 3
  survey_abbrev <- "SYN HS"
  nd_3 <- spatiotemporal_grid(dat, ssid, survey_abbrev, dummy_year)
  unique(nd_3$year)

  ssid <- 4
  survey_abbrev <- "SYN WCVI"
  nd_4 <- spatiotemporal_grid(dat, ssid, survey_abbrev, dummy_year)
  unique(nd_4$year)

  ssid <- 16
  survey_abbrev <- "SYN WCHG"
  nd_16 <- spatiotemporal_grid(dat, ssid, survey_abbrev, dummy_year)
  unique(nd_16$year)

  nd_all <- rbind(nd_1, nd_3, nd_4, nd_16) %>% 
    mutate(year = as.numeric(year))

  newcov <- readRDS("data/new_covariates.rds") %>% dplyr::select(X, Y, trawled, muddy, rocky, mixed, any_rock)
  nd_all <- left_join(nd_all, newcov, by=c("X", "Y"))

  # nd <- nd %>% dplyr::mutate(shallow = ifelse(depth > 35, 0, 1))

  saveRDS(nd_all, file = paste0("data/nd_all_synoptic.rds"))
}
```

# Split biomass by maturity

#### Calculate maturity weighted biomass if not done before
```{r}
 rm(biomass)
 rm(maturity)
# try(biomass <- readRDS(paste0("data/", spp, "/data-by-maturity-", spp, "-", ssid_string, ".rds")))

if (!exists("biomass")) {
  survey_sets <- readRDS(paste0("raw/event-data-", spp, ""))
  fish <- readRDS(paste0("raw/bio-data-", spp, ""))
  bath <- readRDS("data/bathymetry-data")
  # for POP: exclude sample with unusually large fish 
  # sizes are impossible given recorded catch weight
  # fish <- fish[fish$fishing_event_id!=1506954,]

 try({ 
   maturity <- split_catch_maturity(survey_sets, fish, bath,
    survey = survey,
    years = years,
    cutoff_quantile = c(.9995),
    plot = TRUE)
   })

  if(is.null(maturity$maturity)) {
  maturity <- split_catch_maturity(survey_sets, fish, bath,
    survey = c("SYN QCS","SYN HS","SYN WCVI","SYN WCHG"),
    years = years,
    cutoff_quantile = c(.9995),
    plot = TRUE
  )
  maturity$data <- filter(maturity$data, ssid == model_ssid) 
  }

  saveRDS(maturity$data, file = paste0("data/", spp, 
    "/data-by-maturity-", spp, "-", ssid_string, ".rds"
    ))
}
```

#### Save length~weight plots
```{r}
if (exists("maturity")) {
  print(maturity$mass_model)
  ggsave(
    file = paste0("figs/", spp, "/", spp, "-mass-model-", ssid_string, ".png"),
    dpi = 600,
    width = 8, 
    height = 10
  )
}
```

#### Save length at maturity plot 
```{r}
if (exists("maturity")) {
  png(
    file = paste0("figs/", spp, "/", spp, "-maturity-", ssid_string, ".png"), 
    res = 600,
    units = "in",
    width = 8, 
    height = 10
  )
  print(maturity$maturity)
  dev.off()
}

try(print(maturity$maturity))
```

# Build spatiotemporal density model
Load and filter data 
```{r}
biomass <- readRDS(paste0(
    "data/", spp, "/data-by-maturity-", spp, "-", ssid_string, ".rds"
    ))
  covars <- readRDS("data/event-covariates.rds")
  data <- dplyr::left_join(biomass, covars)

  # scale predictors before filtering to ensure mean and SD are global
  data <- data %>% mutate(raw_depth = depth, depth = log(raw_depth)) 
  data <- gfranges::scale_predictors(data, predictors = c(quo(depth)))
       # c(quo(log_depth), quo(mixed), quo(muddy), quo(sandy), quo(rocky), quo(any_rock)))
  data <- data %>% mutate(depth = raw_depth) %>% filter(ssid %in% model_ssid)
```

Run climate independent sdmTMB model
```{r}

  if (region == "Both odd year surveys") {
    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 400)
  }

  if (region == "West Coast Vancouver Island") {
    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 200)
  }

  if (region == "West Coast Haida Gwaii") {
    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 200)
  }

  sdmTMB::plot_spde(spde)

  if (any(names(data) == "adult_density")) {

    adult_formula <- as.formula(paste(
      "adult_density ~ 0 + as.factor(year)", covariates, ""
      ))

    starttime1 <- Sys.time()
    adult_biomass <- sdmTMB::sdmTMB(data,
      adult_formula,
      time_varying = ~ 0 + depth_scaled + depth_scaled2 + trawled,
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = TRUE,
      include_spatial = TRUE,
      enable_priors = priors,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )
    endtime1 <- Sys.time()
    time1 <- round(starttime1 - endtime1)

    saveRDS(adult_biomass, 
      file = paste0("data/", spp, 
        "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
        ))

    imm_formula <- as.formula(paste(
      "imm_density ~ 0 + as.factor(year)", covariates, ""
      ))

    starttime2 <- Sys.time()
    imm_biomass <- sdmTMB::sdmTMB(data, 
      imm_formula,
      time_varying = ~ 0 + depth_scaled + depth_scaled2 + trawled,
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = FALSE,
      include_spatial = TRUE,
      enable_priors = priors,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )
    endtime2 <- Sys.time()
    time2 <- round(starttime2 - endtime2)

    saveRDS(imm_biomass, 
     file = paste0("data/", spp, 
       "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
       ))

  } else {

    starttime3 <- Sys.time()
    dens_formula <- as.formula(paste("density ~ 0 + as.factor(year)", covariates, ""))

    total_biomass <- sdmTMB::sdmTMB(data,
      dens_formula,
      time_varying = ~ 0 + depth_scaled + depth_scaled2 + trawled,
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = TRUE,
      include_spatial = TRUE,
      enable_priors = priors,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )
    endtime3 <- Sys.time()
    time3 <- round(starttime3 - endtime3)

    saveRDS(total_biomass, file = paste0("data/", spp, 
      "/model-total-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
      ))
  }
```

Save predictions for full spatial grid

```{r}
#covs <- gsub("\\+", "-", covariates)
fish <- readRDS(paste0("raw/bio-data-", spp, ""))

rm(adult_predictions)
rm(imm_predictions)
rm(predictions)

if (length(unique(fish$maturity_code)) > 2) {

    nd_all <- readRDS(paste0("data/nd_all_synoptic.rds"))
    adult_biomass <- readRDS(paste0("data/", spp, 
      "/mod-mat-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
    ))

    nd <- nd_all %>% 
      filter(ssid %in% model_ssid) %>% 
      filter(year %in% unique(adult_biomass$data$year))
    nd <- na.omit(nd)
    nd$year <- as.integer(nd$year)

    adult_predictions <- predict(adult_biomass, newdata = nd)
    saveRDS(adult_predictions, file = paste0("data/", spp, 
      "/predictions-", spp, covs, "-", ssid_string, "-mature-biomass-prior-", priors, ".rds"
    ))

    imm_biomass <- readRDS(paste0("data/", spp, 
      "/mod-imm-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
    ))
    imm_predictions <- predict(imm_biomass, newdata = nd)
    saveRDS(imm_predictions, file = paste0("data/", spp, 
      "/predictions-", spp, covs, "-", ssid_string, "-imm-biomass-prior-", priors, ".rds"
    ))
  } else {
    rm(adult_predictions)
    rm(imm_predictions)

    if (!exists(predictions)) {
      nd_all <- readRDS(paste0("data/nd_all_synoptic.rds"))
      nd <- nd_all %>% filter(ssid %in% model_ssid)
      nd <- na.omit(nd)
      nd$year <- as.integer(nd$year)

      total_biomass <- readRDS(paste0("data/", spp, 
        "/model-total-biomass-", spp, covs, "-", ssid_string, "-prior-", priors, ".rds"
      ))
      predictions <- predict(total_biomass, newdata = nd)
      saveRDS(predictions, file = paste0("data/", spp, 
        "/predictions-", spp, covs, "-", ssid_string, "-total-biomass-prior-", priors, ".rds"
      ))
    }
  }
```