### 2019-09-10 analysis/VOCC/2-biomass-trends.Rmd
climate independent biomass model that converges most relaibly is global with trawled and ssid as fixed effects

---
title: "Biomass trends and figures"
author: "Philina English"
date: '2019-07-18'
output: html_document
---

This file calls models built by "1-biomass-models.Rmd" and stiches the predictions saved therein together.

# Choose params

Species
```{r}

# # Species run so far...
# species <- "Arrowtooth Flounder" # temp good, not DO, imm moving deeper?

# species <- "Petrale Sole"
# species <- "English Sole"
# species <- "Dover Sole"
# species <- "Southern Rock Sole"

# species <- "Walleye Pollock" # survey 4 or 16 might work
# species <- "Pacific Cod"

## Sebastes

## TRY BOTH - not schooling
 # species <- "Yelloweye Rockfish" # beautiful curves, with depth change ***
 # species <- "Bocaccio" # good curves, no depth change ... most survey-specific models fail to find minimum ***
#  species <- "Sharpchin Rockfish" # beautiful curves, imm depth change ***
#  species <- "Greenstriped Rockfish" # beautiful curves for adults, no depth change ***
# 
# ## TRY BOTH - schooling
#  species <- "Pacific Ocean Perch" # beautiful curves, no depth change, schooling ***
#  species <- "Silvergray Rockfish" # beautiful curves, no depth change, schooling ***

## TRY JUST TEMPERATURE 
# species <- "Rougheye/Blackspotted Rockfish Complex" # good temp, not DO
# species <- "Redbanded Rockfish" # good temp, not DO, maybe moving shallower
# species <- "Splitnose Rockfish" # good temp, not DO

# species <- "Yellowtail Rockfish" #  good temp for survey 4 only, no DO, schooling
# species <- "Quillback Rockfish" # like warm = overall a bit flat... territorial! 
# species <- "Canary Rockfish" # schooling, winter-birthing, and depth change!, but curves at edges... adult1 looks ok?
#  species <- "Widow Rockfish" # schooling, and depth change, but curves at edges

# species <- "Longspine Thornyhead" # too deep to produce curves
# species <- "Shortspine Thornyhead"
# species <- "Sablefish" # temp might work, no DO, no consistant depth change
# species <- "Lingcod" # curves ok, overall a bit flat
# # species <- "Pacific Hake" # not useful

# species <- "North Pacific Spiny Dogfish" # too flat, but with depth change! note: pooled maturity 
# species <- "Longnose Skate" # do too flat, temp from 4 might be useful, seems to go deeper?
# species <- "Big Skate"
# species <- "Sandpaper Skate"
# species <- "Brown Cat Shark"
# species <- "Spotted Ratfish" # curves not useful because to abundant in shallow
```

Region 
```{r}
region <- "All synoptic surveys"

# region <- "Both odd year surveys"
# region <- "West Coast Vancouver Island"
# region <- "West Coast Haida Gwaii"
```

Choose model details
```{r}
update_spatial_model <- FALSE
priors <- FALSE
# priors <- TRUE
covariates <- ""
covariates <- "+trawled"
covs <- gsub("\\+", "-", covariates)
covariates <- "+trawled+as.factor(ssid)"
 covs <- gsub("\\+", "-", "+trawled+ssid")
# covariates <- "+muddy+mixed+rocky"
# covariates <- "+mixed+rocky"
# covariates <- "+muddy+any_rock"
# covariates <- "+trawled+muddy+rocky+mixed"
# covariates <- "+trawled+mixed+rocky"
# covariates <- "+trawled+mixed"
```

Run all subsequent code...
```{r global_options, include=FALSE}

spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))

# folder to hold figs for this species
dir.create(file.path("figs", spp))
dir.create(file.path("data", spp))

knitr::opts_chunk$set(fig.width=11, fig.height=8.5, 
                      #fig.path=paste0("figs/", spp, "/"),
                      echo=FALSE, warning=FALSE, message=FALSE)
```

```{r, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(gfplot)
library(gfdata)
library(sdmTMB)
library(gfranges)

fish <- readRDS(paste0("raw/bio-data-", spp, ""))

if (region == "All synoptic surveys") {
  survey <- c("SYN QCS", "SYN HS", "SYN WCVI", "SYN WCHG")
  model_ssid <- c(1, 3, 4, 16)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "Both odd year surveys") {
  survey <- c("SYN QCS", "SYN HS")
  model_ssid <- c(1, 3)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "West Coast Vancouver Island") {
  survey <- c("SYN WCVI")
  model_ssid <- c(4)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "West Coast Haida Gwaii") {
  survey <- c("SYN WCHG")
  model_ssid <- c(16)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}
```

Bind together predictions from different models
```{r}
rm(adult_predictions)
rm(imm_predictions)
rm(predictions)

      rm(pred1n3)
  rm(pred4)
rm(pred16)

      rm(pred1n3_imm)
  rm(pred4_imm)
rm(pred16_imm)

    rm(pred1n3_all)
  rm(pred4_all)
rm(pred16_all)
#if (isFALSE(update_spatial_model)) {
if (region != "All synoptic surveys") {

    try(adult_predictions <- readRDS(paste0("data/", spp, 
      "/predictions-", spp, covs, "-", ssid_string, "-mature-biomass-prior-", priors, ".rds")))
    try(imm_predictions <- readRDS(paste0("data/", spp, 
      "/predictions-", spp, covs, "-", ssid_string, "-imm-biomass-prior-", priors, ".rds")))

    if (!exists("adult_predictions")) {
      paste("Maturity split biomass predictions are not available for this survey.")

      try(predictions <- readRDS(paste0("data/", spp, 
        "/predictions-", spp, covs, "-", ssid_string, "-total-biomass-prior-", priors, ".rds")))

      if (!exists("predictions")) {
        paste("Total biomass predictions for this survey are also not available.")
      }
    }

} else {

   try({adult_predictions <- readRDS(paste0("data/", spp, 
      "/predictions-", spp, covs, "-", ssid_string, "-mature-biomass-prior-", priors, ".rds"
    ))})
   try({imm_predictions <- readRDS(paste0("data/", spp, 
      "/predictions-", spp, covs, "-", ssid_string, "-imm-biomass-prior-", priors, ".rds"
    ))})

  if (!exists("adult_predictions")) {
    # OR bind together predictions from different models
   try({pred1n3 <- readRDS(paste0("data/", spp, 
        "/predictions-", spp, covs, "-1n3-mature-biomass-prior-", priors, ".rds"))})
   try({pred4 <- readRDS(paste0("data/", spp, 
        "/predictions-", spp, covs, "-4-mature-biomass-prior-", priors, ".rds"))}) 
   try({pred16 <- readRDS(paste0("data/", spp, 
        "/predictions-", spp, covs, "-16-mature-biomass-prior-", priors, ".rds"))})

    if (!exists("pred1n3")) pred1n3 <- NULL
    if (!exists("pred16")) pred16 <- NULL
    if (!exists("pred4")) pred4 <- NULL
    adult_predictions <- rbind(pred1n3, pred4, pred16)

    try({
      pred1n3_imm <- readRDS(paste0("data/", spp, 
        "/predictions-", spp, covs, "-1n3-imm-biomass-prior-", priors, ".rds"))})
   try({
      pred4_imm <- readRDS(paste0("data/", spp, 
        "/predictions-", spp, covs, "-4-imm-biomass-prior-", priors, ".rds"))})
   try({
      pred16_imm <- readRDS(paste0("data/", spp, 
        "/predictions-", spp, covs, "-16-imm-biomass-prior-", priors, ".rds"))
    })

    if (!exists("pred1n3_imm")) pred1n3_imm <- NULL
    if (!exists("pred16_imm")) pred16_imm <- NULL
    if (!exists("pred4_imm")) pred4_imm <- NULL
    imm_predictions <- rbind(pred1n3_imm, pred4_imm, pred16_imm)

    if (!exists("adult_predictions")) {
      paste("Maturity split biomass predictions are not available.")

      try({predictions <- readRDS(paste0("data/", spp, 
        "/predictions-", spp, covs, "-", ssid_string, "-total-biomass-prior-", priors, ".rds"
      ))})

      if (!exists("predictions")) {

      try({pred1n3_all <- readRDS(paste0("data/", spp, 
        "/predictions-", spp, covs, "-1n3-total-biomass-prior-", priors, ".rds"))})
      try({pred4_all <- readRDS(paste0("data/", spp, 
        "/predictions-", spp, covs, "-4-total-biomass-prior-", priors, ".rds"))})
      try({pred16_all <- readRDS(paste0("data/", spp, 
        "/predictions-", spp, covs, "-16-total-biomass-prior-", priors, ".rds"))})
      try({
      predictions <- rbind(pred1n3_all, pred4_all, pred16_all)
      })

      if (!exists("predictions")) paste("Total biomass predictions are not available.")
      }
    }
  }
}
```

Caluclate variables of interest from predictions
```{r}
if (exists("adult_predictions")) {

  # to convert from kg/m2 to kg/hectare multiply by 10000
  adult_predictions$est_exp <- exp(adult_predictions$est) * 10000
  imm_predictions$est_exp <- exp(imm_predictions$est) * 10000

  adult_predictions$total_bio <- imm_predictions$est_exp + adult_predictions$est_exp
  imm_predictions$prop_imm <- imm_predictions$est_exp / (adult_predictions$est_exp + imm_predictions$est_exp)

  max_raster <- quantile(adult_predictions$est_exp, 0.999)
  max_adult <- signif(max(adult_predictions$est_exp), digits = 2)

  saveRDS(adult_predictions, file = paste0("data/", spp, 
      "/all-predictions-", spp, covs, "-mat-biomass-prior-", priors, ".rds"))
  saveRDS(imm_predictions, file = paste0("data/", spp, 
      "/all-predictions-", spp, covs, "-imm-biomass-prior-", priors, ".rds"))

  model_ssid <- unique(adult_predictions$ssid)

} else {

  predictions$total_bio <- exp(predictions$est) * 10000
  max_raster <- quantile(predictions$total_bio, .999)
  max_bio <- round(max(predictions$total_bio))

  saveRDS(predictions, file = paste0("data/", spp, 
      "/all-predictions-", spp, covs, "-total-biomass-prior-", priors, ".rds"))

  model_ssid <- unique(predictions$ssid)
}
```

# Plots and trends

```{r, error=TRUE}
if (exists("adult_predictions")) {

  p_adult_all <- plot_facet_map(adult_predictions, "est_exp",
    raster_limits = c(0, max_raster),
    legend_position = "none",
    transform_col = fourth_root_power
  ) +
    labs(fill = "kg/ha") +
    ggtitle(paste0("", species, " mature biomass \n(max = ", max_adult, " kg/ha)"))
  print(p_adult_all)

} else {

  p_adult_all <- plot_facet_map(predictions, "total_bio",
    transform_col = fourth_root_power
  ) +
    labs(fill = "kg/ha") +
    ggtitle(paste0("", species, " total biomass \n(max = ", max_bio, " kg/ha)"))

  print(p_adult_all)
}
```

```{r}
if (exists("adult_predictions")) {
  if (1 %in% model_ssid) {
    adult_predictions1 <- adult_predictions %>% filter(ssid == 1)
  } else {
    adult_predictions1 <- NULL
  }

  if (3 %in% model_ssid) {
    adult_predictions3 <- adult_predictions %>% filter(ssid != 16) %>% filter(ssid != 4) 
  } else {
    adult_predictions3 <- NULL
  }

  if (4 %in% model_ssid) {
    adult_predictions4 <- adult_predictions %>% filter(ssid == 4)
  } else {
    adult_predictions4 <- NULL
  }

  if (16 %in% model_ssid) {
    if (any(adult_predictions$ssid == 16)) {
    adult_predictions16 <- adult_predictions %>% filter(ssid == 16)
    } else {
    adult_predictions16 <- NULL
    }
  } else {
    adult_predictions16 <- NULL
  }

  ad_prediction_list <- list(adult_predictions1, adult_predictions3, adult_predictions4, adult_predictions16)
  ad_prediction_list <- ad_prediction_list[!sapply(ad_prediction_list, is.null)]

  p_adult <- list()
  for (i in seq_len(length(model_ssid))) {
    p_adult[[i]] <- plot_facet_map(ad_prediction_list[[i]], "est_exp",
      raster_limits = c(0, max_raster),
      # legend_position = "none"
      transform_col = fourth_root_power
    ) +
      labs(fill = "kg/ha") +
      ggtitle(paste0("", species, " mature biomass \n(max = ", max_adult, " kg/ha)"))
    p_adult[[i]]
  }
  print(p_adult)
}
```

```{r, message=FALSE, warning=FALSE}
#try(out_mature <- readRDS(paste0("data/", spp, "/trends-", spp, covs, "-", ssid_string, "-mature-biomass-priors-", priors, ".rds")))
rm(out_mature)
if (!exists("out_mature")) {
  if (exists("adult_predictions")) {
    predictions <- adult_predictions
  } else {
    try(out_total <- readRDS(paste0(
      "data/", spp, "/trends-", spp, covs, "-", ssid_string, "-total-biomass-priors-", priors, ".rds"
    )))
  }

  if (!exists("out_total")) {
    if (1 %in% model_ssid) {
      out_mature1 <- make_trend_data(predictions,
        ssid = c(1),
        start_time = 2003,
        skip_time = 2004,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = c(1, 1, 1, 1, 2, 2, 2, 2),
        variable_names = "est_exp"
      )
    } else {
      out_mature1 <- NULL
    }

    if (3 %in% model_ssid) {
      out_mature3 <- make_trend_data(predictions,
        ssid = c(3),
        start_time = 2005,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = c(1, 1, 1, 2, 2, 2, 2),
        variable_names = "est_exp"

      )
    } else {
      out_mature3 <- NULL
    }

    if (4 %in% model_ssid) {
      out_mature4 <- make_trend_data(predictions,
        ssid = c(4),
        start_time = 2004,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = c(1, 1, 1, 1, 2, 2, 2, 2),
        variable_names = "est_exp"
      )
    } else {
      out_mature4 <- NULL
    }

    if (16 %in% model_ssid) {
      out_mature16 <- make_trend_data(predictions,
        ssid = c(16),
        start_time = 2004,
        skip_time = 2007,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = c(1, 1, 1, 2, 2, 2, 2),
        variable_names = "est_exp"

      )
    } else {
      out_mature16 <- NULL
    }

    out_mature <- rbind(out_mature1, out_mature3, out_mature4, out_mature16)

    saveRDS(out_mature, file = paste0(
      "data/", spp, "/trends-", spp, covs, "-", ssid_string, "-mature-biomass-priors-", priors, ".rds"
    ))

    if (!exists("adult_predictions")) {
      out_total <- out_mature
      saveRDS(out_total, file = paste0(
        "data/", spp, "/trends-", spp, covs, "-", ssid_string, "-total-biomass-priors-", priors, ".rds"
      ))
    }
  }
}
```

```{r }
if (exists("out_mature")) {

trawl_footprint <- sf::st_read(dsn="data/trawl-footprint",layer="Trawl_footprint")
proj.to <- "+proj=utm +zone=9 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs"
trawl_footprint <- sf::st_transform(trawl_footprint, crs = proj.to)

current_biomass <- plot_vocc(out_mature,
    fill_col = "est_exp_e",
    fill_label = "Mean mature \nbiomass (kg/ha)\npost-2010",
    raster_alpha = 1,
    viridis_option = "C",
    vec_aes = NULL,
    transform_col = fourth_root_power
  )
current_biomass <- current_biomass +
   geom_sf(data= trawl_footprint, aes(geometry=geometry/1000), colour = "gray", fill = "gray", alpha= 0.2, lwd=0.25, inherit.aes = FALSE) +
   coord_sf(xlim = c(180,820), ylim = c(5350,6050), datum = NULL) +
   xlab("") + 
   ylab("") +
   theme(panel.grid.major = element_line(colour = "transparent")) +
   ggtitle(paste("Predicted", species, "density with trawl footprint overlain"))
current_biomass
}
```

```{r, error=TRUE}
if (exists("out_mature")) {
  data <- out_mature
  biotrend_ad <- plot_vocc(data,
    fill_col = "units_per_decade",
    fill_label = "Mature \nbiomass \ntrend \n(kg/ha/decade)",
    raster_alpha = 1,
    vec_aes = NULL,
    white_zero = TRUE,    
    high_fill = "deepskyblue3",#"lightseagreen", # limegreen",
    low_fill = "red", # "deeppink3",# "steel blue 4",
    transform_col = fourth_root_power,
    legend_position = c(0.2, 0.25)
  ) + ggtitle(paste0(
    "", species, " ", min(unique(adult_predictions[["year"]])), "-", max(unique(adult_predictions[["year"]]))
  ))
  biotrend_ad
} else {
  biotrend_ad <- plot_vocc(out_total,
    fill_col = "units_per_decade",
    fill_label = "Biomass \ntrend \n(kg/ha/decade)",
    raster_alpha = 1,
    vec_aes = NULL,
    white_zero = TRUE,    
    high_fill = "deepskyblue3", # "greenyellow", # limegreen", 
    low_fill = "red", #"steel blue 4",
    transform_col = fourth_root_power,
    legend_position = c(0.2, 0.25)
  ) + ggtitle(paste0(
    "", species, " ", min(unique(predictions[["year"]])), "-", max(unique(predictions[["year"]]))
  ))
  biotrend_ad
}
```

```{r, error=TRUE}
if (exists("imm_predictions")) {
  p_imm_all <- plot_facet_map(imm_predictions, "est_exp",
    transform_col = fourth_root_power,
    raster_limits = c(0, max_raster)
  ) +
    labs(fill = "kg/ha") +
    ggtitle(paste0("", species, " immature biomass \n "))

  p_imm_all
} else {
  p_imm_all <- grid::grid.rect(gp = grid::gpar(col = "white"))
}
```

```{r}
if (exists("adult_predictions")) {
  if (1 %in% model_ssid) {
    imm_predictions1 <- imm_predictions %>% filter(ssid == 1)
  } else {
    imm_predictions1 <- NULL
  }

  if (3 %in% model_ssid) {
    imm_predictions3 <- imm_predictions %>% filter(ssid != 16) %>% filter(ssid != 4) 
  } else {
    imm_predictions3 <- NULL
  }

  if (4 %in% model_ssid) {
    imm_predictions4 <- imm_predictions %>% filter(ssid == 4)
  } else {
    imm_predictions4 <- NULL
  }

  if (16 %in% model_ssid) {
    imm_predictions16 <- imm_predictions %>% filter(ssid == 16)
  } else {
    imm_predictions16 <- NULL
  }

  im_prediction_list <- list(imm_predictions1, imm_predictions3, imm_predictions4, imm_predictions16)
  im_prediction_list <- im_prediction_list[!sapply(im_prediction_list, is.null)]

  p_imm <- list()
  for (i in seq_len(length(model_ssid))) {
    p_imm[[i]] <- plot_facet_map(im_prediction_list[[i]], "est_exp",
      raster_limits = c(0, max_raster),
      # legend_position = "none",
      transform_col = fourth_root_power
    ) +
      labs(fill = "kg/ha") +
      ggtitle(paste0("", species, " immature biomass \n(max = ", max_adult, " kg/ha)"))

    p_imm[[i]]
  }
  print(p_imm)
}
```

```{r, message=FALSE, warning=FALSE}
#try(out_imm <- readRDS(paste0("data/", spp, "/trends-", spp, covs, "-", ssid_string, "-immature-biomass-priors-", priors, ".rds")))
rm(out_imm)
if (!exists("out_imm")) {
  if (exists("imm_predictions")) {
    predictions <- imm_predictions

    if (1 %in% model_ssid) {
      out_im1 <- make_trend_data(predictions,
        ssid = c(1),
        start_time = 2003,
        skip_time = 2004,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = c(1, 1, 1, 1, 2, 2, 2, 2),
        variable_names = "est_exp"

      )
    } else {
      out_im1 <- NULL
    }

    if (3 %in% model_ssid) {
      out_im3 <- make_trend_data(predictions,
        ssid = c(3),
        start_time = 2005,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = c(1, 1, 1, 2, 2, 2, 2),
        variable_names = "est_exp"

      )
    } else {
      out_im3 <- NULL
    }

    if (4 %in% model_ssid) {
      out_im4 <- make_trend_data(predictions,
        ssid = c(4),
        start_time = 2004,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = c(1, 1, 1, 1, 2, 2, 2, 2),
        variable_names = "est_exp"

      )
    } else {
      out_im4 <- NULL
    }

    if (16 %in% model_ssid) {
      out_im16 <- make_trend_data(predictions,
        ssid = c(16),
        start_time = 2004,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = c(1, 1, 1, 1, 2, 2, 2, 2),
        variable_names = "est_exp"

      )
    } else {
      out_im16 <- NULL
    }

    out_imm <- rbind(out_im1, out_im3, out_im4, out_im16)
    saveRDS(out_imm, file = paste0(
      "data/", spp, "/trends-", spp, covs, "-", ssid_string, "-immature-biomass-priors-", priors, ".rds"
    ))
  } else {
    rm(out_imm)
  }
}
```

```{r}
if (exists("out_imm")) {
  data <- out_imm
  biotrend_im <- plot_vocc(data,
    fill_col = "units_per_decade",
    fill_label = "Immature \nbiomass \ntrend \n(kg/ha/decade)",
    raster_alpha = 1,
    vec_aes = NULL,
    white_zero = TRUE, 
    high_fill = "deepskyblue3",# "greenyellow", # limegreen",
    low_fill = "red", #"steel blue 4",
    transform_col = fourth_root_power,
    legend_position = c(0.2, 0.25)
  ) + ggtitle(paste0(
    "", species, " ", min(unique(imm_predictions[["year"]])), "-", max(unique(imm_predictions[["year"]]))
  ))
  biotrend_im
} else {
  biotrend_im <- grid::grid.rect(gp = grid::gpar(col = "white"))
}
```

```{r, error=TRUE}
if (exists("imm_predictions")) {
  p_prop_imm <- plot_facet_map(imm_predictions, "prop_imm") +
    labs(fill = "% \nby mass") +
    ggtitle(paste0("", species, " proportion immature"))
  print(p_prop_imm)
} else {
  p_prop_imm <- grid::grid.rect(gp = grid::gpar(col = "white"))
}
```

Trend plots
```{r}
png(
  file = paste0("figs/", spp, "/trends-", spp, covs, "-", ssid_string, "-priors-", priors, ".png"),
  res = 600,
  units = "in",
  width = 11,
  height = 8.5
)
gridExtra::grid.arrange(
  # p_adult, p_imm,
  biotrend_ad, biotrend_im,
  nrow = 1
)
dev.off()
```

Spatiotemporal biomass plots
```{r}
biomass_plots <- c(p_adult, p_imm)

png(
  file = paste0("figs/", spp, "/biomass-", spp, covs, "-", ssid_string, "-priors-", priors, ".png"),
  res = 600,
  units = "in",
  width = 16,
  height = 8.5
)
gridExtra::grid.arrange(
  grobs = biomass_plots,
  nrow = 2
)
dev.off()
```

Plot medley
```{r}
png(
  file = paste0("figs/", spp, "/exploratory-", spp, covs, "-", ssid_string, "-priors-", priors, ".png"),
  res = 600,
  units = "in",
  width = 9.5,
  height = 12.5
)

gridExtra::grid.arrange(
  current_biomass,
  p_prop_imm,
  p_adult_all, p_imm_all,
  biotrend_ad, biotrend_im,
  nrow = 3
)
dev.off()
```