### 2019-10-08 analysis/VOCC/4-vocc-BACI-multiyear.Rmd

---
title: "BACI of `r paste(params$species)` biomass for `r paste(params$region)` VOCC"
author: "Philina English"
date: "`r Sys.Date()`"
output: html_document 
params: 
    species: "Pacific Ocean Perch" 
    region: "West Coast Vancouver Island"
    age: "imm"
    biomass_threshold: "50"
    climate: "temperature"
    threshold: 0.25
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.keep = "all", fig.show = "asis",
  fig.width = 11, fig.height = 8.5,
  echo = FALSE, warning = FALSE, message = FALSE
)
options(scipen = 999)
options(digits = 4)
```

```{r pkgs, cache=FALSE, warning=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
library(gfplot)
library(sdmTMB)
library(gfranges)
ggplot2::theme_set(gfplot::theme_pbs())
```

#### Set parameters
```{r }
species <- params$species
age <- params$age
region <- params$region
biomass_threshold <- params$biomass_threshold
climate <- params$climate

paste("species =", species)
paste("age =", age)
paste("region =", region)
paste("biomass_threshold =", biomass_threshold)
paste("climate =", climate)
```

#### Set default cell sizes and scale
```{r, echo=FALSE}
input_cell_size <- 2
dist_intercept <- 0 # input_cell_size / 2
scale_fac <- 1
delta_t_step <- 2
skip_time <- NULL
```

```{r, message=FALSE, warning=FALSE, include=FALSE}
if (region == "both odd year surveys") {
  survey <- c("SYN QCS", "SYN HS")
  model_ssid <- c(1, 3)
  ssid_string <- paste0(model_ssid, collapse = "n")

  if(climate=="temperature") {
    start_years <- list("2005" = 2005, "2007" = 2007, "2009" = 2009, "2011" = 2011, "2013" = 2013, "2015" = 2015)
  } else {
  start_years <- list("2009" = 2009, "2011" = 2011, "2013" = 2013, "2015" = 2015)
  }
  years <- NULL
}

if (region == "West Coast Vancouver Island") {
  survey <- c("SYN WCVI")
  model_ssid <- c(4)
  ssid_string <- paste0(model_ssid, collapse = "n")
  if(climate=="temperature") {
    start_years <- list("2004" = 2004, "2006" = 2006, "2008" = 2008, "2010" = 2010, "2012" = 2012, "2014" = 2014, "2016" = 2016)
  } else {
  start_years <- list("2008" = 2008, "2010" = 2010, "2012" = 2012, "2016" = 2016) # "2014" = 2014, # removed due to extreme DO values
  }
  years <- NULL
}

if (region == "West Coast Haida Gwaii") {
  survey <- c("SYN WCHG")
  model_ssid <- c(16)
  ssid_string <- paste0(model_ssid, collapse = "n")
  if(climate=="temperature") {
    start_years <- list("2006" = 2006, "2008" = 2008, "2010" = 2010, "2012" = 2012, "2014" = 2014, "2016" = 2016)
  } else {
  start_years <- list("2008" = 2008, "2010" = 2010, "2012" = 2012, "2014" = 2014, "2016" = 2016)
  }
  years <- NULL
}
```

#### Get species-specific optimal values from mountain plots for most abundant year
```{r, include=FALSE}
spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))

all_params <- readRDS(paste0("data/root-values-by-species-by-age.rds"))
#browser()
all_params <- all_params %>% filter(covs == "x2016-log-both-trawledFALSE") %>% mutate(model_type = ifelse(grepl("total", model), "total", "split"))

if (age == "imm") {
  all_params <- filter(all_params, age == "imm")
} else {
  all_params <- filter(all_params, age != "imm")
  if (any(all_params$model_type == "split")) {
    all_params <- filter(all_params, model_type != "total")
  }
}

# rm(root_params)
root_params <- all_params[all_params$species == species, ]
glimpse(root_params)

if (climate == "temperature") {
  temp_threshold <- params$threshold

  # if more than one model option in dataframe will pick the one with highest peak density
  if (biomass_threshold == "75") {
    .upper <- root_params[rev(order(root_params$do_maxy)), ]$upper_t_75
  }
  if (biomass_threshold == "50") {
    .upper <- root_params[rev(order(root_params$t_maxy)), ]$upper_t_50
  }
  if (biomass_threshold == "25") {
    .upper <- root_params[rev(order(root_params$t_maxy)), ]$upper_t_25
  }
  .upper <- na.omit(.upper)

  upper_thresholds <- round(.upper[1], digits = 2)
  if (is.na(upper_thresholds)) {
    stop(print("Temp maximum not modelled"))
  }

  upper_model <- root_params[rev(order(root_params$t_maxy)), ]$model

  lower_thresholds <- NA
  lower_model <- NULL
  variable_names <- c("temp")

  lower_change <- c(Inf)
  upper_change <- temp_threshold
  indices <- c(1, 2)
  delta_t_total <- 2
}

if (climate == "do") {
  do_threshold <- params$threshold
  root_params <- filter(root_params, lower_do_25 > 0.4)

  if (biomass_threshold == "75") {
    .lower <- root_params[rev(order(root_params$do_maxy)), ]$lower_do_75
  }
  if (biomass_threshold == "50") {
    .lower <- root_params[rev(order(root_params$do_maxy)), ]$lower_do_50
  }
  if (biomass_threshold == "25") {
    .lower <- root_params[rev(order(root_params$do_maxy)), ]$lower_do_25
  }

  .lower <- na.omit(.lower)
  lower_thresholds <- round(.lower[1], digits = 2)
  if (is.na(lower_thresholds)) {
    stop(print("DO minimum not modelled"))
  }
  lower_model <- root_params[rev(order(root_params$do_maxy)), ]$model

  upper_thresholds <- NA
  upper_model <- NULL
  variable_names <- c("do_est")

  lower_change <- do_threshold
  upper_change <- c(Inf)
  indices <- c(1, 2)
  delta_t_total <- 2
}

if (climate == "do and temperature") {
  temp_threshold <- params$threshold
  do_threshold <- params$threshold

  root_params <- filter(root_params, lower_do_25 > 0.5)
  if (biomass_threshold == "75") {
    .lower <- root_params[rev(order(root_params$do_maxy)), ]$lower_do_75
  }
  if (biomass_threshold == "50") {
    .lower <- root_params[rev(order(root_params$do_maxy)), ]$lower_do_50
  }
  if (biomass_threshold == "25") {
    .lower <- root_params[rev(order(root_params$do_maxy)), ]$lower_do_25
  }
  .lower <- na.omit(.lower)
  lower_thresholds <- round(.lower[1], digits = 2)

  if (biomass_threshold == "75") {
    .upper <- root_params[rev(order(root_params$do_maxy)), ]$upper_t_75
  }
  if (biomass_threshold == "50") {
    .upper <- root_params[rev(order(root_params$t_maxy)), ]$upper_t_50
  }
  if (biomass_threshold == "25") {
    .upper <- root_params[rev(order(root_params$t_maxy)), ]$upper_t_25
  }
  .upper <- na.omit(.upper)
  upper_thresholds <- round(.upper[1], digits = 2)

  variable_names <- c("do_est", "temp")
  lower_thresholds <- c(lower_thresholds, NA)
  upper_thresholds <- c(NA, upper_thresholds)
  lower_change <- c(do_threshold, Inf)
  upper_change <- c(Inf, temp_threshold)
  upper_model <- root_params[rev(order(root_params$t_maxy)), ]$model
  lower_model <- root_params[rev(order(root_params$do_maxy)), ]$model

  indices <- c(1, 2)
  delta_t_total <- 2
}

trim_thresholds <- c(lower_thresholds, upper_thresholds)
trim_models <- c(lower_model, upper_model)

min_string <- paste0(lower_change, collapse = "-")
max_string <- paste0(upper_change, collapse = "-")

optimal_string <- paste0(na.omit(trim_thresholds), collapse = "-")
climate_string <- gsub(" ", "-", gsub("\\/", "-", tolower(climate)))

glimpse(root_params)
```

# Calculate climate vectors
```{r }
vocc_by_year <- lapply(start_years, function(start_time) {
  try({
    vocc <- readRDS(paste0(
      "data/vocc-", ssid_string, "-", climate_string, "-",
      min_string, "to", max_string, "-", start_time, "-", delta_t_total, ".rds"
    ))
  })

  if (!exists("vocc")) {

    if (climate=="temperature") {
      climate_predictions <- readRDS("data/predicted_temp_allyears.rds") %>% mutate(temp = est)

    } else {
      pred_temp_do <- readRDS(here::here("analysis/tmb-sensor-explore/models/predicted-DO.rds"))

      if (length(variable_names) > 1) {
        climate_predictions <- replicate(length(variable_names), pred_temp_do, simplify = FALSE)
      } else {
        climate_predictions <- pred_temp_do
      }
    }
    end_time <- start_time + 2
    vocc <- make_vector_data(climate_predictions,
      variable_names = variable_names,
      ssid = model_ssid,
      start_time = start_time,
      end_time = end_time,
      skip_time = skip_time,
      input_cell_size = input_cell_size,
      scale_fac = scale_fac,
      delta_t_total = delta_t_total,
      delta_t_step = delta_t_step,
      indices = indices,
      min_thresholds = lower_change,
      max_thresholds = upper_change,
      round_fact = 10
    )

    vocc$var_1_min <- lower_change[1]
    vocc$var_2_min <- lower_change[2]
    vocc$var_1_max <- upper_change[1]
    vocc$var_2_max <- upper_change[2]

    saveRDS(vocc, file = paste0(
      "data/vocc-", ssid_string, "-", climate_string, "-",
      min_string, "to", max_string, "-", start_time, "-", delta_t_total, ".rds"
    ))
  }
  vocc
})
```

### Plot all possible vectors
```{r }
if (!exists("vocc")) {
  all_vector_plots <- lapply(vocc_by_year, function(vocc) {
    start_time <- vocc$start_time[1]

    if (climate == "temperature") {
      vocc$diff <- vocc$units_per_decade / 10 * delta_t_total

      vocc_plot <- plot_vocc(vocc,
        vec_aes = "distance",
        min_vec_plotted = input_cell_size,
        NA_label = ".",
        fill_col = "diff",
        fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
        raster_alpha = 1,
        raster_limits = c(-4, 3),
        white_zero = TRUE,
        vec_alpha = 0.25,
        axis_lables = FALSE,
        transform_col = no_trans
      ) + ggtitle(climate)
    }

    if (climate == "do") {
      vocc$diff <- vocc$units_per_decade / 10 * delta_t_total

      vocc_plot <- plot_vocc(vocc,
        vec_aes = "distance",
        min_vec_plotted = input_cell_size,
        NA_label = ".",
        fill_col = "diff",
        fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
        raster_alpha = 1,
        white_zero = TRUE,
        high_fill = "Steel Blue 4",
        low_fill = "Red 3",
        vec_alpha = 0.25,
        axis_lables = FALSE,
        transform_col = no_trans
      ) + ggtitle(climate)
    }

    if (climate == "do and temperature") {
      vocc$temp_diff <- vocc$temp.units_per_decade / 10 * delta_t_total

      vocc_plot_temp <- plot_vocc(vocc,
        vec_aes = "distance",
        min_vec_plotted = input_cell_size,
        NA_label = ".",
        fill_col = "temp_diff",
        fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
        raster_alpha = 1,
        white_zero = TRUE,
        vec_alpha = 0.25,
        axis_lables = FALSE,
        transform_col = no_trans
      ) + ggtitle("temperature")

      vocc$do_diff <- vocc$do_est.units_per_decade / 10 * delta_t_total

      vocc_plot_do <- plot_vocc(vocc,
        vec_aes = "distance",
        min_vec_plotted = input_cell_size,
        NA_label = ".",
        fill_col = "do_diff",
        fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
        raster_alpha = 1,
        white_zero = TRUE,
        high_fill = "Steel Blue 4",
        low_fill = "Red 3",
        vec_alpha = 0.25,
        axis_lables = FALSE,
        transform_col = no_trans
      ) + ggtitle("do")

      vocc_plot <- gridExtra::grid.arrange(
        vocc_plot_temp, vocc_plot_do,
        ncol = 2 # ,
        # top = grid::textGrob(paste("do decline of", lower_change[1], "or temperature increase of", upper_change[2], " over ", start_time, "-", start_time + delta_t_total, ""))
      )
    }
    vocc_plot
  })

  png(
    file = paste0(
      "figs/vocc-", climate_string, min_string, "to", max_string, "-", ssid_string, ".png"
    ),
    res = 600,
    units = "in",
    width = 8.5,
    height = 11
  )
  gridExtra::grid.arrange(
    grobs = all_vector_plots,
    ncol = 2,
    top = grid::textGrob(paste("Vectors for a DO decline of", lower_change[1], "and/or temperature increase of", upper_change[2], ""))
  )
  dev.off()
}
#print(all_vector_plots)
invisible(lapply(all_vector_plots, print))
```

### Filter vectors to match optimal range for chosen species
```{r}
rm(trimmed_vectors)
trimmed_vectors <- lapply(vocc_by_year, function(vocc) {
  vectors <- trim_vector_data(vocc, variable_names,
    lower_change, upper_change,
    lower_thresholds, upper_thresholds,
    cell_size = input_cell_size, dist_intercept = dist_intercept,
    max_dist = 120
  )
})

all_vectors <- do.call("rbind", trimmed_vectors)

## If VOCC change is a useful column?
# all_vectors$var_1_changeA <- all_vectors$var_1_min
# all_vectors$var_2_changeA <- all_vectors$var_2_min
# all_vectors$var_1_changeB <- all_vectors$var_1_max
# all_vectors$var_2_changeB <- all_vectors$var_2_max

all_vectors$var_1_min <- lower_thresholds[1]
all_vectors$var_2_min <- lower_thresholds[2]
all_vectors$var_1_max <- upper_thresholds[1]
all_vectors$var_2_max <- upper_thresholds[2]
```

### Plot remaining climate vectors for regions where change exceeds thresholds
```{r echo=FALSE, warning=FALSE}
if (climate == "temperature") {
  all_vectors$diff <- all_vectors$units_per_decade / 10 * delta_t_total
  raster_min <- min(all_vectors$diff)
  raster_max <- max(all_vectors$diff)
}

if (climate == "do") {
  all_vectors$diff <- all_vectors$units_per_decade / 10 * delta_t_total
  raster_min <- min(all_vectors$diff)
  raster_max <- max(all_vectors$diff)
}

if (climate == "do and temperature") {
  all_vectors$temp_diff <- all_vectors$temp.units_per_decade / 10 * delta_t_total
  all_vectors$do_diff <- all_vectors$do_est.units_per_decade / 10 * delta_t_total
  raster_min_temp <- min(all_vectors$temp_diff)
  raster_max_temp <- max(all_vectors$temp_diff)
  raster_min_do <- min(all_vectors$do_diff)
  raster_max_do <- max(all_vectors$do_diff)
}

trimmed_vector_plots <- lapply(trimmed_vectors, function(vectors) {
  try({
    start_time <- vectors$start_time[1]

    if (climate == "temperature") {
      vectors$diff <- vectors$units_per_decade / 10 * delta_t_total

      vocc_plot <- plot_vocc(vectors,
        vec_aes = "distance",
        min_vec_plotted = input_cell_size,
        NA_label = ".",
        fill_col = "diff",
        fill_label = paste(
          "Change from\n",
          start_time, "to", start_time + delta_t_total, ""
        ),
        raster_alpha = 1,
        raster_limits = c(raster_min, raster_max),
        white_zero = TRUE,
        vec_alpha = 0.25,
        axis_lables = FALSE,
        transform_col = no_trans
      ) + ggtitle(climate)
    }

    if (climate == "do") {
      vectors$diff <- vectors$units_per_decade / 10 * delta_t_total

      vocc_plot <- plot_vocc(vectors,
        vec_aes = "distance",
        min_vec_plotted = input_cell_size,
        NA_label = ".",
        fill_col = "diff",
        fill_label = paste(
          "Change from\n",
          start_time, "to", start_time + delta_t_total, ""
        ),
        raster_alpha = 1,
        raster_limits = c(raster_min, raster_max),
        white_zero = TRUE,
        high_fill = "Steel Blue 4",
        low_fill = "Red 3",
        vec_alpha = 0.25,
        axis_lables = FALSE,
        transform_col = no_trans
      ) + ggtitle(climate)
    }

    if (climate == "do and temperature") {
      vectors$temp_diff <- vectors$temp.units_per_decade / 10 * delta_t_total

      vocc_plot_temp <- plot_vocc(vectors,
        vec_aes = "distance",
        min_vec_plotted = input_cell_size,
        NA_label = ".",
        fill_col = "temp_diff",
        fill_label = paste(
          "Change from\n",
          start_time, "to", start_time + delta_t_total, ""
        ),
        raster_alpha = 1,
        raster_limits = c(raster_min_temp, raster_max_temp),
        white_zero = TRUE,
        vec_alpha = 0.25,
        axis_lables = FALSE,
        transform_col = no_trans
      ) + ggtitle("temperature")

      vectors$do_diff <- vectors$do_est.units_per_decade / 10 * delta_t_total

      vocc_plot_do <- plot_vocc(vectors,
        vec_aes = "distance",
        min_vec_plotted = input_cell_size,
        NA_label = ".",
        fill_col = "do_diff",
        fill_label = paste(
          "Change from\n",
          start_time, "to", start_time + delta_t_total, ""
        ),
        raster_alpha = 1,
        raster_limits = c(raster_min_do, raster_max_do),
        white_zero = TRUE,
        high_fill = "Steel Blue 4",
        low_fill = "Red 3",
        vec_alpha = 0.25,
        axis_lables = FALSE,
        transform_col = no_trans
      ) + ggtitle("do")

      vocc_plot <- gridExtra::grid.arrange(
        vocc_plot_temp, vocc_plot_do,
        ncol = 2,
        top = grid::textGrob(paste(
          "DO of ", optimal_values[1], "less", min_thresholds[1],
          "or temperature of ", optimal_values[2], "plus", max_thresholds[2],
          " over ", start_time, "-", start_time + delta_t_total, ""
        ))
      )
    }
    vocc_plot
  })
})

#print(trimmed_vector_plots)
invisible(lapply(trimmed_vector_plots, print))
```

```{r}
png(
  file = paste0(
    "figs/", spp,
    "/vocc-", min_string, "to", max_string, "-",
    climate_string, "-", spp, "-", age, "-", ssid_string, "-perc_", biomass_threshold, "-", optimal_string, "-allpairs-test.png"
  ),
  res = 600,
  units = "in",
  width = 8.5,
  height = 11
)
gridExtra::grid.arrange(
  grobs = trimmed_vector_plots,
  ncol = 2,
  top = grid::textGrob(paste("Vectors for cells exceeding", climate, "threshold of", optimal_string, "\n based on", biomass_threshold, "percent of maximum predicted", species, "biomass (", trim_models, ")"), gp = grid::gpar(fontsize = 16))
)
dev.off()
```

# Add species biomass data

Get all biomass predictions 
```{r echo=FALSE, warning=FALSE}
priors <- FALSE
covs <- "-trawled-ssid"
rm(biomass_predictions)
if (age == "imm") {
  biomass_predictions <- readRDS(paste0(
    "data/", spp, "/all-predictions-", spp, covs,
    "-imm-biomass-prior-", priors, ".rds"
  ))
}

if (age == "mature") {
  biomass_predictions <- readRDS(paste0(
    "data/", spp, "/all-predictions-", spp, covs,
    "-mat-biomass-prior-", priors, ".rds"
  ))
}
```

Extract biomass values to match spatial and temporal range of climate vectors
```{r}
# alldata <- lapply(biomass_list, function(biomass_predictions) {

biomass_by_year <- list()
for (j in seq_along(start_years)) {
  start_time <- start_years[[j]]
  end_time <- start_time + 2

  biomass_by_year[[j]] <- make_trend_data(biomass_predictions,
    ssid = model_ssid,
    start_time = start_time,
    skip_time = skip_time,
    end_time = end_time,
    input_cell_size = input_cell_size,
    scale_fac = scale_fac,
    delta_t_total = delta_t_total,
    delta_t_step = delta_t_step,
    indices = indices,
    variable_names = "est_exp"
  )
}

biomass_change <- do.call("rbind", biomass_by_year)

biomass_covs <- biomass_predictions %>%
  select(X, Y, depth, trawled, muddy, any_rock, omega_s) %>%
  distinct()

biomass_change <- biomass_change %>%
  rename(start_density = est_exp_s, end_density = est_exp_e) %>%
  dplyr::select(icell, x, y, start_time, start_density, end_density)
# dplyr::select(-N, -time_step, - units_per_decade, -slope)

biomass <- left_join(biomass_change, biomass_covs, by = c("x" = "X", "y" = "Y"))

# keep just one row for each cell that exceeded the end climate threshold
# FIXME: this code could/maybe should be rewritten with an if_else mutate and then distinct...
source_cells <- all_vectors %>%
  filter(distance > (-dist_intercept)) %>%
  distinct(icell, start_time, .keep_all = TRUE) %>%
  select(-target_X, -target_Y, -target_values, -tid)

source_biomass <- left_join(source_cells, biomass,
  by = c("icell", "x", "y", "start_time")
) %>%
  mutate(cell_type = "source") %>%
  filter(!is.na(start_density))

# identify all possible control cells have been given a distance == - dist_intercept
control_cells <- all_vectors %>%
  filter(distance == (-dist_intercept)) %>%
  distinct(icell, start_time, .keep_all = TRUE) %>%
  mutate(distance = (-dist_intercept)) %>%
  select(-target_X, -target_Y, -target_values, -tid)

control_biomass <- left_join(control_cells, biomass,
  by = c("icell", "x", "y", "start_time")
) %>%
  mutate(cell_type = "control") %>%
  filter(!is.na(start_density))

alldata <- rbind(source_biomass, control_biomass)
alldata <- mutate(alldata, treat = ifelse(cell_type == "source", 1, 0)) %>%
  select(-mean_target_X, -mean_target_Y) %>%
  rename(vect_dist = distance)
alldata$treat <- as.factor(alldata$treat)

if(is.na(alldata$var_1_max[2])) {
  alldata <- alldata %>% select(-var_1_max)
}
if(is.na(alldata$var_1_min[2])) {
  alldata <- alldata %>% select(-var_1_min)
}
if(is.na(alldata$var_2_max[2])) {
  alldata <- alldata %>% select(-var_2_max)
}
if(is.na(alldata$var_2_min[2])) {
  alldata <- alldata %>% select(-var_2_min)
}
# alldata
# })
```

# Select matched control cells

```{r}
rm(vocc_by_sp)
# try({
#   vocc_by_sp <- readRDS(paste0(
#     "data/BACI-", species, "-", age, "-", ssid_string, "-", climate_string, "-", optimal_string, "-", min_string, "to", max_string, "-multiyear-all.rds"
#   ))
# })

# FIXME: should the cells be matched by substrate even when biomass not modelled using substrate?

# alldata[ , colSums(is.na(alldata)) == 0]
# alldata <- alldata %>% select(-var_1_min)

match_formula <- paste0("treat ~ trawled + x + y + depth_st")
# depth.st + omega_s + muddy.st + any_rock.st +
for (j in (variable_names)) {
  match_formula <- paste0("", match_formula, " + ", j, "_s.st")
  start_value.st <- paste0("", j, "_s.st")
  start_value <- paste0("", j, "_s")

  alldata <- alldata %>% mutate((!!start_value.st) :=
    arm::rescale(alldata[[start_value]]))
  # (UQ(rlang::sym(paste0(variable_names[j], "_s")))))
}

# if (!exists("vocc_by_sp")) {
# matched <- lapply(alldata, function(alldata) {
## trim non-overlapping values to reduce size of dataset
control_min <- quantile(alldata$start_density[alldata$treat == "0"], 0.01)
source_min <- quantile(alldata$start_density[alldata$treat == "1"], 0.01)
# control_max <- quantile(alldata$start_density[alldata$treat=="0"], 1)
# source_max <- quantile(alldata$start_density[alldata$treat=="1"], 1)

group_min <- max(control_min, source_min)
# group_max <- min(control_max, source_max)
alldata <- filter(alldata, start_density >= group_min)
# %>% filter(start_density <= group_max)

# need to be standardized? seems not
alldata$depth2 <- alldata$depth^2
alldata$depth_st <- arm::rescale(log(alldata$depth))
# alldata$muddy.st <- arm::rescale(alldata$muddy)
# alldata$any_rock.st <- arm::rescale(alldata$any_rock)

# bin some variables for exact matching
alldata$depth.st <- round(arm::rescale(log(alldata$depth)) * 5) / 5
alldata$start_density.st <- round(arm::rescale(log(alldata$start_density)) * 5) / 5

## Using optimization ... doesn't work
# alldata$start_density_st <- round(arm::rescale(log(alldata$start_density)), 1)
# fit_formula <-
#   paste0("treat ~ start_density_st + depth.st + muddy.st + any_rock.st")
# # for (j in (variable_names)) {
# #   fit_formula <- paste0("", fit_formula, " + ", j, "_s")
# #   }
# # options("optmatch_max_problem_size" = Inf)
# matched <- MatchIt::matchit(as.formula(fit_formula),
#     method = "optimal", data = alldata)

# # Using exact match and rounding
# # Choose data order if not using random
# alldata <- alldata[(order(alldata$y)), ]
# if (order == "increasing") alldata <- alldata[rev(order(alldata$start_density)), ]
# if (order == "decreasing") alldata <- alldata[rev(order(alldata$start_density)), ]
# set.seed(99999)
matched <- MatchIt::matchit(as.formula(match_formula),
  exact = c("start_density.st", "start_time"),
  method = "nearest", m.order = "random",
  distance = "probit",
  discard = "both",
  data = alldata
)
# matched
# })
# }
```

Print diagnostics for matching
```{r echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
if (!exists("vocc_by_sp")) {
  match_plots <- plot(matched, interactive = F)
  summary(matched)
}
```

Prep matched data for analysis
```{r}
# if (!exists("vocc_by_sp")) {
# vocc_by_sp <- lapply(matched, function(matched) {
mdata <- MatchIt::match.data(matched, distance = "pscore")

sources <- mdata[row.names(matched$match.matrix), ]
sources$origobs <- row.names(sources)
sources$matchobs <- matched$match.matrix[, 1]

controls <- mdata[matched$match.matrix, ]
controls$origobs <- row.names(controls)
controls$matchobs <- matched$match.matrix[, 1]

mdata <- rbind(sources, controls) %>%
  tidyr::drop_na() %>%
  filter(matchobs != "-1")

matched_time0 <- mdata %>% mutate(after = 0, density = start_density)
# %>% dplyr::select(-start_density, -end_density)
matched_time1 <- mdata %>% mutate(after = 1, density = end_density)
# %>% dplyr::select(-start_density, -end_density)

for (j in seq_along(variable_names)) {
  climate_var <- paste0(variable_names[j])
  matched_time0 <- matched_time0 %>%
    mutate((!!climate_var) := (UQ(rlang::sym(paste0(variable_names[j], "_s")))))
  matched_time1 <- matched_time1 %>%
    mutate((!!climate_var) := (UQ(rlang::sym(paste0(variable_names[j], "_e")))))
}

vocc_by_sp <- rbind(matched_time0, matched_time1)

# calculate standardized climate variables
for (j in seq_along(variable_names)) {
  mean.var <- paste0("mean.", variable_names[j])
  sd.var <- paste0("sd.", variable_names[j])
  std.var <- paste0("std_", variable_names[j])
  std.var2 <- paste0("std_", variable_names[j], "2")

  vocc_by_sp <- mutate(
    vocc_by_sp,
    (!!mean.var) := mean(UQ(rlang::sym(paste0(variable_names[j]))), na.rm = TRUE),
    (!!sd.var) := sd(UQ(rlang::sym(paste0(variable_names[j]))), na.rm = TRUE),
    (!!std.var) := ((UQ(rlang::sym(paste0(variable_names[j]))) -
      mean(UQ(rlang::sym(paste0(variable_names[j]))), na.rm = TRUE)) /
      (sd(UQ(rlang::sym(paste0(variable_names[j]))), na.rm = TRUE)) * 2),
    (!!std.var2) := ((UQ(rlang::sym(paste0(variable_names[j]))) -
      mean(UQ(rlang::sym(paste0(variable_names[j]))), na.rm = TRUE)) /
      (sd(UQ(rlang::sym(paste0(variable_names[j]))), na.rm = TRUE) * 2))^2
  )
}

# rename spatial coordinates to avoid naming conflict with y~x notation
data <- vocc_by_sp %>%
  rename(X = x, Y = y)
# %>% rename(distance = vect_dist) %>%
# select(-start_density.st, -depth.st, -muddy.st, -any_rock.st, -omega_s.st)

# other variable prep
data$log_density <- log(data$density)
data$log_depth <- log(data$depth)
data$std_log_depth <- arm::rescale(log(data$depth))
data$vect_dist2 <- data$vect_dist^2
data <- mutate(data, time = ifelse(after == 1, "end", "begin"))
data$matchobs <- as.factor(data$matchobs)
data$origobs <- as.factor(data$origobs)
data$start_time <- as.factor(data$start_time)

#  data
# })
vocc_by_sp <- list()
vocc_by_sp[[1]] <- data
vocc_by_sp[[3]] <- match_formula
vocc_by_sp[[4]] <- match_plots
saveRDS(vocc_by_sp, file = paste0(
  "data/", spp,
  "/BACI-", spp, "-", age, "-", ssid_string, "-", climate_string, "-", optimal_string, "-", min_string, "to", max_string, "-multiyear-all.rds"
))
# }
```

## Save spp dataframe
```{r}
# biomass_threshold <- "50"
optimal_string <- paste0(na.omit(trim_thresholds), collapse = "-")
# climate <- "temperature"
# try({
#   vocc_by_sp <- readRDS(paste0(
#     "data/", spp,
#     "/BACI-", spp, "-", age, "-", ssid_string, "-", climate_string, "-", optimal_string, "-", min_string, "to", max_string, "-multiyear-all.rds"
#   ))
# })

n <- nrow(vocc_by_sp[[1]])
species <- rep(species, n)
mature <- rep(age, n)
ssid <- rep(ssid_string, n)
data <- vocc_by_sp[[1]]
df <- cbind(species, mature, ssid, data)
species <- species[[1]]

# folder to hold vocc data
biomass_threshold_string <- paste0("perc_", biomass_threshold)

dir.create(file.path("data/log-both"))
dir.create(file.path("data/log-both", climate_string))
dir.create(file.path("data/log-both", climate_string, biomass_threshold_string))

if (climate == "temperature") {
  dir.create(file.path("data/log-both", climate_string, biomass_threshold_string, max_string))
  dir.create(file.path("data/log-both", climate_string, biomass_threshold_string, max_string, age))
}
if (climate == "temperature") {
  saveRDS(df, file = paste0("data/log-both/", climate_string, "/", "perc_", biomass_threshold, "/", max_string, "/", age, "/vocc-", climate_string, "-", max_string, "-", spp, "-", age, "-", optimal_string, "-", ssid_string, "-all.rds"))
}

if (climate == "do") {
  dir.create(file.path("data/log-both", climate_string, biomass_threshold_string, min_string))
  dir.create(file.path("data/log-both", climate_string, biomass_threshold_string, min_string, age))
}
if (climate == "do") {
  saveRDS(df, file = paste0("data/log-both/", climate_string, "/", "perc_", biomass_threshold, "/", min_string, "/", age, "/vocc-", climate_string, "-", min_string, "to", max_string, "-", spp, "-", age, "-", optimal_string, "-", ssid_string, "-all.rds"))
}

if (climate == "do and temperature") {
  dir.create(file.path("data/log-both", climate_string, biomass_threshold_string, age))
}

if (climate == "do and temperature") {
  saveRDS(df, file = paste0("data/log-both/", climate_string, "/", "perc_", biomass_threshold, "/", age, "/vocc-", climate_string, "-", min_string, "to", max_string, "-", spp, "-", age, "-", optimal_string, "-", ssid_string, "-all.rds"))
}
```

```{r}
try({
  vocc_by_sp <- readRDS(paste0(
    "data/", spp,
    "/BACI-", spp, "-", age, "-", ssid_string, "-", climate_string, "-",
    min_string, "to", max_string, "-multiyear-all.rds"
  ))
})
data <- vocc_by_sp[[1]]
```

### Plot matches
```{r}
# FIXME: SAVE PLOTS?

dat_summary <- data %>%
  group_by(start_time, cell_type) %>%
  tally() %>%
  filter(cell_type == "source")
data_depth <- data %>%
  group_by(start_time) %>%
  mutate(mean_depth = mean(depth))

p1 <- ggplot(
  data,
  # ggplot(filter(data, as.numeric(matchobs) <= 550),
  aes(y = Y, x = X, shape = as.factor(treat), colour = as.factor(matchobs))
) +
  geom_point(alpha = 0.8, size = 1) +
  scale_shape_manual(values = c(0, 15)) +
  scale_colour_viridis_d(option = "C", guide = FALSE) +
  facet_wrap(~start_time) +
  theme_bw()
p1

lapply(variable_names, function(j) {
  ggplot(data, aes(y = log_density, x = UQ(rlang::sym(paste(j))), colour = time)) +
    geom_point(alpha = 0.35) +
    scale_colour_viridis_d(option = "C") +
    facet_grid(start_time ~ cell_type) +
    theme_bw()
})

p2 <- ggplot(data, aes(std_log_depth, fill = cell_type)) +
  geom_density(alpha = 0.5) +
  scale_fill_viridis_d(option = "C") +
  facet_wrap(~start_time) +
  theme_bw()
p2

# ggplot(data, aes(y=log_density, x=depth, colour=cell_type)) +
#   geom_point(alpha=0.35) +
#   scale_colour_viridis_d(option="C") +
#     facet_grid(start_time~time) +
#   theme_bw()

p3 <- ggplot(data, aes(log_density, fill = cell_type)) +
  geom_density(alpha = 0.5) +
  scale_fill_viridis_d(option = "C") +
  facet_grid(time ~ start_time, scales = "fixed") +
  theme_bw() + theme(legend.position = "none")
p3

p4 <- ggplot(data, aes(log_density, fill = cell_type)) +
  geom_histogram(alpha = 0.45) +
  scale_fill_viridis_d(option = "C") +
  facet_grid(time ~ start_time) +
  theme_bw() + theme(legend.position = "none")
p4

dat_summary <- data %>%
  group_by(start_time, cell_type) %>%
  tally() %>%
  filter(cell_type == "source")

p5 <- ggplot(data, aes(y = log_density, x = cell_type, fill = cell_type)) +
  geom_boxplot(alpha = 0.5) +
  scale_fill_viridis_d(option = "C") +
  geom_text(
    data = dat_summary,
    aes(cell_type, min(data$log_density) + 1, label = n), hjust = 1.25
  ) +
  facet_grid(time ~ start_time) +
  theme_bw() + theme(legend.position = "none")
p5
```

```{r}
png(
  file = paste0(
    "figs/", spp,
    "/matches-", min_string, "to", max_string, "-",
    climate_string, "-", spp, "-", age, "-", ssid_string, "-",
    climate_string, "-perc_", biomass_threshold, "-", optimal_string, ".png"
  ),
  res = 600,
  units = "in",
  width = 8.5,
  height = 11
)
gridExtra::grid.arrange(
  p1, p2, p3, p4, p5,
  ncol = 2,
  top = grid::textGrob(paste("Matched pairs for", climate, "threshold of", optimal_string, "\n based on", biomass_threshold, "percent of maximum predicted", species, "biomass (", age, ")"), gp = grid::gpar(fontsize = 16))
)
dev.off()
```