### 2019-10-21 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
update climate models/predictions

---
output: html_document
editor_options: 
  chunk_output_type: inline
---
# Fit spatial temporal models to bottom temperature

```{r setup}
library(tidyverse)
library(sdmTMB)
library(future)
library(dplyr)

plan(multiprocess)
```

Load data
```{r}
all_sensor <- readRDS(here::here("analysis/tmb-sensor-explore/data/dat-sensor-trawl-meanSST-all.rds"))
# all_sensor <- readRDS(here::here("analysis/tmb-sensor-explore/data/dat-sensor-trawl-processed.rds")) # no SST
# glimpse(all_sensor)
# 
# replace obviously faulty depth data with NA
#all_sensor$depth_m[all_sensor$depth_m < 10] <- NA # 5,139

all_depth <- all_sensor %>%
  #filter(depth_max>10) %>% # maybe we should filter incase missing depth values also indicate a sensor fail?
  # filter missing location data and trial year
  dplyr::filter(!is.na(latitude), !is.na(longitude)) %>%

  # convert lat and lon to UTMs
  dplyr::mutate(X = longitude, Y = latitude) %>%
  gfplot:::ll2utm(., utm_zone = 9) %>%

  # interpolate missing depth values
  dplyr::rename(depth = depth_m) %>%
  gfplot:::interp_survey_bathymetry()

unique(all_depth$data$year)

d_trawl <- all_depth$data %>% 
  # remove obvious sensor fails, d_trawl %>% filter(depth_max<10), fishing event ids = 481861, 2179089
  filter(depth_max>10) %>% 
  # filter(ssid != 1) %>%
  # filter(ssid != 3) %>%
  # filter(ssid != 4) %>%
  # filter(ssid != 16) %>%
  gfplot:::scale_survey_predictors()

hist(d_trawl$temperature_c)

d_trawl <- d_trawl[!is.na(d_trawl$depth), ]
d_trawl <- d_trawl[!is.na(d_trawl$temperature_c), ]
#d_trawl <- d_trawl[!is.na(d_trawl$SST), ]
#d_trawl <- d_trawl[!is.na(d_trawl$meanSST), ]
d_trawl <- mutate(d_trawl,
  DOY = lubridate::yday(date),
  DOY_scaled = arm::rescale(lubridate::yday(date)),
  DOY_scaled2 = arm::rescale(lubridate::yday(date))^2,
  #depth_scaled = arm::rescale(log(depth_m)),
  SST_scaled = arm::rescale(meanSST),
  #depth_scaled2 = depth_scaled^2,
  depth_scaled3 = depth_scaled^3,
  shallow = ifelse(depth>35,0,1),
  deep = ifelse(depth<35,0,1)
)
# d_trawl[is.na(d_trawl$shallow), ]
# d_trawl[is.na(d_trawl$SST_scaled), ]

glimpse(d_trawl)

#d_trawl <- d_trawl %>% dplyr::filter(year>2003)
```

Explore SST 
```{r}
ggplot(d_trawl, aes_string("X", "Y", colour = "SST")) +
    geom_point() +
    facet_wrap(~year) +
    coord_fixed() +
    scale_color_viridis_c()
```
```{r}
ggplot(d_trawl, aes(DOY, SST, colour = -depth)) + geom_point() + facet_wrap(~year) 
```

```{r}
ggplot(dplyr::filter(d_trawl, depth <35), aes(SST,temperature_c, colour = -depth)) + geom_point()
ggplot(dplyr::filter(d_trawl, depth <35), aes(SST,temperature_c, colour = DOY)) + geom_point()

```

Explore mean SST 
```{r}
ggplot(d_trawl, aes_string("X", "Y", colour = "meanSST")) +
    geom_point() +
    facet_wrap(~year) +
    coord_fixed() +
    scale_color_viridis_c()
```

```{r}
ggplot(dplyr::filter(d_trawl1, shallow==1), aes(SST_scaled,temperature_c, colour = -depth)) + geom_point()
```

EXLORE TEMPERATURE at DEPTH

```{r}
ggplot(d_trawl, aes_string("X", "Y", colour = "temperature_c")) +
    geom_point() +
    facet_wrap(~year) +
    coord_fixed() +
    scale_color_viridis_c()
```

```{r}
ggplot(d_trawl, aes(log(depth),temperature_c, colour = depth_max)) + geom_point() + facet_wrap(~ssid)
```

```{r}
ggplot(d_trawl, aes(DOY, temperature_c, colour = -depth, alpha=0.5)) + geom_point() + facet_wrap(~year) 
```

```{r}
library(gbm)

m <- gbm( temperature_c ~ depth_scaled + X + Y, data = d_trawl, n.trees = 2000, interaction.depth = 3, shrinkage = 0.02)
m2 <- gbm( temperature_c ~ SST_scaled + depth_scaled + X + Y, data = d_trawl, n.trees = 2000, interaction.depth = 3, shrinkage = 0.02)

plot(m,i.var=2)
plot(m,i.var=3)
plot(m,i.var=4)
plot(m,i.var=5)

plot(m2,i.var=1)
plot(m2,i.var=2)
plot(m,i.var=3)
plot(m,i.var=4)

plot(m2,i.var=1:2)
plot(m2,i.var=c(1,3))
plot(m2,i.var=c(1,4))
d_trawl$r <- predict(m, n.trees = 2000) - d_trawl1$temperature_c

ggplot(d_trawl, aes(depth_scaled, r)) +
  geom_point(alpha=0.4) +
  ylim(-8,8) +
  geom_smooth()
```

### Functions for model exploration:
```{r}
# aic <- function(m) {
#   k <- length(m$model$par)
#   nll <- m$model$objective
#   2 * k - 2 * (-nll)
# }

generate_cv_args <- function(
                             formula = formula,
                             time_varying = time_varying,
                             ar1_fields = c(FALSE), # simiple, fast defaults
                             include_spatial = c(FALSE), # simiple, fast defaults
                             n_knots = c(100), # simiple, fast defaults
                             seed = c(999)) {
  arguments <- expand.grid(
    formula = formula,
    time_varying = time_varying,
    include_spatial = include_spatial,
    ar1_fields = ar1_fields,
    n_knots = n_knots,
    seed = seed
  )
  as.data.frame(arguments)
}

sdmTMB_cv_batch <- function(cv_args, data, ...) {
  cv_batch <- purrr::pmap(cv_args, sdmTMB_cv, data = d_trawl, ...)
  cv_out <- cv_args
  cv_out$cv_loglik <- c()
  cv_out$all_converged <- c()
  cv_out$max_grad <- c()
  for (i in seq_len(nrow(cv_args))) {
    cv_out$cv_loglik[[i]] <- round(cv_batch[[i]]$sum_loglik, 1)
    cv_out$all_converged[[i]] <- all(cv_batch[[i]]$converged)
    cv_out$max_grad[[i]] <- max(cv_batch[[i]]$max_grad)
    if (is.null(cv_out$time_varying[[i]])) {
      cv_out$time_varying[[i]] <- as.character(NA)
    }
  }
  cv_out$formula <- as.character(cv_out$formula)
  cv_out$time_varying <- as.character(cv_out$time_varying)

  cv_out <- cv_out %>%
    group_by(seed) %>%
    mutate(., max_ll_per_seed = max(cv_loglik)) %>%
    mutate(., delta_ll = cv_loglik - max_ll_per_seed) %>%
    ungroup()
  row.names(cv_out) <- NULL
  list(summary = as.data.frame(cv_out), cv_model_set = cv_batch)
}

# # this code keeps breaking... working, fails, worked again, then failed again...
# plot_cv_meshes <- function(sdmTMB_cv_output) {
#   cv_output <- sdmTMB_cv_output
#   k <- max(cv_output$data$cv_fold) # should probably rename cv_fold to fold_id
#   if (k == 2 | k == 5 | k == 10) {
#     op <- graphics::par(
#       mfrow = c((sqrt(k)), ceiling(sqrt(k))),
#       mar = c(1, 1, 1, 1)
#       )
#   } else {
#     op <- graphics::par(
#       mfrow = c(ceiling(sqrt(k)), ceiling(sqrt(k))),
#       mar = c(1, 1, 1, 1)
#     )
#   }
#   for (i in seq_len(k)){
#     plot_spde(cv_output$models[[i]]$spde)
# #FIXME: Error in xy.coords(x, y, xlabel, ylabel, log) :
# #  'x' is a list, but does not have components 'x' and 'y'
#   }
#   graphics::par(op)
# }
```

Short example: 
```{r, eval=FALSE}
cv_args <- generate_cv_args(
  formula = list(
    temperature_c ~ 0 + as.factor(year) + SST_scaled + depth_scaled + depth_scaled2 + depth_scaled3
  ),
  time_varying = list(NULL),
  ar1_fields = c(FALSE),
  include_spatial = c(FALSE),
  n_knots = c(135, 140, 145, 150),
  seed = c(1)
)

cv <- sdmTMB_cv_batch(cv_args, d_trawl, k_folds = 4, time = "year", x = "X", y = "Y")
glimpse(cv$summary)
write.csv(cv$summary, file = "cv-output/cv-test.csv")
cv_test <- read_csv(file = "cv-output/cv-test.csv")
View(cv$summary)
# plot_cv_meshes(cv$cv_model_set[[1]])
# plot_cv_meshes(cv$cv_model_set[[2]])
```

# 10-fold cross-validation

## Just compare fixed slope models with simplest random effect structure
```{r}
cv_args <- generate_cv_args(
  formula = list(
    temperature_c ~ 0 + SST_scaled*shallow + depth_scaled,
    temperature_c ~ 0 + SST_scaled*shallow + depth_scaled + depth_scaled2,
    temperature_c ~ 0 + SST_scaled*shallow + depth_scaled + depth_scaled2 + depth_scaled3,
    temperature_c ~ 1 + SST_scaled + depth_scaled,
    temperature_c ~ 1 + SST_scaled + depth_scaled + depth_scaled2,
    temperature_c ~ 1 + SST_scaled + depth_scaled + depth_scaled2 + depth_scaled3,
    temperature_c ~ 0 + as.factor(year) + depth_scaled,
    temperature_c ~ 0 + as.factor(year) + depth_scaled + depth_scaled2,
    temperature_c ~ 0 + as.factor(year) + depth_scaled + depth_scaled2 + depth_scaled3,
    temperature_c ~ 0 + as.factor(year) + SST_scaled + depth_scaled,
    temperature_c ~ 0 + as.factor(year) + SST_scaled + depth_scaled + depth_scaled2,
    temperature_c ~ 0 + as.factor(year) + SST_scaled + depth_scaled + depth_scaled2 + depth_scaled3,
    temperature_c ~ 0 + as.factor(year) + SST_scaled * depth_scaled,
    temperature_c ~ 0 + as.factor(year) + SST_scaled * depth_scaled + SST_scaled * depth_scaled2,
    temperature_c ~ 0 + as.factor(year) + SST_scaled * depth_scaled + SST_scaled * depth_scaled2 + SST_scaled * depth_scaled3
  ),
  time_varying = list(NULL),
  ar1_fields = c(FALSE),
  include_spatial = c(FALSE),
  n_knots = c(100),
  seed = c(1111, 2222, 3333, 4444)
)

non_time_varying <- sdmTMB_cv_batch(cv_args, d_trawl, time = "year", x = "X", y = "Y")
View(non_time_varying$summary)
write.csv(non_time_varying$summary, file = "cv-output/cv-non-time-varying.csv")
saveRDS(non_time_varying, file = "cv-output/cv-non-time-varying.rds")
glimpse(non_time_varying$summary)
```

## Compare time-varying slope models still with simplest random effect structure
All of top models for any seed contained year and none showed a linear effect of depth
All are better than fixed slope models
```{r}
cv_args <- generate_cv_args(
  formula = list(
    #temperature_c ~ 0 + SST_scaled*shallow,
    temperature_c ~ 0 + as.factor(year) + SST_scaled*shallow,
    #temperature_c ~ 1 + SST_scaled,
    temperature_c ~ 0 + as.factor(year) + SST_scaled,
    temperature_c ~ 0 + as.factor(year)
  ),
  time_varying = list(
    ~ 0 + depth_scaled + depth_scaled2,
    ~ 0 + depth_scaled + depth_scaled2 + depth_scaled3
  ),
  ar1_fields = c(FALSE),
  include_spatial = c(FALSE),
  n_knots = c(100), # keep same n_knots
  seed = c(1111, 2222, 3333, 4444) # keep same seeds
)

time_varying <- sdmTMB_cv_batch(cv_args, d_trawl, time = "year", x = "X", y = "Y")

View(time_varying$summary)
write.csv(time_varying$summary, file = "cv-output/cv-time-varying.csv")
saveRDS(time_varying, file = "cv-output/cv-time-varying.rds")
glimpse(time_varying$summary)
```

## Now vary random effect configuation for top model(s) 
a separate spatial field is in all models with highest likelihoods for each seed
AR1 doesn't make much difference (differences between years appear independent)
```{r}
cv_args <- generate_cv_args(
  formula = list(
    temperature_c ~ 0 + as.factor(year) + SST_scaled*shallow,
    temperature_c ~ 0 + as.factor(year) + SST_scaled,
    temperature_c ~ 0 + as.factor(year)
  ),
  time_varying = list(
    ~ 0 + depth_scaled + depth_scaled2
  ),
  ar1_fields = c(FALSE, TRUE),
  include_spatial = c(FALSE, TRUE),
  n_knots = c(100),
  seed = c(1111, 2222, 3333, 4444)
)

vary_rf <- sdmTMB_cv_batch(cv_args, d_trawl, time = "year", x = "X", y = "Y")

View(vary_rf$summary)
write.csv(vary_rf$summary, file = "cv-output/cv-varying-rf.csv")
saveRDS(vary_rf, file = "cv-output/cv-varying-rf.rds")
glimpse(vary_rf$summary)
```

## Change knot number

```{r}
cv_args <- generate_cv_args(
    formula = list(
    # temperature_c ~ 0 + as.factor(year) + SST_scaled,
    temperature_c ~ 0 + as.factor(year)
  ),
  time_varying = list(
    ~ 0 + depth_scaled + depth_scaled2
  ),
  ar1_fields = c(FALSE,TRUE),
  include_spatial = c(TRUE),
  n_knots = c(140,180,220,250),
  #seed = c(1111, 2222, 3333, 4444) 
  seed = c(5555, 6666, 7777, 8888, 9999)
)

more_knots <- sdmTMB_cv_batch(cv_args, d_trawl, time = "year", x = "X", y = "Y")

View(more_knots$summary)
write.csv(more_knots$summary, file = "cv-output/cv-more-knots2.csv")
saveRDS(more_knots, file = "cv-output/cv-more-knots2.rds")
glimpse(more_knots$summary)
```

## Change seeds 
Also increased knot number too 
Models without SST_scaled are usually better

```{r}
cv_args <- generate_cv_args(
    formula = list(
    temperature_c ~ 0 + as.factor(year) + SST_scaled*shallow,
    temperature_c ~ 0 + as.factor(year) + SST_scaled,
    temperature_c ~ 0 + as.factor(year)
  ),
  time_varying = list(
    ~ 0 + depth_scaled + depth_scaled2
  ),
  ar1_fields = c(FALSE, TRUE),
  include_spatial = c(TRUE, FALSE),
  n_knots = c(130),
  seed = c(5555, 6666, 7777, 8888, 9999)
)

more_seeds <- sdmTMB_cv_batch(cv_args, d_trawl, time = "year", x = "X", y = "Y")

View(more_seeds$summary)
write.csv(more_seeds$summary, file = "cv-output/cv-more-seeds.csv")
saveRDS(more_seeds, file = "cv-output/cv-more-seeds.rds")
glimpse(more_seeds$summary)
```

Load saved batches:
```{r, eval=FALSE}
# non_time_varying <- readRDS(file = "cv-output/cv-non-time-varying.rds")
# time_varying <- readRDS(file = "cv-output/cv-time-varying.rds")
# vary_rf <- readRDS(file = "cv-output/cv-varying-rf.rds")
# knot_refining <- readRDS(file = "cv-output/cv-refine-knots.rds")

non_time_varying <- read.csv(file = "cv-output/cv-non-time-varying.csv")
time_varying <- read.csv(file = "cv-output/cv-time-varying.csv")
vary_rf <- read.csv(file = "cv-output/cv-varying-rf.csv")
non_time_varying1 <- read.csv(file = "cv-output/cv-non-time-varying1.csv")
time_varying1 <- read.csv(file = "cv-output/cv-time-varying1.csv")
vary_rf1 <- read.csv(file = "cv-output/cv-varying-rf1.csv")
refine_knots1 <- read.csv(file = "cv-output/cv-refine-knots1.csv")
more_seeds <- read.csv(file = "cv-output/cv-more-seeds.csv")
more_seeds1 <- read.csv(file = "cv-output/cv-more-seeds1.csv")

all_cv_loglik <- rbind(
  non_time_varying,
  time_varying,
  vary_rf,
  non_time_varying1,
  time_varying1,
  vary_rf1,
  refine_knots1,
  more_seeds,
  more_seeds1,
  deparse.level = 1, make.row.names = TRUE
)

by_seed <- all_cv_loglik %>%
  group_by(seed) %>%
  mutate(., max_ll_per_seed = max(cv_loglik)) %>%
  mutate(., delta_ll = cv_loglik - max_ll_per_seed)

View(by_seed)
all_cv_loglik <- by_seed %>% ungroup()

write.csv(all_cv_loglik, file = "all_cv_loglik.csv")
```

Current top model

```{r}
spde <- make_spde(d_trawl$X, d_trawl$Y, n_knots = 400)
plot_spde(spde)
```

```{r}
m_temp_rw <- sdmTMB(d_trawl,
  temperature_c ~ 0 + as.factor(year),
  time_varying = ~ 0 + depth_scaled + depth_scaled2,
  ar1_fields = TRUE, include_spatial = TRUE, # changed AR1_field to true
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = FALSE
)

saveRDS(m_temp_rw, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-400kn.rds")))

```

```{r}
predictions_rw <- predict(m_temp_rw)
predictions_rw$residuals <- residuals(m_temp_rw)
cor(predictions_rw$est, predictions_rw$temperature_c)^2 #
# aic(m_temp3_rw)

```

Residual plots:
```{r}
ggplot(predictions_rw, aes(est, residuals)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)
```

```{r}
ggplot(predictions_rw, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)
```

```{r}
ggplot(predictions_rw, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth()
```

Does a fixed interaction with SST_scaled do any better?

```{r eval=FALSE}
m_temp3 <- sdmTMB(d_trawl,
  temperature_c ~ 0 + as.factor(year) + SST_scaled*depth_scaled + SST_scaled*depth_scaled2 ,
  ar1_fields = FALSE, include_spatial = TRUE,
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = TRUE
)

predictions3 <- predict(m_temp3)
predictions3$residuals <- residuals(m_temp3)
cor(predictions3$est, predictions3$temperature_c)^2 #
```

```{r eval=FALSE}
ggplot(predictions3, aes(est, residuals)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)
```

```{r eval=FALSE}
ggplot(predictions3, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~year)
```

```{r eval=FALSE}
ggplot(predictions3, aes(depth_scaled, residuals)) +
  geom_point() +
  geom_smooth()
```

#FIXME: DON'T HAVE SST TEMP FOR PREDICITON GRID YET...

## Plotting predictions from saved temperature model

```{r}
nd_all <- readRDS(paste0("../VOCC/data/nd_all_synoptic.rds")) %>% filter(year<2019)
unique(nd_all$year)
glimpse(nd_all)
predict_temp <- predict(m_temp_rw, newdata = nd_all)
saveRDS(predict_temp, file = "../VOCC/data/predicted_temp_allyears_revised")
p_nd <- predict_temp
#p_nd <- readRDS("../VOCC/data/predicted_temp_allyears.rds")
# View(p_nd)
```

```{r}
plot_map_raster <- function(dat, column = "est") {
    ggplot(dat, aes_string("X", "Y", fill = column)) +
    geom_raster() +
    coord_fixed()+
gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank())
}

plot_map_raster(p_nd, "est") +
  scale_fill_viridis_c(trans = "log") +
  facet_wrap(~year) +
  ggtitle("Prediction (fixed effects + all random effects)")
```

```{r}
p_nd1 <- p_nd %>% filter(ssid!=16) %>% filter(ssid!=4)

breaks <- scales::trans_breaks(no_trans[["transform"]], no_trans[["inverse"]],n = 6)
labels <- function(x) format(x, digits = 2, scientific = FALSE)

pred_1 <- plot_map_raster(p_nd1, "est") +
  scale_fill_viridis_c(labels = labels, 
    #trans = "sqrt",
    option = "C",
    #breaks = c(0,2,4,6,8), 
    limits = c(5, 11), 
    na.value = "red") +
  facet_wrap(~year) +
  labs(fill = "°C") + 
  theme(legend.position = "none") +
  ggtitle("Prediction (fixed effects + all random effects)") 
pred_1
```

```{r}
p_nd4 <- p_nd %>% filter(ssid==4)
pred_4 <- plot_map_raster(p_nd4, "est") +
  scale_fill_viridis_c(labels = labels, 
     #trans = "sqrt",
    option = "C",
    #breaks = c(0,2,4,6,8), 
    limits = c(5, 11), 
    na.value = "black")+
  facet_wrap(~year) + 
  theme(legend.position = c(.85,.12)) +
  labs(fill = "°C")
  #ggtitle(" ")
pred_4
```

```{r}
plot_map_raster(p_nd, "est_non_rf") +
  ggtitle("Prediction (fixed effects only)") +
  facet_wrap(~year) +
  scale_fill_viridis_c()
```

```{r}
plot_map_raster(p_nd, "est_rf") +
  ggtitle("Prediction (all random effects only)") +
  facet_wrap(~year) +
  scale_fill_viridis_c()
```

```{r}
#glimpse(p_nd)

omega_s_all <- plot_map_raster(p_nd, "omega_s") +
  ggtitle("Spatial random effects only") +
    scale_fill_gradient2( low = "Steel Blue 4",
 high = "Red 3", limits = c(-1.5,1.5),
    na.value = "red") +  labs(fill = "") + 
gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = c(.82,.75))

omega_s_all
```

```{r}
 plot_map_raster(p_nd, "epsilon_st") +
  ggtitle("Spatiotemporal random effects only") +
  facet_wrap(~year) +   scale_fill_gradient2( low = "Steel Blue 4",
 high = "Red 3", limits = c(-1.5,1.5),
    na.value = "red")
```

```{r}
p_nd1 <- p_nd %>% filter(ssid!=16) %>% filter(ssid!=4)
st_1n3 <- plot_map_raster(p_nd1, "epsilon_st") +
  ggtitle("Spatiotemporal random effects") +
  facet_wrap(~year) +
   scale_fill_gradient2( low = "Steel Blue 4",
 high = "Red 3", limits = c(-1.5,1.5),
    na.value = "red") +
  labs(fill = "epsilon") + 
  gfplot::theme_pbs() + 
  theme(
  legend.position = "none", 
  axis.title.x = element_blank(), axis.title.y = element_blank())
st_1n3
```

```{r}
p_nd4 <- p_nd %>% filter(ssid==4) #%>% filter(ssid!=4)
st_4 <- plot_map_raster(p_nd4, "epsilon_st") +
  #ggtitle(" ") +
  facet_wrap(~year) +
     scale_fill_gradient2( low = "Steel Blue 4",
 high = "Red 3", limits = c(-1.5,1.5),
    na.value = "red") +
          labs(fill = " ") + 
gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = c(.85,.15))
st_4
```

```{r}
p_nd16 <- p_nd %>% filter(ssid==16) #%>% filter(ssid!=4)
st_16 <- plot_map_raster(p_nd16, "epsilon_st") +
  #ggtitle(" ") +
  facet_wrap(~year) +
   scale_fill_gradient2( low = "Steel Blue 4",
 high = "Red 3", limits = c(-1.5,1.5),
    na.value = "red") +
  xlim(65,345) +
    ylim(5800,6100) +
gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = "none")
st_16
```

```{r}
png(
  file = "figs/temp-predictions.png", 
  res = 600,
  units = "in",
  width = 11, 
  height = 10
)

gridExtra::grid.arrange(
  grobs = list(pred_1, st_1n3, omega_s_all, pred_4, st_4, st_16 ),
  widths = c(2, 2, 1.75),
  heights = c(2, 1.35),
  layout_matrix = rbind(c(1, 2, 3),
                        c(4, 5, 6)),
  top = grid::textGrob("Bottom temperature estimates from sdmTMB model")
)

# gridExtra::grid.arrange( nrow = 2, heights = c(2,1.5), 
#   top = grid::textGrob("Dissolved oxygen estimates from sdmTMB model")) 
dev.off()
```

```{r}
temp_cog <- get_cog(p_nd)

temp_cog %>%
  reshape2::dcast(year ~ coord, value.var = "est") %>%
  ggplot(aes(X, Y, colour = year)) + geom_path(arrow = arrow()) +
  scale_color_viridis_c()
```

```{r}
temp_cog %>%
  ggplot(aes(year, est, ymin = lwr, ymax = upr)) +
  geom_line() +
  geom_ribbon(alpha = 0.2) +
  facet_wrap(~coord, scales = "free_y")
```