### 2019-11-08 analysis/VOCC/3-bioclimatic-trends.Rmd
plot data from biomass model with climate variables

---
title: "`r paste(params$species)` biomass trends and figures"
author: "Philina English"
date: "`r Sys.Date()`"
output: html_document
params: 
    species: "Redbanded Rockfish"
    region: "All synoptic surveys"
    start_year: 2008
    bind_predictions: FALSE
    covs: "-log-both-noAR1" # covariate string used to name model rds 
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set( fig.width = 9, fig.height = 6.5, #fig.width=11, fig.height=8.5, 
                      #fig.path=paste0("figs/", spp, "/"),
                      echo=FALSE, warning=FALSE, message=FALSE)
library(dplyr)
library(ggplot2)
library(gfplot)
library(gfdata)
library(sdmTMB)
library(gfranges)
```

```{r params}
species <- params$species
region <- params$region
start_year <- params$start_year

bind_predictions <- params$bind_predictions
covs <- params$covs
priors <- FALSE

paste("region =", region)
paste("ssids modelled separately =", bind_predictions)
paste("model label =", covs)
paste("priors =", priors)

spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))

if (region == "Both odd year surveys") {
  survey <- c("SYN QCS", "SYN HS")
  model_ssid <- c(1, 3)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "West Coast Vancouver Island") {
  survey <- c("SYN WCVI")
  model_ssid <- c(4)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "West Coast Haida Gwaii") {
  survey <- c("SYN WCHG")
  model_ssid <- c(16)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "All synoptic surveys") {
  survey <- c("SYN QCS", "SYN HS", "SYN WCVI", "SYN WCHG")
  model_ssid <- c(1, 3, 4, 16)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}
```

Save predictions for full spatial grid
```{r}
#fish <- readRDS(paste0("raw/bio-data-", spp, ""))

rm(adult_predictions)
rm(imm_predictions)
rm(predictions)

maturity <- readRDS(paste0(
    "data/", spp, "/data-by-maturity-", spp, "-", ssid_string, ".rds"
    ))

if (any(names(maturity) == "adult_density")) {
#if (length(unique(fish$maturity_code)) > 2) {
  nd_all <- readRDS(paste0("data/predicted-DO-new.rds"))
  adult_biomass <- readRDS(paste0("data/", spp, 
        "/mod-mat-biomass-", spp,"-fixed", covs, "-", ssid_string, "-prior-", priors, ".rds"))
  #nd$year <- as.integer(nd$year)
  nd <- nd_all %>%
    filter(ssid %in% model_ssid) %>%
    filter(year %in% unique(adult_biomass$data$year)) %>% 
    mutate(
      do_mlpl_scaled = (log_do - adult_biomass$data$do_mlpl_mean[1]) /(adult_biomass$data$do_mlpl_sd[1]*2), 
      do_mlpl_scaled2 = do_mlpl_scaled^2)
  nd <- na.omit(nd)

  adult_predictions_obj <- predict(adult_biomass, newdata = nd, return_tmb_object = TRUE)
  adult_predictions <- adult_predictions_obj$data

  adult_predictions$est_rw_i <- adult_predictions_obj$obj$report()$proj_rw_i
  # test_eta <- adult_predictions_obj$obj$report()$proj_eta
  # test_fe <- adult_predictions_obj$obj$report()$proj_fe
  # 
  # test_diff <- test_eta - test_fe
  # range(adult_predictions$est_rw_i)
  # report <- adult_predictions_obj$obj$report()

  saveRDS(adult_predictions, file = paste0("data/", spp,
    "/predictions-", spp, covs, "-", ssid_string, "-mature-bioclimatic-prior-", priors, ".rds"))

  imm_biomass <- readRDS(paste0("data/", spp, 
        "/mod-imm-biomass-", spp,"-fixed", covs, "-", ssid_string, "-prior-", priors, ".rds"
  ))

    nd <- nd_all %>%
    filter(ssid %in% model_ssid) %>%
    filter(year %in% unique(imm_biomass$data$year))%>% 
    mutate(do_mlpl_scaled = (log_do - imm_biomass$data$do_mlpl_mean[1]) /(imm_biomass$data$do_mlpl_sd[1]*2), do_mlpl_scaled2 = do_mlpl_scaled^2)
  nd <- na.omit(nd)
  #nd$year <- as.numeric(nd$year)

  imm_predictions_obj <- predict(imm_biomass, newdata = nd, return_tmb_object = TRUE)
  imm_predictions <- imm_predictions_obj$data

  imm_predictions$est_rw_i <- imm_predictions_obj$obj$report()$proj_rw_i

  saveRDS(imm_predictions, file = paste0("data/", spp,
    "/predictions-", spp, covs, "-", ssid_string, "-imm-bioclimatic-prior-", priors, ".rds"))

} else {

    total_biomass <- readRDS(paste0("data/", spp, 
      "/model-total-biomass-", spp, "-fixed", covs, "-", ssid_string, "-prior-", priors, ".rds"))

    nd_all <- readRDS(paste0("data/predicted-DO-new.rds"))
    nd <- nd_all %>% 
      filter(ssid %in% model_ssid) %>%
      filter(year %in% unique(total_biomass$data$year))%>% 
    mutate(do_mlpl_scaled = (log_do - total_biomass$data$do_mlpl_mean) /(total_biomass$data$do_mlpl_sd*2), do_mlpl_scaled2 = do_mlpl_scaled^2)
    nd <- na.omit(nd)
    nd$year <- as.integer(nd$year)

    predictions <- predict(total_biomass, newdata = nd)
    saveRDS(predictions, file = paste0("data/", spp,
      "/predictions-", spp, covs, "-", ssid_string, "-total-bioclimatic-prior-", priors, ".rds"))
}
```

Load model predictions
```{r results = "hide"}
rm(adult_predictions)
rm(imm_predictions)
rm(predictions)

rm(pred1n3)
rm(pred4)
rm(pred16)

rm(pred1n3_imm)
rm(pred4_imm)
rm(pred16_imm)

rm(pred1n3_all)
rm(pred4_all)
rm(pred16_all)

if (region != "All synoptic surveys") {
  try(adult_predictions <- readRDS(paste0("data/", spp,
    "/predictions-", spp, covs, "-1n3n4n16-mature-bioclimatic-prior-", priors, ".rds"
  )) %>% filter(ssid %in% model_ssid))
  try(imm_predictions <- readRDS(paste0("data/", spp,
    "/predictions-", spp, covs, "-1n3n4n16-imm-bioclimatic-prior-", priors, ".rds"
  )) %>% filter(ssid %in% model_ssid))

  if (!exists("adult_predictions")) {
    paste("Maturity split biomass predictions are not available for this survey.")

    try(predictions <- readRDS(paste0("data/", spp,
      "/predictions-", spp, covs, "-1n3n4n16-total-bioclimatic-prior-", priors, ".rds"
    )) %>% filter(ssid %in% model_ssid))

    if (!exists("predictions")) {
      paste("Total biomass predictions for this survey are also not available.")
    }
  }
} else {

  try({adult_predictions <- readRDS(paste0("data/", spp,
    "/predictions-", spp, covs, "-", ssid_string, "-mature-bioclimatic-prior-", priors, ".rds"
    ))
  })
  try({imm_predictions <- readRDS(paste0("data/", spp,
    "/predictions-", spp, covs, "-", ssid_string, "-imm-bioclimatic-prior-", priors, ".rds"
    ))
  })

  # OR bind together predictions from different models
  if (bind_predictions) {
    try({pred1n3 <- readRDS(paste0("data/", spp,
      "/predictions-", spp, covs, "-1n3-mature-bioclimatic-prior-", priors, ".rds"
      ))
    })
    try({pred4 <- readRDS(paste0("data/", spp,
      "/predictions-", spp, covs, "-4-mature-bioclimatic-prior-", priors, ".rds"
      ))
    })
    try({pred16 <- readRDS(paste0("data/", spp,
      "/predictions-", spp, covs, "-16-mature-bioclimatic-prior-", priors, ".rds"
      ))
    })

    if (!exists("pred1n3")) pred1n3 <- NULL
    if (!exists("pred16")) pred16 <- NULL
    if (!exists("pred4")) pred4 <- NULL
    adult_predictions <- rbind(pred1n3, pred4, pred16)

    try({pred1n3_imm <- readRDS(paste0("data/", spp,
      "/predictions-", spp, covs, "-1n3-imm-bioclimatic-prior-", priors, ".rds"
      ))
    })
    try({pred4_imm <- readRDS(paste0("data/", spp,
      "/predictions-", spp, covs, "-4-imm-bioclimatic-prior-", priors, ".rds"))
    })
    try({pred16_imm <- readRDS(paste0("data/", spp,
        "/predictions-", spp, covs, "-16-imm-bioclimatic-prior-", priors, ".rds"))
    })

    if (!exists("pred1n3_imm")) pred1n3_imm <- NULL
    if (!exists("pred16_imm")) pred16_imm <- NULL
    if (!exists("pred4_imm")) pred4_imm <- NULL
    imm_predictions <- rbind(pred1n3_imm, pred4_imm, pred16_imm)
  }
  }

  if (!exists("adult_predictions")) {
      paste("Maturity split biomass predictions are not available.")

      try({predictions <- readRDS(paste0("data/", spp,
        "/predictions-", spp, covs, "-", ssid_string, "-total-bioclimatic-prior-", priors, ".rds"))
      })

      if (!exists("predictions")) {
        try({pred1n3_all <- readRDS(paste0("data/", spp,
          "/predictions-", spp, covs, "-1n3-total-bioclimatic-prior-", priors, ".rds"))
        })
        try({pred4_all <- readRDS(paste0("data/", spp,
          "/predictions-", spp, covs, "-4-total-bioclimatic-prior-", priors, ".rds"))
        })
        try({pred16_all <- readRDS(paste0("data/", spp,
          "/predictions-", spp, covs, "-16-total-bioclimatic-prior-", priors, ".rds"))
        })
        try({
          predictions <- rbind(pred1n3_all, pred4_all, pred16_all)
        })

        if (!exists("predictions")) paste("Total biomass predictions are not available.")
      }
    }
```

Caluclate variables of interest from predictions
```{r}
rm(max_raster)
rm(max_adult)
rm(max_bio)

if (exists("adult_predictions")) {
  # to convert from kg/m2 to kg/hectare multiply by 10000
  adult_predictions$est_exp <- exp(adult_predictions$est) * 10000
  imm_predictions$est_exp <- exp(imm_predictions$est) * 10000
  adult_predictions$bioclimatic <- exp(adult_predictions$est - adult_predictions$epsilon_st) * 10000
  imm_predictions$bioclimatic <- exp(imm_predictions$est - imm_predictions$epsilon_st) * 10000

  adult_predictions$total_bio <- imm_predictions$est_exp + adult_predictions$est_exp
  imm_predictions$prop_imm <- 
  imm_predictions$est_exp / (adult_predictions$est_exp + imm_predictions$est_exp)

  saveRDS(adult_predictions, file = paste0(
    "data/", spp,
    "/all-predictions-", spp, covs, "-mat-bioclimatic-prior-", priors, ".rds"
  ))
  saveRDS(imm_predictions, file = paste0(
    "data/", spp,
    "/all-predictions-", spp, covs, "-imm-bioclimatic-prior-", priors, ".rds"
  ))

  adult_predictions <- filter(adult_predictions, year >= start_year)
  imm_predictions <- filter(imm_predictions, year >= start_year)

  max_raster <- quantile(adult_predictions$est_exp, 0.999)
  max_adult <- signif(max(adult_predictions$est_exp), digits = 2)

  model_ssid <- unique(adult_predictions$ssid)
} else {
  predictions <- filter(predictions, year >= start_year)

  predictions$est_exp <- exp(predictions$est) * 10000
  predictions$bioclimatic <- exp(predictions$est - predictions$epsilon_st) * 10000

  predictions$total_bio <- exp(predictions$est) * 10000
  max_raster <- quantile(predictions$total_bio, .999)
  #max_bio <- round(max(predictions$total_bio))
  max_bio <- signif(max(predictions$est_exp), digits = 2)

  saveRDS(predictions, file = paste0(
    "data/", spp,
    "/all-predictions-", spp, covs, "-total-bioclimatic-prior-", priors, ".rds"
  ))

  model_ssid <- unique(predictions$ssid)
}
```

## Mature (or total) biomass predictions
```{r, error=TRUE}
if (exists("adult_predictions")) {
  p_adult_all <- plot_facet_map(adult_predictions, "est_exp",
    raster_limits = c(0, max_raster),
    legend_position = "none",
    transform_col = fourth_root_power
  ) +
    labs(fill = "kg/ha") +
    ggtitle(paste0("", species, " mature biomass \n(max = ", max_adult, " kg/ha)"))
  print(p_adult_all)
} else {
  p_adult_all <- plot_facet_map(predictions, "total_bio",
    transform_col = fourth_root_power
  ) +
    labs(fill = "kg/ha") +
    ggtitle(paste0("", species, " total biomass \n(max = ", max_bio, " kg/ha)"))

  print(p_adult_all)
}
```

### Split by survey (SSID)
```{r}
if (exists("adult_predictions")) predictions <- adult_predictions

  if (1 %in% model_ssid) {
    adult_predictions1 <- predictions %>% filter(ssid == 1)
  } else {
    adult_predictions1 <- NULL
  }

  if (3 %in% model_ssid) {
    adult_predictions3 <- predictions %>%
      filter(ssid != 16) %>%
      filter(ssid != 4)
  } else {
    adult_predictions3 <- NULL
  }

  if (4 %in% model_ssid) {
    adult_predictions4 <- predictions %>% filter(ssid == 4)
  } else {
    adult_predictions4 <- NULL
  }

  if (16 %in% model_ssid) {
    if (any(predictions$ssid == 16)) {
      adult_predictions16 <- predictions %>% filter(ssid == 16)
    } else {
      adult_predictions16 <- NULL
    }
  } else {
    adult_predictions16 <- NULL
  }

  ad_prediction_list <- list(adult_predictions1, adult_predictions3, adult_predictions4, adult_predictions16)
  ad_prediction_list <- ad_prediction_list[!sapply(ad_prediction_list, is.null)]

  p_adult <- list()
  if(exists("adult_predictions")) {
  for (i in seq_len(length(model_ssid))) {
    p_adult[[i]] <- plot_facet_map(ad_prediction_list[[i]], "est_exp",
      raster_limits = c(0, max_raster),
      # legend_position = "none"
      transform_col = fourth_root_power
    ) +
      labs(fill = "kg/ha") +
      ggtitle(paste0("", species, " mature biomass \n(max = ", max_adult, " kg/ha)"))
    p_adult[[i]]
  }
  } else {
    for (i in seq_len(length(model_ssid))) {
    p_adult[[i]] <- plot_facet_map(ad_prediction_list[[i]], "est_exp",
      raster_limits = c(0, max_raster),
      # legend_position = "none"
      transform_col = fourth_root_power
    ) +
      labs(fill = "kg/ha") +
      ggtitle(paste0("", species, " total biomass \n(max = ", max_bio, " kg/ha)"))
    p_adult[[i]]  
    }
  }
  print(p_adult)
```

```{r, message=FALSE, warning=FALSE}
rm(out_mature)
rm(out_total)
# try(out_mature <- readRDS(paste0("data/", spp, "/trends-", spp, covs, "-", ssid_string, "-mature-bioclimatic-priors-", priors, ".rds")))

if (!exists("out_mature")) {
  if (exists("adult_predictions")) {
    predictions <- adult_predictions
  } else {
   try(out_total <- readRDS(paste0("data/", spp, "/trends-", spp, covs, "-", ssid_string, "-total-bioclimatic-priors-", priors, ".rds")))
  }

  if (!exists("out_total")) {
    if (1 %in% model_ssid) {

      .predictions <- filter(predictions, ssid == 1)
      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)

      out_mature1 <- make_trend_data(predictions,
        ssid = c(1),
        start_time = start_year,
        skip_time = 2004,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = indices,
        variable_names = "est_exp"
      )
    } else {
      out_mature1 <- NULL
    }

    if (3 %in% model_ssid) {
      .predictions <- filter(predictions, ssid == 3)
      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)

      out_mature3 <- make_trend_data(predictions,
        ssid = c(3),
        start_time = start_year,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = indices,
        variable_names = "est_exp"
      )
    } else {
      out_mature3 <- NULL
    }

    if (4 %in% model_ssid) {

      .predictions <- filter(predictions, ssid == 4)
      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)

      out_mature4 <- make_trend_data(predictions,
        ssid = c(4),
        start_time = start_year,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = indices,
        variable_names = "est_exp"
      )
    } else {
      out_mature4 <- NULL
    }

    if (16 %in% model_ssid) {

      .predictions <- filter(predictions, ssid == 16)
      #unique(.predictions$year)
      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)

      out_mature16 <- make_trend_data(predictions,
        ssid = c(16),
        start_time = start_year,
        skip_time = 2007,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = indices,
        variable_names = "est_exp"
      )
    } else {
      out_mature16 <- NULL
    }

    out_mature <- rbind(out_mature1, out_mature3, out_mature4, out_mature16)

    if (exists("adult_predictions")) {
            saveRDS(out_mature, file = paste0("data/", spp, "/trends-", 
        spp, covs, "-", ssid_string, "-mature-bioclimatic-priors-", priors, ".rds"))

         } else {
      out_total <- out_mature
      rm(out_mature)
      saveRDS(out_total, file = paste0("data/", spp, "/trends-", 
        spp, covs, "-", ssid_string, "-total-bioclimatic-priors-", priors, ".rds"))
    }
  }
}
```

### Recent biomass with trawl footprint
```{r }
if (!exists("out_mature")) {out_recent <- out_total} else {out_recent <- out_mature}

  trawl_footprint <- sf::st_read(dsn = "data/trawl-footprint", layer = "Trawl_footprint")
  proj.to <- "+proj=utm +zone=9 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs"
  trawl_footprint <- sf::st_transform(trawl_footprint, crs = proj.to)

  current_biomass <- plot_vocc(out_recent,
    fill_col = "est_exp_e",
    fill_label = "Mean\nbiomass (kg/ha)\npost-2010",
    raster_alpha = 1,
    viridis_option = "C",
    vec_aes = NULL,
    transform_col = fourth_root_power
  )
  current_biomass <- current_biomass +
    geom_sf(data = trawl_footprint, aes(geometry = geometry / 1000), colour = "gray", fill = "gray", alpha = 0.2, lwd = 0.25, inherit.aes = FALSE) +
    coord_sf(xlim = c(180, 820), ylim = c(5350, 6050), datum = NULL) +
    xlab("") +
    ylab("") +
    theme(panel.grid.major = element_line(colour = "transparent")) +
    ggtitle(paste("Predicted", species, "density with trawl footprint overlain"))
  current_biomass

```

## Trends in biomass density
```{r, error=TRUE}
if (exists("out_mature")) {
  data <- out_mature
  biotrend_ad <- plot_vocc(data,
    fill_col = "units_per_decade",
    fill_label = "Mature \nbiomass \ntrend \n(kg/ha/decade)",
    raster_alpha = 1,
    vec_aes = NULL,
    white_zero = TRUE,
    high_fill = "deepskyblue3", # "lightseagreen", # limegreen",
    low_fill = "red", # "deeppink3",# "steel blue 4",
    transform_col = fourth_root_power,
    #raster_limits = c(0, quantile(data$units_per_decade, 0.99999)),
    legend_position = c(0.2, 0.25)
  ) + ggtitle(paste0(
    "", species, " ", min(unique(as.numeric(adult_predictions[["year"]]))), "-", max(unique(as.numeric(adult_predictions[["year"]])))
  ))
  biotrend_ad
} else {
  biotrend_ad <- plot_vocc(out_total,
    fill_col = "units_per_decade",
    fill_label = "Biomass \ntrend \n(kg/ha/decade)",
    raster_alpha = 1,
    vec_aes = NULL,
    white_zero = TRUE,
    high_fill = "deepskyblue3", # "greenyellow", # limegreen",
    low_fill = "red", # "steel blue 4",
    transform_col = fourth_root_power,
    legend_position = c(0.2, 0.25)
  ) + ggtitle(paste0(
    "", species, " ", min(unique(as.numeric(predictions[["year"]]))), "-", max(unique(as.numeric(predictions[["year"]])))
  ))
  biotrend_ad
}
```

# Immature biomass density
### if available
```{r, error=TRUE}
if (exists("imm_predictions")) {
  p_imm_all <- plot_facet_map(imm_predictions, "est_exp",
    transform_col = fourth_root_power,
    raster_limits = c(0, max_raster)
  ) +
    labs(fill = "kg/ha") +
    ggtitle(paste0("", species, " immature biomass \n "))

  p_imm_all
} else {
  p_imm_all <- grid::grid.rect(gp = grid::gpar(col = "white"))
}
```

```{r}
if (exists("adult_predictions")) {
  if (1 %in% model_ssid) {
    imm_predictions1 <- imm_predictions %>% filter(ssid == 1)
  } else {
    imm_predictions1 <- NULL
  }

  if (3 %in% model_ssid) {
    imm_predictions3 <- imm_predictions %>%
      filter(ssid != 16) %>%
      filter(ssid != 4)
  } else {
    imm_predictions3 <- NULL
  }

  if (4 %in% model_ssid) {
    imm_predictions4 <- imm_predictions %>% filter(ssid == 4)
  } else {
    imm_predictions4 <- NULL
  }

  if (16 %in% model_ssid) {
    imm_predictions16 <- imm_predictions %>% filter(ssid == 16)
  } else {
    imm_predictions16 <- NULL
  }

  im_prediction_list <- list(imm_predictions1, imm_predictions3, imm_predictions4, imm_predictions16)
  im_prediction_list <- im_prediction_list[!sapply(im_prediction_list, is.null)]

  p_imm <- list()
  for (i in seq_len(length(model_ssid))) {
    p_imm[[i]] <- plot_facet_map(im_prediction_list[[i]], "est_exp",
      raster_limits = c(0, max_raster),
      # legend_position = "none",
      transform_col = fourth_root_power
    ) +
      labs(fill = "kg/ha") +
      ggtitle(paste0("", species, " immature biomass \n(max = ", max_adult, " kg/ha)"))

    p_imm[[i]]
  }
  print(p_imm)
}
```

```{r, message=FALSE, warning=FALSE}
# try(out_imm <- readRDS(paste0("data/", spp, "/trends-", spp, covs, "-", ssid_string, "-immature-bioclimatic-priors-", priors, ".rds")))
rm(out_imm)
if (!exists("out_imm")) {
  if (exists("imm_predictions")) {
    predictions <- imm_predictions

    if (1 %in% model_ssid) {

      .predictions <- filter(predictions, ssid == 1)
      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)
      out_im1 <- make_trend_data(predictions,
        ssid = c(1),
        start_time = start_year,
        skip_time = 2004,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = indices,
        variable_names = "est_exp"
      )
    } else {
      out_im1 <- NULL
    }

    if (3 %in% model_ssid) {
      .predictions <- filter(predictions, ssid == 3)
      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)

      out_im3 <- make_trend_data(predictions,
        ssid = c(3),
        start_time = start_year,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = indices,
        variable_names = "est_exp"
      )
    } else {
      out_im3 <- NULL
    }

    if (4 %in% model_ssid) {
      .predictions <- filter(predictions, ssid == 4)
      unique(.predictions$year)
      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)

      out_im4 <- make_trend_data(predictions,
        ssid = c(4),
        start_time = start_year,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = indices,
        variable_names = "est_exp"
      )
    } else {
      out_im4 <- NULL
    }

    if (16 %in% model_ssid) {
      .predictions <- filter(predictions, ssid == 16)
      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)

      out_im16 <- make_trend_data(predictions,
        ssid = c(16),
        start_time = start_year,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = indices,
        variable_names = "est_exp"
      )
    } else {
      out_im16 <- NULL
    }

    out_imm <- rbind(out_im1, out_im3, out_im4, out_im16)
    saveRDS(out_imm, file = paste0(
      "data/", spp, "/trends-", spp, covs, "-", ssid_string, "-immature-bioclimatic-priors-", priors, ".rds"
    ))
  } else {
    rm(out_imm)
  }
}
```

```{r}
if (exists("out_imm")) {
  data <- out_imm
  biotrend_im <- plot_vocc(data,
    fill_col = "units_per_decade",
    fill_label = "Immature \nbiomass \ntrend \n(kg/ha/decade)",
    raster_alpha = 1,
    vec_aes = NULL,
    white_zero = TRUE,
    high_fill = "deepskyblue3", # "greenyellow", # limegreen",
    low_fill = "red", # "steel blue 4",
    transform_col = fourth_root_power,
    legend_position = c(0.2, 0.25)
  ) + ggtitle(paste0(
    "", species, " ", min(unique(imm_predictions[["year"]])), "-", max(unique(imm_predictions[["year"]]))
  ))
  biotrend_im
} else {
  biotrend_im <- grid::grid.rect(gp = grid::gpar(col = "white"))
}
```

```{r, error=TRUE}
if (exists("imm_predictions")) {
  p_prop_imm <- plot_facet_map(imm_predictions, "prop_imm") +
    labs(fill = "% \nby mass") +
    ggtitle(paste0("", species, " proportion immature"))
  print(p_prop_imm)
} else {
  p_prop_imm <- grid::grid.rect(gp = grid::gpar(col = "white"))
}
```

Save trend plots
```{r results = "hide"}
png(
  file = paste0("figs/", spp, "/trends-", spp, covs, "-", ssid_string, "-priors-", priors, ".png"),
  res = 600,
  units = "in",
  width = 11,
  height = 8.5
)
gridExtra::grid.arrange(
  # p_adult, p_imm,
  biotrend_ad, biotrend_im,
  nrow = 1
)
dev.off()
```

Save spatiotemporal biomass plots
```{r results = "hide"}
if(exists("adult_predictions")) { biomass_plots <- c(p_adult, p_imm) } else {
  biomass_plots <- c(p_adult)
}

png(
  file = paste0("figs/", spp, "/bioclimatic-biomass-", spp, covs, "-", ssid_string, "-priors-", priors, ".png"),
  res = 600,
  units = "in",
  width = 16,
  height = 8.5
)
gridExtra::grid.arrange(
  grobs = biomass_plots,
  nrow = 2
)
dev.off()
```

Save plot medley
```{r results = "hide"}
png(
  file = paste0("figs/", spp, "/bioclimatic-exploratory-", spp, covs, "-", ssid_string, "-priors-", priors, ".png"),
  res = 600,
  units = "in",
  width = 9.5,
  height = 12.5
)

gridExtra::grid.arrange(
  current_biomass,
  p_prop_imm,
  p_adult_all, p_imm_all,
  biotrend_ad, biotrend_im,
  nrow = 3
)
dev.off()
```

```{r}
# plot_map_raster <- function(dat, column = "est") {
#     ggplot(dat, aes_string("X", "Y", fill = column)) +
#     geom_raster() +
#     coord_fixed()+
# gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank())
# }

 # plot_map_raster(adult_predictions, "epsilon_st") +
 #  ggtitle("Spatiotemporal random effects only") +
 #  facet_wrap(~year) +
 #  scale_fill_gradient2( high = "Steel Blue 4",
 #  low = "Red 3") + coord_fixed()

 if (exists("adult_predictions")) {
  # p_st <- list()
  # for (i in seq_len(length(model_ssid))) {
  #   p_st[[i]] <- plot_map_raster(ad_prediction_list[[i]], "epsilon_st") +
  #     ggtitle("Spatiotemporal random effects only") +
  #     facet_wrap(~year) +
  #     scale_fill_gradient2(high = "#66C2A5", #mid = "black", 
  #       low = "#5E4FA2"#, trans = fourth_root_power
  #     )
  #   p_st[[i]]
  #  }
  # print(p_st)

  p_adult_st <- list()
  for (i in seq_len(length(model_ssid))) {
    p_adult_st[[i]] <-plot_facet_map(ad_prediction_list[[i]], "epsilon_st",
    white_zero = TRUE,
    high_fill = "#66C2A5",
    low_fill = "#5E4FA2"
    #raster_limits = c(0, max_raster),
    #legend_position = "none",
    #transform_col = fourth_root_power
  ) +
    #labs(fill = "kg/ha") +
    ggtitle(paste0("", species, " mature spatiotemporal variation"))
    p_adult_st[[i]]
  }
  print(p_adult_st)
#} else {
  # p_adult_st <- plot_facet_map(predictions, "epsilon_st",
  #   transform_col = fourth_root_power
  # ) +
  #   #labs(fill = "kg/ha") +
  #   ggtitle(paste0("", species, "total spatiotemporal variation"))
  # 
  # print(p_adult_st)
}
```

```{r}
adult_predictions$year.f <- as.factor(adult_predictions$year)

glimpse(adult_predictions)

data <- mutate(adult_predictions, cell_id = paste0(X,Y)) %>% 
  group_by(cell_id) %>% 
  mutate(cell_max = max(bioclimatic)) %>% group_by(year.f) %>% mutate(ann_biomass = sum(est_exp))
unique(data$ann_biomass)
 glimpse(data)

# start_biomass <- filter(adult_predictions, year == min(as.numeric(adult_predictions$year))) 
#     bio5perc <- sum(adult_predictions$start_density)*0.05
#     s <- sort(annual_biomass$start_density)
#     bio_sum <- cumsum(s)
#     lower_density_threshold <- s[which(bio_sum>bio5perc)[1]]

threshold <- quantile(data$cell_max, 0.2)
library(forcats)
violin_data <- data %>% filter (cell_max > threshold) %>% mutate(year = fct_reorder(year.f, desc(ann_biomass)))
 glimpse(violin_data)

ggplot(violin_data) + 
  geom_hline(aes(yintercept=0)) +
  geom_violin(aes(fct_reorder(year.f, (ann_biomass)), epsilon_st, fill = year.f)) +   
  geom_text(aes(fct_reorder(year.f, (ann_biomass)), y=2.8, label = round(ann_biomass)), check_overlap = TRUE) +
  scale_fill_viridis_d(option = "C") +
  ylab("Spatiotemporal variation") +
  xlab("Year") +
  #coord_flip() +
  theme(legend.position="none") +
  ggtitle(paste("Local", species, "abundance deviations not explained by climate and overall abundance"))
```

```{r}
# data <- adult_predictions %>% group_by (year) %>% 
#   mutate(mean_temp = mean(temp)) %>% 
#   filter (est_exp > quantile(est_exp, 0.05)) %>% 
#   filter (do_est > 1.5)
# 
# ggplot(data, aes(temp, epsilon_st)) + 
#   geom_point(alpha = 0.025) + 
#   geom_smooth(method="lm", formula = y~poly(x,2)) +
#   facet_wrap(~year) +
#   scale_color_viridis_d(option = "C") +
#   guides(color = FALSE) +
#   ylab("Spatiotemporal variation") +
#   ggtitle(paste("Local", species, "abundance deviations not explained by climate and overall abundance"))

```

Save spatiotemporal residual plots
```{r results = "hide"}
# if(exists("adult_predictions")) { resid_plots <- c(p_st, p_imm) } else {
#   biomass_plots <- c(p_adult)
# }

png(
  file = paste0("figs/", spp, "/bioclimatic-residuals-", spp, covs, "-", ssid_string, "-priors-", priors, ".png"),
  res = 600,
  units = "in",
  width = 16,
  height = 8.5
)
gridExtra::grid.arrange(
  grobs = p_adult_st,
  nrow = 2
)
dev.off()
```

```{r}
 if (exists("adult_predictions")) {
  p_bioclim <- list()
  for (i in seq_len(length(model_ssid))) {
    p_bioclim[[i]] <- plot_facet_map(ad_prediction_list[[i]], "bioclimatic",
      raster_limits = c(0, max_raster),
      # legend_position = "none",
      transform_col = fourth_root_power
    ) +
      labs(fill = "kg/ha") +
      ggtitle(paste0("", species, " predicted biomass based on climate \nand fixed spatial variation"))
    p_bioclim[[i]]
  }
 }
  print(p_bioclim)

```

Save bioclimatic predictions plots
```{r results = "hide"}
# if(exists("adult_predictions")) { resid_plots <- c(p_st, p_imm) } else {
#   biomass_plots <- c(p_adult)
# }

png(
  file = paste0("figs/", spp, "/bioclimatic-predictions-", spp, covs, "-", ssid_string, "-priors-", priors, ".png"),
  res = 600,
  units = "in",
  width = 16,
  height = 8.5
)
gridExtra::grid.arrange(
  grobs = p_bioclim,
  nrow = 2
)
dev.off()
```

```{r}
if (exists("imm_predictions")) {
  p_prop_imm <- plot_facet_map(imm_predictions, "prop_imm") +
    labs(fill = "% \nby mass") +
    ggtitle(paste0("", species, " proportion immature"))
  print(p_prop_imm)
} else {
  p_prop_imm <- grid::grid.rect(gp = grid::gpar(col = "white"))
}
```

```{r}
# Velocity of climate change ---------------------------------------------

d <- ad_prediction_list[[2]]

d$layer <- d$bioclimatic

# Just to confirm the trends we're seeing are real,
# mod = mgcv::gam(est ~ s(year, k=5) + te(X,Y) + log(depth),
#   data=d)
# plot(mod)

# This parameter controls how the original 2-km projection is aggregated.
# for example, a value of 5 means that the raster would be reprojected to 10km grid
scale_fac <- 1

# create a RasterBrick
# raster for each year
rlist <- list()
for (i in 1:length(unique(d$year))) {
  rlist[[i]] <- raster::rasterFromXYZ(dplyr::filter(d, year == unique(d$year)[i]) %>%
    dplyr::select(X, Y, layer))
  rlist[[i]] <- raster::aggregate(rlist[[i]], fact = scale_fac)
}
# stack rasters into layers -> rasterbrick
rstack <- raster::stack(rlist[[1]], rlist[[2]])
for (i in 3:length(rlist)) {
  rstack <- raster::stack(rstack, rlist[[i]])
}
rbrick <- raster::brick(rstack)

# Then calculate the trend per pixel:
slopedat <- vocc::calcslope(rbrick)
# Then get the mean temperature for a time period and calculate the spatial gradient:
allyears <- rep(1, raster::nlayers(rbrick))
mnsst <- raster::stackApply(rbrick, indices = allyears, fun = mean)
spatx <- vocc::spatialgrad(mnsst)
# Now we can calculate the VoCC:
velodf <- vocc::calcvelocity(spatx, slopedat, y_dist = 1)

# Mapping it again is straightforward:
rtrend <- rgrad_lon <- rgrad_lat <- rvocc <- angle <- magn <- raster::raster(rbrick)
rgrad_lat[spatx$icell] <- spatx$NS # latitude shift, NS
rgrad_lon[spatx$icell] <- spatx$WE # longitude shift, WE
rtrend[slopedat$icell] <- -1 * slopedat$slope
rvocc[velodf$icell] <- velodf$velocity

# convert to data frames for ggplot
rtrend_df <- as.data.frame(raster::rasterToPoints(rtrend)) %>%
  dplyr::rename(trend = layer)
rgradlat_df <- as.data.frame(raster::rasterToPoints(rgrad_lat)) %>%
  dplyr::rename(gradNS = layer)
rgradlon_df <- as.data.frame(raster::rasterToPoints(rgrad_lon)) %>%
  dplyr::rename(gradWE = layer)
rvocc_df <- as.data.frame(raster::rasterToPoints(rvocc)) %>%
  dplyr::rename(velocity = layer)

# create ggquiver plots. need dataframe of lon, lat, delta_lon, delta_lat, trend, velocity
df <- dplyr::left_join(rtrend_df, rgradlat_df, by = c("x", "y")) %>%
  dplyr::left_join(rgradlon_df, by = c("x", "y")) %>%
  dplyr::left_join(rvocc_df, by = c("x", "y"))

gtrend <- ggplot(df, aes(x, y, fill = -trend)) +
  geom_raster() +
  scale_fill_gradient2(low = scales::muted("blue"), high = scales::muted("red")) +
  xlab("Longitude") + ylab("Latitude") + ggtitle("Trend") + coord_fixed()

# spatial gradient plot
quantile_cutoff <- 0.05 # for plotting
df <- dplyr::mutate(df,
  u_velo = trend / gradWE,
  v_velo = trend / gradNS,
  ulow = quantile(u_velo, quantile_cutoff),
  uhi = quantile(u_velo, 1 - quantile_cutoff),
  vlow = quantile(v_velo, quantile_cutoff),
  vhi = quantile(v_velo, 1 - quantile_cutoff),
  u_velo = ifelse(u_velo < ulow, ulow, u_velo),
  u_velo = ifelse(u_velo > uhi, uhi, u_velo),
  v_velo = ifelse(v_velo < vlow, vlow, v_velo),
  v_velo = ifelse(v_velo > vhi, vhi, v_velo)
) %>%
  dplyr::select(-ulow, -uhi, -vlow, -vhi)

# gradient plot with ggquiver
ggrad <- ggplot(df) +
  ggquiver::geom_quiver(aes(x, y, u = gradNS, v = gradWE)) +
  xlab("Longitude") + ylab("Latitude") + ggtitle("Gradient") + coord_fixed()

# velocity plot
gvocc <- ggplot(mutate(df, `Trend\n(kg/decade)` = -trend)) +
  ggquiver::geom_quiver(aes(x, y, u = u_velo, v = v_velo,
    colour = `Trend\n(kg/decade)`),
    vecsize = 2) +
  scale_colour_gradient2(low = scales::muted("blue"), high = scales::muted("red")) +
  xlab("UTM") + ylab("UTM") + #ggtitle("Velocity") +
  gfplot::theme_pbs()

# coast <- gfplot:::load_coastline(range(surv$longitude) + c(-1, 1),
#   range(surv$latitude) + c(-1, 1),
#   utm_zone = 9
# )

# isobath <- gfplot:::load_isobath(range(surv$longitude) + c(-5, 5),
#   range(surv$latitude) + c(-5, 5),
#   bath = c(100, 200, 300, 500), utm_zone = 9
# )

# library(ggnewscale)

gvocc <- gvocc +
  labs(colour = "Local\nclimate trend\n(°C/decade)") +
  ggnewscale::new_scale_color() +
# geom_path(
#   data = isobath, aes_string(
#     x = "X", y = "Y",
#     group = "paste(PID, SID)", colour = "PID"
#   ),
#   inherit.aes = FALSE, lwd = 0.4, alpha = 0.4
# ) +
  scale_colour_continuous(low = "grey70", high = "grey10") +
  # labs(colour = "Depth") +
  guides(colour = FALSE) +
  coord_fixed(xlim = range(df$x) + c(-3, 3), ylim = range(df$y) + c(-3, 3))

# gvocc <- gvocc + geom_polygon(
#   data = coast, aes_string(x = "X", y = "Y", group = "PID"),
#   fill = "grey87", col = "grey70", lwd = 0.2
# )

```