### 2019-11-28 analysis/VOCC/3-climate-models-fixed-depth.Rmd
model for bioclimatic estimates without fixed spatial field, use AR1, and fixed response to depth

---
title: "Spatiotemporal models of `r paste(params$species)` biomass with climate covariates"
author: "Philina English"
date: "`r Sys.Date()`"
output: html_document
params: 
    species: "Redbanded Rockfish"
    region: "All synoptic surveys"
    #region: "Both odd year surveys"
    log_temp: TRUE # should temperature be log()
    model_w_do: TRUE
    priors: FALSE
    knots: 500
    covariates: "+trawled" # additional non-climate variables #+as.factor(ssid)
    covs: "-AR1-spatial" #
    #covs: "-log-both-fixed-depth" #"-log-both-noAR1" # string describing model covariates
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.width=11, fig.height=8.5, 
                     # fig.path=paste0("figs/", spp, "/"),
                      echo=FALSE, warning=FALSE, message=FALSE)
library(dplyr)
library(ggplot2)
library(gfplot)
library(gfdata)
library(sdmTMB)
library(gfranges)
```

```{r params}
species <- params$species
region <- params$region
covariates <- params$covariates 
covs <- params$covs
log_temp <- params$log_temp
model_w_do <- params$model_w_do
priors <- params$priors
knots <- params$knots

paste("region =", region)
paste("model covariates =", covariates)
paste("model label =", covs)
paste("log(temperature) =", log_temp)
paste("model_w_do =", model_w_do)
paste("priors =", priors)
paste("knots =", knots)

spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))

# folder to hold figs for this species
dir.create(file.path("figs", spp))
dir.create(file.path("data", spp))

if (region == "Both odd year surveys") {
  survey <- c("SYN QCS", "SYN HS")
  model_ssid <- c(1, 3)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "West Coast Vancouver Island") {
  survey <- c("SYN WCVI")
  model_ssid <- c(4)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "West Coast Haida Gwaii") {
  survey <- c("SYN WCHG")
  model_ssid <- c(16)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "All synoptic surveys") {
  survey <- c("SYN QCS", "SYN HS", "SYN WCVI", "SYN WCHG")
  model_ssid <- c(1, 3, 4, 16)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}
```

Combine biomass and sensor data
```{r}
  biomass <- readRDS(paste0(
    "data/", spp, "/data-by-maturity-", spp, "-", ssid_string, ".rds"
    ))
  covars <- readRDS("data/event-covariates.rds")
  data <- dplyr::left_join(biomass, covars) %>% mutate(depth_akima = depth) %>% select(-depth)

  CTD <- readRDS("../tmb-sensor-explore/data/all-sensor-data-processed.rds")

  min(CTD$temperature_c, na.rm = T) #  
  mintemp <- 2.6

  data <- dplyr::left_join(data, CTD)

  # scale predictors before filtering to ensure mean and SD are global
  data <- data %>% mutate(raw_depth = depth_bath, depth = log(raw_depth))

  if (log_temp) {data <- mutate(data, temp = log(temperature_c))
  } else {
    data <- mutate(data, temp = temperature_c)
  }

if (model_w_do) {
  data <- filter(data, !(year == 2016 & ssid == 4) ) # 2016 do an outlier for survey 4
  data <- data %>% mutate(exclude = if_else(do_mlpl>8, 1, 0)) %>% 
      filter(exclude != 1) %>% # added on july 24 after first round of models run...
      filter(year > 2007) # 2007 DO data is flawed

    # scale predictors before filtering to ensure mean and SD are global
  data <- data %>% mutate(raw_do = do_mlpl, do_mlpl = log(do_mlpl))

  do_data <- scale_predictors(data, predictors = c(quo(depth), quo(temp), quo(do_mlpl))) 
       # c(quo(log_depth), quo(mixed), quo(muddy), quo(sandy), quo(rocky), quo(any_rock)))

  data <- do_data %>% mutate(depth = raw_depth) %>% #filter(year != 2016) %>%
    filter(ssid %in% model_ssid) %>% 
    filter(!is.na(depth_scaled)) %>% filter(!is.na(do_mlpl)) %>% filter(!is.na(temp))

  if (any(names(data) == "adult_density")) { 
  data <- data %>% select(X, Y, X10, Y10,
      adult_density, imm_density,
      depth_scaled, depth_scaled2, depth_mean, depth_sd, 
      do_mlpl_scaled, do_mlpl_scaled2,  do_mlpl_mean, do_mlpl_sd, 
      temp_scaled, temp_scaled2, temp_mean, temp_sd, 
      trawled, any_rock, muddy, sandy, year, ssid, fishing_event_id, 
      depth, do_mlpl, raw_do, temp, temperature_c) 
 } else {
    data <- data %>% select(X, Y, X10, Y10,
      density,
      depth_scaled, depth_scaled2, depth_mean, depth_sd, 
      do_mlpl_scaled, do_mlpl_scaled2,  do_mlpl_mean, do_mlpl_sd, 
      temp_scaled, temp_scaled2, temp_mean, temp_sd, 
      trawled, any_rock, muddy, sandy, year, ssid, fishing_event_id, 
      depth, do_mlpl, raw_do, temp, temperature_c) 
 }
} else {
  data <- scale_predictors(data, predictors = c(quo(depth), quo(temp))) 
       # c(quo(log_depth), quo(mixed), quo(muddy), quo(sandy), quo(rocky), quo(any_rock)))

  data <- data %>% mutate(depth = raw_depth) %>% #filter(year != 2016) %>%
    filter(ssid %in% model_ssid) %>% 
    filter(!is.na(depth_scaled)) %>% filter(!is.na(temp))

if (any(names(data) == "adult_density")) { 
  data <- data %>% select(X, Y, X10, Y10,
      adult_density, imm_density,
      depth_scaled, depth_scaled2, depth_mean, depth_sd, 
      temp_scaled, temp_scaled2, temp_mean, temp_sd, 
      trawled, any_rock, muddy, sandy, year, ssid, fishing_event_id, 
      depth, temp, temperature_c) 
 } else {
    data <- data %>% select(X, Y, X10, Y10,
      density,
      depth_scaled, depth_scaled2, depth_mean, depth_sd, 
      temp_scaled, temp_scaled2, temp_mean, temp_sd, 
      trawled, any_rock, muddy, sandy, year, ssid, fishing_event_id, 
      depth, temp, temperature_c) 
 }
}

# d <- data %>% filter(year==2011) #%>% filter(depth> 400) 
# = plot((adult_density)~(depth), data=d)

```

```{r}
  if (region == "Both odd year surveys") {
    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 400)
  }

  if (region == "West Coast Vancouver Island") {
    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 200)
  }

  if (region == "West Coast Haida Gwaii") {
    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 200)
  }

  if (region == "All synoptic surveys") {
    spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = knots)
  }

  sdmTMB::plot_spde(spde)
```

Run sdmTMB model
```{r}
tictoc::tic()
  if (any(names(data) == "adult_density")) {

    if (model_w_do) {
    adult_formula <- as.formula(paste(
      "adult_density ~ 0 + as.factor(year) + do_mlpl_scaled + do_mlpl_scaled2 + temp_scaled + temp_scaled2 + depth_scaled + depth_scaled2", covariates, ""
      ))
    } else {
      adult_formula <- as.formula(paste(
      "adult_density ~ 0 + as.factor(year) + temp_scaled + temp_scaled2 + depth_scaled + depth_scaled2", covariates, ""
      ))
    }

    adult_biomass_d <- sdmTMB(data = data,
      adult_formula,
      #time_varying = ~ trawled, #0 + depth_scaled + depth_scaled2, 
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = TRUE, 
      include_spatial = FALSE, # default = TRUE
      reml = TRUE,
      enable_priors = priors,
      #control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )

    saveRDS(adult_biomass_d, 
      file = paste0("data/", spp, 
        "/mod-mat-biomass-", spp,"-fixed", covs, "-", ssid_string, "-prior-", priors, ".rds"
        ))

    if (model_w_do) {
    imm_formula <- as.formula(paste(
      "imm_density ~ 0 + as.factor(year) + do_mlpl_scaled + do_mlpl_scaled2 + temp_scaled + temp_scaled2 + depth_scaled + depth_scaled2", covariates, ""
      ))
    } else {
         imm_formula <- as.formula(paste(
      "imm_density ~ 0 + as.factor(year) + temp_scaled + temp_scaled2 + depth_scaled + depth_scaled2", covariates, ""
      ))
    }

    imm_biomass_d <- sdmTMB::sdmTMB(data = data, 
      imm_formula,
      #time_varying = ~ trawled, # 0 + depth_scaled + depth_scaled2, 
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      ar1_fields = TRUE,
      include_spatial = FALSE, # default = TRUE
      reml = TRUE,
      enable_priors = priors,
      #control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )

    saveRDS(imm_biomass_d, 
      file = paste0("data/", spp, 
        "/mod-imm-biomass-", spp,"-fixed", covs, "-", ssid_string, "-prior-", priors, ".rds"
        ))

  } else {

    dens_formula <- as.formula(paste("density ~ 0 + as.factor(year) + do_mlpl_scaled + do_mlpl_scaled2 + temp_scaled + temp_scaled2", covariates, ""))

    total_biomass_d <- sdmTMB::sdmTMB(data = data,
      dens_formula,
      time_varying = ~ 0 + depth_scaled + depth_scaled2, 
      time = "year", spde = spde,
      family = tweedie(link = "log"),
      #ar1_fields = TRUE,
      include_spatial = FALSE, # default = TRUE
      enable_priors = priors,
      control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
      silent = FALSE
    )

    saveRDS(total_biomass_d, file = paste0("data/", spp, 
      "/model-total-biomass-", spp, "-fixed", covs, "-", ssid_string, "-prior-", priors, ".rds"
      ))
  }
tictoc::toc()
```