### 2020-01-01 analysis/VOCC/4-vocc-biotic.Rmd
batch calculation of biotic variability and  gradient velocities

---
title: "`r paste(params$species)` climatic and biotic velocities for `r paste(params$region)` (immature: `r paste(params$immature)`"
author: "Philina English"
date: "`r Sys.Date()`"
output: html_document 
params: 
    species: "Arrowtooth Flounder" 
    immature: FALSE
    region: "both odd year surveys"
    #region: "West Coast Vancouver Island"
    covs: "-tv-depth-only"
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.keep = "all", fig.show = "asis", dpi = 600,
  fig.width = 6, fig.height = 5,
  echo = FALSE, warning = FALSE, message = FALSE
)
options(scipen = 999)
options(digits = 4)
```

```{r pkgs, cache=FALSE, warning=FALSE, message=FALSE}
library("gfranges")
library("dplyr")
library("ggplot2")

ggplot2::theme_set(gfplot::theme_pbs())
```

#### Set parameters
```{r }
species <- params$species
region <- params$region
covs <- params$covs
immature <- params$immature

spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))

paste("species =", species)
paste("region =", region)
paste("covs =", covs)
```

#### Set cell sizes and scale
```{r, echo=TRUE}
dist_intercept <- 0 # input_cell_size / 2
input_cell_size <- 2
scale_fac <- 2 # means that the raster is reprojected to _ X original grid (2 km)
```

```{r, message=FALSE, warning=FALSE, include=FALSE}
if (region == "both odd year surveys") {
  survey <- c("SYN QCS", "SYN HS")
  model_ssid <- c(1, 3)
  ssid_string <- paste0(model_ssid, collapse = "n")
}

if (region == "West Coast Vancouver Island") {
  survey <- c("SYN WCVI")
  model_ssid <- c(4)
  ssid_string <- paste0(model_ssid, collapse = "n")
}

if (region == "West Coast Haida Gwaii") {
  survey <- c("SYN WCHG")
  model_ssid <- c(16)
  ssid_string <- paste0(model_ssid, collapse = "n")
}
```

#### Get species-specific optimal values from mountain plots for most abundant year
```{r, include=FALSE}
if (immature) {
  age <- "immature"
  d_all <-readRDS(paste0("data/", spp,
  "/predictions-", spp, covs, "-1n3n4n16-imm-biomass-prior-FALSE.rds")) %>% filter(year > 2007)

  } else {
    age <- "mature" 
      try({
        d_all <-readRDS(paste0("data/", spp,
        "/predictions-", spp, covs, "-1n3n4n16-total-biomass-prior-FALSE.rds")) %>% filter(year > 2007) 
      })
      try({
        d_all <-readRDS(paste0("data/", spp,
        "/predictions-", spp, covs, "-1n3n4n16-mature-biomass-prior-FALSE.rds")) %>% filter(year > 2007) 
      })

   }

nd_all <- readRDS(paste0("data/predicted-DO-new.rds")) %>% select (X, Y, year, temp, do_est)

d_all <- left_join(d_all, nd_all, by=c("X","Y", "year"))

glimpse(d_all)
```

#### SET TIME FRAME
```{r}
multiyear <- TRUE

if (multiyear){
 #unique(d_all$year)
 d <- d_all %>% filter(ssid %in% model_ssid)
 length(unique(d$year))
 years_sampled <- length(unique(d$year))
 year_range <- (years_sampled -1)*2
 start_year <- 2007
 if (years_sampled==5) indices <- c(1, 2, 2, 2, 2)
 if (years_sampled==6) indices <- c(1, 2, 2, 2, 2, 2)
} else{ 
d <- d_all %>% filter(year >= start_year) %>% filter(year <= start_year+3)
unique(d$year)
indices <- c(1,2)
year_range <- 2
}

skip_time <- NULL
delta_t_step <- year_range
delta_t_total <- year_range
```

#### TRIM environment to range sampled
```{r}
#d <- d %>% filter(do_est > 0.23) %>% filter(do_est < 7.91) # full range
d <- d %>% filter(do_est > 0.28) %>% filter(do_est < 7.06) # 0.005 and 0.995
#d <- d %>% filter(temp > 2.61) %>% filter(temp < 14.31) # full range
d <- d %>% filter(temp > 3.07) %>% filter(temp < 11.3) # 0.005 and 0.995
```

# Calculate climate velocity
### Dissolved Oxygen
```{r }
gf0 <- vocc_gradient_calc(d, "do_est",
  scale_fac = scale_fac,
  # all layers with max indice will be averaged for spatial gradient
  # if default (NULL) will use all years
  # indices = indices,
  divisor = 1,
  quantile_cutoff = 0.05
)

gf0 <- gf0 %>%
  mutate(
    trend_per_year = trend,
    trend_per_decade = trend_per_year * 10,
    velocity = velocity
  ) %>% dplyr::mutate(X = x, Y = y) %>%
   gfplot:::utm2ll(., utm_zone = 9)

grad0 <- plot_vocc(gf0,
  #theme_black = TRUE,
  coast = TRUE,
  vec_aes = NULL,
  #grad_vec_aes = "velocity",
  #vec_lwd_range = c(0.5, 0.5),
  NA_label = "NA",
  fill_col = "velocity",
  fill_label = "gradient velocity",
  raster_alpha = 1,
  # raster_limits = c(raster_min, raster_max),
  na_colour = "black",
  high_fill = "Steel Blue 4",
  low_fill = "Red 3",
  white_zero = TRUE,
  vec_alpha = 0.35,
  axis_lables = FALSE,
  transform_col = sqrt # no_trans # 
) + labs(
  title = "do velocity",
  subtitle = paste0(min(d$year), "-", max(d$year))
)

grad0
```

### Temperature
```{r}
gf2 <- vocc_gradient_calc(d, "temp",
  scale_fac = scale_fac,
  # all layers with max indice will be averaged for spatial gradient
  # if default (NULL) will use all years
  # indices = indices,
  divisor = 1,
  quantile_cutoff = 0.05
)

gf2 <- gf2 %>%
  mutate(
    trend_per_year = trend ,
    trend_per_decade = trend_per_year * 10,
    velocity = velocity
  ) %>%
  dplyr::mutate(X = x, Y = y)

grad2 <- plot_vocc(gf2,
  # theme = "black",
  coast = TRUE,
  vec_aes = NULL,
  #grad_vec_aes = "velocity",
  #vec_lwd_range = c(0.5, 0.5),
  NA_label = "NA",
  fill_col = "velocity",
  fill_label = "gradient velocity",
  raster_alpha = 1,
  # raster_limits = c(raster_min, raster_max),
  na_colour = "white",
  white_zero = TRUE,
  vec_alpha = 0.35,
  axis_lables = FALSE,
  transform_col = sqrt # no_trans # log10 # fourth_root_power # sqrt # 
) + labs(
  title = "temp velocity",
  subtitle = paste0(min(d$year), "-", max(d$year))
)

grad2
```

### Gradient-based BIOTIC VECTORS
```{r}
#indices <- c(1,2,3,4,5)

gf3 <- vocc_gradient_calc(d, "est",
  scale_fac = scale_fac,
  #indices = indices,
  divisor = 1,
  log_space = TRUE,
  quantile_cutoff = 0.025
)

gf3 <- gf3 %>%
  mutate(
    CV_formula = sqrt(exp(cv^2)-1),
    CV = sd/exp(mean), 
    mean_biomass = exp(mean)*10000,
    trend_per_year = exp(trend),
    trend_per_decade = trend_per_year * 10,
    velocity = velocity
  ) %>%
  dplyr::mutate(X = x, Y = y)

bio5perc <- sum(gf3$mean_biomass, na.rm = TRUE) * 0.0001
s <- sort(gf3$mean_biomass)

bio_sum <- cumsum(s)
lower_density_threshold <- s[which(bio_sum >= bio5perc)[1]]

gf3t <- filter(gf3, mean_biomass > lower_density_threshold)

grad3 <- gfranges::plot_vocc(gf3t,
  # theme = "black",
  coast = TRUE,
  vec_aes = NULL,
  #grad_vec_aes = "velocity",
  #vec_lwd_range = c(0.5, 0.5),
  NA_label = "NA",
  fill_col = "velocity",
  fill_label = "gradient velocity",
  raster_alpha = 1,
 # raster_limits = c(raster_min, raster_max),
  na_colour = "black",
  # white_zero = TRUE,
  high_fill = "Steel Blue 4",
  low_fill = "Red 3",
  axis_lables = FALSE,
  transform_col =  sqrt # no_trans #
) + labs(
  title = paste0(species, " (", age, ")"), #
  subtitle = paste0("Biotic velocity ", min(d$year), "-", max(d$year))
)
grad3
```

```{r}

grad4 <- gfranges::plot_vocc(gf3t,
  # theme = "black",
  coast = TRUE,
  vec_aes = NULL,
  #grad_vec_aes = "velocity",
  #vec_lwd_range = c(0.5, 0.5),
  NA_label = "NA",
  fill_col = "mean_biomass",
  fill_label = "mean",
  raster_alpha = 1,
  # raster_limits = c(raster_min, raster_max),
  na_colour = "black",
  # white_zero = TRUE,
  high_fill = "Steel Blue 4",
  low_fill = "Red 3",
  axis_lables = FALSE,
  transform_col =  sqrt#
) + labs(
  title = paste0(species, " (", age, ")"), #
  subtitle = paste0("Mean biomass ", min(d$year), "-", max(d$year))
)
grad4

```

```{r}
grad5 <- gfranges::plot_vocc(gf3t,
  # theme = "black",
  coast = TRUE,
  vec_aes = NULL,
  #grad_vec_aes = "velocity",
  #vec_lwd_range = c(0.5, 0.5),
  NA_label = "NA",
  fill_col = "trend_per_year",
  fill_label = "biomass trend",
  raster_alpha = 1,
  #raster_limits = c(0, 10),
  na_colour = "black",
  white_zero = TRUE,
  high_fill = "Steel Blue 4",
  low_fill = "Red 3",
  axis_lables = FALSE,
  transform_col = sqrt #
) + labs(
  title = paste0(species, " (", age, ")"), #
  subtitle = paste0("Biomass trend ", min(d$year), "-", max(d$year))
)
grad5

```

```{r}
grad6 <- gfranges::plot_vocc(gf3t,
  # theme = "black",
  coast = TRUE,
  vec_aes = NULL,
  #grad_vec_aes = "velocity",
  #vec_lwd_range = c(0.5, 0.5),
  NA_label = "NA",
  fill_col = "cv",
  fill_label = "sd in log space",
  raster_alpha = 1,
  #raster_limits = c(0, 7),
  na_colour = "black",
  # white_zero = TRUE,
  high_fill = "Steel Blue 4",
  low_fill = "Red 3",
  axis_lables = FALSE,
  transform_col = sqrt #
) + labs(
  title = paste0(species, " (", age, ")"), #
  subtitle = paste0("Biotic variability (cv from vocc_gradient_calc funcition) ", min(d$year), "-", max(d$year))
)
grad6
```

```{r}
raster_max <- quantile(gf3t$CV, 0.90, na.rm = TRUE)
grad7 <- gfranges::plot_vocc(gf3t,
  # theme = "black",
  coast = TRUE,
  vec_aes = NULL,
  #grad_vec_aes = "velocity",
  #vec_lwd_range = c(0.5, 0.5),
  NA_label = "NA",
  fill_col = "CV",
  fill_label = "CV",
  raster_alpha = 1,
  raster_limits = c(0, raster_max),
  na_colour = "black",
  # white_zero = TRUE,
  high_fill = "Steel Blue 4",
  low_fill = "Red 3",
  axis_lables = FALSE,
  transform_col = sqrt #
) + labs(
  title = paste0(species, " (", age, ")"), #
  subtitle = paste0("CV in estimated biomass ", min(d$year), "-", max(d$year))
)
grad7
```

### Save gradient-based VOCCs
```{r}

biotic_values <- gf3t %>% rename(
    biotic_vel = velocity, 
    biotic_trend = trend_per_year, 
    biotic_CV = CV,
    sd_est = cv,
    biotic_grad = gradient) %>% 
  mutate(sd_biomass = sd*10000,
    mean_biomass = exp(mean)*10000) %>%
  select(x, y, biotic_vel, biotic_trend, biotic_grad, biotic_CV, sd_est, sd_biomass, mean_biomass) %>% 
  mutate(start_year = !!start_year)

DO_values <- gf0 %>% rename( 
    DO_vel = velocity, 
    DO_trend = trend_per_year, 
    DO_grad = gradient) %>% 
  select(x, y, DO_vel, DO_trend, DO_grad)

temp_values <- gf2 %>% rename( 
    temp_vel = velocity, 
    temp_trend = trend_per_year, 
    temp_grad = gradient) %>% 
  select(x, y, temp_vel, temp_trend, temp_grad)

vocc <- left_join(biotic_values, DO_values) # add DO
vocc  <- left_join(vocc , temp_values) # add temp
vocc  <- mutate(vocc, species = !!species, method = "gradient")

write.csv(vocc, file = paste0(
  "data/biotic_test/biotic-vocc-", spp, "-", age, covs, "-", ssid_string, "-", min(d$year), "-", max(d$year), ".csv"
))

```