### 2020-02-06 analysis/VOCC/old-code/4-vocc-by-spp.Rmd

---
title: "Linking biomass change with VOCC"
author: "Philina English"
date: '2019-08-14'
output: html_document
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 11, fig.height = 8.5,
  echo = FALSE, warning = FALSE, message = FALSE
)
options(scipen = 999)
options(digits = 4)
```

#### Set species
```{r echo=TRUE}
# # Species run so far...
# species <- "Arrowtooth Flounder"
# species <- "Pacific Cod"
# species <- "Sablefish"
species <- "Silvergray Rockfish"
# species <- "Lingcod"
# species <- "North Pacific Spiny Dogfish" # note: using all data for maturity thresholds
# species <- "Quillback Rockfish"
# species <- "Pacific Ocean Perch"
# species <- "Yelloweye Rockfish"
```

#### Get species-specific optimal values from mountain plots for most abundant year
```{r, echo=FALSE}
spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))

params <- readRDS(paste0("data/optimal-values-by-species.rds"))
params <- params[rev(order(as.Date(params$date))), ]
params <- params[!duplicated(params$species), ]

optimal_do <- round(params[params$species == species, ]$optimal_do, digits = 2)
optimal_temp <- round(params[params$species == species, ]$optimal_temp, digits = 2)

optimal_values <- c(optimal_do, optimal_temp)
```

#### Set vector type
```{r echo=TRUE}
 climate <- "temperature"
# climate <- "do"
# climate <- "do and temperature"
```

#### Make climate vectors for biannual changes between surveys
```{r echo=TRUE}
trend <- FALSE
start_time <- 2012
end_time <- 2015
temp_threshold <- c(0.1)
do_threshold <- c(0.1)
```

# OR

#### Make climate vectors for average changes and lower threshold values
```{r echo=TRUE}
# trend <- TRUE
# temp_trend_threshold <- c(0.25)
# do_trend_threshold <- c(0.2)
```

#### Set Region 
```{r echo=TRUE}
# region <- "West Coast Vancouver Island"
region <- "both odd year surveys"
# region <- "West Coast Haida Gwaii"
```

Run all subsequent code...
```{r, include=FALSE}
library(dplyr)
library(ggplot2)
# library(gfplot)
# library(gfdata)
library(sdmTMB)
library(gfranges)
```

```{r, message=FALSE, warning=FALSE, include=FALSE}
if (region == "both odd year surveys") {
  survey <- c("SYN QCS", "SYN HS")
  model_ssid <- c(1, 3)
  ssid_string <- paste0(model_ssid, collapse = "n")
  trend_indices_do <- c(1, 1, 1, 2, 2)
  trend_indices_temp <- c(1, 1, 1, 2, 2)
  years <- NULL
}

if (region == "West Coast Vancouver Island") {
  survey <- c("SYN WCVI")
  model_ssid <- c(4)
  ssid_string <- paste0(model_ssid, collapse = "n")
  trend_indices_do <- c(1, 1, 1, 2, 2)
  trend_indices_temp <- c(1, 1, 1, 2, 2, 2)
  years <- NULL
}

if (region == "West Coast Haida Gwaii") {
  survey <- c("SYN WCHG")
  model_ssid <- c(16)
  ssid_string <- paste0(model_ssid, collapse = "n")
  trend_indices_do <- c(1, 1, 1, 2, 2)
  trend_indices_temp <- c(1, 1, 1, 2, 2)
  years <- NULL
}
```

```{r, include=FALSE}
# Set default cell sizes and scale
input_cell_size <- 2
scale_fac <- 1
delta_t_step <- 2
skip_time <- NULL

if (climate == "temperature") {
  variable_names <- c("temp")
  min_thresholds <- c(Inf)
  delta_threshold <- temp_threshold
  optimal_values <- optimal_temp

  if (trend) {
    max_thresholds <- temp_trend_threshold
    start_time <- 2008 # change if can add back in prior 2008 data
    end_time <- NULL
    delta_t_total <- 6 # change if can add back in prior 2008 data
    indices <- trend_indices_temp
  } else {
    max_thresholds <- temp_threshold
    indices <- c(1, 2)
    delta_t_total <- 2
  }
}

if (climate == "do") {
  variable_names <- c("do_est")
  max_thresholds <- c(Inf)
  delta_threshold <- -1 * (do_threshold)
  optimal_values <- optimal_do

  if (trend) {
    min_thresholds <- do_trend_threshold
    start_time <- 2008
    skip_time <- 2016
    end_time <- NULL
    delta_t_total <- 6
    indices <- trend_indices_do
  } else {
    min_thresholds <- do_threshold
    indices <- c(1, 2)
    delta_t_total <- 2
  }
}

if (climate == "do and temperature") {
  variable_names <- c("do_est", "temp")

  if (trend) {
    min_thresholds <- c(do_trend_threshold, Inf)
    max_thresholds <- c(Inf, temp_trend_threshold)
    start_time <- 2008
    skip_time <- 2016
    end_time <- NULL
    delta_t_total <- 6
    indices <- trend_indices_do
  } else {
    min_thresholds <- c(do_threshold, Inf)
    max_thresholds <- c(Inf, temp_threshold)
    indices <- c(1, 2)
    delta_t_total <- 2
  }
}

min_string <- paste0(min_thresholds, collapse = "-")
max_string <- paste0(max_thresholds, collapse = "-")

optimal_string <- paste0(optimal_values, collapse = "-")
climate_string <- gsub(" ", "-", gsub("\\/", "-", tolower(climate)))
```

# Calculate climate vectors

```{r}
rm(vocc)
try({
  vocc <- readRDS(paste0(
    "data/vocc-", ssid_string, "-", climate_string, "-",
    min_string, "to", max_string, "-", start_time, "-", delta_t_total, ".rds"
  ))
})

if (!exists("vocc")) {
  pred_temp_do <- readRDS(here::here("analysis/tmb-sensor-explore/models/predicted-DO.rds"))

  if (length(variable_names) > 1) {
    climate_predictions <- replicate(length(variable_names), pred_temp_do, simplify = FALSE)
  } else {
    climate_predictions <- pred_temp_do
  }

  starttime <- Sys.time()
  vocc <- make_vector_data(climate_predictions,
    variable_names = variable_names,
    ssid = model_ssid,
    start_time = start_time,
    end_time = end_time,
    skip_time = skip_time,
    input_cell_size = input_cell_size,
    scale_fac = scale_fac,
    delta_t_total = delta_t_total,
    delta_t_step = delta_t_step,
    indices = indices,
    min_thresholds = min_thresholds,
    max_thresholds = max_thresholds,
    round_fact = 10
  )
  endtime <- Sys.time()
  time_vocc <- round(starttime - endtime)

  vocc$var_1_min <- min_thresholds[1]
  vocc$var_2_min <- min_thresholds[2]
  vocc$var_1_max <- max_thresholds[1]
  vocc$var_2_max <- max_thresholds[2]

  saveRDS(vocc, file = paste0(
    "data/vocc-", ssid_string, "-", climate_string, "-",
    min_string, "to", max_string, "-", start_time, "-", delta_t_total, ".rds"
  ))
}
# glimpse(vocc)
```

Plot all possible vectors
```{r echo=FALSE}
if (climate == "temperature") {
  vocc$diff <- vocc$units_per_decade / 10 * delta_t_total

  vocc_plot <- plot_vocc(vocc,
    vec_aes = "distance",
    min_vec_plotted = input_cell_size,
    NA_label = ".",
    fill_col = "diff",
    fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
    raster_alpha = 1,
    white_zero = TRUE,
    vec_alpha = 0.25,
    axis_lables = FALSE,
    transform_col = no_trans 
  ) + ggtitle(climate)
}

if (climate == "do") {
  vocc$diff <- vocc$units_per_decade / 10 * delta_t_total

  vocc_plot <- plot_vocc(vocc,
    vec_aes = "distance",
    min_vec_plotted = input_cell_size,
    NA_label = ".",
    fill_col = "diff",
    fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
    raster_alpha = 1,
    white_zero = TRUE,
    high_fill = "Steel Blue 4",
    low_fill = "Red 3",
    vec_alpha = 0.25,
    axis_lables = FALSE,
    transform_col = no_trans 
  ) + ggtitle(climate)
}

if (climate == "do and temperature") {
  vocc$temp_diff <- vocc$temp.units_per_decade / 10 * delta_t_total

  vocc_plot_temp <- plot_vocc(vocc,
    vec_aes = "distance",
    min_vec_plotted = input_cell_size,
    NA_label = ".",
    fill_col = "temp_diff",
    fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
    raster_alpha = 1,
    white_zero = TRUE,
    vec_alpha = 0.25,
    axis_lables = FALSE,
    transform_col = no_trans 
  ) + ggtitle("temperature")

  vocc$do_diff <- vocc$do_est.units_per_decade / 10 * delta_t_total

  vocc_plot_do <- plot_vocc(vocc,
    vec_aes = "distance",
    min_vec_plotted = input_cell_size,
    NA_label = ".",
    fill_col = "do_diff",
    fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
    raster_alpha = 1,
    white_zero = TRUE,
    high_fill = "Steel Blue 4",
    low_fill = "Red 3",
    vec_alpha = 0.25,
    axis_lables = FALSE,
    transform_col = no_trans 
  ) + ggtitle("do")

  vocc_plot <- gridExtra::grid.arrange(
    vocc_plot_temp, vocc_plot_do,
    ncol = 2,
    top = grid::textGrob(paste("do of ", optimal_values[1], "less", min_thresholds[1], "or temperature of ", optimal_values[2], "plus", max_thresholds[2], " over ", start_time, "-", start_time + delta_t_total, ""))
  )
}
vocc_plot
```

Filter vectors to match optimal range for chosen species
```{r}
vectors <- trim_vector_data(vocc, variable_names, optimal_values, min_thresholds, max_thresholds,
  cell_size = input_cell_size, dist_intercept = input_cell_size,
  max_dist = 120
)

if (length(vectors[, 2]) == 0) {
  other_coefs <- readRDS(paste0("data/", climate_string, "-vocc-by-spp-sdmTMB-coefs.RDS"))
  coefs <- other_coefs[1, ]
  coefs$var <- "NA"
  coefs$beta <- "NA"
  coefs$se <- "NA"
  coefs$model <- "NA"
  coefs$species <- spp
  coefs$lifestage <- "NA"
  coefs$climate <- climate
  coefs$min_thresholds <- min_string
  coefs$max_thresholds <- max_string
  coefs$optimum <- optimal_string
  coefs$start_time <- start_time
  coefs$timespan <- delta_t_total
  coefs$region <- region
  coefs$ssid_string <- ssid_string

  coefs <- rbind(other_coefs, coefs) %>% distinct()
  saveRDS(coefs, file = paste0("data/", climate_string, "-vocc-by-spp-sdmTMB-coefs.RDS"))

  stop("No vectors. Conditions saved to 'data/vocc-byspp-sdmTMB-coefs'.")
}
# glimpse(vocc)
# glimpse(vectors)
```

### Plot remaining climate vectors for regions where change exceeds thresholds
```{r echo=FALSE, warning=FALSE}

if (climate == "temperature") {
  vectors$diff <- vectors$units_per_decade / 10 * delta_t_total

  vocc_plot <- plot_vocc(vectors,
    vec_aes = "distance",
    min_vec_plotted = input_cell_size,
    NA_label = ".",
    fill_col = "diff",
    fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
    raster_alpha = 1,
    white_zero = TRUE,
    vec_alpha = 0.25,
    axis_lables = FALSE,
    transform_col = no_trans 
  ) + ggtitle(climate)
}

if (climate == "do") {
  vectors$diff <- vectors$units_per_decade / 10 * delta_t_total

  vocc_plot <- plot_vocc(vectors,
    vec_aes = "distance",
    min_vec_plotted = input_cell_size,
    NA_label = ".",
    fill_col = "diff",
    fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
    raster_alpha = 1,
    white_zero = TRUE,
    high_fill = "Steel Blue 4",
    low_fill = "Red 3",
    vec_alpha = 0.25,
    axis_lables = FALSE,
    transform_col = no_trans 
  ) + ggtitle(climate)
}

if (climate == "do and temperature") {
  vectors$temp_diff <- vectors$temp.units_per_decade / 10 * delta_t_total

  vocc_plot_temp <- plot_vocc(vectors,
    vec_aes = "distance",
    min_vec_plotted = input_cell_size,
    NA_label = ".",
    fill_col = "temp_diff",
    fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
    raster_alpha = 1,
    white_zero = TRUE,
    vec_alpha = 0.25,
    axis_lables = FALSE,
    transform_col = no_trans 
  ) + ggtitle("temperature")

  vectors$do_diff <- vectors$do_est.units_per_decade / 10 * delta_t_total

  vocc_plot_do <- plot_vocc(vectors,
    vec_aes = "distance",
    min_vec_plotted = input_cell_size,
    NA_label = ".",
    fill_col = "do_diff",
    fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
    raster_alpha = 1,
    white_zero = TRUE,
    high_fill = "Steel Blue 4",
    low_fill = "Red 3",
    vec_alpha = 0.25,
    axis_lables = FALSE,
    transform_col = no_trans 
  ) + ggtitle("do")

  vocc_plot <- gridExtra::grid.arrange(
    vocc_plot_temp, vocc_plot_do,
    ncol = 2,
    top = grid::textGrob(paste("DO of ", optimal_values[1], "less", min_thresholds[1], "or temperature of ", optimal_values[2], "plus", max_thresholds[2], " over ", start_time, "-", start_time + delta_t_total, ""))
  )
}
vocc_plot
```
# Add species biomass data

#### Get all biomass predictions 
```{r, echo=FALSE, warning=FALSE}
priors <- FALSE
# priors <- TRUE

covariates <- "+muddy+any_rock"

covs <- gsub("\\+", "-", covariates)

rm(adult_predictions)
rm(imm_predictions)
rm(predictions)

try({
  adult_predictions <- readRDS(paste0(
    "data/", spp,
    "/all-predictions-", spp, covs, "-mat-biomass-prior-", priors, ".rds"
  ))
})

try({
  imm_predictions <- readRDS(paste0(
    "data/", spp,
    "/all-predictions-", spp, covs, "-imm-biomass-prior-", priors, ".rds"
  ))
})

try({
  predictions <- readRDS(paste0(
    "data/", spp,
    "/all-predictions-", spp, covs, "-total-biomass-prior-", priors, ".rds"
  ))
})

if (!exists("adult_predictions")) adult_predictions <- NULL
if (!exists("imm_predictions")) imm_predictions <- NULL
if (!exists("predictions")) predictions <- NULL

biomass_list <- list(
  adult = adult_predictions,
  imm = imm_predictions,
  total = predictions
)

biomass_list <- biomass_list[!sapply(biomass_list, is.null)]

# glimpse(biomass_list)
```

#### Extract biomass values to match spatial and temporal range of climate vectors

```{r}
vocc_by_sp <- lapply(biomass_list, function(i) {
  biomass_change <- make_trend_data(i,
    ssid = model_ssid,
    start_time = start_time,
    skip_time = skip_time,
    end_time = end_time,
    input_cell_size = input_cell_size,
    scale_fac = scale_fac,
    delta_t_total = delta_t_total,
    delta_t_step = delta_t_step,
    indices = indices,
    variable_names = "est_exp"
  )
  # browser()
  vectors <- vectors %>%
    mutate(
      start_cell = icell,
      vector_id = paste0(id, "_to_", tid),
      distance2 = distance * distance,
      source_end_value = e, 
      target_value = target_values
    )

  start <- biomass_change %>%
    rename(start_density = est_exp_s, end_density = est_exp_e) %>%
    dplyr::select(icell, x, y, start_density, end_density)

  vector_source <- left_join(vectors, start, by = c("icell", "x", "y")) %>%
    mutate(
      cell_type = "source", end_climate = source_end_value,
      source_density = start_density, source_end_density = end_density
    )

  end <- biomass_change %>%
    rename(
      start_density = est_exp_s,
      end_density = est_exp_e,
      target_X = x, target_Y = y
    ) %>%
    dplyr::select(start_density, end_density, target_X, target_Y)

  vector_target <- left_join(vectors, end, by = c("target_X", "target_Y")) %>%
    mutate(cell_type = "target", end_climate = target_value)

  source_density <- biomass_change %>%
    rename(
      source_density = est_exp_s,
      source_end_density = est_exp_e
    ) %>%
    dplyr::select(source_density, source_end_density, x, y)

  vector_target <- left_join(vector_target, source_density, by = c("x", "y"))

  # combine source and target cell data
  vocc_by_sp <- rbind(vector_source, vector_target) %>%
    mutate(cell_type = as.factor(cell_type)) %>%
    filter(!is.na(icell)) %>%
    filter(!is.na(source_density)) %>%
    filter(!is.na(end_density)) %>%

    # remove potential source cells with lowest 10% quantile of fish biomass
    filter(source_density > quantile(source_density, 0.1)) %>%

    # calculate percent change in biomass
    mutate(
      log_change = log(end_density / start_density),
      perc_change = end_density / start_density,
      log_source_density = source_density
    )
  # browser()
  end_climate <- purrr::map_df(
    strsplit(vocc_by_sp$end_climate, " "),
    function(x) data.frame(t(as.numeric(x)))
  )

  names(end_climate) <- paste0("end_", variable_names)

  vocc_by_sp <- cbind(vocc_by_sp, end_climate) %>%
    rename(end_climate_chr = end_climate)

  # get mean and sd for densities without cell duplication
  icells <- vocc_by_sp %>% distinct(icell, .keep_all = TRUE)

  mean.log.density <- mean(icells$log_source_density, na.rm = TRUE)
  sd.log.density <- sd(icells$log_source_density, na.rm = TRUE)
  mean.change <- mean(icells$perc_change, na.rm = TRUE)
  sd.change <- sd(icells$perc_change, na.rm = TRUE)

  for (j in seq_along(variable_names)) {
    mean.var <- paste0("mean.", variable_names[j])
    sd.var <- paste0("sd.", variable_names[j])

    icells <- mutate(
      icells,
      (!!mean.var) := mean(UQ(rlang::sym(paste0(variable_names[j], "_e"))), na.rm = TRUE),
      (!!sd.var) := sd(UQ(rlang::sym(paste0(variable_names[j], "_e"))), na.rm = TRUE)
    )
  }

  # calculate standardized density variables
  vocc_by_sp <- vocc_by_sp %>%
    mutate(
      std_source_density = (log_source_density - mean.log.density) / (sd.log.density * 2),
      std_source_density2 = (log_source_density - mean.log.density) / (sd.log.density * 2)^2,
      mean_source_density = mean.log.density, sd_density = sd.log.density,
      mean_change = mean.change, sd_change = sd.change
    )

  # calculate standardized climate variables
  for (j in seq_along(variable_names)) {
    mean.var <- paste0("mean.", variable_names[j])
    sd.var <- paste0("sd.", variable_names[j])
    std.var <- paste0("std.", variable_names[j])
    std.var2 <- paste0("std.", variable_names[j], "2")

    vocc_by_sp <- mutate(
      vocc_by_sp,
      (!!mean.var) := icells[1, mean.var],
      (!!sd.var) := icells[1, sd.var],
      (!!std.var) := (UQ(rlang::sym(paste0(variable_names[j], "_e"))) - icells[1, mean.var]) / (icells[1, sd.var]*2),
      (!!std.var2) := ((UQ(rlang::sym(paste0(variable_names[j], "_e"))) - icells[1, mean.var]) / icells[1, sd.var])^2
    )
  }

  # rename spatial coordinates for inclusion in models and plots without naming conflicts
  vocc_by_sp <- vocc_by_sp %>%
    rename(X = x, Y = y)

  vocc_by_sp
})
```

## Adult plots

Difference between target and source cells split by nearest cells and further cells
```{r}
targets_only <- vocc_by_sp[["adult"]] %>% filter(cell_type == "target")
targets_only$change_diff <- targets_only$log_change - log(targets_only$source_end_density / targets_only$source_density)

ggplot(filter(targets_only, distance < 2), aes(change_diff, fill = "green")) +
  geom_histogram(bins = 60) +
  geom_histogram(
    data = filter(targets_only, distance >= 2), aes(change_diff, fill = "blue"),
    bins = 60, alpha = 0.5, inherit.aes = FALSE
  ) +
  geom_vline(xintercept = 0, color = "black", size = 0.5) +
  scale_fill_manual("Distance", labels = c("Further cells", "Nearest cells"), values = c("blue" = "blue", "green" = "green"), guide = guide_legend(reverse = TRUE)) +
  ggtitle(paste("Log adult", species, "biomass change \n(", climate, "-based Target-Source cells in", region, ")"))

ggplot(filter(targets_only, distance < 2), aes(change_diff)) +
  geom_density(aes(fill = "green")) +
  geom_density(data = filter(targets_only, distance >= 2), aes(change_diff, fill = "blue"), alpha = 0.5, inherit.aes = FALSE) +
  geom_vline(xintercept = 0, color = "black", size = 0.5) +
  scale_fill_manual("Distance", labels = c("Further cells", "Nearest cells"), values = c("blue" = "blue", "green" = "green"), guide = guide_legend(reverse = TRUE)) +
  ggtitle(paste("Log adult", species, "biomass change \n(", climate, "-based Target-Source cells in", region, ")"))
```

Difference between target and source cells split by starting source biomass
```{r}
targets_only <- vocc_by_sp[["adult"]] %>% filter(cell_type == "target")
targets_only$change_diff <- targets_only$log_change - log(targets_only$source_end_density / targets_only$source_density)

range(targets_only$source_density)

ggplot(targets_only, aes(change_diff)) +
  geom_histogram(data = filter(targets_only, source_density >= quantile(targets_only$source_density, 0.25)), aes(change_diff, fill = "green"), bins = 60, alpha = 0.7, inherit.aes = FALSE) +
  geom_histogram(data = filter(targets_only, source_density < quantile(targets_only$source_density, 0.25)), aes(change_diff, fill = "blue"), bins = 60, alpha = 0.7, inherit.aes = FALSE) +
  geom_histogram(data = filter(targets_only, source_density > quantile(targets_only$source_density, 0.75)), aes(change_diff, fill = "yellow"), bins = 60, alpha = 0.9, inherit.aes = FALSE) +
  geom_vline(xintercept = 0, color = "black", size = 0.5) +
  scale_fill_manual("Source biomass", labels = c("<25th quantile", ">25th quantile", ">75th quantile"), values = c("blue" = "blue", "green" = "green", "yellow" = "yellow"), guide = guide_legend(reverse = FALSE)) +
  ggtitle(paste("Log adult", species, "biomass change \n(", climate, "-based Target-Source cells in", region, ")"))

ggplot(targets_only, aes(change_diff)) +
  geom_density(data = filter(targets_only, source_density >= quantile(targets_only$source_density, 0.25)), aes(change_diff, fill = "green"), alpha = 0.9, inherit.aes = FALSE) +
  geom_density(data = filter(targets_only, source_density < quantile(targets_only$source_density, 0.25)), aes(change_diff, fill = "blue"), alpha = 0.3, inherit.aes = FALSE) +
  geom_density(data = filter(targets_only, source_density > quantile(targets_only$source_density, 0.75)), aes(change_diff, fill = "yellow"), alpha = 0.6, inherit.aes = FALSE) +
  geom_vline(xintercept = 0, color = "black", size = 0.5) +
  scale_fill_manual("Source biomass", labels = c("<25th quantile", ">25th quantile", ">75th quantile"), values = c("blue" = "blue", "green" = "green", "yellow" = "yellow"), guide = guide_legend(reverse = FALSE)) +
  ggtitle(paste("Log adult", species, "biomass change \n(", climate, "-based Target-Source cells in", region, ")"))
```

Check final climate-biomass relationship
```{r}
data <- vocc_by_sp[["adult"]]

mature_plot <- lapply(variable_names, function(j) {
  ggplot(data, aes(
    x = UQ(rlang::sym(paste0("end_", j))), y = end_density, colour = cell_type, size = start_density,
    group = cell_type
  )) +
    geom_point(alpha = 0.5) +
    geom_smooth(aes(x = UQ(rlang::sym(paste0("end_", j))), y = end_density),
      # method="lm", formula = y ~ x + I(x^2) + I(x^3),
      inherit.aes = FALSE
    ) + # + I(x^3)
    # scale_size_continuous(guide = "none") +
    scale_y_continuous(trans = log10) +
    xlab(paste("Final value of", j, "for cells")) +
    ylab(paste("Final biomass")) +
    scale_colour_viridis_d(begin = 0.1, end = 0.5) +
    gfplot::theme_pbs() + theme(legend.position = c(0.8, 0.7)) +
    ggtitle(paste("Mature", species, "for", region, "and", climate, "based VOCC"))
})
print(mature_plot)
```

Check pattens in distance and end climate
```{r}
data <- vocc_by_sp[["adult"]]
# data <- filter(vocc_by_sp[["adult"]],  distance < 90) # distance > 1 &

mature_plot2 <- lapply(variable_names, function(j) {
  ggplot(data, aes(
    x = (distance + 1), y = log_change,
    group = cell_type, size = start_density,
    colour = UQ(rlang::sym(paste0("end_", j)))
  )) +
    geom_point(alpha = 0.5) +
    # geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
    geom_smooth(method = "lm", se = FALSE) + # formula = y~x+I(x^2),
    scale_size_continuous(guide = "none") +
    scale_x_continuous(trans = log10) +
    facet_wrap(~cell_type) + # , scales = "free") +
    ylab(paste("Log change in biomass over", vectors$timespan[1], "years")) +
    xlab(paste("Distance traveled to stay within critical range of", j, "")) +
    scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", j)) +
    # scale_colour_viridis_d(begin = 0.1, end = 0.5) +
    gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.2)) +
    ggtitle(paste("Mature", species, "for", region, "and", climate, "based VOCC"))
})
mature_plot2
```

Check pattens in log source biomass and end climate
```{r warnings=FALSE}
data <- vocc_by_sp[["adult"]]
# data <- filter(vocc_by_sp[["adult"]], distance < 50)
# data <- filter(vocc_by_sp[["adult"]], distance > 4 & distance < 50)

mature_plot3 <- lapply(variable_names, function(j) {
  ggplot(data, aes(
    x = source_density, y = log_change,
    colour = UQ(rlang::sym(paste0("end_", j))),
    alpha = 1 / (distance + input_cell_size), # size = 1/distance,
    group = cell_type
  )) +
    geom_jitter(width = 0.01) +
    geom_smooth(method = "lm", se = FALSE) +
    facet_wrap(~cell_type) + # , scales = "free") +
    scale_alpha_continuous(guide = "none") +
    scale_size_continuous(guide = "none") +
    scale_x_continuous(trans = log10) +
    # scale_y_continuous(trans = sqrt) +
    ylab(paste("Log change in biomass over", vectors$timespan[1], "years")) +
    xlab(paste("Source density (dark points are closer to source)")) +
    scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", j)) +
    gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) +
    ggtitle(paste("Mature", species, "for", region, "and", climate, "based VOCC"))
})
mature_plot3
```

## Adult models

### Boosted regression
```{r}
data <- vocc_by_sp[["adult"]]

gbm_formula <- paste0("log_change ~ std_source_density + distance + cell_type + X + Y") # + cell_type * distance2

for (j in (variable_names)) {
  gbm_formula <- paste0(gbm_formula, " + std.", j)
}

mboost <- gbm::gbm(
  as.formula(gbm_formula),
  data = data,
  distribution = "gaussian",
  n.trees = 2000,
  interaction.depth = 3,
  shrinkage = 0.02
)
```

```{r}
summary(mboost)
```

```{r}
plot(mboost, 3)
plot(mboost, 1)
plot(mboost, 2)
plot(mboost, 6)
if (length(variable_names) > 1) plot(mboost, 7)
if (length(variable_names) > 2) plot(mboost, 8)

plot(mboost, c(1, 6, 3))
if (length(variable_names) > 1) plot(mboost, c(1, 7, 3))
if (length(variable_names) > 2) plot(mboost, c(1, 8, 3))

plot(mboost, c(2, 6, 3))
if (length(variable_names) > 1) plot(mboost, c(2, 7, 3))
if (length(variable_names) > 2) plot(mboost, c(2, 8, 3))
```

Other gbm plots
```{r eval=FALSE}
plot(mboost, c(1, 2))
plot(mboost, c(1, 3))
plot(mboost, c(1, 6))

if (length(variable_names) > 1) plot(mboost, c(1, 7))
if (length(variable_names) > 2) plot(mboost, c(1, 8))

plot(mboost, c(2, 3))
plot(mboost, c(2, 6))
if (length(variable_names) > 1) plot(mboost, c(2, 7))
if (length(variable_names) > 2) plot(mboost, c(2, 8))

plot(mboost, c(4, 5))
```

### Regular mixed-model

```{r}
data <- vocc_by_sp[["adult"]] # %>% filter(distance < 50)
glimpse(data)

length(unique(data$icell))
length(unique(data$vector_id))

model_formula <- paste0("log_change ~ cell_type * std_source_density + cell_type * distance + cell_type * std_source_density2") # + cell_type * distance2

for (j in (variable_names)) {
  model_formula <- paste0("", model_formula, " + std.", j, "+ std.", j, "2")
}

model_formula <- paste0("", model_formula, "+ (1 | vector_id) + (1 | icell)")

test <- lmerTest::lmer(as.formula(model_formula), REML = F, data = data)

summary(test)
# plot(test2)

# library(sjPlot)
# library(sjlabelled)
# library(sjmisc)
sjPlot::plot_model(test, title = paste("Log adult", species, "biomass change"))
```

```{r}
beta <- lme4::fixef(test)
se <- arm::se.fixef(test)
var <- names(beta)

coefs <- as.data.frame(cbind(beta, se))
coefs$ci <- coefs$se * 1.96
var[1] <- "intercept"
row.names(coefs) <- NULL
coefs <- cbind(var, coefs)
coefs$model <- model_formula
coefs$species <- spp
coefs$lifestage <- "adult"
coefs$climate <- climate
coefs$min_thresholds <- min_thresholds
coefs$max_thresholds <- max_thresholds
coefs$optimum <- optimal_string
coefs$start_time <- start_time
coefs$timespan <- delta_t_total
coefs$region <- region
coefs$ssid_string <- ssid_string

rm(other_coefs)
try({
  other_coefs <- readRDS(paste0("data/vocc-", climate_string, "-by-spp-lmer-coefs.RDS"))
})

if (exists("other_coefs")) {
  coefs <- rbind(other_coefs, coefs) %>%
    distinct() %>%
    mutate_if(is.numeric, funs(signif(., digits = 4)))
}

saveRDS(coefs, file = paste0("data/vocc-", climate_string, "-by-spp-lmer-coefs.RDS"))
```

Robust lmm
```{r eval=FALSE}
library(lme4)
data <- vocc_by_sp[["adult"]]

test3 <- robustlmm::rlmer(as.formula(model_formula), data = data)

summary(test3)
plot(test3, ask = FALSE)
```

Maybe try GAM?
```{r}
# library()
#
# model_formula <- paste0("log_change ~ cell_type + s(std_source_density) + s(distance) ") # + cell_type * distance2
#
# for (j in (variable_names) ){
#   model_formula <- paste0("", model_formula, " + s(std.", j, ") + "
#   )
# }
#
# model_formula <- paste0("", model_formula, "+ (1 | vector_id) + (1 | icell)")
#
#
# test2 <- gamm4::....(as.formula(model_formula), REML = F, data = data)
#
```

### sdmTMB

```{r}
data <- vocc_by_sp[["adult"]]
# Xmax <- quantile(data$X, 0.9995)
# Xmin <- quantile(data$X, 0.005)
# std_max <- quantile(data$log_change, 0.9995)
#
# data <- data %>% filter ( X >= Xmin & X <= Xmax )

xy <- as.data.frame(cbind(data$X, data$Y))

if (region == "both odd year surveys") {
  n_knots <- 200
}

if (region == "West Coast Vancouver Island") {
  n_knots <- 100
}

# if (region == "West Coast Haida Gwaii") {
#   spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 60)
# }

if (n_knots > nrow(unique(xy))) {
  n_knots <- nrow(unique(xy))
}

spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = n_knots)
sdmTMB::plot_spde(spde)
```

```{r}
data <- vocc_by_sp[["adult"]]

tmb_formula <- paste0("log_change ~ cell_type * std_source_density + cell_type * distance") # + cell_type * distance2 cell_type * std_source_density2 +

for (j in (variable_names)) {
  tmb_formula <- paste0("", tmb_formula, " + std.", j, "+ std.", j, "2")
}

mtmb <- sdmTMB(
  data = data,
  as.formula(tmb_formula),
  spde = spde, # control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
  silent = TRUE
)
mtmb
```

```{r}
p <- predict(mtmb)
p$residuals <- residuals(mtmb)
# idata$resids <- p$residuals
# glimpse(p)
ggplot(p, aes(est, residuals)) +
  geom_point(alpha = 0.4) +
  ylim(-8, 8) +
  geom_smooth()

plot_map_raster <- function(dat, column = "est") {
  ggplot(dat, aes_string("X", "Y", fill = column)) +
    geom_raster() +
    coord_fixed() +
    gfplot::theme_pbs() +
    scale_fill_gradient2(trans=fourth_root_power) +
    theme(axis.title.x = element_blank(), axis.title.y = element_blank())
}

plot_map_raster(p, "est")
plot_map_raster(p, "est_rf")
plot_map_raster(p, "omega_s")
plot_map_raster(p, "residuals")
```

# Immature

### Immature plots

Difference between target and source cells split by nearest cells and further cells
```{r}
targets_only <- vocc_by_sp[["imm"]] %>% filter(cell_type == "target")
targets_only$change_diff <- targets_only$log_change - log(targets_only$source_end_density / targets_only$source_density)

ggplot(filter(targets_only, distance < 2), aes(change_diff, fill = "green")) +
  geom_histogram(bins = 60) +
  geom_histogram(
    data = filter(targets_only, distance >= 2), aes(change_diff, fill = "blue"),
    bins = 60, alpha = 0.5, inherit.aes = FALSE
  ) +
  scale_fill_manual("Distance", labels = c("Further cells", "Nearest cells"), values = c("blue" = "blue", "green" = "green"), guide = guide_legend(reverse = TRUE)) +
  ggtitle(paste("Log immature", species, "biomass change \n(Target-Source cells in", region, ")"))

ggplot(filter(targets_only, distance < 2), aes(change_diff)) +
  geom_density(aes(fill = "green")) +
  geom_density(data = filter(targets_only, distance >= 2), aes(change_diff, fill = "blue"), alpha = 0.5, inherit.aes = FALSE) +
  geom_vline(xintercept = 0, color = "black", size = 0.5) +
  scale_fill_manual("Distance", labels = c("Further cells", "Nearest cells"), values = c("blue" = "blue", "green" = "green"), guide = guide_legend(reverse = TRUE)) +
  ggtitle(paste("Log immature", species, "biomass change \n(Target-Source cells in", region, ")"))
```

Difference between target and source cells split by starting source biomass
```{r}
targets_only <- vocc_by_sp[["imm"]] %>% filter(cell_type == "target")
targets_only$change_diff <- targets_only$log_change - log(targets_only$source_end_density / targets_only$source_density)
# range(targets_only$source_density)

ggplot(targets_only, aes(change_diff)) +
  geom_histogram(data = filter(targets_only, source_density >= quantile(targets_only$source_density, 0.25)), aes(change_diff, fill = "green"), bins = 60, alpha = 0.7, inherit.aes = FALSE) +
  geom_histogram(data = filter(targets_only, source_density < quantile(targets_only$source_density, 0.25)), aes(change_diff, fill = "blue"), bins = 60, alpha = 0.7, inherit.aes = FALSE) +
  geom_histogram(data = filter(targets_only, source_density > quantile(targets_only$source_density, 0.75)), aes(change_diff, fill = "yellow"), bins = 60, alpha = 0.9, inherit.aes = FALSE) +
  geom_vline(xintercept = 0, color = "black", size = 0.5) +
  scale_fill_manual("Source biomass", labels = c("<25th quantile", ">25th quantile", ">75th quantile"), values = c("blue" = "blue", "green" = "green", "yellow" = "yellow"), guide = guide_legend(reverse = FALSE)) +
  ggtitle(paste("Log immature", species, "biomass change \n(Target-Source cells in", region, ")"))

ggplot(targets_only, aes(change_diff)) +
  geom_density(data = filter(targets_only, source_density >= quantile(targets_only$source_density, 0.25)), aes(change_diff, fill = "green"), alpha = 0.9, inherit.aes = FALSE) +
  geom_density(data = filter(targets_only, source_density < quantile(targets_only$source_density, 0.25)), aes(change_diff, fill = "blue"), alpha = 0.3, inherit.aes = FALSE) +
  geom_density(data = filter(targets_only, source_density > quantile(targets_only$source_density, 0.75)), aes(change_diff, fill = "yellow"), alpha = 0.6, inherit.aes = FALSE) +
  geom_vline(xintercept = 0, color = "black", size = 0.5) +
  scale_fill_manual("Source biomass", labels = c("<25th quantile", ">25th quantile", ">75th quantile"), values = c("blue" = "blue", "green" = "green", "yellow" = "yellow"), guide = guide_legend(reverse = FALSE)) +
  ggtitle(paste("Log immature", species, "biomass change \n(Target-Source cells in", region, ")"))
```

Check final climate-biomass relationship
```{r}
data <- vocc_by_sp[["imm"]]
# glimpse(data)

immature_plot <- lapply(variable_names, function(j) {
  ggplot(data, aes(
    x = UQ(rlang::sym(paste0("end_", j))), y = (end_density),
    colour = cell_type, size = start_density,
    group = cell_type
  )) +
    geom_point(alpha = 0.5) +
    geom_smooth(aes(x = UQ(rlang::sym(paste0("end_", j))), y = end_density),
      # method="lm", formula = y ~ x + I(x^2) + I(x^3),
      inherit.aes = FALSE
    ) + # + I(x^3)
    # scale_size_continuous(guide = "none") +
    scale_y_continuous(trans = log10) +
    xlab(paste("Final value of", j, "for cells")) +
    ylab(paste("Final biomass")) +
    scale_colour_viridis_d(begin = 0.1, end = 0.5) +
    gfplot::theme_pbs() + theme(legend.position = c(0.8, 0.8)) +
    ggtitle(paste("Immature", species, "for", region, "and", climate, "based VOCC"))
})
immature_plot
```

Check pattens in distance and end climate
```{r}
data <- vocc_by_sp[["imm"]]
# data <- filter(vocc_by_sp[["imm"]],  distance < 90) # distance > 1 &
# data <- filter(vocc_by_sp[["imm"]], distance > 1 ) #  &

immature_plot2 <- lapply(variable_names, function(j) {
  ggplot(data, aes(
    x = (distance + 1), y = log_change,
    group = cell_type, size = start_density,
    colour = UQ(rlang::sym(paste0("end_", j)))
  )) +
    geom_point(alpha = 0.5) +
    # geom_jitter(width = 0.05, alpha = 0.45) + #, size = 2
    geom_smooth(method = "lm") + # , formula = y~x+I(x^2)
    scale_size_continuous(guide = "none") +
    scale_x_continuous(trans = log10) +
    facet_wrap(~cell_type) + # , scales = "free") +
    ylab(paste("Log change in biomass over", vectors$timespan[1], "years")) +
    xlab(paste("Distance traveled to stay within critical range of", j, "")) +
    scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", j)) +
    # scale_colour_viridis_d(begin = 0.1, end = 0.5) +
    gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.7)) +
    ggtitle(paste("Immature", species, "for", region, "and", climate, "based VOCC"))
})
immature_plot2
```

Check pattens in log source biomass and end climate
```{r warnings=FALSE}
data <- vocc_by_sp[["imm"]]
# data <- filter(vocc_by_sp[["imm"]], distance < 50)
# data <- filter(vocc_by_sp[["imm"]], distance > 4 & distance < 50)

immature_plot3 <- lapply(variable_names, function(j) {
  ggplot(data, aes(
    x = source_density, y = log_change,
    colour = UQ(rlang::sym(paste0("end_", j))),
    alpha = 1 / (distance + input_cell_size), # size = 1/distance,
    group = cell_type
  )) +
    geom_jitter(width = 0.01) +
    geom_smooth(method = "lm") +
    facet_wrap(~cell_type) + # , scales = "free") +
    scale_alpha_continuous(guide = "none") +
    scale_size_continuous(guide = "none") +
    scale_x_continuous(trans = log10) +
    ylab(paste("Log change in biomass over", vectors$timespan[1], "years")) +
    xlab(paste("Source density (dark points are closer to source)")) +
    scale_colour_viridis_c(begin = 0.1, end = 0.9, name = paste("end", j)) +
    gfplot::theme_pbs() + theme(legend.position = c(0.9, 0.8)) +
    ggtitle(paste("Immature", species, "for", region, "and", climate, "based VOCC"))
})
immature_plot3
```

## Immature models

### Boosted regression
```{r}
idata <- vocc_by_sp[["imm"]]

im <- gbm::gbm(
  as.formula(gbm_formula),
  data = idata,
  distribution = "gaussian",
  n.trees = 2000,
  interaction.depth = 3,
  shrinkage = 0.02
)
```

```{r}
summary(im)
```

```{r}
plot(im, 3)
plot(im, 1)
plot(im, 2)
plot(im, 6)
if (length(variable_names) > 1) plot(im, 7)
if (length(variable_names) > 2) plot(im, 8)

plot(im, c(1, 6, 3))
if (length(variable_names) > 1) plot(im, c(1, 7, 3))
if (length(variable_names) > 2) plot(im, c(1, 8, 3))

plot(im, c(2, 6, 3))
if (length(variable_names) > 1) plot(im, c(2, 7, 3))
if (length(variable_names) > 2) plot(im, c(2, 8, 3))
```

Other gbm plots
```{r eval=FALSE}
plot(im, c(1, 2))
plot(im, c(1, 3))
plot(im, c(1, 6))

if (length(variable_names) > 1) plot(im, c(1, 7))
if (length(variable_names) > 2) plot(im, c(1, 8))

plot(im, c(2, 3))
plot(im, c(2, 6))
if (length(variable_names) > 1) plot(im, c(2, 7))
if (length(variable_names) > 2) plot(im, c(2, 8))

plot(im, c(4, 5))
```

### Regular mixed-model

```{r}
idata <- vocc_by_sp[["imm"]]

itest <- lmerTest::lmer(as.formula(model_formula), REML = F, data = idata)

summary(itest)

# library(sjlabelled)
# library(sjmisc)
sjPlot::plot_model(itest, title = paste("Log immature", species, "biomass change"))
```

```{r}
beta <- lme4::fixef(itest)
se <- arm::se.fixef(itest)
var <- names(beta)

coefs <- as.data.frame(cbind(beta, se))
coefs$ci <- coefs$se * 1.96
var[1] <- "intercept"
row.names(coefs) <- NULL
coefs <- cbind(var, coefs)
coefs$model <- model_formula
coefs$species <- spp
coefs$lifestage <- "immature"
coefs$climate <- climate
coefs$min_thresholds <- min_thresholds
coefs$max_thresholds <- max_thresholds
coefs$optimum <- optimal_string
coefs$start_time <- start_time
coefs$timespan <- delta_t_total
coefs$region <- region
coefs$ssid_string <- ssid_string

rm(other_coefs)
try({
  other_coefs <- readRDS(paste0("data/vocc-", climate_string, "-by-spp-lmer-coefs.RDS"))
})

if (exists("other_coefs")) {
  coefs <- rbind(other_coefs, coefs) %>%
    distinct() %>%
    mutate_if(is.numeric, funs(signif(., digits = 4)))
}

coeds <- coefs %>% distinct()

saveRDS(coefs, file = paste0("data/vocc-", climate_string, "-by-spp-lmer-coefs.RDS"))
```

Robust lmm
```{r eval=FALSE}
library(lme4)

test3 <- robustlmm::rlmer(as.formula(model_formula), data = idata)

summary(test3)
plot(test3, ask = FALSE)
```

### sdmTMB

```{r}
idata <- vocc_by_sp[["imm"]]

xy <- as.data.frame(cbind(idata$X, idata$Y))

if (region == "both odd year surveys") {
  n_knots <- 200
}

if (region == "West Coast Vancouver Island") {
  n_knots <- 100
}

# if (region == "West Coast Haida Gwaii") {
#   spde <- sdmTMB::make_spde(data$X, data$Y, n_knots = 60)
# }

if (n_knots > nrow(unique(xy))) {
  n_knots <- nrow(unique(xy))
}

ispde <- sdmTMB::make_spde(idata$X, idata$Y, n_knots = n_knots)
sdmTMB::plot_spde(ispde)
```

```{r}
idata <- vocc_by_sp[["imm"]]

# tmb_formula <- paste0("log_change ~ cell_type * std_source_density + cell_type * distance + distance2") # + cell_type * std_source_density2
#
# for (j in (variable_names) ){
#   tmb_formula <- paste0("", tmb_formula, " + std.", j, "+ std.", j, "2"
#   )
# }

itmb <- sdmTMB(
  data = idata,
  as.formula(tmb_formula),
  spde = ispde, # control = sdmTMBcontrol(step.min = 0.01, step.max = 1),
  silent = TRUE
)
itmb
```

```{r}
ip <- predict(itmb)
ip$residuals <- residuals(itmb)
# idata$resids <- p$residuals
# glimpse(p)
ggplot(ip, aes(est, residuals)) +
  geom_point(alpha = 0.4) +
  ylim(-8, 8) +
  geom_smooth()

plot_map_raster <- function(dat, column = "est") {
  ggplot(dat, aes_string("X", "Y", fill = column)) +
    geom_raster() +
    coord_fixed() +
    gfplot::theme_pbs() +
    scale_fill_gradient2(trans=fourth_root_power) +
    theme(axis.title.x = element_blank(), axis.title.y = element_blank())
}

plot_map_raster(ip, "est")
plot_map_raster(ip, "est_rf")
plot_map_raster(ip, "omega_s")
plot_map_raster(ip, "residuals")
```

```{r}
m <- mtmb

varnames <- colnames(m$tmb_data$X_ij)

coefs <- summary(m$sd_report)
coefs <- as.data.frame(coefs)
names(coefs) <- c("beta", "se")
coefs$ci <- coefs$se * 1.96
var <- varnames
var[1] <- "intercept"
coefs <- coefs[1:length(var), ]
row.names(coefs) <- NULL

coefs <- cbind(var, coefs)
# coefs$var
# coefs$beta
# coefs$se
coefs$model <- model_formula
coefs$species <- spp
coefs$lifestage <- "adult"
coefs$climate <- climate
coefs$min_thresholds <- min_thresholds
coefs$max_thresholds <- max_thresholds
coefs$optimum <- optimal_string
coefs$start_time <- start_time
coefs$timespan <- delta_t_total
coefs$region <- region
coefs$ssid_string <- ssid_string

rm(other_coefs)
try({
  other_coefs <- readRDS(paste0("data/", climate_string, "-vocc-by-spp-sdmTMB-coefs.RDS"))
})

if (exists("other_coefs")) {
  coefs <- rbind(other_coefs, coefs) %>%
    distinct() %>%
    mutate_if(is.numeric, funs(signif(., digits = 4)))
}
saveRDS(coefs, file = paste0("data/", climate_string, "-vocc-by-spp-sdmTMB-coefs.RDS"))
```

```{r}
m <- itmb

varnames <- colnames(m$tmb_data$X_ij)

coefs <- summary(m$sd_report)
coefs <- as.data.frame(coefs)
names(coefs) <- c("beta", "se")
# coefs$beta <- signif(coefs$beta, digits = 3)
# coefs$se <- signif(coefs$se, digits = 3)
coefs$ci <- signif(coefs$se * 1.96, digits = 3)
var <- varnames
var[1] <- "intercept"
coefs <- coefs[1:length(var), ]
row.names(coefs) <- NULL
coefs <- cbind(var, coefs)
coefs$species <- spp
coefs$lifestage <- "immature"
coefs$climate <- climate
coefs$min_thresholds <- min_thresholds
coefs$max_thresholds <- max_thresholds
coefs$optimum <- optimal_string
coefs$start_time <- start_time
coefs$timespan <- delta_t_total
coefs$region <- region
coefs$ssid_string <- ssid_string
coefs$model <- model_formula

try({
  other_coefs <- readRDS(paste0("data/vocc-", climate_string, "-by-spp-sdmTMB-coefs.RDS"))
})

if (exists("other_coefs")) {
  coefs <- rbind(other_coefs, coefs) %>%
    distinct() %>%
    mutate_if(is.numeric, funs(signif(., digits = 4)))
}

# glimpse(coefs)

saveRDS(coefs, file = paste0("data/vocc-", climate_string, "-by-spp-sdmTMB-coefs.RDS"))
```

```{r eval=FALSE}
lmer_coefs <- readRDS(paste0("data/vocc-", climate_string, "-by-spp-lmer-coefs.RDS"))

lmer_coefs$beta_upper <- lmer_coefs$beta + lmer_coefs$ci
lmer_coefs$beta_lower <- lmer_coefs$beta - lmer_coefs$ci

# View(lmer_coefs)

beta_target <- lmer_coefs %>%
  filter(ssid_string == "1n3") %>%
  filter(var == "cell_typetarget")

View(beta_target)
```