### 2020-02-07 analysis/VOCC/BACI/4-vocc-BACI-render.Rmd

---
title: "BACI of biomass change for VOCC source cells for all years"
author: "Philina English"
date: '2019-08-24'
output: html_document
params: 
    species: "Yelloweye Rockfish" 
    region: "West Coast Vancouver Island"
    age: "mature"
    biomass_threshold: "25"
    climate: "do"
    threshold: c(0.25) 
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(
  fig.width = 11, fig.height = 8.5,
  echo = FALSE, warning = FALSE, message = FALSE
)
options(scipen = 999)
options(digits = 4)
```

```{r pkgs, cache=FALSE, warning=FALSE, message=FALSE}
library("dplyr")
library("ggplot2")
library("gfplot")
library("sdmTMB")
library("gfranges")
```

#### Set parameters
```{r echo=TRUE}
species <- params$species
age <- params$age
region <- params$region
biomass_threshold <- params$biomass_threshold
climate <- params$climate
```

#### Set default cell sizes and scale
```{r, echo=FALSE}
global_only <- TRUE
input_cell_size <- 2
dist_intercept <- input_cell_size/2
scale_fac <- 1
delta_t_step <- 2
skip_time <- NULL
```

# Run all subsequent code...

```{r, message=FALSE, warning=FALSE, include=FALSE}
if (region == "both odd year surveys") {
  survey <- c("SYN QCS", "SYN HS")
  model_ssid <- c(1, 3)
  ssid_string <- paste0(model_ssid, collapse = "n")
  start_years <- list("2009"=2009, "2011"=2011, "2013"=2013, "2015"=2015) 
  # maybe remove 2011 & 2015 once we can add "2017"=2017 to avoid start and end values overlapping
  years <- NULL
}

if (region == "West Coast Vancouver Island") {
  survey <- c("SYN WCVI")
  model_ssid <- c(4)
  ssid_string <- paste0(model_ssid, collapse = "n")
  start_years <- list("2008"=2008, "2010"=2010, "2012"=2012, "2014"=2014, "2016"=2016) 
  # prob need to remove 2010 & 2014
  years <- NULL
}

if (region == "West Coast Haida Gwaii") {
  survey <- c("SYN WCHG")
  model_ssid <- c(16)
  ssid_string <- paste0(model_ssid, collapse = "n")
  start_years <- list("2008"=2008, "2010"=2010, "2012"=2012, "2014"=2014, "2016"=2016) 
  # prob need to remove 2010 & 2014
  years <- NULL
}
```

#### Get species-specific optimal values from mountain plots for most abundant year
```{r, include=FALSE}
spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))

root_params <- readRDS(paste0("data/root-values-by-species-by-age.rds"))

root_params <- root_params %>% mutate(model_type = ifelse(grepl("total", model), "total", "split"))

if (age == "imm") {
  root_params <- filter(root_params, age == "imm")
} else {
  root_params <- filter(root_params, age != "imm")
  if (any(root_params$model_type == "split")) { root_params <- filter(root_params, model_type != "total") }
}

root_params <- root_params[root_params$species == species, ] 
glimpse(root_params)

if (global_only) {
 root_params <- root_params %>% mutate(global = ifelse(grepl("_", model), "F", "T"))
 root_params <- filter(root_params, global == "T")
} else {
 if (model_ssid == 4) {
   root_params <- root_params %>% mutate(survey4 = ifelse(grepl("4", model), "T", "F"))
   root_params <- filter(root_params, survey4 == "T")
 }
}

if (climate == "temperature") {

temp_threshold <- params$threshold

  if (biomass_threshold == "75") {
  .lower <- root_params[rev(order(root_params$do_maxy)), ]$upper_t_75 }
  if (biomass_threshold == "50") {
  .upper <- root_params[rev(order(root_params$t_maxy)), ]$upper_t_50 }
  if (biomass_threshold == "25") {
  .upper <- root_params[rev(order(root_params$t_maxy)), ]$upper_t_25 }
  .upper <- na.omit(.upper)

  upper_thresholds <- round(.upper[1], digits = 2)
   if (is.na(upper_thresholds)) {stop(print("Temp maximum not modelled"))}

  upper_model <- root_params[rev(order(root_params$t_maxy)), ]$model

  lower_thresholds <- NA
  lower_model <- NULL
  variable_names <- c("temp")

  lower_change <- c(Inf)
  upper_change <- temp_threshold
  indices <- c(1, 2)
  delta_t_total <- 2
  }

if (climate == "do") {
  do_threshold <- params$threshold
  root_params <- filter(root_params, lower_do_25 > 0.4)

  if (biomass_threshold == "75") {
  .lower <- root_params[rev(order(root_params$do_maxy)), ]$lower_do_75 }
  if (biomass_threshold == "50") {
  .lower <- root_params[rev(order(root_params$do_maxy)), ]$lower_do_50 }
  if (biomass_threshold == "25") {
  .lower <- root_params[rev(order(root_params$do_maxy)), ]$lower_do_25 }

  .lower <- na.omit(.lower)
  lower_thresholds <- round(.lower[1], digits = 2)
  if (is.na(lower_thresholds)) {stop(print("DO minimum not modelled"))}
  lower_model <- root_params[rev(order(root_params$do_maxy)), ]$model

  upper_thresholds <- NA
  upper_model <- NULL
  variable_names <- c("do_est")

  lower_change <- do_threshold
  upper_change <- c(Inf)
  indices <- c(1, 2)
  delta_t_total <- 2
  }

if (climate == "do and temperature") {
  temp_threshold <- params$threshold
  do_threshold <- params$threshold

  root_params <- filter(root_params, lower_do_25 > 0.5)
  if (biomass_threshold == "75") {
  .lower <- root_params[rev(order(root_params$do_maxy)), ]$lower_do_75 }
  if (biomass_threshold == "50") {
  .lower <- root_params[rev(order(root_params$do_maxy)), ]$lower_do_50 }
  if (biomass_threshold == "25") {
  .lower <- root_params[rev(order(root_params$do_maxy)), ]$lower_do_25 }
  .lower <- na.omit(.lower)
  lower_thresholds <- round(.lower[1], digits = 2)

  if (biomass_threshold == "75") {
  .lower <- root_params[rev(order(root_params$do_maxy)), ]$upper_t_75 }
  if (biomass_threshold == "50") {
  .upper <- root_params[rev(order(root_params$t_maxy)), ]$upper_t_50 }
  if (biomass_threshold == "25") {
  .upper <- root_params[rev(order(root_params$t_maxy)), ]$upper_t_25 }
  .upper <- na.omit(.upper)
  upper_thresholds <- round(.upper[1], digits = 2)

  variable_names <- c("do_est", "temp")
  lower_thresholds <- c(lower_thresholds, NA)
  upper_thresholds <- c(NA, upper_thresholds)
  lower_change <- c(do_threshold, Inf)
  upper_change <- c(Inf, temp_threshold)
  upper_model <- root_params[rev(order(root_params$t_maxy)), ]$model
  lower_model <- root_params[rev(order(root_params$do_maxy)), ]$model

  indices <- c(1, 2)
  delta_t_total <- 2
  }

trim_thresholds <- c(lower_thresholds, upper_thresholds)
trim_models <- c(lower_model, upper_model)

min_string <- paste0(lower_change, collapse = "-")
max_string <- paste0(upper_change, collapse = "-")

optimal_string <- paste0(na.omit(trim_thresholds), collapse = "-")
climate_string <- gsub(" ", "-", gsub("\\/", "-", tolower(climate)))

glimpse(root_params)
```

# Calculate climate vectors
```{r}
vocc_by_year <- lapply(start_years, function(start_time) {

try({
  vocc <- readRDS(paste0(
    "data/vocc-", ssid_string, "-", climate_string, "-",
    min_string, "to", max_string, "-", start_time, "-", delta_t_total, ".rds"
  ))
})

if (!exists("vocc")) {
  pred_temp_do <- readRDS(here::here("analysis/tmb-sensor-explore/models/predicted-DO.rds"))

  if (length(variable_names) > 1) {
    climate_predictions <- replicate(length(variable_names), pred_temp_do, simplify = FALSE)
  } else {
    climate_predictions <- pred_temp_do
  }
  end_time <- start_time + 2
  vocc <- make_vector_data(climate_predictions,
    variable_names = variable_names,
    ssid = model_ssid,
    start_time = start_time,
    end_time = end_time,
    skip_time = skip_time,
    input_cell_size = input_cell_size,
    scale_fac = scale_fac,
    delta_t_total = delta_t_total,
    delta_t_step = delta_t_step,
    indices = indices,
    min_thresholds = lower_change,
    max_thresholds = upper_change,
    round_fact = 10
  )

  vocc$var_1_min <- lower_change[1]
  vocc$var_2_min <- lower_change[2]
  vocc$var_1_max <- upper_change[1]
  vocc$var_2_max <- upper_change[2]

  saveRDS(vocc, file = paste0(
    "data/vocc-", ssid_string, "-", climate_string, "-",
    min_string, "to", max_string, "-", start_time, "-", delta_t_total, ".rds"
  ))
}
 vocc
})  
```

### Plot all possible vectors
```{r echo=FALSE}
all_vector_plots <- lapply(vocc_by_year, function(vocc) {    
start_time <- vocc$start_time[1]

if (climate == "temperature") {
  vocc$diff <- vocc$units_per_decade / 10 * delta_t_total

  vocc_plot <- plot_vocc(vocc,
    vec_aes = "distance",
    min_vec_plotted = input_cell_size,
    NA_label = ".",
    fill_col = "diff",
    fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
    raster_alpha = 1,
    raster_limits = c(-4, 3),
    white_zero = TRUE,
    vec_alpha = 0.25,
    axis_lables = FALSE,
    transform_col = no_trans 
  ) + ggtitle(climate)
}

if (climate == "do") {
  vocc$diff <- vocc$units_per_decade / 10 * delta_t_total

  vocc_plot <- plot_vocc(vocc,
    vec_aes = "distance",
    min_vec_plotted = input_cell_size,
    NA_label = ".",
    fill_col = "diff",
    fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
    raster_alpha = 1,
    white_zero = TRUE,
    high_fill = "Steel Blue 4",
    low_fill = "Red 3",
    vec_alpha = 0.25,
    axis_lables = FALSE,
    transform_col = no_trans 
  ) + ggtitle(climate)
}

if (climate == "do and temperature") {
  vocc$temp_diff <- vocc$temp.units_per_decade / 10 * delta_t_total

  vocc_plot_temp <- plot_vocc(vocc,
    vec_aes = "distance",
    min_vec_plotted = input_cell_size,
    NA_label = ".",
    fill_col = "temp_diff",
    fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
    raster_alpha = 1,
    white_zero = TRUE,
    vec_alpha = 0.25,
    axis_lables = FALSE,
    transform_col = no_trans 
  ) + ggtitle("temperature")

  vocc$do_diff <- vocc$do_est.units_per_decade / 10 * delta_t_total

  vocc_plot_do <- plot_vocc(vocc,
    vec_aes = "distance",
    min_vec_plotted = input_cell_size,
    NA_label = ".",
    fill_col = "do_diff",
    fill_label = paste("Change from\n", start_time, "to", start_time + delta_t_total, ""),
    raster_alpha = 1,
    white_zero = TRUE,
    high_fill = "Steel Blue 4",
    low_fill = "Red 3",
    vec_alpha = 0.25,
    axis_lables = FALSE,
    transform_col = no_trans 
  ) + ggtitle("do")

  vocc_plot <- gridExtra::grid.arrange(
    vocc_plot_temp, vocc_plot_do,
    ncol = 2,
    top = grid::textGrob(paste("do of ", optimal_values[1], "less", min_thresholds[1], "or temperature of ", optimal_values[2], "plus", max_thresholds[2], " over ", start_time, "-", start_time + delta_t_total, ""))
  )
}
vocc_plot
})
print(all_vector_plots)
```

### Filter vectors to match optimal range for chosen species
```{r}
trimmed_vectors <- lapply(vocc_by_year, function(vocc) {

vectors <- trim_vector_data(vocc, variable_names, 
  lower_change, upper_change, 
  lower_thresholds, upper_thresholds,
  cell_size = input_cell_size, dist_intercept = dist_intercept,
  max_dist = 120
)
})

all_vectors <- do.call("rbind", trimmed_vectors)
```

### Plot remaining climate vectors for regions where change exceeds thresholds
```{r echo=FALSE, warning=FALSE}
if (climate == "temperature") {
all_vectors$diff <- all_vectors$units_per_decade / 10 * delta_t_total
raster_min <- min(all_vectors$diff)
raster_max <- max(all_vectors$diff)
}

if (climate == "do") {
all_vectors$diff <- all_vectors$units_per_decade / 10 * delta_t_total
raster_min <- min(all_vectors$diff)
raster_max <- max(all_vectors$diff)
}

if (climate == "do and temperature") {
  all_vectors$temp_diff <- all_vectors$temp.units_per_decade / 10 * delta_t_total
  all_vectors$do_diff <- all_vectors$do_est.units_per_decade / 10 * delta_t_total
  raster_min_temp <- min(all_vectors$temp_diff)
  raster_max_temp <- max(all_vectors$temp_diff)
  raster_min_do <- min(all_vectors$do_diff)
  raster_max_do <- max(all_vectors$do_diff)
}

trimmed_vector_plots <- lapply(trimmed_vectors, function(vectors) {

try({

  start_time <- vectors$start_time[1]

  if (climate == "temperature") {
  vectors$diff <- vectors$units_per_decade / 10 * delta_t_total

  vocc_plot <- plot_vocc(vectors,
    vec_aes = "distance",
    min_vec_plotted = input_cell_size,
    NA_label = ".",
    fill_col = "diff",
    fill_label = paste("Change from\n", 
      start_time, "to", start_time + delta_t_total, ""),
    raster_alpha = 1,
    raster_limits = c(raster_min, raster_max),
    white_zero = TRUE,
    vec_alpha = 0.25,
    axis_lables = FALSE,
    transform_col = no_trans 
  ) + ggtitle(climate)
}

if (climate == "do") {
  vectors$diff <- vectors$units_per_decade / 10 * delta_t_total

  vocc_plot <- plot_vocc(vectors,
    vec_aes = "distance",
    min_vec_plotted = input_cell_size,
    NA_label = ".",
    fill_col = "diff",
    fill_label = paste("Change from\n", 
      start_time, "to", start_time + delta_t_total, ""),
    raster_alpha = 1,
    raster_limits = c(raster_min, raster_max),
    white_zero = TRUE,
    high_fill = "Steel Blue 4",
    low_fill = "Red 3",
    vec_alpha = 0.25,
    axis_lables = FALSE,
    transform_col = no_trans 
  ) + ggtitle(climate)
}

if (climate == "do and temperature") {
  vectors$temp_diff <- vectors$temp.units_per_decade / 10 * delta_t_total

  vocc_plot_temp <- plot_vocc(vectors,
    vec_aes = "distance",
    min_vec_plotted = input_cell_size,
    NA_label = ".",
    fill_col = "temp_diff",
    fill_label = paste("Change from\n", 
      start_time, "to", start_time + delta_t_total, ""),
    raster_alpha = 1,
    raster_limits = c(raster_min_temp, raster_max_temp),
    white_zero = TRUE,
    vec_alpha = 0.25,
    axis_lables = FALSE,
    transform_col = no_trans 
  ) + ggtitle("temperature")

  vectors$do_diff <- vectors$do_est.units_per_decade / 10 * delta_t_total

  vocc_plot_do <- plot_vocc(vectors,
    vec_aes = "distance",
    min_vec_plotted = input_cell_size,
    NA_label = ".",
    fill_col = "do_diff",
    fill_label = paste("Change from\n", 
      start_time, "to", start_time + delta_t_total, ""),
    raster_alpha = 1,
    raster_limits = c(raster_min_do, raster_max_do),
    white_zero = TRUE,
    high_fill = "Steel Blue 4",
    low_fill = "Red 3",
    vec_alpha = 0.25,
    axis_lables = FALSE,
    transform_col = no_trans 
  ) + ggtitle("do")

  vocc_plot <- gridExtra::grid.arrange(
    vocc_plot_temp, vocc_plot_do,
    ncol = 2,
    top = grid::textGrob(paste(
      "DO of ", optimal_values[1], "less", min_thresholds[1], 
      "or temperature of ", optimal_values[2], "plus", max_thresholds[2], 
      " over ", start_time, "-", start_time + delta_t_total, ""))
  )
}
vocc_plot
})
})

print(trimmed_vector_plots)
```

```{r}
png(
  file = paste0("figs/", spp,
    "/vocc-", min_string, "to", max_string, "-", 
    climate_string, "-", spp, "-", age, "-", ssid_string, "-perc_", biomass_threshold, "-", optimal_string, "-allpairs-test.png"
  ),
  res = 600,
  units = "in",
  width = 8.5,
  height = 11
)
gridExtra::grid.arrange(
  grobs = trimmed_vector_plots,
  ncol = 2,
  top = grid::textGrob(paste("Vectors for cells exceeding", climate, "threshold of", optimal_string, "\n based on", biomass_threshold, "percent of maximum predicted", species, "biomass (", trim_models, ")"), gp=grid::gpar(fontsize=16))
)
dev.off()
```

# Add species biomass data

Get all biomass predictions 
```{r echo=FALSE, warning=FALSE}
priors <- FALSE
covs <- "-trawled-ssid"

if (age == "imm") {
  biomass_predictions <- readRDS(paste0(
    "data/", spp, "/all-predictions-", spp, covs, 
    "-imm-biomass-prior-", priors, ".rds"))
}

if (age == "mature") {
  biomass_predictions <- readRDS(paste0(
    "data/", spp, "/all-predictions-", spp, covs, 
    "-mat-biomass-prior-", priors, ".rds"))
}
```

Extract biomass values to match spatial and temporal range of climate vectors
```{r}
#alldata <- lapply(biomass_list, function(biomass_predictions) {

  biomass_by_year <- list()
  for (j in seq_along(start_years)) {
   start_time <- start_years[[j]]
   end_time <- start_time + 2

   biomass_by_year[[j]] <- make_trend_data(biomass_predictions,
    ssid = model_ssid,
    start_time = start_time,
    skip_time = skip_time,
    end_time = end_time,
    input_cell_size = input_cell_size,
    scale_fac = scale_fac,
    delta_t_total = delta_t_total,
    delta_t_step = delta_t_step,
    indices = indices,
    variable_names = "est_exp"
  )
  }

  biomass_change <- do.call("rbind", biomass_by_year)

  biomass_covs <- biomass_predictions %>% 
    select(X, Y, depth, trawled, muddy, any_rock, omega_s) %>% 
    distinct()

  biomass_change <- biomass_change %>% 
    rename(start_density = est_exp_s, end_density = est_exp_e) %>%
    dplyr::select(icell, x, y, start_time, start_density, end_density)
    # dplyr::select(-N, -time_step, - units_per_decade, -slope)

  biomass <- left_join(biomass_change, biomass_covs, by = c("x" = "X", "y"="Y"))

  # keep just one row for each cell that exceeded the end climate threshold 
  source_cells <- all_vectors %>% filter(distance > (-dist_intercept)) %>% 
    distinct(icell, start_time, .keep_all = TRUE) %>% 
    select(-target_X, -target_Y, -target_values, -tid)

  source_biomass <- left_join(source_cells, biomass, 
    by = c("icell", "x", "y", "start_time")) %>%
    mutate(cell_type = "source") %>% filter(!is.na(start_density))

  # identify all possible control cells have been given a distance == - dist_intercept
  control_cells <- all_vectors %>% filter(distance == (-dist_intercept)) %>% 
    distinct(icell, start_time, .keep_all = TRUE) %>% 
    mutate(distance = (- dist_intercept)) %>% 
    select(-target_X, -target_Y, -target_values, -tid)

  control_biomass <- left_join(control_cells, biomass, 
    by = c("icell", "x", "y", "start_time")) %>%
    mutate(cell_type = "control") %>% filter(!is.na(start_density))

  alldata <- rbind(source_biomass, control_biomass) 
  alldata <- mutate(alldata, treat = ifelse(cell_type=="source", 1, 0)) %>% 
    select(-var_2_min, -var_2_max, -mean_target_X, -mean_target_Y) %>% 
    rename(vect_dist = distance)
  alldata$treat <- as.factor(alldata$treat)

 # alldata
#})
```

# Select matched control cells

```{r}
#rm(vocc_by_sp)
# try({
#   vocc_by_sp <- readRDS(paste0(
#     "data/BACI-", species, "-", age, "-", ssid_string, "-", climate_string, "-", optimal_string, "-", min_string, "to", max_string, "-multiyear-all.rds"
#   ))
# })

#FIXME: should the cells be matched by substrate even when biomass not modelled using substrate?

match_formula <- paste0("treat ~ trawled + x + y + depth_st") 
  # depth.st + omega_s + muddy.st + any_rock.st + 
  for (j in (variable_names)) {
    match_formula <- paste0("", match_formula, " + ", j, "_s.st")
    start_value.st <- paste0("", j, "_s.st")
    start_value <- paste0("", j, "_s")

    alldata <- alldata %>% mutate((!!start_value.st) := 
        arm::rescale(alldata[[start_value]]) )
  #(UQ(rlang::sym(paste0(variable_names[j], "_s")))))
    }

if (!exists("vocc_by_sp")){
# matched <- lapply(alldata, function(alldata) {  
  ## trim non-overlapping values to reduce size of dataset 
  control_min <- quantile(alldata$start_density[alldata$treat=="0"], 0.01)
  source_min <- quantile(alldata$start_density[alldata$treat=="1"], 0.01)
  # control_max <- quantile(alldata$start_density[alldata$treat=="0"], 1)
  # source_max <- quantile(alldata$start_density[alldata$treat=="1"], 1)

  group_min <- max(control_min, source_min)
  # group_max <- min(control_max, source_max)
  alldata <- filter(alldata, start_density >= group_min) 
  #%>% filter(start_density <= group_max)

  # need to be standardized? seems not
  alldata$depth2 <- alldata$depth^2
  alldata$start_density.st <- round(arm::rescale(log(alldata$start_density))*5)/5
  alldata$depth_st <- arm::rescale(log(alldata$depth))
  alldata$depth.st <- round(arm::rescale(log(alldata$depth))*5)/5
  alldata$muddy.st <- arm::rescale(alldata$muddy) 
  alldata$any_rock.st <- arm::rescale(alldata$any_rock)

 ## Using optimization ... doesn't work   
  # alldata$start_density_st <- round(arm::rescale(log(alldata$start_density)), 1)
  # fit_formula <- 
  #   paste0("treat ~ start_density_st + depth.st + muddy.st + any_rock.st") 
  # # for (j in (variable_names)) {
  # #   fit_formula <- paste0("", fit_formula, " + ", j, "_s")
  # #   }
  # # options("optmatch_max_problem_size" = Inf)
  # matched <- MatchIt::matchit(as.formula(fit_formula), 
  #     method = "optimal", data = alldata)

  # # Using exact match and rounding
  # # Choose data order if not using random
  # alldata <- alldata[(order(alldata$y)), ]
  # if (order == "increasing") alldata <- alldata[rev(order(alldata$start_density)), ]
  # if (order == "decreasing") alldata <- alldata[rev(order(alldata$start_density)), ]
# set.seed(99999)
  matched <- MatchIt::matchit(as.formula(match_formula), 
    exact = c("start_density.st", "start_time"),
    method = "nearest", m.order = "random",
    distance = "probit",
    discard = "both",
    data = alldata)
  # matched
# })
}
```

Print diagnostics for matching
```{r echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
if (!exists("vocc_by_sp")){
match_plots <- plot(matched, interactive = F)
summary(matched)
}

```

Prep matched data for analysis
```{r}
if (!exists("vocc_by_sp")){
#vocc_by_sp <- lapply(matched, function(matched) {    
  mdata <- MatchIt::match.data(matched, distance = "pscore")

  sources <- mdata[row.names(matched$match.matrix), ]
  sources$origobs <- row.names(sources)   
  sources$matchobs <- matched$match.matrix[,1]

  controls <- mdata[matched$match.matrix, ]
  controls$origobs <- row.names(controls)   
  controls$matchobs <- matched$match.matrix[,1]

  mdata <- rbind(sources, controls) %>% tidyr::drop_na() %>% filter(matchobs!="-1")

  matched_time0 <- mdata %>% mutate(after = 0, density = start_density) 
  #%>% dplyr::select(-start_density, -end_density)
  matched_time1 <- mdata %>% mutate(after = 1, density = end_density) 
  #%>% dplyr::select(-start_density, -end_density)

  for (j in seq_along(variable_names)) {
    climate_var <- paste0(variable_names[j])
    matched_time0 <- matched_time0 %>% 
      mutate((!!climate_var) := (UQ(rlang::sym(paste0(variable_names[j], "_s")))))
    matched_time1 <- matched_time1 %>% 
      mutate((!!climate_var) := (UQ(rlang::sym(paste0(variable_names[j], "_e")))))
   }

  vocc_by_sp <- rbind(matched_time0, matched_time1)

  # calculate standardized climate variables
  for (j in seq_along(variable_names)) {
    mean.var <- paste0("mean.", variable_names[j])
    sd.var <- paste0("sd.", variable_names[j])
    std.var <- paste0("std_", variable_names[j])
    std.var2 <- paste0("std_", variable_names[j], "2")

    vocc_by_sp <- mutate(
      vocc_by_sp,
      (!!mean.var) := mean(UQ(rlang::sym(paste0(variable_names[j]))), na.rm = TRUE),
      (!!sd.var) := sd(UQ(rlang::sym(paste0(variable_names[j]))), na.rm = TRUE),
      (!!std.var) := ((UQ(rlang::sym(paste0(variable_names[j]))) - 
          mean(UQ(rlang::sym(paste0(variable_names[j]))), na.rm = TRUE)) / 
          (sd(UQ(rlang::sym(paste0(variable_names[j]))), na.rm = TRUE))*2),
      (!!std.var2) := (( UQ(rlang::sym(paste0(variable_names[j]))) - 
          mean(UQ(rlang::sym(paste0(variable_names[j]))), na.rm = TRUE) ) / 
          (sd(UQ(rlang::sym(paste0(variable_names[j]))), na.rm = TRUE)*2))^2
    )
  }

  # rename spatial coordinates to avoid naming conflict with y~x notation
  data <- vocc_by_sp %>%
    rename(X = x, Y = y)
    # %>% rename(distance = vect_dist) %>% 
    # select(-start_density.st, -depth.st, -muddy.st, -any_rock.st, -omega_s.st)

  # other variable prep
  data$log_density <-  log(data$density)
  data$log_depth <-  log(data$depth)
  data$std_log_depth <- arm::rescale(log(data$depth))
  data$vect_dist2 <- data$vect_dist^2
  data <- mutate(data, time = ifelse(after==1, "end", "begin"))
  data$matchobs <- as.factor(data$matchobs)
  data$origobs  <- as.factor(data$origobs)   
  data$start_time <- as.factor(data$start_time)

#  data
#})
vocc_by_sp <- list()
vocc_by_sp[[1]] <- data
vocc_by_sp[[3]] <- match_formula  
vocc_by_sp[[4]] <- match_plots
saveRDS(vocc_by_sp, file = paste0(
    "data/", spp,
    "/BACI-", spp, "-", age, "-", ssid_string, "-", climate_string, "-", optimal_string, "-", min_string, "to", max_string, "-multiyear-all.rds"
  ))
}
```

## Save spp dataframe
```{r}
#biomass_threshold <- "50"
optimal_string <- paste0(na.omit(trim_thresholds), collapse = "-")
#climate <- "temperature"
try({
  vocc_by_sp <- readRDS(paste0(
    "data/", spp,
    "/BACI-", spp, "-", age, "-", ssid_string, "-", climate_string, "-", optimal_string, "-", min_string, "to", max_string, "-multiyear-all.rds"
  ))
})

n <- nrow(vocc_by_sp[[1]])
species <- rep(species, n) 
mature <- rep(age, n)
ssid <- rep(ssid_string, n)
data <- vocc_by_sp[[1]]
df<- cbind(species, mature, ssid, data)
species <- species[[1]]

# folder to hold vocc data
biomass_threshold_string <- paste0("perc_", biomass_threshold)

dir.create(file.path("data/_all", climate_string, biomass_threshold_string))

if (climate=="temperature") {
dir.create(file.path("data/_all", climate_string, biomass_threshold_string, max_string))
dir.create(file.path("data/_all", climate_string, biomass_threshold_string, max_string, age))
}
if (climate=="temperature") {
  saveRDS(df, file = paste0("data/_all/", climate_string,"/", "perc_", biomass_threshold,"/", max_string,"/", age, "/vocc-", climate_string, "-", max_string, "-", spp, "-", age, "-", optimal_string, "-", ssid_string, "-all.rds"
  ))
}

if (climate=="do") {
  dir.create(file.path("data/_all", climate_string))  
  dir.create(file.path("data/_all", climate_string, biomass_threshold_string))  
  dir.create(file.path("data/_all", climate_string, biomass_threshold_string, min_string))
  dir.create(file.path("data/_all", climate_string, biomass_threshold_string, min_string, age))  
  }
if (climate=="do") {
  saveRDS(df, file = paste0("data/_all/", climate_string,"/", "perc_", biomass_threshold,"/", min_string,"/", age, "/vocc-", climate_string, "-", min_string, "to", max_string, "-", spp, "-", age, "-", optimal_string, "-", ssid_string, "-all.rds"
  ))
}

if (climate=="do and temperature") {
  dir.create(file.path("data/_all", climate_string))  
  dir.create(file.path("data/_all", climate_string, biomass_threshold_string))  
  dir.create(file.path("data/_all", climate_string, biomass_threshold_string, age))  
}

if (climate=="do and temperature") {
  saveRDS(df, file = paste0("data/_all/", climate_string, "/", "perc_", biomass_threshold,"/", age, "/vocc-", climate_string, "-", min_string, "to", max_string, "-", spp, "-", age, "-", optimal_string, "-", ssid_string, "-all.rds"
  ))
}
```

```{r}
try({
  vocc_by_sp <- readRDS(paste0(
    "data/", spp,
    "/BACI-", spp, "-", age, "-", ssid_string, "-", climate_string, "-",
    min_string, "to", max_string, "-multiyear-all.rds"
  ))
})
data <- vocc_by_sp[[1]]
```

### Plot matches
```{r}
#FIXME: SAVE PLOTS?

dat_summary <- data %>% 
  group_by(start_time, cell_type) %>%
  tally() %>% filter(cell_type=="source")
data_depth <- data %>% 
  group_by(start_time) %>% 
  mutate(mean_depth = mean(depth))

p1 <- ggplot(data, 
#ggplot(filter(data, as.numeric(matchobs) <= 550), 
  aes(y=Y, x=X, shape=as.factor(treat), colour=as.factor(matchobs))) +
  geom_point(alpha=0.8, size = 1) +
  scale_shape_manual(values= c(0, 15)) +
  scale_colour_viridis_d(option="C", guide = FALSE) +
  facet_wrap(~start_time) +
  theme_bw()
p1

lapply(variable_names, function(j) {
ggplot(data, aes(y=log_density, x=UQ(rlang::sym(paste(j))), colour=time)) +
  geom_point(alpha=0.35) +
  scale_colour_viridis_d(option="C") +
  facet_grid(start_time~cell_type) +
  theme_bw()
})

p2 <- ggplot(data, aes(std_log_depth, fill = cell_type)) +
  geom_density(alpha=0.5) +
  scale_fill_viridis_d(option="C") +
  facet_wrap(~start_time) +
  theme_bw()
p2

# ggplot(data, aes(y=log_density, x=depth, colour=cell_type)) +
#   geom_point(alpha=0.35) +
#   scale_colour_viridis_d(option="C") +
#     facet_grid(start_time~time) +
#   theme_bw()

p3 <- ggplot(data, aes(log_density, fill = cell_type)) +
  geom_density(alpha=0.5) +
  scale_fill_viridis_d(option="C") +
  facet_grid(start_time~time, scales = 'fixed') +
  theme_bw() + theme(legend.position='none')
p3

p4 <- ggplot(data, aes(log_density, fill = cell_type)) +
  geom_histogram(alpha=0.45) +
  scale_fill_viridis_d(option="C") +
  facet_grid(time~start_time) +
  theme_bw() + theme(legend.position='none')
p4

dat_summary <- data %>% 
  group_by(start_time, cell_type) %>% 
  tally() %>% filter(cell_type=="source")

p5 <- ggplot(data, aes(y=log_density, x=cell_type, fill=cell_type)) +
  geom_boxplot(alpha=0.5) +
  scale_fill_viridis_d(option="C") +
  geom_text(data = dat_summary,
            aes(cell_type, min(data$log_density)+1, label = n), hjust = 1.25) +
  facet_grid(time~start_time) +
  theme_bw() + theme(legend.position='none')
p5
```

```{r}
png(
  file = paste0("figs/", spp,
    "/matches-", min_string, "to", max_string, "-",
    climate_string, "-", spp, "-", age, "-", ssid_string, "-",
    climate_string, "-perc_", biomass_threshold, "-", optimal_string, ".png"
  ),
  res = 600,
  units = "in",
  width = 8.5,
  height = 11
)
gridExtra::grid.arrange(
  p1,p2,p3,p4,p5,
  ncol = 2,
  top = grid::textGrob(paste("Matched pairs for", climate, "threshold of", optimal_string, "\n based on", biomass_threshold, "percent of maximum predicted", species, "biomass"), gp=grid::gpar(fontsize=16))
)
dev.off()
```