### 2020-02-07 analysis/VOCC/biomass-trends.Rmd
rename with numbers ordering steps

---
title: "`r paste(params$species)` biomass trends and figures"
author: "Philina English"
date: "`r Sys.Date()`"
output: html_document
params: 
    species: "Pacific Ocean Perch" 
    #species: "Redbanded Rockfish"
    region: "All synoptic surveys"
    start_year: 2008
    bind_predictions: FALSE
    covs: "-trawled-ssid-reml"
    #covs: "-log-both-noAR1" #"-trawled-ssid" # covariate string used to name model rds 
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set( fig.width = 9, fig.height = 6.5, #fig.width=11, fig.height=8.5, 
                      #fig.path=paste0("figs/", spp, "/"),
                      echo=FALSE, warning=FALSE, message=FALSE)
library(dplyr)
library(ggplot2)
library(gfplot)
library(gfdata)
library(sdmTMB)
library(gfranges)
```

```{r params}
species <- params$species
region <- params$region
start_year <- params$start_year

bind_predictions <- params$bind_predictions
covs <- params$covs
priors <- FALSE

paste("region =", region)
paste("ssids modelled separately =", bind_predictions)
paste("model label =", covs)
paste("priors =", priors)

spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))

if (region == "All synoptic surveys") {
  survey <- c("SYN QCS", "SYN HS", "SYN WCVI", "SYN WCHG")
  model_ssid <- c(1, 3, 4, 16)
  ssid_string <- paste0(model_ssid, collapse = "n")
}

if (region == "both odd year surveys") {
  survey <- c("SYN QCS", "SYN HS")
  model_ssid <- c(1, 3)
  ssid_string <- paste0(model_ssid, collapse = "n")
}

if (region == "West Coast Vancouver Island") {
  survey <- c("SYN WCVI")
  model_ssid <- c(4)
  ssid_string <- paste0(model_ssid, collapse = "n")
}

if (region == "West Coast Haida Gwaii") {
  survey <- c("SYN WCHG")
  model_ssid <- c(16)
  ssid_string <- paste0(model_ssid, collapse = "n")
}
```

Load model predictions
```{r results = "hide"}
rm(adult_predictions)
rm(imm_predictions)
rm(predictions)

rm(pred1n3)
rm(pred4)
rm(pred16)

rm(pred1n3_imm)
rm(pred4_imm)
rm(pred16_imm)

rm(pred1n3_all)
rm(pred4_all)
rm(pred16_all)

if (region != "All synoptic surveys") {
  try(adult_predictions <- readRDS(paste0("data/", spp,
    "/predictions-", spp, covs, "-1n3n4n16-mature-biomass-prior-", priors, ".rds"
  )) %>% filter(ssid %in% model_ssid))
  try(imm_predictions <- readRDS(paste0("data/", spp,
    "/predictions-", spp, covs, "-1n3n4n16-imm-biomass-prior-", priors, ".rds"
  )) %>% filter(ssid %in% model_ssid))

  if (!exists("adult_predictions")) {
    paste("Maturity split biomass predictions are not available for this survey.")

    try(predictions <- readRDS(paste0("data/", spp,
      "/predictions-", spp, covs, "-1n3n4n16-total-biomass-prior-", priors, ".rds"
    )) %>% filter(ssid %in% model_ssid))

    if (!exists("predictions")) {
      paste("Total biomass predictions for this survey are also not available.")
    }
  }
} else {

  try({adult_predictions <- readRDS(paste0("data/", spp,
    "/predictions-", spp, covs, "-", ssid_string, "-mature-biomass-prior-", priors, ".rds"
    ))
  })
  try({imm_predictions <- readRDS(paste0("data/", spp,
    "/predictions-", spp, covs, "-", ssid_string, "-imm-biomass-prior-", priors, ".rds"
    ))
  })

  # OR bind together predictions from different models
  if (bind_predictions) {
    try({pred1n3 <- readRDS(paste0("data/", spp,
      "/predictions-", spp, covs, "-1n3-mature-biomass-prior-", priors, ".rds"
      ))
    })
    try({pred4 <- readRDS(paste0("data/", spp,
      "/predictions-", spp, covs, "-4-mature-biomass-prior-", priors, ".rds"
      ))
    })
    try({pred16 <- readRDS(paste0("data/", spp,
      "/predictions-", spp, covs, "-16-mature-biomass-prior-", priors, ".rds"
      ))
    })

    if (!exists("pred1n3")) pred1n3 <- NULL
    if (!exists("pred16")) pred16 <- NULL
    if (!exists("pred4")) pred4 <- NULL
    adult_predictions <- rbind(pred1n3, pred4, pred16)

    try({pred1n3_imm <- readRDS(paste0("data/", spp,
      "/predictions-", spp, covs, "-1n3-imm-biomass-prior-", priors, ".rds"
      ))
    })
    try({pred4_imm <- readRDS(paste0("data/", spp,
      "/predictions-", spp, covs, "-4-imm-biomass-prior-", priors, ".rds"))
    })
    try({pred16_imm <- readRDS(paste0("data/", spp,
        "/predictions-", spp, covs, "-16-imm-biomass-prior-", priors, ".rds"))
    })

    if (!exists("pred1n3_imm")) pred1n3_imm <- NULL
    if (!exists("pred16_imm")) pred16_imm <- NULL
    if (!exists("pred4_imm")) pred4_imm <- NULL
    imm_predictions <- rbind(pred1n3_imm, pred4_imm, pred16_imm)
  }
  }

  if (!exists("adult_predictions")) {
      paste("Maturity split biomass predictions are not available.")

      try({predictions <- readRDS(paste0("data/", spp,
        "/predictions-", spp, covs, "-", ssid_string, "-total-biomass-prior-", priors, ".rds"))
      })

      if (!exists("predictions")) {
        try({pred1n3_all <- readRDS(paste0("data/", spp,
          "/predictions-", spp, covs, "-1n3-total-biomass-prior-", priors, ".rds"))
        })
        try({pred4_all <- readRDS(paste0("data/", spp,
          "/predictions-", spp, covs, "-4-total-biomass-prior-", priors, ".rds"))
        })
        try({pred16_all <- readRDS(paste0("data/", spp,
          "/predictions-", spp, covs, "-16-total-biomass-prior-", priors, ".rds"))
        })
        try({
          predictions <- rbind(pred1n3_all, pred4_all, pred16_all)
        })

        if (!exists("predictions")) paste("Total biomass predictions are not available.")
      }
    }
```

Caluclate variables of interest from predictions
```{r}
if (exists("adult_predictions")) {
  # to convert from kg/m2 to kg/hectare multiply by 10000
  adult_predictions$est_exp <- exp(adult_predictions$est) * 10000
  imm_predictions$est_exp <- exp(imm_predictions$est) * 10000

  adult_predictions$total_bio <- imm_predictions$est_exp + adult_predictions$est_exp
  imm_predictions$prop_imm <- 
  imm_predictions$est_exp / (adult_predictions$est_exp + imm_predictions$est_exp)

  saveRDS(adult_predictions, file = paste0(
    "data/", spp,
    "/all-predictions-", spp, covs, "-mat-biomass-prior-", priors, ".rds"
  ))
  saveRDS(imm_predictions, file = paste0(
    "data/", spp,
    "/all-predictions-", spp, covs, "-imm-biomass-prior-", priors, ".rds"
  ))

  adult_predictions <- filter(adult_predictions, year >= start_year)
  imm_predictions <- filter(imm_predictions, year >= start_year)

  max_raster <- quantile(adult_predictions$est_exp, 0.999)
  max_adult <- signif(max(adult_predictions$est_exp), digits = 2)

  model_ssid <- unique(adult_predictions$ssid)
} else {
  predictions <- filter(predictions, year >= start_year)

  predictions$est_exp <- exp(predictions$est) * 10000
  predictions$total_bio <- exp(predictions$est) * 10000
  max_raster <- quantile(predictions$total_bio, .999)
  #max_bio <- round(max(predictions$total_bio))
  max_bio <- signif(max(predictions$est_exp), digits = 2)

  saveRDS(predictions, file = paste0(
    "data/", spp,
    "/all-predictions-", spp, covs, "-total-biomass-prior-", priors, ".rds"
  ))

  model_ssid <- unique(predictions$ssid)
}
```

## Mature (or total) biomass predictions
```{r, error=TRUE}
if (exists("adult_predictions")) {
  p_adult_all <- plot_facet_map(adult_predictions, "est_exp",
    raster_limits = c(0, max_raster),
    legend_position = "none",
    transform_col = fourth_root_power
  ) +
    labs(fill = "kg/ha") +
    ggtitle(paste0("", species, " mature biomass \n(max = ", max_adult, " kg/ha)"))
  print(p_adult_all)
} else {
  p_adult_all <- plot_facet_map(predictions, "total_bio",
    transform_col = fourth_root_power
  ) +
    labs(fill = "kg/ha") +
    ggtitle(paste0("", species, " total biomass \n(max = ", max_bio, " kg/ha)"))

  print(p_adult_all)
}
```

### Split by survey (SSID)
```{r}
if (exists("adult_predictions")) predictions <- adult_predictions

  if (1 %in% model_ssid) {
    adult_predictions1 <- predictions %>% filter(ssid == 1)
  } else {
    adult_predictions1 <- NULL
  }

  if (3 %in% model_ssid) {
    adult_predictions3 <- predictions %>%
      filter(ssid != 16) %>%
      filter(ssid != 4)
  } else {
    adult_predictions3 <- NULL
  }

  if (4 %in% model_ssid) {
    adult_predictions4 <- predictions %>% filter(ssid == 4)
  } else {
    adult_predictions4 <- NULL
  }

  if (16 %in% model_ssid) {
    if (any(predictions$ssid == 16)) {
      adult_predictions16 <- predictions %>% filter(ssid == 16)
    } else {
      adult_predictions16 <- NULL
    }
  } else {
    adult_predictions16 <- NULL
  }

  ad_prediction_list <- list(adult_predictions1, adult_predictions3, adult_predictions4, adult_predictions16)
  ad_prediction_list <- ad_prediction_list[!sapply(ad_prediction_list, is.null)]

  p_adult <- list()
  if(exists("adult_predictions")) {
  for (i in seq_len(length(model_ssid))) {
    p_adult[[i]] <- plot_facet_map(ad_prediction_list[[i]], "est_exp",
      raster_limits = c(0, max_raster),
      # legend_position = "none"
      transform_col = fourth_root_power
    ) +
      labs(fill = "kg/ha") +
      ggtitle(paste0("", species, " mature biomass \n(max = ", max_adult, " kg/ha)"))
    p_adult[[i]]
  }
  } else {
    for (i in seq_len(length(model_ssid))) {
    p_adult[[i]] <- plot_facet_map(ad_prediction_list[[i]], "est_exp",
      raster_limits = c(0, max_raster),
      # legend_position = "none"
      transform_col = fourth_root_power
    ) +
      labs(fill = "kg/ha") +
      ggtitle(paste0("", species, " total biomass \n(max = ", max_bio, " kg/ha)"))
    p_adult[[i]]  
    }
  }
  print(p_adult)
```

```{r, message=FALSE, warning=FALSE}
rm(out_mature)
rm(out_total)
# try(out_mature <- readRDS(paste0("data/", spp, "/trends-", spp, covs, "-", ssid_string, "-mature-biomass-priors-", priors, ".rds")))

if (!exists("out_mature")) {
  if (exists("adult_predictions")) {
    predictions <- adult_predictions
  } else {
   try(out_total <- readRDS(paste0("data/", spp, "/trends-", spp, covs, "-", ssid_string, "-total-biomass-priors-", priors, ".rds")))
  }

  if (!exists("out_total")) {
    if (1 %in% model_ssid) {

      .predictions <- filter(predictions, ssid == 1)
      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)

      out_mature1 <- make_trend_data(predictions,
        ssid = c(1),
        start_time = start_year,
        skip_time = 2004,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = indices,
        variable_names = "est_exp"
      )
    } else {
      out_mature1 <- NULL
    }

    if (3 %in% model_ssid) {
      .predictions <- filter(predictions, ssid == 3)
      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)

      out_mature3 <- make_trend_data(predictions,
        ssid = c(3),
        start_time = start_year,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = indices,
        variable_names = "est_exp"
      )
    } else {
      out_mature3 <- NULL
    }

    if (4 %in% model_ssid) {

      .predictions <- filter(predictions, ssid == 4)
      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)

      out_mature4 <- make_trend_data(predictions,
        ssid = c(4),
        start_time = start_year,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = indices,
        variable_names = "est_exp"
      )
    } else {
      out_mature4 <- NULL
    }

    if (16 %in% model_ssid) {

      .predictions <- filter(predictions, ssid == 16)
      #unique(.predictions$year)
      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)

      out_mature16 <- make_trend_data(predictions,
        ssid = c(16),
        start_time = start_year,
        skip_time = 2007,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = indices,
        variable_names = "est_exp"
      )
    } else {
      out_mature16 <- NULL
    }

    out_mature <- rbind(out_mature1, out_mature3, out_mature4, out_mature16)

    if (exists("adult_predictions")) {
            saveRDS(out_mature, file = paste0("data/", spp, "/trends-", 
        spp, covs, "-", ssid_string, "-mature-biomass-priors-", priors, ".rds"))

         } else {
      out_total <- out_mature
      rm(out_mature)
      saveRDS(out_total, file = paste0("data/", spp, "/trends-", 
        spp, covs, "-", ssid_string, "-total-biomass-priors-", priors, ".rds"))
    }
  }
}
```

### Recent biomass with trawl footprint
```{r }
if (!exists("out_mature")) {out_recent <- out_total} else {out_recent <- out_mature}

  trawl_footprint <- sf::st_read(dsn = "data/trawl-footprint", layer = "Trawl_footprint")
  proj.to <- "+proj=utm +zone=9 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs"
  trawl_footprint <- sf::st_transform(trawl_footprint, crs = proj.to)

  current_biomass <- plot_vocc(out_recent,
    fill_col = "est_exp_e",
    fill_label = "Mean\nbiomass (kg/ha)\npost-2010",
    raster_alpha = 1,
    viridis_option = "C",
    vec_aes = NULL,
    transform_col = fourth_root_power
  )
  current_biomass <- current_biomass +
    geom_sf(data = trawl_footprint, aes(geometry = geometry / 1000), colour = "gray", fill = "gray", alpha = 0.2, lwd = 0.25, inherit.aes = FALSE) +
    coord_sf(xlim = c(180, 820), ylim = c(5350, 6050), datum = NULL) +
    xlab("") +
    ylab("") +
    theme(panel.grid.major = element_line(colour = "transparent")) +
    ggtitle(paste("Predicted", species, "density with trawl footprint overlain"))
  current_biomass

```

## Trends in biomass density
```{r, error=TRUE}
if (exists("out_mature")) {
  data <- out_mature
  biotrend_ad <- plot_vocc(data,
    fill_col = "units_per_decade",
    fill_label = "Mature \nbiomass \ntrend \n(kg/ha/decade)",
    raster_alpha = 1,
    vec_aes = NULL,
    white_zero = TRUE,
    high_fill = "deepskyblue3", # "lightseagreen", # limegreen",
    low_fill = "red", # "deeppink3",# "steel blue 4",
    transform_col = fourth_root_power,
    legend_position = c(0.2, 0.25)
  ) + ggtitle(paste0(
    "", species, " ", min(unique(adult_predictions[["year"]])), "-", max(unique(adult_predictions[["year"]]))
  ))
  biotrend_ad
} else {
  biotrend_ad <- plot_vocc(out_total,
    fill_col = "units_per_decade",
    fill_label = "Biomass \ntrend \n(kg/ha/decade)",
    raster_alpha = 1,
    vec_aes = NULL,
    white_zero = TRUE,
    high_fill = "deepskyblue3", # "greenyellow", # limegreen",
    low_fill = "red", # "steel blue 4",
    transform_col = fourth_root_power,
    legend_position = c(0.2, 0.25)
  ) + ggtitle(paste0(
    "", species, " ", min(unique(predictions[["year"]])), "-", max(unique(predictions[["year"]]))
  ))
  biotrend_ad
}
```

# Immature biomass density
### if available
```{r, error=TRUE}
if (exists("imm_predictions")) {
  p_imm_all <- plot_facet_map(imm_predictions, "est_exp",
    transform_col = fourth_root_power,
    raster_limits = c(0, max_raster)
  ) +
    labs(fill = "kg/ha") +
    ggtitle(paste0("", species, " immature biomass \n "))

  p_imm_all
} else {
  p_imm_all <- grid::grid.rect(gp = grid::gpar(col = "white"))
}
```

```{r}
if (exists("adult_predictions")) {
  if (1 %in% model_ssid) {
    imm_predictions1 <- imm_predictions %>% filter(ssid == 1)
  } else {
    imm_predictions1 <- NULL
  }

  if (3 %in% model_ssid) {
    imm_predictions3 <- imm_predictions %>%
      filter(ssid != 16) %>%
      filter(ssid != 4)
  } else {
    imm_predictions3 <- NULL
  }

  if (4 %in% model_ssid) {
    imm_predictions4 <- imm_predictions %>% filter(ssid == 4)
  } else {
    imm_predictions4 <- NULL
  }

  if (16 %in% model_ssid) {
    imm_predictions16 <- imm_predictions %>% filter(ssid == 16)
  } else {
    imm_predictions16 <- NULL
  }

  im_prediction_list <- list(imm_predictions1, imm_predictions3, imm_predictions4, imm_predictions16)
  im_prediction_list <- im_prediction_list[!sapply(im_prediction_list, is.null)]

  p_imm <- list()
  for (i in seq_len(length(model_ssid))) {
    p_imm[[i]] <- plot_facet_map(im_prediction_list[[i]], "est_exp",
      raster_limits = c(0, max_raster),
      # legend_position = "none",
      transform_col = fourth_root_power
    ) +
      labs(fill = "kg/ha") +
      ggtitle(paste0("", species, " immature biomass \n(max = ", max_adult, " kg/ha)"))

    p_imm[[i]]
  }
  print(p_imm)
}
```

```{r, message=FALSE, warning=FALSE}
# try(out_imm <- readRDS(paste0("data/", spp, "/trends-", spp, covs, "-", ssid_string, "-immature-biomass-priors-", priors, ".rds")))
rm(out_imm)
if (!exists("out_imm")) {
  if (exists("imm_predictions")) {
    predictions <- imm_predictions

    if (1 %in% model_ssid) {

      .predictions <- filter(predictions, ssid == 1)
      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)
      out_im1 <- make_trend_data(predictions,
        ssid = c(1),
        start_time = start_year,
        skip_time = 2004,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = indices,
        variable_names = "est_exp"
      )
    } else {
      out_im1 <- NULL
    }

    if (3 %in% model_ssid) {
      .predictions <- filter(predictions, ssid == 3)
      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)

      out_im3 <- make_trend_data(predictions,
        ssid = c(3),
        start_time = start_year,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = indices,
        variable_names = "est_exp"
      )
    } else {
      out_im3 <- NULL
    }

    if (4 %in% model_ssid) {
      .predictions <- filter(predictions, ssid == 4)
      unique(.predictions$year)
      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)

      out_im4 <- make_trend_data(predictions,
        ssid = c(4),
        start_time = start_year,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = indices,
        variable_names = "est_exp"
      )
    } else {
      out_im4 <- NULL
    }

    if (16 %in% model_ssid) {
      .predictions <- filter(predictions, ssid == 16)
      if(length(unique(.predictions$year))==5) indices <- c(1, 1, 1, 2, 2)
      if(length(unique(.predictions$year))==6) indices <- c(1, 1, 1, 2, 2, 2)
      if(length(unique(.predictions$year))==7) indices <- c(1, 1, 1, 2, 2, 2, 2)
      if(length(unique(.predictions$year))==8) indices <- c(1, 1, 1, 1, 2, 2, 2, 2)

      out_im16 <- make_trend_data(predictions,
        ssid = c(16),
        start_time = start_year,
        input_cell_size = 2,
        scale_fac = 1,
        delta_t_total = 10,
        delta_t_step = 2,
        indices = indices,
        variable_names = "est_exp"
      )
    } else {
      out_im16 <- NULL
    }

    out_imm <- rbind(out_im1, out_im3, out_im4, out_im16)
    saveRDS(out_imm, file = paste0(
      "data/", spp, "/trends-", spp, covs, "-", ssid_string, "-immature-biomass-priors-", priors, ".rds"
    ))
  } else {
    rm(out_imm)
  }
}
```

```{r}
if (exists("out_imm")) {
  data <- out_imm
  biotrend_im <- plot_vocc(data,
    fill_col = "units_per_decade",
    fill_label = "Immature \nbiomass \ntrend \n(kg/ha/decade)",
    raster_alpha = 1,
    vec_aes = NULL,
    white_zero = TRUE,
    high_fill = "deepskyblue3", # "greenyellow", # limegreen",
    low_fill = "red", # "steel blue 4",
    transform_col = fourth_root_power,
    legend_position = c(0.2, 0.25)
  ) + ggtitle(paste0(
    "", species, " ", min(unique(imm_predictions[["year"]])), "-", max(unique(imm_predictions[["year"]]))
  ))
  biotrend_im
} else {
  biotrend_im <- grid::grid.rect(gp = grid::gpar(col = "white"))
}
```

```{r, error=TRUE}
if (exists("imm_predictions")) {
  p_prop_imm <- plot_facet_map(imm_predictions, "prop_imm") +
    labs(fill = "% \nby mass") +
    ggtitle(paste0("", species, " proportion immature"))
  print(p_prop_imm)
} else {
  p_prop_imm <- grid::grid.rect(gp = grid::gpar(col = "white"))
}
```

Save trend plots
```{r results = "hide"}
png(
  file = paste0("figs/", spp, "/trends-", spp, covs, "-", ssid_string, "-priors-", priors, ".png"),
  res = 600,
  units = "in",
  width = 11,
  height = 8.5
)
gridExtra::grid.arrange(
  # p_adult, p_imm,
  biotrend_ad, biotrend_im,
  nrow = 1
)
dev.off()
```

Save spatiotemporal biomass plots
```{r results = "hide"}
if(exists("adult_predictions")) { biomass_plots <- c(p_adult, p_imm) } else {
  biomass_plots <- c(p_adult)
}

png(
  file = paste0("figs/", spp, "/biomass-", spp, covs, "-", ssid_string, "-priors-", priors, ".png"),
  res = 600,
  units = "in",
  width = 16,
  height = 8.5
)
gridExtra::grid.arrange(
  grobs = biomass_plots,
  nrow = 2
)
dev.off()
```

Save plot medley
```{r results = "hide"}
png(
  file = paste0("figs/", spp, "/exploratory-", spp, covs, "-", ssid_string, "-priors-", priors, ".png"),
  res = 600,
  units = "in",
  width = 9.5,
  height = 12.5
)

gridExtra::grid.arrange(
  current_biomass,
  p_prop_imm,
  p_adult_all, p_imm_all,
  biotrend_ad, biotrend_im,
  nrow = 3
)
dev.off()
```

```{r}
plot_map_raster <- function(dat, column = "est") {
    ggplot(dat, aes_string("X", "Y", fill = column)) +
    geom_raster() +
    coord_fixed()+
gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank())
}

 # plot_map_raster(adult_predictions, "epsilon_st") +
 #  ggtitle("Spatiotemporal random effects only") +
 #  facet_wrap(~year) +
 #  scale_fill_gradient2( high = "Steel Blue 4",
 #  low = "Red 3") + coord_fixed()

 if (exists("adult_predictions")) {
  p_st <- list()
  for (i in seq_len(length(model_ssid))) {
    p_st[[i]] <- plot_map_raster(ad_prediction_list[[i]], "epsilon_st") +
      ggtitle("Spatiotemporal random effects only") +
      facet_wrap(~year) +
      scale_fill_gradient2(high = "#66C2A5", #mid = "black", 
        low = "#5E4FA2"#, trans = fourth_root_power
      )
    p_st[[i]]
   }
  print(p_st)

  # p_adult_st <- plot_facet_map(adult_predictions, "epsilon_st",
  #   white_zero = TRUE
  #   #raster_limits = c(0, max_raster),
  #   #legend_position = "none",
  #   #transform_col = fourth_root_power
  # ) +
  #   #labs(fill = "kg/ha") +
  #   ggtitle(paste0("", species, "mature spatiotemporal variation"))
  # print(p_adult_st)
} else {
  # p_adult_st <- plot_facet_map(predictions, "epsilon_st",
  #   transform_col = fourth_root_power
  # ) +
  #   #labs(fill = "kg/ha") +
  #   ggtitle(paste0("", species, "total spatiotemporal variation"))
  # 
  # print(p_adult_st)
}
```

Save spatiotemporal residual plots
```{r results = "hide"}
# if(exists("adult_predictions")) { resid_plots <- c(p_st, p_imm) } else {
#   biomass_plots <- c(p_adult)
# }

png(
  file = paste0("figs/", spp, "/biotic-residuals-", spp, covs, "-", ssid_string, "-priors-", priors, ".png"),
  res = 600,
  units = "in",
  width = 16,
  height = 8.5
)
gridExtra::grid.arrange(
  grobs = p_st,
  nrow = 2
)
dev.off()
```