### 2020-05-15 analysis/VOCC/1-split-biomass-by-maturity.Rmd
add option for sample_id_re to maturity splits with year effect

---
title: "Split `r paste(params$species)` biomass by maturity class"
author: "Philina English"
date: "`r Sys.Date()`"
output: html_document
params: 
    #species: "Bocaccio"
    species: "Threadfin Sculpin"
    #species: "Quillback Rockfish"
    #species: "Yelloweye Rockfish"
    #species: "Shortraker Rockfish"
    # species: "Pacific Ocean Perch"
    region: "All synoptic surveys"
    # region: "Both odd year surveys"
    threshold: 0.05
    split: FALSE
---

```{r global_options, include=FALSE}
species <- params$species
region <- params$region
split <- params$split
threshold <- params$threshold

spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))

knitr::opts_chunk$set(fig.width=11, fig.height=8.5, 
                      fig.path=paste0("figs/", spp, "/"),
                      echo=FALSE, warning=FALSE, message=FALSE)
library(dplyr)
library(ggplot2)
library(gfplot)
library(gfdata)
library(sdmTMB)
library(gfranges)
```

```{r}
spp <- gsub(" ", "-", gsub("\\/", "-", tolower(species)))
paste("region =", region)
paste("threshold =", threshold)
paste("split =", split)

# folder to hold figs for this species
dir.create(file.path("figs", spp))
dir.create(file.path("data", spp))

if (region == "Both odd year surveys") {
  survey <- c("SYN QCS", "SYN HS")
  model_ssid <- c(1, 3)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "West Coast Vancouver Island") {
  survey <- c("SYN WCVI")
  model_ssid <- c(4)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "West Coast Haida Gwaii") {
  survey <- c("SYN WCHG")
  model_ssid <- c(16)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}

if (region == "All synoptic surveys") {
  survey <- c("SYN QCS", "SYN HS", "SYN WCVI", "SYN WCHG")
  model_ssid <- c(1, 3, 4, 16)
  ssid_string <- paste0(model_ssid, collapse = "n")
  years <- NULL
}
```

# Split biomass by maturity

#### Calculate maturity weighted biomass if not done before
```{r}
rm(biomass)
rm(maturity)

#try(biomass <- readRDS(paste0("data/", spp, "/data-by-maturity-", spp, "-", ssid_string, ".rds")))

if (!exists("biomass")) {
  survey_sets <- readRDS(paste0("raw/event-data-", spp, ""))
  fish <- readRDS(paste0("raw/bio-data-", spp, ""))
  bath <- readRDS("data/bathymetry-data")

  # for POP: exclude sample with unusually large fish 
  # sizes are impossible given recorded catch weight
  # fish <- fish[fish$fishing_event_id!=1506954,]

  # for Redbanded: 1 fish found at too shallow a depth... 
  # probably leftover in net?
  # survey_sets <- survey_sets[survey_sets$fishing_event_id!=2536031,]
  # fish <- fish[fish$fishing_event_id!=2536031,]

 try({ 
   maturity <- split_catch_maturity(survey_sets, fish, bath,
    survey = survey,
    years = years,
    year_re = TRUE,
    # sample_id_re = TRUE,
    cutoff_quantile = c(.9995),
    p_threshold = threshold,
    plot = TRUE)
   })

  if(is.null(maturity$maturity)) {
  maturity <- split_catch_maturity(survey_sets, fish, bath,
    survey = c("SYN QCS","SYN HS","SYN WCVI","SYN WCHG"),
    years = years,
    year_re = TRUE,
    # sample_id_re = TRUE,
    cutoff_quantile = c(.9995),
    p_threshold = threshold,
    plot = TRUE
  )
  }

  if (split) {
  maturity$data <- filter(maturity$data, ssid %in% model_ssid)

  if (!is.null(maturity$maturity)) {
  saveRDS(maturity$model, file = paste0("data/", spp, 
    "/maturity-ogive-", spp, "-", ssid_string, ".rds"
    ))
  }
  } else {
  if (is.null(maturity$maturity)) {

    maturity$data <- filter(maturity$data, ssid %in% model_ssid)

  } else { # if maturity model is deemed unuseful
    maturity$data <- filter(maturity$data, ssid %in% model_ssid) %>% 
      select(-adult_density, -imm_density) 
  }
  }

  saveRDS(maturity$data, file = paste0("data/", spp, 
    "/data-by-maturity-", spp, "-", ssid_string, ".rds"
    ))
}

positive_sets <- filter(survey_sets, density_kgpm2 != 0)
prop_pos <- round(nrow(positive_sets)/nrow(survey_sets), digits = 3)

paste("proportion of positive sets =", prop_pos)
```

#### Save length~weight plots
```{r}
if (exists("maturity")) {
  print(maturity$mass_model)
  ggsave(
    file = paste0("figs/", spp, "/", spp, "-mass-model-", ssid_string, ".png"),
    dpi = 600,
    width = 8, 
    height = 10
  )
}
```

#### Save length at maturity plot 
```{r}
  png(
    file = paste0("figs/_maturity_ogives/maturity-", prop_pos,"-", spp, "-", ssid_string, ".png"), 
    res = 600,
    units = "in",
    width = 8, 
    height = 6
  )
  if (exists("maturity")) {
  print(maturity$maturity)
  } else {
  ggplot(fish, aes(length)) + geom_histogram(bins=50) + facet_grid(~sex)  
    }

  dev.off()

try(print(maturity$maturity))
```