### 2020-06-19 analysis/tmb-sensor-explore/2-fit-temperature.Rmd
fitting climate models

---
output: html_document
editor_options: 
  chunk_output_type: inline
---
# Fit spatial temporal models to bottom temperature

```{r setup}
library(tidyverse)
library(sdmTMB)
library(future)
library(dplyr)
```

Load data if not using SST
```{r}
all_sensor <- readRDS(here::here("analysis/tmb-sensor-explore/data/dat-sensor-trawl-processed.rds")) # no SST

all_depth <- all_sensor %>%
  # filter(depth_max>10) %>% # maybe we should filter incase missing depth values also indicate a sensor fail?
  # filter missing location data and trial year
  dplyr::filter(!is.na(latitude), !is.na(longitude)) %>%

  # convert lat and lon to UTMs
  dplyr::mutate(X = longitude, Y = latitude) %>%
  gfplot:::ll2utm(., utm_zone = 9) %>%

  # interpolate missing depth values
  dplyr::rename(depth = depth_m) %>%
  gfplot:::interp_survey_bathymetry()

d_trawl <- all_depth$data %>% 
  # remove obvious sensor fails, d_trawl %>% filter(depth_max<10), fishing event ids = 481861, 2179089
  filter(depth_max>10) %>% 
  gfplot:::scale_survey_predictors()

d_trawl <- d_trawl[!is.na(d_trawl$depth), ]
d_trawl <- d_trawl[!is.na(d_trawl$temperature_c), ]

d_trawl <- mutate(d_trawl,
  DOY = lubridate::yday(date),
  DOY_scaled = arm::rescale(lubridate::yday(date)),
  DOY_scaled2 = arm::rescale(lubridate::yday(date))^2)
```

EXLORE TEMPERATURE at DEPTH
```{r}
ggplot(d_trawl, aes_string("X", "Y", colour = "temperature_c")) +
    geom_point() +
    facet_wrap(~year) +
    coord_fixed() +
    scale_color_viridis_c()
```

```{r}
ggplot(d_trawl, aes(DOY, temperature_c, colour = -depth, alpha=0.5)) + geom_point() + facet_wrap(~year) 
```

Current temperature model
```{r}
spde <- make_spde(d_trawl$X, d_trawl$Y, n_knots = 500)
plot_spde(spde)
```

```{r eval = F}
m_temp_rw <- sdmTMB(
  temperature_c ~ 0 + as.factor(year), #+ DOY_scaled + as.factor(ssid)
  time_varying = ~ 0 + depth_scaled + depth_scaled2,
  data = d_trawl,
  ar1_fields = TRUE, # changed AR1_field to true
  include_spatial = TRUE, 
  time = "year", spde = spde, family = gaussian(link = "identity"),
  silent = FALSE
)

saveRDS(m_temp_rw, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-500kn.rds")))
```

Check predictions against raw data
```{r}
predictions <- predict(m_temp_rw)
predictions$residuals <- residuals(m_temp_rw)
cor(predictions$est, predictions$temperature_c)^2 
```

Model diagnostics functions
```{r}
get_diag <- function(m, response = "log_do", variable = "depth_scaled") {
  predictions <- predict(m)
  predictions$residuals <- residuals(m)

  plot_map <- function(dat, column = "est") {
    ggplot(dat, aes_string("X", "Y", colour = column)) +
      geom_point() +
      coord_fixed()
  }

  g <- plot_map(predictions, "est") +
    scale_colour_viridis_c() +
    ggtitle("Prediction (fixed effects + all random effects)")
  print(g)

  g <- plot_map(predictions, "est_non_rf") +
    ggtitle("Prediction (fixed effects only)") +
    scale_colour_viridis_c()
  print(g)

  g <- plot_map(predictions, "est_rf") +
    ggtitle("All random effects only") +
    scale_colour_gradient2()
  print(g)

  g <- plot_map(predictions, "omega_s") +
    ggtitle("Spatial random effects only") +
    scale_colour_gradient2()
 print(g)

   g <- plot_map(predictions, "epsilon_st") +
    ggtitle("Spatiotemporal random effects only") +
    facet_wrap(~year) +
    scale_colour_gradient2()
 print(g)

  g <- ggplot(predictions, aes_string("est", response)) +
    geom_point() +
    facet_wrap(~year) +
    coord_fixed() +
    geom_abline()
  print(g)

  g <- ggplot(predictions, aes(est, residuals)) +
    geom_point() +
    geom_smooth()
  print(g)

  g <- ggplot(predictions, aes_string(variable, "residuals")) +
    geom_point() +
    geom_smooth() +
    facet_wrap(~year)
  print(g)

  g <- ggplot(predictions, aes(X, Y, colour = residuals)) +
    geom_point() +
    coord_fixed() +
    scale_color_gradient2()+
    facet_wrap(~year)
  print(g)

  aic <- function(m) {
    k <- length(m$model$par)
    nll <- m$model$objective
    2 * k - 2 * (-nll)
  }

  print("R^2:")
  r2 <- cor(predictions$est, predictions[[response]])^2
  print(r2)

  print("")
  print("AIC:")
  print(aic(m))
}
```

```{r eval=FALSE}
m_temp_rw <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-500kn.rds")))
get_diag(m_temp_rw, response = "temperature_c", variable = "depth_scaled")
```

WOULD SST HELP?

Load and prep SST data
```{r eval=FALSE}
all_sensor <- readRDS(here::here("analysis/tmb-sensor-explore/data/dat-sensor-trawl-meanSST-all.rds"))
# all_sensor <- readRDS(here::here("analysis/tmb-sensor-explore/data/dat-sensor-trawl-processed.rds")) # no SST
# glimpse(all_sensor)

all_depth <- all_sensor %>%
  # filter(depth_max>10) %>% # maybe we should filter incase missing depth values also indicate a sensor fail?
  # filter missing location data and trial year
  dplyr::filter(!is.na(latitude), !is.na(longitude)) %>%

  # convert lat and lon to UTMs
  dplyr::mutate(X = longitude, Y = latitude) %>%
  gfplot:::ll2utm(., utm_zone = 9) %>%

  # interpolate missing depth values
  dplyr::rename(depth = depth_m) %>%
  gfplot:::interp_survey_bathymetry()

d_trawl <- all_depth$data %>% 
  # remove obvious sensor fails, d_trawl %>% filter(depth_max<10), fishing event ids = 481861, 2179089
  filter(depth_max>10) %>% 
  gfplot:::scale_survey_predictors()

d_trawl <- d_trawl[!is.na(d_trawl$depth), ]
d_trawl <- d_trawl[!is.na(d_trawl$temperature_c), ]

d_trawl2 <- d_trawl[!is.na(d_trawl$SST), ]
d_trawl2 <- d_trawl2[!is.na(d_trawl2$meanSST), ]

d_trawl2 <- mutate(d_trawl2,
  DOY = lubridate::yday(date),
  DOY_scaled = arm::rescale(DOY),
  DOY_scaled2 = DOY_scaled^2,
  depth_scaled = arm::rescale(log(depth)),
  SST_scaled = arm::rescale(meanSST),
  depth_scaled3 = depth_scaled^3,
  shallow = ifelse(depth>35,0,1),
  deep = ifelse(depth<35,0,1)
)

d_trawl2[is.na(d_trawl2$shallow), ]
d_trawl2[is.na(d_trawl2$SST_scaled), ]

# glimpse(d_trawl2)
```

Explore SST on survey day
```{r eval=FALSE}
ggplot(d_trawl2, aes_string("X", "Y", colour = "SST")) +
    geom_point() +
    facet_wrap(~year) +
    coord_fixed() +
    scale_color_viridis_c()
```

Is bottom temp related to SST in shallower samples?
```{r eval=FALSE}
ggplot(dplyr::filter(d_trawl2, depth <35), aes(SST,temperature_c, colour = -depth)) + geom_point()
ggplot(dplyr::filter(d_trawl2, depth <35), aes(SST,temperature_c, colour = DOY)) + geom_point()
```

Explore mean SST 
```{r eval=FALSE}
ggplot(d_trawl2, aes_string("X", "Y", colour = "meanSST")) +
    geom_point() +
    facet_wrap(~year) +
    coord_fixed() +
    scale_color_viridis_c()
```

```{r eval=FALSE}
ggplot(dplyr::filter(d_trawl2, shallow==1), aes(SST_scaled,temperature_c, colour = -depth)) + geom_point()
```

Does a fixed interaction with SST_scaled do any better?
```{r}
spde2 <- make_spde(d_trawl2$X, d_trawl2$Y, n_knots = 500)
plot_spde(spde2)
```

```{r eval=FALSE}
m_temp3 <- sdmTMB(
  temperature_c ~ 0 + as.factor(year) + SST_scaled * shallow + DOY_scaled * shallow + DOY_scaled2 * shallow,
  data = d_trawl2,
  time_varying = ~ 0 + depth_scaled + depth_scaled2,
  ar1_fields = TRUE, include_spatial = TRUE,
  time = "year", spde = spde2, family = gaussian(link = "identity"),
  silent = TRUE
)
```

```{r eval=FALSE}
predictions3 <- predict(m_temp3)
predictions3$residuals <- residuals(m_temp3)
cor(predictions3$est, predictions3$temperature_c)^2 #
```

## Plotting predictions from saved temperature model

```{r }
m_temp_rw <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-500kn.rds")))
nd_all <- readRDS(here::here("analysis/VOCC/data/nd_all_synoptic.rds")) %>% filter(year<2019)

unique(nd_all$year)
glimpse(nd_all)
pred_temp <- predict(m_temp_rw, newdata = nd_all)

# saveRDS(pred_temp, file = "../VOCC/data/predicted_temp_allyears_revised2")
```

```{r}
p_nd <- readRDS("../VOCC/data/predicted_temp_allyears_revised2")

plot_map_raster <- function(dat, column = "est") {
    ggplot(dat, aes_string("X", "Y", fill = column)) +
    geom_raster() +
    coord_fixed()+
gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank())
}

plot_map_raster(p_nd, "est") +
  scale_fill_viridis_c(trans = sqrt) +
  facet_wrap(~year) +
  ggtitle("Prediction (fixed effects + all random effects)")
```

```{r}
plot_facet_map(p_nd, "est", viridis_option = "D",
    raster_limits = c(min(p_nd$est), quantile(p_nd$est, 0.99)),
    transform_col = sqrt
  )
```

```{r}
p_nd1 <- p_nd %>% filter(ssid!=16) %>% filter(ssid!=4)

breaks <- scales::trans_breaks(no_trans[["transform"]], no_trans[["inverse"]],n = 6)
labels <- function(x) format(x, digits = 2, scientific = FALSE)

pred_1 <- plot_map_raster(p_nd1, "est") +
  scale_fill_viridis_c(labels = labels, 
    option = "C",
    limits = c(4, 11),
    na.value = "red") +
  facet_wrap(~year) +
  labs(fill = "°C") + 
  ggtitle("Prediction (fixed effects + all random effects)") 
pred_1
```

```{r}
p_nd4 <- p_nd %>% filter(ssid==4)
pred_4 <- plot_map_raster(p_nd4, "est") +
  scale_fill_viridis_c(labels = labels, 
    option = "C",
    limits = c(5, 11), 
    na.value = "black")+
  facet_wrap(~year) + 
  theme(legend.position = c(.85,.12)) +
  labs(fill = "°C")
pred_4
```

```{r}
plot_map_raster(p_nd, "est_non_rf") +
  ggtitle("Prediction (fixed effects only)") +
  facet_wrap(~year) +
  scale_fill_viridis_c()
```

```{r}
plot_map_raster(p_nd, "est_rf") +
  ggtitle("Prediction (all random effects only)") +
  facet_wrap(~year) +
  scale_fill_viridis_c()
```

```{r}
omega_s_all <- plot_map_raster(p_nd, "omega_s") +
  ggtitle("Spatial random effects only") +
    scale_fill_gradient2( low = "Steel Blue 4",
 high = "Red 3", limits = c(-1.5,1.5),
    na.value = "red") +  labs(fill = "") + 
gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = c(.82,.75))

omega_s_all
```

```{r}
 plot_map_raster(p_nd, "epsilon_st") +
  ggtitle("Spatiotemporal random effects only") +
  facet_wrap(~year) +   scale_fill_gradient2( low = "Steel Blue 4",
 high = "Red 3", limits = c(-1.5,1.5),
    na.value = "red")
```

```{r}
p_nd1 <- p_nd %>% filter(ssid!=16) %>% filter(ssid!=4)

(st_1n3 <- plot_map_raster(p_nd1, "epsilon_st") +
  ggtitle("Spatiotemporal random effects") +
  facet_wrap(~year) +
   scale_fill_gradient2( low = "Steel Blue 4",
 high = "Red 3", limits = c(-1.5,1.5), trans = sqrt,
    na.value = "red") +
  labs(fill = "epsilon") + 
  gfplot::theme_pbs() + 
  theme(
  legend.position = "none", 
  axis.title.x = element_blank(), axis.title.y = element_blank()))

# png(
#   file = "figs/temp-predictions-1n3.png", 
#   res = 400,
#   units = "in",
#   width = 8, 
#   height = 8
# ) 
# st_1n3
# dev.off()

```

```{r}
p_nd4 <- p_nd %>% filter(ssid==4) #%>% filter(ssid!=4)
st_4 <- plot_map_raster(p_nd4, "epsilon_st") +
  #ggtitle(" ") +
  facet_wrap(~year) +
     scale_fill_gradient2( low = "Steel Blue 4",
 high = "Red 3", limits = c(-1.5,1.5), trans = sqrt,
    na.value = "red") +
          labs(fill = " ") + 
gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = c(.85,.15))
st_4
```

```{r}
p_nd16 <- p_nd %>% filter(ssid==16) #%>% filter(ssid!=4)
st_16 <- plot_map_raster(p_nd16, "epsilon_st") +
  #ggtitle(" ") +
  facet_wrap(~year) +
   scale_fill_gradient2( low = "Steel Blue 4",
 high = "Red 3", limits = c(-1.5,1.5),
    na.value = "red") +
  xlim(65,345) +
    ylim(5800,6100) +
gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = "none")
st_16
```

```{r, eval= F}
png(
  file = "figs/temp-predictions.png", 
  res = 600,
  units = "in",
  width = 11, 
  height = 10
)

gridExtra::grid.arrange(
  grobs = list(pred_1, st_1n3, omega_s_all, pred_4, st_4, st_16 ),
  widths = c(2, 2, 1.75),
  heights = c(2, 1.35),
  layout_matrix = rbind(c(1, 2, 3),
                        c(4, 5, 6)),
  top = grid::textGrob("Bottom temperature estimates from sdmTMB model")
)

# gridExtra::grid.arrange( nrow = 2, heights = c(2,1.5), 
#   top = grid::textGrob("Dissolved oxygen estimates from sdmTMB model")) 
dev.off()
```

BOOSTED REGRESSION MODEL
```{r, eval = FALSE}
library(gbm)

m <- gbm( temperature_c ~ depth_scaled + X + Y, data = d_trawl, n.trees = 2000, interaction.depth = 3, shrinkage = 0.02)
m2 <- gbm( temperature_c ~ SST_scaled + depth_scaled + X + Y, data = d_trawl, n.trees = 2000, interaction.depth = 3, shrinkage = 0.02)

plot(m,i.var=2)
plot(m,i.var=3)
plot(m,i.var=4)
plot(m,i.var=5)

plot(m2,i.var=1)
plot(m2,i.var=2)
plot(m,i.var=3)
plot(m,i.var=4)

plot(m2,i.var=1:2)
plot(m2,i.var=c(1,3))
plot(m2,i.var=c(1,4))
d_trawl$r <- predict(m, n.trees = 2000) - d_trawl1$temperature_c

ggplot(d_trawl, aes(depth_scaled, r)) +
  geom_point(alpha=0.4) +
  ylim(-8,8) +
  geom_smooth()
```

CENTER OF GRAVITY
```{r, eval = F}
p_cog <- predict(m_temp_rw, newdata = nd_all, return_tmb_object = TRUE)

temp_cog <- get_cog(p_cog)

temp_cog %>%
  reshape2::dcast(year ~ coord, value.var = "est") %>%
  ggplot(aes(X, Y, colour = year)) + geom_path(arrow = arrow()) +
  scale_color_viridis_c()
```

```{r}
temp_cog %>%
  ggplot(aes(year, est, ymin = lwr, ymax = upr)) +
  geom_line() +
  geom_ribbon(alpha = 0.2) +
  facet_wrap(~coord, scales = "free_y")
```