### 2020-06-24 analysis/VOCC/4-fishing-trend.Rmd
recalculate fishing trend and add depth to 4x4 grids

---
title: "Fishing trends"
output: html_notebook
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(
  fig.keep = "all", fig.show = "asis", dpi = 600,
  fig.width = 6, fig.height = 5,
  echo = FALSE, warning = FALSE, message = FALSE
)
options(scipen = 999)
options(digits = 4)
```

```{r pkgs, cache=FALSE, warning=FALSE, message=FALSE}
library("gfranges")
library("dplyr")
library("ggplot2")

ggplot2::theme_set(gfplot::theme_pbs())
```

#### Set cell sizes and scale
```{r, echo=TRUE}
scale_fac <- 2 # means that the raster is reprojected to _ X original grid (2 km)
```

```{r}
d_all <- readRDS(paste0(
    "data/sablefish/predictions-sablefish-tv-depth-only-1n3n4n16-mature-biomass-prior-FALSE.rds"
  )) %>% filter(year >= 2008)

all_years <- unique(d_all$year)
length(all_years)
grid <- select(d_all, X,Y, depth, ssid) %>% unique()

grid_by_year <- data.frame(matrix(ncol=0, nrow= nrow(grid) * length(all_years)))
grid_by_year$ssid <- rep(grid$ssid, times = length(all_years))
grid_by_year$X <- rep(grid$X, times = length(all_years))
grid_by_year$Y <- rep(grid$Y, times = length(all_years))
grid_by_year$depth <- rep(grid$depth, times = length(all_years))
grid_by_year$year <- rep(all_years, each = nrow(grid))
```

Calculate mean and trend in log fishing pressure
```{r }
fishing <- readRDS("data/_fishing_effort/fishing-effort-grid.rds")

d <- left_join(grid_by_year, fishing, by = c("X", "Y", "year"))

d$log_effort[is.na(d$log_effort)] <- -2.3
d$effort[is.na(d$effort)] <- 0

### BOTH ODD YEAR ###
d1n3 <- d %>% filter(ssid %in% c(1,3))

ssid1n3 <- vocc_gradient_calc(d1n3, "log_effort",
  scale_fac = scale_fac,
  quantile_cutoff = 0.05
)

depth1n3 <- vocc_gradient_calc(d1n3, "depth",
  scale_fac = scale_fac,
  quantile_cutoff = 0.05
)
ssid1n3$depth <- depth1n3$mean

d4 <- d %>% filter(ssid %in% c(4))

### WCVI ###
ssid4 <- vocc_gradient_calc(d4, "log_effort",
  scale_fac = scale_fac,
  quantile_cutoff = 0.05
)
depth4 <- vocc_gradient_calc(d4, "depth",
  scale_fac = scale_fac,
  quantile_cutoff = 0.05
)
ssid4$depth <- depth4$mean

### WCHG ###
d16 <- d %>% filter(ssid %in% c(16))

ssid16 <- vocc_gradient_calc(d16, "log_effort",
  scale_fac = scale_fac,
  quantile_cutoff = 0.05
)
depth16 <- vocc_gradient_calc(d16, "depth",
  scale_fac = scale_fac,
  quantile_cutoff = 0.05
)
ssid16$depth <- depth16$mean

fished <- rbind(ssid1n3, ssid4, ssid16)

fished <- fished %>%
  mutate(
    fishing_trend = trend,
    mean_effort = exp(mean),
    fishing_vel = velocity,
    fishing_grad = gradient
  ) %>%
  dplyr::mutate(X = x, Y = y) %>%
  gfplot:::utm2ll(., utm_zone = 9)

# fished$squashed_effort <- collapse_outliers(fished$mean_effort, c(0, 0.99))
# hist(fished$squashed_effort)

# fished <- left_join(fished, grid)
fished <- fished %>%
  select(X, Y, x, y, fishing_trend, mean_effort, fishing_vel, fishing_grad, depth)

saveRDS(fished, file = paste0("data/fishing-effort-change-w-depth.rds"))
```