### 2020-11-14 analysis/tmb-sensor-explore/2-fit-temperature.Rmd

---
output: html_document
editor_options: 
  chunk_output_type: inline
---
# Fit spatial temporal models to bottom temperature

```{r setup}
library(tidyverse)
library(sdmTMB)
# library(future)
library(dplyr)
library(gfranges)

plot_map_raster <- function(dat, column = "est") {
    ggplot(dat, aes_string("X", "Y", fill = column)) +
    geom_raster() +
    coord_fixed()+
gfplot::theme_pbs() + theme(
  axis.title = element_blank(), 
  axis.ticks = element_blank(),
  axis.text = element_blank()
  )
}
```

Model diagnostics function
```{r}
get_diag <- function(m, response = "temperature_c", variable = "depth_scaled", colour_var = "depth", start_year = 2007, ssids = c(1,3,4,16)) {
  predictions <- predict(m)
  predictions$residuals <- residuals(m)

  predictions <- predictions %>% filter (year >= start_year)

  plot_map <- function(dat, column = "est") {
    ggplot(dat, aes_string("X", "Y", colour = column)) +
      geom_point() +
      coord_fixed()+
      gfplot::theme_pbs() 
  }

  g <- plot_map(predictions, "est") +
    scale_colour_viridis_c() +
    ggtitle("Prediction (fixed effects + all random effects)")
  print(g)

  g <- plot_map(predictions, "est_non_rf") +
    ggtitle("Prediction (fixed effects only)") +
    scale_colour_viridis_c()
  print(g)

  g <- plot_map(predictions, "est_rf") +
    ggtitle("All random effects only") +
    scale_colour_gradient2()
  print(g)

  g <- plot_map(predictions, "omega_s") +
    ggtitle("Spatial random effects only") +
    scale_colour_gradient2()
  print(g)

  g <- plot_map(predictions, "epsilon_st") +
    ggtitle("Spatiotemporal random effects only") +
    facet_wrap(~year) +
    scale_colour_gradient2()
  print(g)

  g <- ggplot(predictions, aes_string("est", response)) +
    geom_point(alpha = 0.2) +
    facet_wrap(~year) +
    coord_fixed() +
    geom_abline()+
    gfplot::theme_pbs() 
  print(g)

  g <- ggplot(predictions, aes(est, residuals)) +
    geom_point(alpha = 0.2) +
    geom_smooth()+
    gfplot::theme_pbs() 
  print(g)

  g <- ggplot(predictions, aes_string(variable, "residuals", colour = colour_var)) +
    geom_point(alpha = 0.4, size = 1.2) +
    geom_smooth(colour="grey", size = 1.2) +
    scale_colour_viridis_c(option = "B", direction = -1, begin = 0, end = 0.7, 
      limits= c(min(predictions[colour_var]), max(predictions[colour_var]))) + #, trans= sqrt
    # facet_wrap(~year, scales = "free_x")+
    gfplot::theme_pbs() + theme(
      axis.text.y = element_text(size = 14),
      axis.text.x = element_blank(), axis.ticks.x = element_blank())
  print(g)

  g <- ggplot(predictions, aes_string(variable, "residuals", colour = colour_var)) +
    geom_point(alpha = 0.4, size = 0.7) +
    geom_smooth(colour="grey", size = 0.7) +
    scale_colour_viridis_c(option = "B", direction = -1, begin = 0, end = 0.7, 
      limits= c(min(predictions[colour_var]), max(predictions[colour_var]))) + #, trans= sqrt
    facet_wrap(~year, scales = "free_x")+
    gfplot::theme_pbs() + theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
  print(g)

    g <- ggplot(filter(predictions, ssid %in% ssids), aes_string(variable, "residuals", colour = colour_var)) +
    geom_point(alpha = 0.4, size = 1) +
    geom_smooth(colour="grey", size = 1) +
    scale_colour_viridis_c(option = "B", direction = -1, begin = 0, end = 0.7, 
      limits= c(min(predictions[colour_var]), max(predictions[colour_var]))) + #, trans= sqrt
    facet_wrap(~ssid, scales = "free_x")+
    gfplot::theme_pbs() 
  print(g)

  g <- ggplot(predictions, aes(X, Y, colour = residuals)) +
    geom_point(alpha = 0.2) +
    coord_fixed() +
    scale_color_gradient2() +
    facet_wrap(~year)+
    gfplot::theme_pbs() 
  print(g)

  aic <- function(m) {
    k <- length(m$model$par)
    nll <- m$model$objective
    2 * k - 2 * (-nll)
  }

  print("R^2:")
  r2 <- cor(predictions$est, predictions[[response]])^2
  print(r2)

  print("")
  print("AIC:")
  print(aic(m))
}
```

Load data if not using SST
```{r}
all_sensor <- readRDS(here::here("analysis/tmb-sensor-explore/data/dat-sensor-trawl-processed.rds")) # no SST

all_depth <- all_sensor %>%
  # filter(depth_max>10) %>% # maybe we should filter incase missing depth values also indicate a sensor fail?
  # filter missing location data and trial year
  dplyr::filter(!is.na(latitude), !is.na(longitude)) %>%

  # convert lat and lon to UTMs
  dplyr::mutate(X = longitude, Y = latitude) %>%
  gfplot:::ll2utm(., utm_zone = 9) %>%

  # interpolate missing depth values
  dplyr::rename(depth = depth_m) %>%
  gfplot:::interp_survey_bathymetry()

d_trawl <- all_depth$data %>% 
  # remove obvious sensor fails, d_trawl %>% filter(depth_max<10), fishing event ids = 481861, 2179089
  filter(depth_max>10) %>% 
  gfplot:::scale_survey_predictors()

d_trawl <- d_trawl[!is.na(d_trawl$depth), ]
d_trawl <- d_trawl[!is.na(d_trawl$temperature_c), ]

d_trawl <- mutate(d_trawl,
  DOY = lubridate::yday(date),
  DOY_scaled = arm::rescale(lubridate::yday(date)),
  DOY_scaled2 = arm::rescale(lubridate::yday(date))^2)
```

EXLORE TEMPERATURE at DEPTH
```{r}
ggplot(d_trawl, aes_string("X", "Y", colour = "temperature_c")) +
    geom_point() +
    facet_wrap(~year) +
    coord_fixed() +
    scale_color_viridis_c()
```

```{r}
ggplot(d_trawl, aes(DOY, temperature_c, colour = -depth, alpha=0.5)) + geom_point() + facet_wrap(~year) 
```

Current temperature model
```{r}
spde <- sdmTMB::make_mesh(d_trawl, c("X", "Y"), n_knots = 800, type = "kmeans")
plot(spde)
```
Original model code
```{r eval = F}
# spde <- make_spde(d_trawl$X, d_trawl$Y, n_knots = 800)
# m_temp_rw <- sdmTMB(
#   temperature_c ~ 0 + as.factor(year), #+ DOY_scaled + as.factor(ssid)
#   time_varying = ~ 0 + depth_scaled + depth_scaled2,
#   data = d_trawl,
#   ar1_fields = TRUE, # changed AR1 to true
#   include_spatial = TRUE, 
#   time = "year", spde = spde, 
#   family = gaussian(link = "identity"),
#   nlminb_loops = 2,
#   newton_steps = 1,
#   silent = FALSE
# )
# saveRDS(m_temp_rw, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-800kn.rds")))
```

New CV model code
```{r eval = F}
m_temp_cv <- sdmTMB_cv(
  temperature_c ~ 0 + as.factor(year), #+ DOY_scaled + as.factor(ssid)
  time_varying = ~ 0 + depth_scaled + depth_scaled2,
  data = d_trawl,
  ar1_fields = TRUE, # changed AR1 to true
  include_spatial = TRUE, 
  time = "year", spde = spde, 
  family = gaussian(link = "identity"),
  k_folds = 5,
  # nlminb_loops = 2,
  # newton_steps = 1,
  silent = FALSE
)
saveRDS(m_temp_cv, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-800kn-cv.rds")))

m_temp_cv <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-800kn-cv.rds")))
m_temp_cv$fold_loglik
m_temp_cv$sum_loglik
m_temp_cv$max_gradients
```

Check predictions against raw data
```{r}
# m_temp_rw <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-500kn-2.rds"))
m_temp_rw <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-800kn.rds"))
predictions <- predict(m_temp_rw)
predictions$residuals <- residuals(m_temp_rw)
cor(predictions$est, predictions$temperature_c)^2 
sum(abs(predictions$residuals))

# 2908.879-2789.322
# 119.557/2908.879 # 4% reduction in abs errors
```

```{r eval=F}
# test against ROMS
events_roms <- readRDS("data/roms_temp_by_events2.rds") %>% select(fishing_event_id, roms, year)

predictions <- left_join(predictions, events_roms)

cor(predictions$roms, predictions$est, use = "pairwise.complete.obs")
cor(predictions$roms, predictions$temperature_c, use = "pairwise.complete.obs")

ggplot(predictions, aes(roms, est, colour = -depth)) + geom_point(alpha=0.3) + 
  facet_wrap(~year) + coord_cartesian(xlim = c(2,13), ylim = c(2,13)) +
  geom_abline(intercept = 0, slope = 1) +
  gfplot::theme_pbs()

ggplot(predictions, aes(roms, temperature_c, colour = DOY)) + geom_point(alpha=0.3) +
  facet_wrap(~year) + coord_cartesian(xlim = c(2,13), ylim = c(2,13)) +
  geom_abline(intercept = 0, slope = 1) +
  scale_colour_viridis_c() +
  gfplot::theme_pbs()

ggplot(predictions, aes(DOY, (roms-temperature_c), colour = -depth)) + geom_point(alpha=0.3) +
  facet_wrap(~year) + #coord_cartesian(xlim = c(2,13), ylim = c(2,13)) +
  geom_abline(intercept = 0, slope = 0) +
  scale_colour_viridis_c() +
  gfplot::theme_pbs()

ggplot(predictions, aes(DOY, (est-roms), colour = -depth)) + geom_point(alpha=0.3) +
  facet_wrap(~year) + #coord_cartesian(xlim = c(2,13), ylim = c(2,13)) +
  geom_abline(intercept = 0, slope = 0) +
  scale_colour_viridis_c() +
  gfplot::theme_pbs()

p_nd <- readRDS((here::here("analysis/VOCC/data/predicted-DO-2020-06-20-more2016-roms.rds")))

cor(p_nd$roms_temp, p_nd$temp, use = "pairwise.complete.obs")

ggplot(p_nd, aes(roms_temp, temp, colour = -depth)) + geom_point(alpha=0.04) + 
  facet_wrap(~year) + coord_cartesian(xlim = c(2,13), ylim = c(2,13)) +
  geom_abline(intercept = 0, slope = 1) +
  gfplot::theme_pbs()

```

```{r eval=FALSE}
m_temp_rw <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-500kn-2.rds")))
get_diag(m_temp_rw, response = "temperature_c", variable = "depth_scaled")
```

WOULD SST HELP?

Load and prep SST data
```{r eval=FALSE}
# all_sensor <- readRDS(here::here("analysis/tmb-sensor-explore/data/dat-sensor-trawl-meanSST-all.rds"))
# # all_sensor <- readRDS(here::here("analysis/tmb-sensor-explore/data/dat-sensor-trawl-processed.rds")) # no SST
# # glimpse(all_sensor)
# 
# all_depth <- all_sensor %>%
#   # filter(depth_max>10) %>% # maybe we should filter incase missing depth values also indicate a sensor fail?
#   # filter missing location data and trial year
#   dplyr::filter(!is.na(latitude), !is.na(longitude)) %>%
# 
#   # convert lat and lon to UTMs
#   dplyr::mutate(X = longitude, Y = latitude) %>%
#   gfplot:::ll2utm(., utm_zone = 9) %>% 
# 
#   # interpolate missing depth values
#   dplyr::rename(depth = depth_m) %>%
#   gfplot:::interp_survey_bathymetry()

# d_trawl <- all_depth$data %>% 
#   # remove obvious sensor fails, d_trawl %>% filter(depth_max<10), fishing event ids = 481861, 2179089
#   filter(depth_max>10) %>% 
#   gfplot:::scale_survey_predictors()

olddata <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-400kn.rds"))) 
d_trawl <- olddata$data

d_trawl <- d_trawl[!is.na(d_trawl$depth), ]
d_trawl <- d_trawl[!is.na(d_trawl$temperature_c), ]

d_trawl2 <- d_trawl[!is.na(d_trawl$SST), ]
d_trawl2 <- d_trawl2[!is.na(d_trawl2$meanSST), ]

d_trawl2 <- mutate(d_trawl2,
  DOY = lubridate::yday(date),
  DOY_scaled = arm::rescale(DOY),
  DOY_scaled2 = DOY_scaled^2,
  depth_scaled = arm::rescale(log(depth)),
  SST_scaled = arm::rescale(meanSST),
  depth_scaled3 = depth_scaled^3,
  shallow = ifelse(depth>35,0,1),
  deep = ifelse(depth<35,0,1)
)

# d_trawl2[is.na(d_trawl2$shallow), ]
# d_trawl2[is.na(d_trawl2$SST_scaled), ]
```

Explore SST on survey day
```{r eval=FALSE}
ggplot(d_trawl2, aes_string("X", "Y", colour = "SST")) +
    geom_point() +
    facet_wrap(~year) +
    coord_fixed() +
    scale_color_viridis_c()
```

Is bottom temp related to SST in shallower samples?
```{r eval=FALSE}
ggplot(dplyr::filter(d_trawl2, depth <35), aes(SST,temperature_c, colour = -depth)) + geom_point()
ggplot(dplyr::filter(d_trawl2, depth <35), aes(SST,temperature_c, colour = DOY)) + geom_point()
```

Explore mean SST 
```{r eval=FALSE}
ggplot(d_trawl2, aes_string("X", "Y", colour = "meanSST")) +
    geom_point() +
    facet_wrap(~year) +
    coord_fixed() +
    scale_color_viridis_c()
```

```{r eval=FALSE}
ggplot(dplyr::filter(d_trawl2, shallow==1), aes(SST_scaled,temperature_c, colour = -depth)) + geom_point()
```

Does a fixed interaction with SST_scaled do any better?
Rerun main model with SST dataset for comparisons
```{r}
spde <- make_mesh(d_trawl2, c("X", "Y"), n_knots = 800)

m_temp_new <- sdmTMB(
  temperature_c ~ 0 + as.factor(year), 
  time_varying = ~ 0 + depth_scaled + depth_scaled2,
  data = d_trawl2, spde = spde,
  ar1_fields = TRUE, 
  include_spatial = TRUE,
  time = "year", 
  family = gaussian(link = "identity"),
  nlminb_loops = 2,
  newton_steps = 1,
  silent = FALSE
)
saveRDS(m_temp_new, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-800kn-sstdata.rds")))
m_temp_new
AIC(m_temp_new)
```

```{r}
spde <- make_mesh(d_trawl2, c("X", "Y"), n_knots = 800)

m_temp_sst <- sdmTMB(
  temperature_c ~ 0 + as.factor(year) + SST_scaled * shallow, 
  data = d_trawl2, spde = spde,
  time_varying = ~ 0 + depth_scaled + depth_scaled2,
  ar1_fields = TRUE, 
  include_spatial = TRUE,
  time = "year", 
  family = gaussian(link = "identity"),
  nlminb_loops = 2,
  newton_steps = 1,
  silent = TRUE
)

saveRDS(m_temp_sst, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-800kn-SST.rds")))
m_temp_sst
AIC(m_temp_sst)
```

```{r}
m_temp_new <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-800kn-sstdata.rds")))
m_temp_sst <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-800kn-SST.rds")))
get_diag(m_temp_new, response = "temperature_c")
get_diag(m_temp_sst, response = "temperature_c")
```

# try with fewer knots
```{r eval = F}
spde3 <- sdmTMB::make_mesh(d_trawl2, c("X", "Y"), n_knots = 400, type = "kmeans")

m_temp1aic <- sdmTMB(
  temperature_c ~ 0 + as.factor(year),
  time_varying = ~ 0 + depth_scaled + depth_scaled2,
  data = d_trawl2, spde = spde3,
  ar1_fields = TRUE, 
  include_spatial = TRUE,
  time = "year", 
  family = gaussian(link = "identity"),
  nlminb_loops = 2,
  newton_steps = 1,
  silent = FALSE
)
saveRDS(m_temp1aic, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-400kn-new.rds")))

m_temp1aic <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-400kn-new.rds")))
# m_temp1aic<- sdmTMB:::update_model(m_temp1aic)
AIC(m_temp1aic)
```

Run CV on main model
```{r}
# spde2 <- make_spde(d_trawl2$X, d_trawl2$Y, n_knots = 800)
spde2 <- sdmTMB::make_mesh(d_trawl2, c("X", "Y"), n_knots = 200, type = "kmeans")
plot(spde2)
```

```{r}
m_temp1 <- sdmTMB_cv(
  temperature_c ~ 0 + as.factor(year), 
  time_varying = ~ 0 + depth_scaled + depth_scaled2,
  data = d_trawl2, spde = spde3, 
  ar1_fields = TRUE, 
  include_spatial = TRUE, 
  time = "year", 
  family = gaussian(link = "identity"),
  k_folds = 20,
  # nlminb_loops = 2,
  # newton_steps = 1,
  silent = FALSE
)
# saveRDS(m_temp1, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-400kn-cv20.rds")))
saveRDS(m_temp1, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-150kn-cv20.rds")))

# m_temp1 <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-300kn-cv.rds")))
m_temp1$fold_loglik
m_temp1$sum_loglik
m_temp1$max_gradients
m_temp1$pdHess
```

Check if DOY helps model
```{r eval=FALSE}
spde3 <- sdmTMB::make_mesh(d_trawl2, c("X", "Y"), n_knots = 400, type = "kmeans")

m_temp2aic <- sdmTMB(
  temperature_c ~ 0 + as.factor(year) + DOY_scaled + DOY_scaled2,
  data = d_trawl2, spde = spde3, 
  time_varying = ~ 0 + depth_scaled + depth_scaled2,
  ar1_fields = TRUE, 
  include_spatial = TRUE,
  time = "year", 
  family = gaussian(link = "identity"),
  nlminb_loops = 2,
  newton_steps = 1,
  # k_folds = 20,
  silent = TRUE
)

saveRDS(m_temp2aic, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-DOY-400kn.rds")))
m_temp2aic <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-DOY-400kn.rds")))
m_temp2aic
AIC(m_temp2aic)

# m_temp2cv <- sdmTMB_cv(
#   temperature_c ~ 0 + as.factor(year) + DOY_scaled + DOY_scaled2,
#   data = d_trawl2, spde = spde2, 
#   time_varying = ~ 0 + depth_scaled + depth_scaled2,
#   ar1_fields = TRUE, 
#   include_spatial = TRUE,
#   time = "year", 
#   family = gaussian(link = "identity"),
#   # nlminb_loops = 2,
#   # newton_steps = 1,
#   k_folds = 20,
#   silent = TRUE
# )
# 
# saveRDS(m_temp2, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-DOY-cv20.rds")))
# m_temp2cv <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-DOY-cv20.rds")))
# m_temp2cv
# m_temp2cv$fold_loglik
# m_temp2cv$sum_loglik
# m_temp2cv$max_gradients
```

Check if SST helps model 
```{r}
m_temp3aic <- sdmTMB(
  temperature_c ~ 0 + as.factor(year) + SST_scaled * shallow, 
  data = d_trawl2, spde = spde3, 
  time_varying = ~ 0 + depth_scaled + depth_scaled2,
  ar1_fields = TRUE, 
  include_spatial = TRUE,
  time = "year", 
  family = gaussian(link = "identity"),
  nlminb_loops = 2,
  newton_steps = 1,
  silent = TRUE
)
saveRDS(m_temp3aic, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-SST-400kn.rds")))
# m_temp3aic <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-SST-400kn.rds"))
m_temp3aic <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-SST-400kn.rds")))
AIC(m_temp1aic, m_temp2aic, m_temp3aic)

m_temp3aic
```

Check if SST and DOY help model
```{r eval=FALSE}
m_temp4aic <- sdmTMB(
  temperature_c ~ 0 + as.factor(year) + SST_scaled * DOY_scaled * shallow, 
  data = d_trawl2, spde = spde3, 
  time_varying = ~ 0 + depth_scaled + depth_scaled2,
  ar1_fields = TRUE, 
  include_spatial = TRUE,
  time = "year", 
  family = gaussian(link = "identity"),
  nlminb_loops = 2,
  newton_steps = 1,
  silent = TRUE
)
saveRDS(m_temp4aic, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-SST-DOY-400kn.rds")))

m_temp4aic <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-SST-DOY-400kn.rds"))
# m_temp4aic
AIC(m_temp1aic, m_temp2aic,  m_temp3aic, m_temp4aic)
```

```{r eval=FALSE}
m_temp3 <- sdmTMB_cv(
  temperature_c ~ 0 + as.factor(year) + SST_scaled * shallow, 
  data = d_trawl2, spde = spde2, 
  time_varying = ~ 0 + depth_scaled + depth_scaled2,
  ar1_fields = TRUE, 
  include_spatial = TRUE,
  time = "year", 
  family = gaussian(link = "identity"),
  # nlminb_loops = 2,
  # newton_steps = 1,
  k_folds = 20,
  silent = TRUE
)

# 500 knot early version
# saveRDS(m_temp3, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-SST.rds")))

# new 800 knot version for cross validation
saveRDS(m_temp3, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-SST-cv20.rds")))

# m_temp3 <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-SST-cv20.rds"))
# m_temp3
m_temp3$fold_loglik
m_temp3$sum_loglik
m_temp3$max_gradients
```

Check if SST and DOY help model using cv
```{r eval=FALSE}
m_temp4 <- sdmTMB_cv(
  temperature_c ~ 0 + as.factor(year) + SST_scaled * shallow + SST_scaled * DOY_scaled * shallow, 
  data = d_trawl2, spde = spde2, 
  time_varying = ~ 0 + depth_scaled + depth_scaled2,
  ar1_fields = TRUE, 
  include_spatial = TRUE,
  time = "year", 
  family = gaussian(link = "identity"),
  # nlminb_loops = 2,
  # newton_steps = 1,
  k_folds = 20,
  silent = TRUE
)

# 500 knot early version
# saveRDS(m_temp3, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-SST.rds")))

# new 800 knot version for cross validation
saveRDS(m_temp4, file = (here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-SST-DOY-cv20.rds")))

# m_temp4 <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-SST-DOY-cv20.rds"))
# m_temp4
m_temp4$fold_loglik
m_temp4$sum_loglik
m_temp4$max_gradients
m_temp4$pdHess
```

```{r}
m_temp_sst <- readRDS(here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-SST.rds"))

m_temp_sst <- sdmTMB:::update_model(m_temp_sst)

```

```{r eval=FALSE}
predictions3 <- predict(m_temp3)
predictions3$residuals <- residuals(m_temp3)
cor(predictions3$est, predictions3$temperature_c)^2 #
```

```{r eval=FALSE}
get_diag(m_temp_new, response = "temperature_c", variable = "depth_scaled", colour_var = "depth", start_year = 2002)
get_diag(m_temp_new, response = "temperature_c", variable = "DOY", colour_var = "depth", start_year = 2002, ssids = c(1,3,4, 16))
```

## Plotting predictions from saved temperature model

```{r }
# m_temp_rw <- readRDS((here::here("analysis/tmb-sensor-explore/models/model-temp-all-years-500kn.rds")))
nd_all <- readRDS(here::here("analysis/VOCC/data/nd_all_synoptic.rds")) %>% filter(year<2019)

unique(nd_all$year)
glimpse(nd_all)
p_nd <- predict(m_temp_rw, newdata = nd_all)

# saveRDS(p_nd, file = "../VOCC/data/predicted_temp_allyears_800kn.rds")
```

```{r}
p_nd <- readRDS("../VOCC/data/predicted_temp_allyears_800kn.rds")

plot_map_raster(p_nd, "est") +
  scale_fill_viridis_c(trans = sqrt) +
  facet_wrap(~year) +
  ggtitle("Prediction (fixed effects + all random effects)")
```

```{r}
plot_facet_map(p_nd, "est", viridis_option = "D",
    raster_limits = c(min(p_nd$est), quantile(p_nd$est, 0.99)),
    transform_col = sqrt
  )
```

```{r}
p_nd1 <- p_nd %>% filter(ssid!=16) %>% filter(ssid!=4)

breaks <- scales::trans_breaks(no_trans[["transform"]], no_trans[["inverse"]],n = 6)
labels <- function(x) format(x, digits = 2, scientific = FALSE)

pred_1 <- plot_map_raster(p_nd1, "est") +
  scale_fill_viridis_c(labels = labels, 
    option = "C",
    limits = c(4, 11),
    na.value = "red") +
  facet_wrap(~year) +
  labs(fill = "°C") + 
  ggtitle("Prediction (fixed effects + all random effects)") 
pred_1
```

Time-varying depth 
```{r eval}
# pd <- expand.grid(
#   depth_scaled = seq(min(m_temp_rw$data$depth_scaled),
#     max(m_temp_rw$data$depth_scaled), length.out = 50),
#   year = unique(m_temp_rw$data$year)
# )
# pd$depth_scaled2 <- pd$depth_scaled^2
# 
# # p <- predict(m_temp_rw, newdata = pd, se_fit = TRUE, re_form = NA, xy_cols = c("X", "Y"))
# p <- predict(m_temp_new, newdata = pd, se_fit = TRUE, re_form = NA)
# 
# p$Depth <- exp(p$depth_scaled*0.661 + 5.07)
# p$Temperature <- p$est
# # 
# p$lowCI <- (p$est - 1.96 * p$est_se)
# p$highCI <- (p$est + 1.96 * p$est_se)
# # 
# saveRDS(p, here::here("analysis/tmb-sensor-explore/data/depth-tv-m-temp-rw-800-kn.rds"))
p <- readRDS(here::here("analysis/tmb-sensor-explore/data/depth-tv-m-temp-rw-800-kn.rds"))

ggplot(filter(p, Depth >15 & year >2007), aes(Depth, Temperature,
  ymin = lowCI, 
  ymax = highCI, 
  group = as.factor(year), fill=year, colour = year)) +
  geom_line(size = 0.5, alpha =0.85) +
  # geom_smooth(method= "loess", size = 0.4, alpha =0.85, se =F ) +
  # geom_smooth(span = 0.5, size = 0.4, alpha =0.85, se =F ) +
  scale_fill_viridis_c(option = "C") +
  scale_colour_viridis_c(option = "C") +
  # ylab("DO") +
  geom_ribbon(alpha = 0.2, colour = NA) +
  # scale_y_continuous(limits = c(0,4.5)) +
  scale_x_reverse(limits = c(600,20)) +
  ylim( 2, 15) +
  coord_flip() +
  # scale_x_continuous(limits = c(18,210)) +
  gfplot::theme_pbs() +
  theme(legend.position = c(.8,.4))
ggsave(filename = here::here("analysis/tmb-sensor-explore/figs/depth-tv-m-temp-rw-800-kn.png"), width = 3, height = 2.5)

# ggplot(filter(p, Depth >15 & year == 2018), aes(Depth, Temperature,
#   ymin = (Temperature - 1.96 * est_se), 
#   ymax = (Temperature + 1.96 * est_se))) +
#   geom_line(size = 1.5, alpha =0.85) + 
#   # scale_color_viridis_c(option = "C") +
#   geom_ribbon(alpha=0.2) +
#   scale_x_reverse(limits = c(210,20)) +
#   ylim(5, 15) +
#   coord_flip() +
#   gfplot::theme_pbs() + 
#   theme(legend.position = c(.8,.4))

```

```{r}
p_nd4 <- p_nd %>% filter(ssid==4)
pred_4 <- plot_map_raster(p_nd4, "est") +
  scale_fill_viridis_c(labels = labels, 
    option = "C",
    limits = c(5, 11), 
    na.value = "black")+
  facet_wrap(~year) + 
  theme(legend.position = c(.85,.12)) +
  labs(fill = "°C")
pred_4
```

```{r}
plot_map_raster(p_nd, "est_non_rf") +
  ggtitle("Prediction (fixed effects only)") +
  facet_wrap(~year) +
  scale_fill_viridis_c()
```

```{r}
plot_map_raster(p_nd, "est_rf") +
  ggtitle("Prediction (all random effects only)") +
  facet_wrap(~year) +
  scale_fill_viridis_c()
```

```{r}
(omega_s_all <- plot_map_raster(p_nd, "omega_s") +
  ggtitle("Spatial random effects only") +
    scale_fill_gradient2( low = "Steel Blue 4",
 high = "Red 3", limits = c(-1.5,1.5),
    na.value = "red") +  labs(fill = "") + theme(legend.position = c(0.2, 0.3))
  )

ggsave(filename = here::here("analysis/tmb-sensor-explore/figs/omegas-temp-rw-800-kn.png"), width = 3, height = 3)

omega_s_all

```

```{r}
 plot_map_raster(p_nd, "epsilon_st") +
  ggtitle("Spatiotemporal random effects only") +
  facet_wrap(~year) +   scale_fill_gradient2( low = "Steel Blue 4",
 high = "Red 3", limits = c(-1.5,1.5),
    na.value = "red")
```

```{r}
p_nd <- readRDS("../VOCC/data/predicted_temp_allyears_800kn.rds")

p_nd1 <- p_nd %>% filter(ssid!=16) %>% filter(ssid!=4) %>% filter(year>2007)

(st_1n3 <- plot_map_raster(p_nd1, "epsilon_st") +
  ggtitle("Spatiotemporal random effects") +
  facet_grid(~year) +
   scale_fill_gradient2( low = "Steel Blue 4",
 high = "Red 3", limits = c(-1.5,1.5), #trans = sqrt,
    na.value = "red") +
  labs(fill = "epsilon") + 
  theme(
  legend.position = "none"))

ggsave(filename = here::here("analysis/tmb-sensor-explore/figs/temp-epsilon-1n3.png"), width =6, height = 2)

# png(
#   file = "figs/temp-predictions-1n3.png", 
#   res = 400,
#   units = "in",
#   width = 8, 
#   height = 8
# ) 
# st_1n3
# dev.off()

```

```{r}
p_nd <- readRDS("../VOCC/data/predicted_temp_allyears_800kn.rds")

p_nd4 <- p_nd %>% filter(ssid==4) %>% filter(year>2007)#%>% filter(ssid!=4)
(st_4 <- plot_map_raster(p_nd4, "epsilon_st") +
  #ggtitle(" ") +
  facet_grid(~year) +
     scale_fill_gradient2( low = "Steel Blue 4",
 high = "Red 3", limits = c(-1.5,1.5), #trans = sqrt,
    na.value = "red") +
          labs(fill = " ") + theme( legend.position = "none") )
ggsave(filename = here::here("analysis/tmb-sensor-explore/figs/temp-epsilon-4.png"), width = 6, height = 2)

st_4
```

```{r}
p_nd <- readRDS("../VOCC/data/predicted_temp_allyears_800kn.rds")

p_nd16 <- p_nd %>% filter(ssid==16) %>% filter(year>2007)#%>% filter(ssid!=4)
(st_16 <- plot_map_raster(p_nd16, "epsilon_st") +
  #ggtitle(" ") +
  facet_grid(~year) +
   scale_fill_gradient2( low = "Steel Blue 4",
 high = "Red 3", limits = c(-1.5,1.5),
    na.value = "red") +
  xlim(20,360) + ylim(5850,6050)+ 
  # xlim(65,345) + ylim(5800,6100)+ 
    theme(panel.border = element_blank(), legend.position = "none") )

ggsave(filename = here::here("analysis/tmb-sensor-explore/figs/temp-epsilon-16.png"), width = 6, height = 1)
st_16
```

```{r, eval= F}
png(
  file = "figs/temp-predictions.png", 
  res = 600,
  units = "in",
  width = 11, 
  height = 10
)

gridExtra::grid.arrange(
  grobs = list(pred_1, st_1n3, omega_s_all, pred_4, st_4, st_16 ),
  widths = c(2, 2, 1.75),
  heights = c(2, 1.35),
  layout_matrix = rbind(c(1, 2, 3),
                        c(4, 5, 6)),
  top = grid::textGrob("Bottom temperature estimates from sdmTMB model")
)

# gridExtra::grid.arrange( nrow = 2, heights = c(2,1.5), 
#   top = grid::textGrob("Dissolved oxygen estimates from sdmTMB model")) 
dev.off()
```

BOOSTED REGRESSION MODEL
```{r, eval = FALSE}
library(gbm)

m <- gbm( temperature_c ~ depth_scaled + X + Y, data = d_trawl, n.trees = 2000, interaction.depth = 3, shrinkage = 0.02)
m2 <- gbm( temperature_c ~ SST_scaled + depth_scaled + X + Y, data = d_trawl, n.trees = 2000, interaction.depth = 3, shrinkage = 0.02)

plot(m,i.var=2)
plot(m,i.var=3)
plot(m,i.var=4)
plot(m,i.var=5)

plot(m2,i.var=1)
plot(m2,i.var=2)
plot(m,i.var=3)
plot(m,i.var=4)

plot(m2,i.var=1:2)
plot(m2,i.var=c(1,3))
plot(m2,i.var=c(1,4))
d_trawl$r <- predict(m, n.trees = 2000) - d_trawl1$temperature_c

ggplot(d_trawl, aes(depth_scaled, r)) +
  geom_point(alpha=0.4) +
  ylim(-8,8) +
  geom_smooth()
```

CENTER OF GRAVITY
```{r, eval = F}
p_cog <- predict(m_temp_rw, newdata = nd_all, return_tmb_object = TRUE)

temp_cog <- get_cog(p_cog)

temp_cog %>%
  reshape2::dcast(year ~ coord, value.var = "est") %>%
  ggplot(aes(X, Y, colour = year)) + geom_path(arrow = arrow()) +
  scale_color_viridis_c()
```

```{r}
temp_cog %>%
  ggplot(aes(year, est, ymin = lwr, ymax = upr)) +
  geom_line() +
  geom_ribbon(alpha = 0.2) +
  facet_wrap(~coord, scales = "free_y")
```